### 2022-04-21 docs/pap.Rmd
add pap

---
title: "Pre-Analysis Plan: Ideological Biases of the Crowd?"
date: "`r format(Sys.time(), '%d %B, %Y - %X (%Z)')`"
output: 
  pdf_document:
    toc: yes
    number_sections: true
editor_options: 
  chunk_output_type: console
bibliography: references.bib
urlcolor: blue
linkcolor: blue
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \floatsetup[table]{capposition=top}
   - \usepackage{booktabs}
   - \usepackage{amsmath}
   - \usepackage{longtable}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{threeparttable}
   - \usepackage{threeparttablex}
   - \usepackage[normalem]{ulem}
   - \usepackage{makecell}
   - \usepackage{xcolor}
   - \usepackage[utf8]{inputenc}
---
```{r setup, include=FALSE}
## include this at top of your RMarkdown file for pretty output
## make sure to have the printr package installed: install.packages('printr')
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
#library(printr)
library(tidyverse)
library(kableExtra)
options(kableExtra.latex.load_packages = FALSE)
```

\newpage

# Expectations

**Ideological Bias hypothesis** (_H1a_): The larger the ideological distance between respondent and the party, the less likely respondents code stances according to the party’s position.

**Ideological Bias hypothesis** (_H1b_): The effect of H1a is stronger for the underspecified condition.

**Ideological Overinterpretation hypothesis** (_H2a_):  The larger the ideological distance between respondent and the party, the more likely respondents interpret underspecified sentences as stance.

**Political Knowledge Overinterpretation hypothesis** (_H2b_): The more political knowledge, the more likely people interpret underspecified sentences as stance.

**Masking Solution hypothesis** (_H3a_):  Masking reduces the effect of respondents’ ideological position for coding stances according to the party’s position.

**Masking Solution hypothesis** (_H3b_): Masking reduces the effect of respondents’ level of political knowledge for coding stances according to the party’s position.

# Research Desing and Protocol

## Sample
We  will conduct this survey experiment in the Netherlands in April 2022.
The sample, recruited through [KiesKompas](https://www.kieskompas.nl/en/), will consist of 2,000 participants (based on the power analysis presented in Figure \ref{fig:power1}) of 18 years and older.
Kieskompas works with non-random opt-in respondents.
Therefore, I measure many demographic background variables (see [Section 3.2](#control-variables)).
Balance checks will be conducted to demonstrate whether certain categories are over represented in a certain experimental group.
The study has been approved by the [Research Ethics Review Committee](https://fsw.vu.nl/nl/onderzoek/research-ethics-review/index.aspx) of the _Vrije Universiteit Amsterdam_ (see the approval [here](https://github.com/MarikenvdVelden/compromise-punish/blob/main/docs/2022-3-30-59.pdf)).
To ensure good quality of our data, one attention check (discussed in more detail in [Section 3.3](#attention-checks)) is included.

## Experimental Protocol
The study is conducted online and in Dutch.
Participants are told that they are taking part in a survey to get an overview of how Dutch people form their views on politics.
After reading an informed consent message participants are forwarded to the main questionnaire (or the survey will be terminated if they do not agree to the consent form).

First, participants complete a set of pre-treatment variables (i.e. vote recall, issue positions of the issues used in the experiment (Climate, EU, Immigration, and Tax), ideology, and political knowledge).
This block ends with the attention check included in this survey.
When participants fail this attention check, a warning appears asking them to read the question again
carefully  and  to  answer  again.
Therefter, participants see the experimental stimuli.
The stimuli in the experiment are to show respondents fully specified or underspecified sentences on the issue position of a political actor or a masked political actor (see Section Treatment for more details).

## Power Analysis
As detailed in [Section 4](#analysis), we conduct an OLS regression for the four sentences
each of the dependent variables (i.e. sen 2--4): a) "The people, not politicians, should make the most important political decisions"; b) "The politicians in the German Bundestag need to follow the will of the people"; and c) "The political differences between elites and the people are greater than the differences among citizens".
The manipulation, i.e. the civic conception (i.e. _Bürger_) or th ethnic conception (i.e. _Volk_) as well as respondents national identity and far right party ID are the main independent variables.

To calculate power for the hypotheses, the R package `DeclareDesign` is used [@blair2019declaring].
The effect sizes are between $b = 0.2$ and $b = 0.1$ -- i.e. a small effect visualized by the purple and blue lines in Figure \ref{fig:power1}.
The hypothesis are directional, Figure \ref{fig:power1} therefore displays one-tailed tests with $\alpha = 0.05$.
The power analysis shows that testing hypotheses 1 and 2 requires a sample size of $2,000$ participants, who all annotate the four sentences, i.e. $8,000$ participants in total, (x-axis) to reach very high levels of power, i.e. $>95\%$, (indicated by the black dashed line in Figure \ref{fig:power}).

```{r power, fig.align = 'center', out.width = "100%", fig.cap = "\\label{fig:power1}Power Analysis", fig.topcaption=TRUE}
source(here::here("docs/poweranalysis.R"))
p
```

# Measures

## Dependent Variables
We rely on the three items in the populism scale that mention the people: a) "The people, not politicians, should make the most important political decisions"; b) "The politicians in the German Bundestag need to follow the will of the people"; and c) "The political differences between elites and the people are greater than the differences among citizens". All items are measured on a 5-point Likert-scale: `very much disagree`, `disagree`, `neutral`, `agree`, `very much agree`.

See Table \ref{tab:dv} for the exact phrasing of the question.

```{r dv, topcaption=TRUE}
d <- tibble(Variable = c("Item 2 (DV1)",
                         "Item 3 (DV2)",
                         "Item 4 (DV3)"), 
            `Wording ENG` = c("The people, not politicians, should make the most important political decisions","he politicians in the German Bundestag need to follow the will of the people", "The political differences between elites and the people are greater than the differences among citizens"),
            `Wording DE` = c("[Die Bürger/Das Volk], und nicht die Politiker, sollte(n) die wichtigsten politischen Entscheidungen treffen", "Die Abgeordneten des Deutschen Bundestags müssen dem Willen [der Bürger/des Volkes] Folge leisten", "Die politischen Unterschiede zwischen den Eliten und [den Bürgern/dem Volk] sind größer als die Unterschiede (zwischen Bürgern)"))

kbl(d, booktabs =T, caption = "\\label{tab:dv}Survey Questions - DV") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "7cm") %>%
  column_spec(3, width = "7cm")

```

## Treatment
Respondents are randomly assigned to the populism items translated into German using either the civic conception (i.e. _Bürger_) or the ethnic conception (i.e. _Volk_).

&nbsp;

## Control Variables

As control variables, the following _demographics_ are measured: gender, age, education, geographical region, level of urbanness, vote choice in the 2021 parliamentary elections, employment, and income.
For the analysis, only variables that are unbalanced over the experimental conditions will be included.
Table \ref{tab:demographics} gives an overview of the questions asked in the survey as well as their English translations.

- _Gender_ is measured as `sex`. The answer categories are `Male` (value of `1`), `Female` (value of `0`), and `No answer` (value of `999`).

- _Age_ is measured using 6 categories: `17 or younger`, `18--29`, `30--39`, `40--49`, `50--59`, `60--74`.

- _Education_ is measured as the highest successfully completed level of education, recoded into four categories: `low`, `middle`, `high`, and `none`.
I create dummy variables for each level of education with the lowest category as base category.

- _Urbanness_ Respondents are asked for what the type of location they live: `big city`, `suburb of a large city`, `middle-szed or small city`, `rural village`, `detached house in the countryside`, `don't know/won't say`.

- _Vote Recall_ Respondents were asked which party they voted for in the last election with their second vote. Answering categories were: `CDU/CSU`, `SPD`, `AfD`, `FDP`, `DIE LINKE`, `BÜNDNIS 90/DIE GRÜNEN`, `Other party:`, `I didn't vote/wasn't eligble to vote`, `Don't know/won't say`.

- _Employment_ Respondents were asked which category of employment -- `Full-time employed`, `Part-time employed`, `Entrepreneur`, `Unemployed and searching for a job`, `Unemployed and not searching for a job or incapacitated`, `Housewife/Househusband or else`, `Retired`, `Student or full-time education` -- applied most to them.

- _Income_ Respondents were questioned on their monthly income in bins of €500 -- `€500 or less`, `€501-€1000`, `€1001-€1500`, `€1501-€2000`, `€2001-€2500`, `€2501-€3000`, `€3001-€3500`, `€3501-€4000`, `€4501-€7500`, `€7501 or more` -- as well as giving them the options of `won't  say` and `don't know`.

- _Geographical region_ is measured using the same questions as the [German Longitudinal Election Study](https://www.gesis.org/wahlen/gles), asking respondents where they are born and currently live, differentiating between the different German provinces (_Bundesl{\"a}nder_): Baden-W{\"u}rtemberg, Bavaria, Berlin, Brandenburg, Bremen, Hamburg, Hessen, Lower Saxony, Mecklenburg-Western Pomerania, North Rhine-
Westphalia, Rhineland-Palatinate, Saarland, Saxony, Saxony-Anholt, Schleswig-Holstein, and Thuringia.

```{r demographics, topcaption=TRUE}
d <- tibble(Variable = c("gender (D1)", "age (D2)", "education (D3)",
                         "geographical-region (D4)", "vote-recall2 (D6)", "employment (D7)",
                         "income (D8)", "living-place (D9)", "birth-place (D10)"), 
            `Wording ENG` = c("Which gender do you feel you belong to?", "To which of the following age groups do you belong?",
                              "What is your highest educational qualification?", 
                             "Which of the following categories best describes where you live?",
                             "Which party did you vote for with your second vote in the last federal election in September 2021?",
                             "Now we would like to ask you something about your employment. Which of this list applies to you?",
                              "What is the total monthly net income of your household? This refers to the sum that remains after deducting taxes and social security contributions.",
                             "In which federal state or on the territory of which present federal state do you currently live?",
                              "And in which federal state or on the territory of which present federal state were you born?"),
            `Wording DE` = c("Welchem Geschlecht fühlen Sie sich zugehörig?", 
                             "Zu welcher der nachfolgenden Altersgruppen gehören Sie?",
                             "Was ist Ihr höchster Bildungsabschluss?",
                             "Welche der folgenden Kategorien beschreibt am besten, wo Sie wohnen?",
                             "Welche Partei haben Sie bei der letzten Bundestagswahl im September 2021 mit Ihrer Zweitstimme gewählt?",
                             "Nun möchten wir Sie gerne etwas zu Ihrer Erwerbstätigkeit fragen. Was von dieser Liste trifft auf Sie zu?",
                             "Wie hoch ist das monatliche Netto-Einkommen Ihres Haushaltes insgesamt? Gemeint ist die Summe, die nach Abzug von Steuern und Sozialversicherungsbeiträgen übrig bleibt.",
                             "Im welchen Bundesland bzw. auf dem Gebiet welches heutigen Bundeslandes wohnen Sie derzeit?",
                             "Und in welchem Bundesland bzw. auf dem Gebiet welches heutigen Bundeslandes wurden Sie geboren?"))

kbl(d, booktabs =T, caption = "\\label{tab:demographics}Survey Questions - Demographics") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "7cm") %>%
  column_spec(3, width = "7cm")

```

In addition, pre-treatment, respondents' ideological position, political knowledge, interest, and attitude towards national identity are measured (see Table \ref{tab:pret}).
Those variables will only be included in the analyses if balance checks indicate they are necessary.
Moreover, the variables will be used to explore heterogeneous relationships.

- _Ideological position_ is measured using an 11-point scale ranging from left (`0`) to right (`10`).

- _Exclusive national identity_ is measured with four statements ("being born in Germany", "having German ancestors", "being able to speak German", "adhering to German traditions and customs") on a 5-point Likert-scale: `very much disagree`, `disagree`, `neutral`, `agree`, `very much agree`. If `Cronbach's` $\alpha =>0.8$, we will construct an additive scale for this measure.
When `Cronbach's` $\alpha <0.8$, we will add the statement as seperate dependent variables in the multiverse analysis.

```{r pret, topcaption=TRUE}
d <- tibble(Variable = c("political-knowledge1 (PT4)", "political-knowledge2 (PT5)", "political-knowledge3 (PT6)", "political-interest (PT7)","RILE (PT8)", "Exclusive national identity (HT4)", "Exclusive national identity (HT4-1)", "Exclusive national identity (HT4-2)", "Exclusive national identity (HT4-3)", "Exclusive national identity (HT4-4)" ), 
            `Wording ENG` = c("In the federal election, you have two votes, a first vote and a second vote. How is that actually, which of the two votes is decisive for the distribution of seats in the Bundestag?",
                             "Now we would like to know from you, from what percentage of the second votes a party can send deputies in any case in the Bundestag?",
                             "And can you say approximately what the current unemployment rate is in Germany? Is it lower or higher than 10 percent?", 
                             "Once speaking in general terms: How interested are you in politics - very strongly, strongly, moderately, less strongly, or not at all?",
                            "In politics, people often talk about 'left' and 'right.' Where would you classify yourself?",
                            "Some people think that the following points are important to be truly German. Others do not think they are important. How important do you think the following points are to being German?", "being born in Germany", "having German ancestors", "being able to speak German", "adhering to German traditions and customs"),
            `Wording DE` = c("Bei der Bundestagswahl haben Sie ja zwei Stimmen, eine Erststimme und eine Zweitstimme. Wie ist das eigentlich, welche der beiden Stimmen ist ausschlaggebend für die Sitzverteilung im Bundestag?",
                             "Jetzt möchten wir gerne von Ihnen wissen, ab wie viel Prozent der Zweitstimmen eine Partei auf jeden Fall Abgeordnete in den Bundestag entsenden kann? ",
                             "Und können Sie ungefähr sagen, wie hoch die derzeitige Arbeitslosenquote in Deutschland ist? Ist sie niedriger oder höher als 10 Prozent?", 
                             "Einmal ganz allgemein gesprochen: Wie stark interessieren Sie sich für Politik – sehr stark, stark, mittelmäßig, weniger stark oder überhaupt nicht?",
                            "In der Politik reden die Leute häufig von 'links' und 'rechts'. Wo würden Sie sich selbst einordnen?",
                            "Manche Leute meinen, dass die folgenden Punkte wichtig sind, um wirklich deutsch zu sein. Andere halten diese nicht für wichtig. Für wie wichtig halten Sie die folgenden Punkte, um deutsch zu sein?", "in Deutschland geboren sein", "deutsche Vorfahren haben", "deutsch sprechen können", "sich an deutsche Traditionen und Gepflogenheiten halten"))

kbl(d, booktabs =T, caption = "\\label{tab:pret}Survey Questions - PreTreatment Questions") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "7cm") %>%
  column_spec(3, width = "7cm")
```

- _Political knowledge_ is measured with three items. First, respondents are asked which of the two votes is determinative for the seat distribution in parliament: `first vote`, `second vote`, `both`, `don't know`. Secondly, respondents are asked about the electoral threshold. Last, respondents are asked whether or not the unemployment rate is over ten procent.

- _Political interest_ is measured by asking people how strongly they are interested in politics using a 5-point Likert-scale (`very strong`, `strong`, `medium`, `less strong`, `not at all`) and a `don't know/won't say` option.

## Attention Checks
We include two attention checks in the survey.
The first one is after the demographic covariates, the second one is asked just before respondents enter the round of the experimental treatments.
The attention checks are taken from @berinsky2014separating and adapted to the German context by the authors.
If a respondents fails the first attention check, a warning appears and the respondent can only continue with the survey once the respondent has correctly answered the question correctly.
The second attention check also has a warning -- meaning that respondents have to select two options -- but if they fail to correctly pass the check, they are excluded.
Each excluded respondent due to failing an attention check is replaced with another ``good respondent".

**Attention Check 1**  When a big news story breaks people often go online to get up-to-the-minute details on what is going on.
We want to know which websites people trust to get this information.
We also want to know if people are paying attention to the question.
To show that you have read this much, please ignore the question and select BILD-Zeitung and Süddeutsche Zeitung as your two answers.
When there is a big news story, which is the one news website you would visit first? (Please only choose one).
Eight (German) news outlets are provided to choose from.
Respondents pass the attention check if they select _BILD-Zeitung_ and _Süddeutsche Zeitung_.

**Attention Check 2**:
We would like to get a sense of your general preferences.
Most  modern  theories  of  decision  making  recognize  that  decisions  do  not take  place  in  a  vacuum.
Individual  preferences  and  knowledge,  along  with  situational variables can greatly impact the decision process.
To demonstrate that you've read this much,  just  go  ahead  and  select  both  red  and  green  among  the  alternatives  below,  no matter what your favourite color is.  Yes, ignore the question below and select both of those options.
What is your favourite color?
Six colors are provided to choose from, respondents pass the attention check if they select red and green.

## Exclusion Criteria
Participants are required to respond to each question.
Participants who fail the second attention check will be excluded but replaced by another participant.

# Analysis
We test the hypotheses formulated in [Section 1](#expectations) by fitting linear multivariate regressions separately for the two issues.
In each model, we will estimate the coefficient for whether the party compromised or remained steadfast (H1), whether or not the coalition talks continued (H2), and the interaction of compromise and the outcome (H3).
We will only add control variables in the analyses that are unbalanced, as explained in [Section 4.1](#balance-checks).

## Balance Checks
We will conduct a balance test based on demographics (age, gender, education, geographical region, level of urbanness,employment, and income), vote choice in the 2021 parliamentary elections, ideological self-placement, political knowledge, political interest, party identity strength, and positions on and salience of the three issues, using the `cobalt` R package [@greifer2021]. 
If the groups are unbalanced on one of these variables -- i.e. standardized mean differences $<0.05$ -- I will add the covariates to the analyses.
I will use the code below to conduct the balance tests (see [here](https://github.com/MarikenvdVelden/compromise-punish/tree/main/src/analysis/balance-test.R) for the R script).

```{r balance, eval=FALSE, echo = TRUE}
covs <- d %>%
  mutate(treatment = populism) %>%
  select(populism, D1:D10, PT1:PT8, HT4)

balanced <-bal.tab(treatment ~ factor(D1) + D2 +
                     factor(D3) + D4 + D5 + 
                     factor(D6) + D7 + D8 + D9 +
                     D10 + PT4 + PT5 + PT6 + PT7 + 
                     PT8 + HT4,
                     data = covs,
          thresholds = c(m = 0.05))[[1]] 
```

## Hypothesis 1
We test the _Ethnic Conception hypothesis_, we will conduct a difference in means for the group who received the ethnic and the civic question wording, for all populist items containing a reference to _the people_ -- see code chunck below.

```{r h1, eval = FALSE, echo = TRUE}
source(here::here("src/lib/functions.R"))

df %>% 
  mutate(ethnic == if_else(populist=="Volk", "Ethnic", "Civic")) %>%
  select(`POST-2`, `POST-3`, `POST-4`, ethnic) %>%
  gather(key = variable, value = value, -ethnic) %>%
  group_by(ethnic, variable) %>%
  summarise(value = list(value)) %>% 
  spread(ethnic, value) %>% 
  group_by(variable) %>% 
  mutate(p_value = t.test(unlist(Ethnic), unlist(Civic))$p.value,
         t_value = t.test(unlist(Ethnic), unlist(Civic))$statistic)
```

## Hypothesis 2a

We test the _Exclusive National Identity hypothesis_ using the equation and the code chunk below.
The coefficient of $\beta_{1}$ denotes the difference in populist attitude for civic and ethnic conceptions of the people.
The coefficient of $\beta_{2}$ denotes the difference in populist attitude for different levels of national identity.
The coefficient of $\beta_{3}$ is the interaction of the  $\beta_{1}$ and  $\beta_{2}$.
If the Average Marginal Effect of $\beta_{1}$ is positive and statistically significant for higher values of $\beta_{2}$ (i.e. stronger national identity), then H2a will be confirmed.

$$\begin{aligned}
    \hat{Y}_{item} = \beta_{0} + \beta_{1}Ethnic  +
    \beta_{2}National \ Identity + \beta_{3} Ethnic * National \ Identity + \nonumber \\
    \beta_{4}Unbalanced \ Covariates + \varepsilon
  \end{aligned}$$

```{r h2a, eval = FALSE, echo = TRUE}
source(here::here("src/lib/functions.R"))

df <- d %>% 
  mutate(ethnic == if_else(populist=="Volk", "Ethnic", "Civic"))  %>%
  select(`POST-2`, ethnic, HT4)
h2a_1 <- lm(`POST-2` ~  ethnic * HT4,
                   data = df)

df <- d %>% 
  mutate(ethnic == if_else(populist=="Volk", "Ethnic", "Civic"))  %>%
  select(`POST-3`, ethnic, HT4)
h2a_2 <- lm(`POST-3` ~  ethnic * HT4,
                   data = df)

df <- d %>% 
  mutate(ethnic == if_else(populist=="Volk", "Ethnic", "Civic"))  %>%
  select(`POST-4`, ethnic, HT4)
h2a_3 <- lm(`POST-4` ~  ethnic * HT4,
                   data = df)
```

## Hypothesis 2b

We test the _Far Right Party ID hypothesis_ using the equation and the code chunk below.
The coefficient of $\beta_{1}$ denotes the difference in populist attitude for civic and ethnic conceptions of the people.
The coefficient of $\beta_{2}$ denotes the difference in populist attitude for whether or not a respondent has voted for a far right party.
The coefficient of $\beta_{3}$ is the interaction of the  $\beta_{1}$ and  $\beta_{2}$.
If the Average Marginal Effect of $\beta_{1}$ is positive and statistically significant for higher values of $\beta_{2}=1$ (i.e. voted for a radical right party), then H2b will be confirmed.

$$\begin{aligned}
    \hat{Y}_{item} = \beta_{0} + \beta_{1}Ethnic  +
    \beta_{2}Far \ Right \ Party \ ID + \beta_{3} Ethnic * Far \ Right \ Party \ ID + \nonumber\\
    \beta_{4}Unbalanced \ Covariates + \varepsilon
  \end{aligned}$$

```{r h2b, eval = FALSE, echo = TRUE}
source(here::here("src/lib/functions.R"))

df <- d %>% 
  mutate(afd_vote == if_else(`D6-3`==1, 1, 0),
         ethnic == if_else(populist=="Volk", "Ethnic", "Civic"))  %>%
  select(`POST-2`, ethnic, afd_vote)
h2a_1 <- lm(`POST-2` ~  ethnic * afd_vote,
                   data = df)

df <- d %>% 
  mutate(afd_vote == if_else(`D6-3`==1, 1, 0),
         ethnic == if_else(populist=="Volk", "Ethnic", "Civic"))  %>%
  select(`POST-3`, ethnic, afd_vote)
h2a_2 <- lm(`POST-3` ~  ethnic * afd_vote,
                   data = df)

df <- d %>% 
  mutate(afd_vote == if_else(`D6-3`==1, 1, 0),
         ethnic == if_else(populist=="Volk", "Ethnic", "Civic"))  %>%
  select(`POST-4`, ethnic, afd_vote)
h2a_3 <- lm(`POST-4` ~  ethnic * afd_vote,
                   data = df)
```

## Statistical Significance
All the hypotheses are directional, and therefore all of the tests will be one-tailed.
We will use an $\alpha$-value of $0.05$ as the value for statistical significance in all models above.

\newpage
# References

### 2022-04-21 docs/poweranalysis.R

@@ -1,0 +1,51 @@
+# Power Analysis
+rm(list = ls())
+library(tidyverse)
+library(DeclareDesign)
+library(scales)
+library(here)
+
+#H1 + H2: pooled over parties
+N <- 100
+assignment_prob <- 0.25 # 1 out of 4: mask vs. unmask, specified vs. unspecified 
+treatment_effect <- 0.1
+
+population <- declare_population(N = N, u = rnorm(N))
+potential_outcomes <- declare_potential_outcomes(Y_Z_0 = u, Y_Z_1 = Y_Z_0 + treatment_effect)
+estimand <- declare_inquiry(ATE = mean(Y_Z_1 - Y_Z_0))
+assignment <- declare_assignment(prob = assignment_prob, legacy = TRUE)
+reveal_Y <- declare_reveal()
+estimator <- declare_estimator(Y ~ Z, inquiry = estimand, term="Z", model=lm)
+two_arm_design <- population + potential_outcomes + estimand + 
+  assignment + reveal_Y + estimator
+designs <- redesign(design=two_arm_design, N=seq(4000,12000,1000), treatment_effect=seq(0.1,0.4,0.1))
+
+alpha <- .05
+my_diagnosands <- declare_diagnosands(power.onetailed=mean(p.value<alpha), keep_defaults=TRUE)
+diagnosis1 <- diagnose_designs(designs,diagnosands=my_diagnosands)
+diagnosis <- diagnosis1$diagnosands_df
+diagnosis1  <- diagnosis %>% 
+  select(N, treatment_effect, power.onetailed) %>%
+  mutate(id = "Power-Analysis")
+
+p <- diagnosis1 %>%
+  filter(treatment_effect<0.3) %>% 
+  ggplot(aes(x=N, y=power.onetailed, group=factor(treatment_effect), colour=factor(treatment_effect))) +
+  geom_line() +
+  geom_hline(yintercept=0.95, linetype="dashed") +
+  geom_hline(yintercept=0.8, linetype="dashed", color = "grey") +
+  geom_vline(xintercept = 8000,linetype="dotdash", color = "darkgrey") +
+  labs(x = "Number of Respondents", y = "Power: One-Tailed",
+       title = "Pooled Issue Analysis",
+       subtitle = "Various Levels of Treatment Effect") +
+  scale_y_continuous(labels=percent) +
+  theme_minimal() +
+  scale_color_viridis_d() +
+  theme(plot.title = element_text(hjust = 0.5),
[…]

### 2022-06-02 docs/pap.Rmd
add pap

@@
 Balance checks will be conducted to demonstrate whether certain categories are over represented in a certain experimental group.
-The study has been approved by the [Research Ethics Review Committee](https://fsw.vu.nl/nl/onderzoek/research-ethics-review/index.aspx) of the _Vrije Universiteit Amsterdam_ (see the approval [here](https://github.com/MarikenvdVelden/compromise-punish/blob/main/docs/2022-3-30-59.pdf)).
+The study has been approved by the [Research Ethics Review Committee](https://fsw.vu.nl/nl/onderzoek/research-ethics-review/index.aspx) of the _Vrije Universiteit Amsterdam_ (see the approval [here](https://github.com/MarikenvdVelden/bias-experiment/blob/master/docs/2022-3-30-59.pdf)).
 To ensure good quality of our data, one attention check (discussed in more detail in [Section 3.3](#attention-checks)) is included.
@@
 ## Power Analysis
-As detailed in [Section 4](#analysis), we conduct an OLS regression for the four sentences
-each of the dependent variables (i.e.
-sen 2--4): a) "The people, not politicians, should make the most important political decisions"; b) "The politicians in the German Bundestag need to follow the will of the people"; and c) "The political differences between elites and the people are greater than the differences among citizens".
-The manipulation, i.e. the civic conception (i.e. _Bürger_) or th ethnic conception (i.e. _Volk_) as well as respondents national identity and far right party ID are the main independent variables.
+As detailed in [Section 4](#analysis), we conduct an OLS regression for the four sentences.
+The manipulation, i.e.
+the masking condition vs.
+the political actor, as well as the specification of the sentence (i.e.
+_fully specified_ or _underspecified_) as well as respondents ideological distance to the party and their levels of political knowledge are the main independent variables.
 
@@
 ## Dependent Variables
-We rely on the three items in the populism scale that mention the people: a) "The people, not politicians, should make the most important political decisions"; b) "The politicians in the German Bundestag need to follow the will of the people"; and c) "The political differences between elites and the people are greater than the differences among citizens".
-All items are measured on a 5-point Likert-scale: `very much disagree`, `disagree`, `neutral`, `agree`, `very much agree`.
+We rely on whether or not a party's (implied) stance is coded according to the party’s position (H1 and H3) as well as whether or not the statement is coded as a stance at all (H3).
+For each issue, we ask the respondent `what is according to the sentence above the position of [ACTOR]?`, with the answer categories: `in favor`, `against`, `no stance`, `don't know`.
 
-See Table \ref{tab:dv} for the exact phrasing of the question.
+## Treatment
+Respondents are randomly assigned to either view a political party as an actor, or a masked condition, where they see `X` as an actor; simultaneously, respondents see either a fully specified sentence or a underspecified sentence, in which one needs additional information to interpret the positon on an actor.
+Table \ref{tab:conditions} gives an overview of the variations in treatment  in the survey as well as their English translations.
 
-```{r dv, topcaption=TRUE}
-d <- tibble(Variable = c("Item 2 (DV1)",
-                         "Item 3 (DV2)",
-                         "Item 4 (DV3)"), 
-            `Wording ENG` = c("The people, not politicians, should make the most important political decisions","he politicians in the German Bundestag need to follow the will of the people", "The political differences between elites and the people are greater than the differences among citizens"),
-            `Wording DE` = c("[Die Bürger/Das Volk], und nicht die Politiker, sollte(n) die wichtigsten politischen Entscheidungen treffen", "Die Abgeordneten des Deutschen Bundestags müssen dem Willen [der Bürger/des Volkes] Folge leisten", "Die politischen Unterschiede zwischen den Eliten und [den Bürgern/dem Volk] sind größer als die Unterschiede (zwischen Bürgern)"))
+```{r conditions, topcaption=TRUE}
+d <- tibble(Condition = c(rep("Specified", 4),
+                          rep("Underspecified", 4)), 
+            `Wording ENG` = c("[PVV/X] says immigration should be made harder.",
+                              "[GreenLeft/X] says nitrogen emissions need to be reduced.",
+                              "[Labour Party/X] says tax rate should go up for highest earners.",
+                              "[Forum for Democracy/X] says that membership in the European Union has been especially bad for the Netherlands so far.".
+                              "[PVV/X] says many immigrants are coming this way.",
+                              "[GreenLeft/X] says nitrogen policy must be different.",
+                              "[Labour Party/X] says tax system must be changed.",
+                              "[Forum for Democracy/X] says the Netherlands should have a different role in the European Union."),
+            `Wording NL` = c("[PVV/X] zegt dat immigratie moeilijker gemaakt moet worden.",
+                             "[GroenLinks/X] zegt dat stikstofuitstoot meer tegengegaan moet worden.",
+                             "[PvdA/X] zegt dat het belastingtarief voor de hoogste inkomens omhoog moet.",
+                             "[Forum voor Democratie/X] zegt dat het lidmaatschap van de Europese Unie tot nu toe vooral slecht geweest voor Nederland is.",
+                             "[PVV/X] zegt dat veel immigranten deze kant op komen.",
+                             "[GroenLinks/X] zegt dat het stikstofbeleid anders moet.",
+                             "[PvdA/X] zegt dat het belastingstelsel moet worden aangepast.",
+                             "[Forum voor Democratie/X] zegt dat Nederland een andere rol in de Europese Unie moet hebben."))
 
-kbl(d, booktabs =T, caption = "\\label{tab:dv}Survey Questions - DV") %>%
+kbl(d, booktabs =T, caption = "\\label{tab:conditions}Survey Questions - Experimental Conditions") %>%
   kable_styling(latex_options = c("striped", "hold_position"),
@@
   column_spec(3, width = "7cm")
-
 ```
-
-## Treatment
-Respondents are randomly assigned to the populism items translated into German using either the civic conception (i.e. _Bürger_) or the ethnic conception (i.e. _Volk_).
-
-&nbsp;
 
@@
 
-As control variables, the following _demographics_ are measured: gender, age, education, geographical region, level of urbanness, vote choice in the 2021 parliamentary elections, employment, and income.
+As control variables, the following _demographics_ are measured post-treatment: gender, age, education, geographical region, level of urbanness, employment, and income.
 For the analysis, only variables that are unbalanced over the experimental conditions will be included.
@@
 
-- _Urbanness_ Respondents are asked for what the type of location they live: `big city`, `suburb of a large city`, `middle-szed or small city`, `rural village`, `detached house in the countryside`, `don't know/won't say`.
-
-- _Vote Recall_ Respondents were asked which party they voted for in the last election with their second vote.
-Answering categories were: `CDU/CSU`, `SPD`, `AfD`, `FDP`, `DIE LINKE`, `BÜNDNIS 90/DIE GRÜNEN`, `Other party:`, `I didn't vote/wasn't eligble to vote`, `Don't know/won't say`.
-
 - _Employment_ Respondents were asked which category of employment -- `Full-time employed`, `Part-time employed`, `Entrepreneur`, `Unemployed and searching for a job`, `Unemployed and not searching for a job or incapacitated`, `Housewife/Househusband or else`, `Retired`, `Student or full-time education` -- applied most to them.
@@
 
-- _Geographical region_ is measured using the same questions as the [German Longitudinal Election Study](https://www.gesis.org/wahlen/gles), asking respondents where they are born and currently live, differentiating between the different German provinces (_Bundesl{\"a}nder_): Baden-W{\"u}rtemberg, Bavaria, Berlin, Brandenburg, Bremen, Hamburg, Hessen, Lower Saxony, Mecklenburg-Western Pomerania, North Rhine-
-Westphalia, Rhineland-Palatinate, Saarland, Saxony, Saxony-Anholt, Schleswig-Holstein, and Thuringia.
+- _Geographical region_ is measured using the _Nielsen districs_, dividing the Netherlands into 1) the 3 major cities plus suburbs, Amsterdam (plus Diemen, Ouder-Amstel, Landsmeer, Amstelveen), Rotterdam (plus Schiedam, Capelle aan den IJssel, Krimpen aan den IJssel, Nederlek, Ridderkerk, Barendrecht, Albrandswaard) and The Hague (plus Leidschendam, Voorburg, Rijswijk, Wassenaar, Wateringen); 2) West (Noord-Holland, Zuid-Holland and Utrecht (excluding the major cities and their suburbs; 3) North (Groningen, Friesland and Drenthe), 4) East (Overijssel, Gelderland and Flevoland); and South (Zeeland, Noord-Brabant and Limburg).
 
 ```{r demographics, topcaption=TRUE}
-d <- tibble(Variable = c("gender (D1)", "age (D2)", "education (D3)",
-                         "geographical-region (D4)", "vote-recall2 (D6)", "employment (D7)",
-                         "income (D8)", "living-place (D9)", "birth-place (D10)"), 
-            `Wording ENG` = c("Which gender do you feel you belong to?", "To which of the following age groups do you belong?",
-                              "What is your highest educational qualification?", 
-                             "Which of the following categories best describes where you live?",
-                             "Which party did you vote for with your second vote in the last federal election in September 2021?",
-                             "Now we would like to ask you something about your employment. Which of this list applies to you?",
-                              "What is the total monthly net income of your household? This refers to the sum that remains after deducting taxes and social security contributions.",
-                             "In which federal state or on the territory of which present federal state do you currently live?",
-                              "And in which federal state or on the territory of which present federal state were you born?"),
-            `Wording DE` = c("Welchem Geschlecht fühlen Sie sich zugehörig?", 
-                             "Zu welcher der nachfolgenden Altersgruppen gehören Sie?",
-                             "Was ist Ihr höchster Bildungsabschluss?",
-                             "Welche der folgenden Kategorien beschreibt am besten, wo Sie wohnen?",
-                             "Welche Partei haben Sie bei der letzten Bundestagswahl im September 2021 mit Ihrer Zweitstimme gewählt?",
-                             "Nun möchten wir Sie gerne etwas zu Ihrer Erwerbstätigkeit fragen. Was von dieser Liste trifft auf Sie zu?",
-                             "Wie hoch ist das monatliche Netto-Einkommen Ihres Haushaltes insgesamt? Gemeint ist die Summe, die nach Abzug von Steuern und Sozialversicherungsbeiträgen übrig bleibt.",
-                             "Im welchen Bundesland bzw. auf dem Gebiet welches heutigen Bundeslandes wohnen Sie derzeit?",
-                             "Und in welchem Bundesland bzw. auf dem Gebiet welches heutigen Bundeslandes wurden Sie geboren?"))
+d <- tibble(Variable = c("gender (D1)", "age (D2)", 
+                         "region (D3)", "job (D5)", "income (D6)",
+                         "education (D7)"), 
+            `Wording ENG` = c("What is your gender?", "What is your year of birth?",
+                             "In which region do you live?",
+                             "What applies most to you? Are you...",
+                              "Can you indicate which income groups your net monthly household income falls into?",
+                             "What is your highest level of education?", ),
+            `Wording NL` = c("Wat is uw geslacht?", "Wat is uw geboortejaar?", 
+                             "In welke regio woont u?",
+                             "Wat is het meest op u van toepassing? Bent u...",
+                             "Kunt u aangeven in welke inkomensgroepen uw netto maandelijks huishoudinkomen valt?", 
+                             "Wat is uw hoogst GENOTEN opleiding",))
 
@@
   column_spec(3, width = "7cm")
-
 ```
 
-In addition, pre-treatment, respondents' ideological position, political knowledge, interest, and attitude towards national identity are measured (see Table \ref{tab:pret}).
+In addition, pre-treatment, respondents' ideological position, political knowledge, vote recall, and position on the issues migration, climate, tax, and EU are measured (see Table \ref{tab:pret}).
 Those variables will only be included in the analyses if balance checks indicate they are necessary.
@@
 
-- _Exclusive national identity_ is measured with four statements ("being born in Germany", "having German ancestors", "being able to speak German", "adhering to German traditions and customs") on a 5-point Likert-scale: `very much disagree`, `disagree`, `neutral`, `agree`, `very much agree`.
-If `Cronbach's` $\alpha =>0.8$, we will construct an additive scale for this measure.
-When `Cronbach's` $\alpha <0.8$, we will add the statement as seperate dependent variables in the multiverse analysis.
+- _Vote Recall_ Respondents were asked which party they voted for in the 2021 parliamentary elections.
+The options were 1) all parties that were elected into parliament -- `Bij1`, `BoerBurgerBeweging`, `CDA`, `ChristenUnie`, `D66`, `Denk`, `Forum for Democracy`, `JA21`, `GroenLinks`, `PvdA`, `Animal Rights Party`, `PVV`, `SGP`, `SP`, `VOLT`,`VVD`, `50Plus Party` -- 2) another party; 3) blanco vote; and 4) a `Don't know` option.
 
 ```{r pret, topcaption=TRUE}
-d <- tibble(Variable = c("political-knowledge1 (PT4)", "political-knowledge2 (PT5)", "political-knowledge3 (PT6)", "political-interest (PT7)","RILE (PT8)", "Exclusive national identity (HT4)", "Exclusive national identity (HT4-1)", "Exclusive national identity (HT4-2)", "Exclusive national identity (HT4-3)", "Exclusive national identity (HT4-4)" ),
-            `Wording ENG` = c("In the federal election, you have two votes, a first vote and a second vote.
-How is that actually, which of the two votes is decisive for the distribution of seats in the Bundestag?",
-                             "Now we would like to know from you, from what percentage of the second votes a party can send deputies in any case in the Bundestag?",
-                             "And can you say approximately what the current unemployment rate is in Germany? Is it lower or higher than 10 percent?", 
-                             "Once speaking in general terms: How interested are you in politics - very strongly, strongly, moderately, less strongly, or not at all?",
-                            "In politics, people often talk about 'left' and 'right.' Where would you classify yourself?",
-                            "Some people think that the following points are important to be truly German.
-Others do not think they are important.
-How important do you think the following points are to being German?", "being born in Germany", "having German ancestors", "being able to speak German", "adhering to German traditions and customs"),
-            `Wording DE` = c("Bei der Bundestagswahl haben Sie ja zwei Stimmen, eine Erststimme und eine Zweitstimme.
-Wie ist das eigentlich, welche der beiden Stimmen ist ausschlaggebend für die Sitzverteilung im Bundestag?",
-                             "Jetzt möchten wir gerne von Ihnen wissen, ab wie viel Prozent der Zweitstimmen eine Partei auf jeden Fall Abgeordnete in den Bundestag entsenden kann? ",
-                             "Und können Sie ungefähr sagen, wie hoch die derzeitige Arbeitslosenquote in Deutschland ist? Ist sie niedriger oder höher als 10 Prozent?", 
-                             "Einmal ganz allgemein gesprochen: Wie stark interessieren Sie sich für Politik – sehr stark, stark, mittelmäßig, weniger stark oder überhaupt nicht?",
-                            "In der Politik reden die Leute häufig von 'links' und 'rechts'. Wo würden Sie sich selbst einordnen?",
-                            "Manche Leute meinen, dass die folgenden Punkte wichtig sind, um wirklich deutsch zu sein.
-Andere halten diese nicht für wichtig.
-Für wie wichtig halten Sie die folgenden Punkte, um deutsch zu sein?", "in Deutschland geboren sein", "deutsche Vorfahren haben", "deutsch sprechen können", "sich an deutsche Traditionen und Gepflogenheiten halten"))
+d <- tibble(Variable = c("vote recall (PT1)", "introduction",
+                         "migration (PT2)", "climate (PT3)",
+                         "tax (PT4)", "EU (PT5)", "RILE (PT6)",
+                         "political-knowledge1 (PT7_1)", 
+                         "political-knowledge2 (PT7_2)", 
+                         "political-knowledge3 (PT7_3)", 
+                         "political-knowledge4 (PT7_4)", 
+                         "political-knowledge5 (PT7_5)",
+                         "political-knowledge5 (PT7_6)"), 
+            `Wording ENG` = c( "Which party did you vote for during the last parliamentary elections of March 2021?",
+                               "To what extent do you agree with the following statements:",
+                               "There are too many immigrants in the Netherlands.",
+                               "The climate crisis is exaggerated.",
+                               "The tax rate for the highest earners should go up",
+                               ""
+                               "In politics there is often talk about left and right. Where would you put yourself on this scale?",
+                               "Which parties currently form the coalition government?",
+                               "Which of the following is the current Finance Minister in the Rutte cabinet?",
+                               "Which political party was the second-largest party in the last Lower House elections on March 17?",
+                               "By whom are members of the Senate elected?",
+                               "Who makes or is making new laws in the Netherlands?",
+                               "If the government and the House of Representatives pass a law that is not in line with the Dutch Constitution, may the courts disapprove it?"),
+            `Wording NL` = c("Op welke politieke partij heeft u bij de vorige Tweede Kamerverkiezingen van maart 2021 gestemd?",
+                             "In hoeverre bent u het eens met onderstaande stellingen:",
+                             "Er zijn te veel immigranten in Nederland.",
+                             "De klimaatcrisis wordt overdreven.",
+                             "Het belastingtarief voor de hoogste inkomens moet omhoog",
+                             "Het lidmaatschap van de Europese Unie is tot nu toe vooral slecht geweest voor Nederland",
+                             "In de politiek wordt vaak gesproken over links en rechts. Waar zou u uzelf op deze schaal plaatsen?",
+                             "Welke partijen vormen op dit moment de coalitieregering?",
+                            "Wie van de volgende personen is de huidige minister van Financiën in het kabinet Rutte?",
+                            "Welke politieke partij was de op één na grootste partij tijdens de afgelopen Tweede Kamerverkiezingen van 17 maart?",
+                            "Door wie worden de leden van de Eerste Kamer gekozen?",
+                            "Wie maakt of maken in Nederland nieuwe wetten?",
+                            "Als de regering en de Tweede Kamer een wet aannemen die niet overeenkomt met de Nederlandse Grondwet, mag de rechter deze wet dan afkeuren?"))
 
@@
 
-- _Political knowledge_ is measured with three items.
-First, respondents are asked which of the two votes is determinative for the seat distribution in parliament: `first vote`, `second vote`, `both`, `don't know`.
-Secondly, respondents are asked about the electoral threshold.
-Last, respondents are asked whether or not the unemployment rate is over ten procent.
+- _Political knowledge_ is measured with six items from the DPES.
 
-- _Political interest_ is measured by asking people how strongly they are interested in politics using a 5-point Likert-scale (`very strong`, `strong`, `medium`, `less strong`, `not at all`) and a `don't know/won't say` option.
+- _Position on migration_ is measured by asking people whether or not there are too many immigrants in the Netherlands using a 5-point Likert-scale (`fully disagree`, `disagree`, `neutral`, `agree`, `fully agree`).
 
-## Attention Checks
-We include two attention checks in the survey.
-The first one is after the demographic covariates, the second one is asked just before respondents enter the round of the experimental treatments.
+- _Position on climate_ is measured by asking people whether or not the climate crisis is exaggerated using a 5-point Likert-scale (`fully disagree`, `disagree`, `neutral`, `agree`, `fully agree`).
+
+- _Position on tax_ is measured by asking people whether or not the  tax rate for the highest earners should go up using a 5-point Likert-scale (`fully disagree`, `disagree`, `neutral`, `agree`, `fully agree`).
+
+- _Position on the EU_ is measured by asking people whether or not membership in the European Union has been especially bad for the Netherlands so far using a 5-point Likert-scale (`fully disagree`, `disagree`, `neutral`, `agree`, `fully agree`).
+
+## Attention Check
+We include one attention checks in the survey.
+This is asked just before respondents enter the round of the experimental treatments.
 The attention checks are taken from @berinsky2014separating and adapted to the German context by the authors.
 If a respondents fails the first attention check, a warning appears and the respondent can only continue with the survey once the respondent has correctly answered the question correctly.
-The second attention check also has a warning -- meaning that respondents have to select two options -- but if they fail to correctly pass the check, they are excluded.
-Each excluded respondent due to failing an attention check is replaced with another ``good respondent".
 
-**Attention Check 1**  When a big news story breaks people often go online to get up-to-the-minute details on what is going on.
+**Attention Check**  When a big news story breaks people often go online to get up-to-the-minute details on what is going on.
 We want to know which websites people trust to get this information.
 We also want to know if people are paying attention to the question.
-To show that you have read this much, please ignore the question and select BILD-Zeitung and Süddeutsche Zeitung as your two answers.
-When there is a big news story, which is the one news website you would visit first? (Please only choose one).
-Eight (German) news outlets are provided to choose from.
-Respondents pass the attention check if they select _BILD-Zeitung_ and _Süddeutsche Zeitung_.
-
-**Attention Check 2**:
-We would like to get a sense of your general preferences.
-Most  modern  theories  of  decision  making  recognize  that  decisions  do  not take  place  in  a  vacuum.
-Individual  preferences  and  knowledge,  along  with  situational variables can greatly impact the decision process.
-To demonstrate that you've read this much,  just  go  ahead  and  select  both  red  and  green  among  the  alternatives  below,  no matter what your favourite color is.
-Yes, ignore the question below and select both of those options.
-What is your favourite color?
-Six colors are provided to choose from, respondents pass the attention check if they select red and green.
+To show that you have read this much, please ignore the question and select Volkskrant and Metro as your two answers.
+When there is a big news story, which is the one news website you would visit first? (Please
+only choose one).
+Eight (Dutch) news outlets are provided to choose from.
+Respondents pass the attention check if they select _de Volkskrant_ and _Metro_.
 
@@
 Participants are required to respond to each question.
-Participants who fail the second attention check will be excluded but replaced by another participant.
 
 # Analysis
-We test the hypotheses formulated in [Section 1](#expectations) by fitting linear multivariate regressions separately for the two issues.
-In each model, we will estimate the coefficient for whether the party compromised or remained steadfast (H1), whether or not the coalition talks continued (H2), and the interaction of compromise and the outcome (H3).
+We test the hypotheses formulated in [Section 1](#expectations) by fitting linear multivariate regressions pooling the issues.
+In each model, we will estimate the coefficient for whether the actor in the sentence is masked or not, as well as for whether the sentence is fully specified or underspecified.
+We will add ideological distance from the party mentioned as well as political knowledge as interactions with the experimental conditions.
 We will only add control variables in the analyses that are unbalanced, as explained in [Section 4.1](#balance-checks).
@@
 ## Balance Checks
-We will conduct a balance test based on demographics (age, gender, education, geographical region, level of urbanness,employment, and income), vote choice in the 2021 parliamentary elections, ideological self-placement, political knowledge, political interest, party identity strength, and positions on and salience of the three issues, using the `cobalt` R package [@greifer2021].
-If the groups are unbalanced on one of these variables -- i.e. standardized mean differences $<0.05$ -- I will add the covariates to the analyses.
-I will use the code below to conduct the balance tests (see [here](https://github.com/MarikenvdVelden/compromise-punish/tree/main/src/analysis/balance-test.R) for the R script).
+We will conduct a balance test based on demographics (age, gender, education, geographical region, level of urbanness,employment, and income), vote choice in the 2021 parliamentary elections, ideological self-placement, political knowledge, and positions on the issues, using the `cobalt` R package [@greifer2021].
+If the groups are unbalanced on one of these variables -- i.e. standardized mean differences $<0.05$ -- we will add the covariates to the analyses.
+We will use the code below to conduct the balance tests (see [here](https://github.com/MarikenvdVelden/compromise-punish/tree/main/src/analysis/balance-test.R) for the R script).
 
@@
 covs <- d %>%
-  mutate(treatment = populism) %>%
-  select(populism, D1:D10, PT1:PT8, HT4)
+  mutate(treatment = sentence) %>%
+  select(sentence, D1:D7, PT1:PT&)
 
@@
                      factor(D3) + D4 + D5 + 
-                     factor(D6) + D7 + D8 + D9 +
-                     D10 + PT4 + PT5 + PT6 + PT7 + 
-                     PT8 + HT4,
+                     factor(D6) + D7 + PT1 + 
+                     PT2 + PT3 + PT4 + PT5 + 
+                     PT6 + PT7,
                      data = covs,
@@
 ## Hypothesis 1
-We test the _Ethnic Conception hypothesis_, we will conduct a difference in means for the group who received the ethnic and the civic question wording, for all populist items containing a reference to _the people_ -- see code chunck below.
+To test the _Ideological Bias hypothesis_, we will conduct a multilevel model, with respondents clustered in issues, see code chunck below.
+Using the pooled data we will estimate a within groups fixed effects model. 
+$Y\hat{Y}_{r, i, t}$ in Equation \ref{eq:pooled} denotes the  evaluation  of  a stance by respondent $r$, during issue $i$ and at experimental round $t$ -- ranging from round 1 to round 4.
+The standard errors are clustered at the individual level.\
+
+$$\begin{equation}\label{eq:pooled}
+    \hat{stance-correct}_{r, i, t} =\beta_{0} + \beta_{1}masked_{r, i, t} +
+    \beta_{2}specification_{r, i, t} + 
+    \beta_{3}ideological-distance-to-party_{r, i, t} +
+    \beta_{4}political-knowledge_{r, i, t} +
+    \alpha_{i} + \gamma_{t} + \varepsilon_{r, i, t}
+\end{equation}$$
+
+We consider evidence for H1a when the effect of $\beta_{3}$ is postive and statistically significant.
+If the Average Marginal Effect of $\beta_{3}$ is positive and statistically siginificant when $\beta_{2} = 1$ (i.e.
+condition is underspecidied), the second _Ideological Bias hypothesis_ (H1b) will be confirmed.
 
 ```{r h1, eval = FALSE, echo = TRUE}
-source(here::here("src/lib/functions.R"))
+h1a <- lmer(stance_correct ~ masked + specification + 
+              ideological_distance_to_party + political_knowledge +
+              (1 | issue), data= d)
 
-df %>% 
-  mutate(ethnic == if_else(populist=="Volk", "Ethnic", "Civic")) %>%
-  select(`POST-2`, `POST-3`, `POST-4`, ethnic) %>%
-  gather(key = variable, value = value, -ethnic) %>%
-  group_by(ethnic, variable) %>%
-  summarise(value = list(value)) %>% 
-  spread(ethnic, value) %>% 
-  group_by(variable) %>% 
-  mutate(p_value = t.test(unlist(Ethnic), unlist(Civic))$p.value,
-         t_value = t.test(unlist(Ethnic), unlist(Civic))$statistic)
+h1b <- lmer(stance_correct ~ masked + 
+              specification * ideological_distance_to_party + 
+              political_knowledge + (1 | issue), data= d)
 ```
 
-## Hypothesis 2a
+## Hypothesis 2
+To test the _Overinterpretation Hypotheses_, we will conduct a multilevel model, with respondents clustered in issues, see code chunck below.
+Using the pooled data we will estimate a within groups fixed effects model. 
+$Y\hat{Y}_{r, i, t}$ in Equation \ref{eq:pooled} denotes the  evaluation  of  a stance by respondent $r$, during issue $i$ and at experimental round $t$ -- ranging from round 1 to round 4.
+The standard errors are clustered at the individual level
 
-We test the _Exclusive National Identity hypothesis_ using the equation and the code chunk below.
-The coefficient of $\beta_{1}$ denotes the difference in populist attitude for civic and ethnic conceptions of the people.
-The coefficient of $\beta_{2}$ denotes the difference in populist attitude for different levels of national identity.
-The coefficient of $\beta_{3}$ is the interaction of the  $\beta_{1}$ and  $\beta_{2}$.
-If the Average Marginal Effect of $\beta_{1}$ is positive and statistically significant for higher values of $\beta_{2}$ (i.e. stronger national identity), then H2a will be confirmed.
+$$\begin{equation}\label{eq:pooled}
+    \hat{stance-correct}_{r, i, t} =\beta_{0} + \beta_{1}masked_{r, i, t} +
+    \beta_{2}specification_{r, i, t} + 
+    \beta_{3}heterogeneous_treatment{r, i, t} +
+    \beta_{4} specification * heterogeneous_treatment_{r, i, t} +
+    \alpha_{i} + \gamma_{t} + \varepsilon_{r, i, t}
+\end{equation}$$
 
-$$\begin{aligned}
-    \hat{Y}_{item} = \beta_{0} + \beta_{1}Ethnic  +
-    \beta_{2}National \ Identity + \beta_{3} Ethnic * National \ Identity + \nonumber \\
-    \beta_{4}Unbalanced \ Covariates + \varepsilon
-  \end{aligned}$$
+When the Average Marginal Effect of $\beta_{3}$ ideological distance is positive and statistically siginificant when $\beta_{2} = 1$ (i.e.
+condition is underspecidied), the _Ideological Overinterpretation hypothesis_ (H2a) will be confirmed.
+When the Average Marginal Effect of $\beta_{3}$ political knowledge is positive and statistically siginificant when $\beta_{2} = 1$ (i.e.
+condition is underspecidied), the _Political Knowledge Overinterpretation hypothesis_ (H2b) will be confirmed.
 
-```{r h2a, eval = FALSE, echo = TRUE}
-source(here::here("src/lib/functions.R"))
+```{r h2, eval = FALSE, echo = TRUE}
+h2a <- lmer(stance ~ masked + specification * ideological_distance_to_party +
+              political_knowledge + (1 | issue), data= d)
 
-df <- d %>% 
-  mutate(ethnic == if_else(populist=="Volk", "Ethnic", "Civic"))  %>%
-  select(`POST-2`, ethnic, HT4)
-h2a_1 <- lm(`POST-2` ~  ethnic * HT4,
-                   data = df)
-
-df <- d %>% 
-  mutate(ethnic == if_else(populist=="Volk", "Ethnic", "Civic"))  %>%
-  select(`POST-3`, ethnic, HT4)
-h2a_2 <- lm(`POST-3` ~  ethnic * HT4,
-                   data = df)
-
-df <- d %>% 
-  mutate(ethnic == if_else(populist=="Volk", "Ethnic", "Civic"))  %>%
-  select(`POST-4`, ethnic, HT4)
-h2a_3 <- lm(`POST-4` ~  ethnic * HT4,
-                   data = df)
+h2b <- lmer(stance ~ masked + 
+              specification * ypolitical_knowledge + 
+              ideological_distance_to_party + (1 | issue), data= d)
 ```
 
-## Hypothesis 2b
+## Hypothesis 3
+To test the _Masking Solution hypotheses_, we will conduct a multilevel model, with respondents clustered in issues, see code chunck below.
+Using the pooled data we will estimate a within groups fixed effects model. 
+$Y\hat{Y}_{r, i, t}$ in Equation \ref{eq:pooled} denotes the  evaluation  of  a stance by respondent $r$, during issue $i$ and at experimental round $t$ -- ranging from round 1 to round 4.
+The standard errors are clustered at the individual level
 
-We test the _Far Right Party ID hypothesis_ using the equation and the code chunk below.
-The coefficient of $\beta_{1}$ denotes the difference in populist attitude for civic and ethnic conceptions of the people.
-The coefficient of $\beta_{2}$ denotes the difference in populist attitude for whether or not a respondent has voted for a far right party.
-The coefficient of $\beta_{3}$ is the interaction of the  $\beta_{1}$ and  $\beta_{2}$.
-If the Average Marginal Effect of $\beta_{1}$ is positive and statistically significant for higher values of $\beta_{2}=1$ (i.e. voted for a radical right party), then H2b will be confirmed.
+$$\begin{equation}\label{eq:pooled}
+    \hat{stance-correct}_{r, i, t} =\beta_{0} + \beta_{1}masked_{r, i, t} +
+    \beta_{2}specification_{r, i, t} + 
+    \beta_{3}heterogeneous_treatment{r, i, t} +
+    \beta_{4} masked * heterogeneous_treatment_{r, i, t} +
+    \alpha_{i} + \gamma_{t} + \varepsilon_{r, i, t}
+\end{equation}$$
 
-$$\begin{aligned}
-    \hat{Y}_{item} = \beta_{0} + \beta_{1}Ethnic  +
-    \beta_{2}Far \ Right \ Party \ ID + \beta_{3} Ethnic * Far \ Right \ Party \ ID + \nonumber\\
-    \beta_{4}Unbalanced \ Covariates + \varepsilon
-  \end{aligned}$$
+When the Average Marginal Effect of $\beta_{3}$ ideological distance is negatiive and statistically siginificant when $\beta_{2} = 1$ (i.e.
+condition is masked), the first _Masking Solution hypothesis_ (H3a) will be confirmed.
+When the Average Marginal Effect of $\beta_{3}$ political knowledge is negative and statistically siginificant when $\beta_{2} = 1$ (i.e.
+condition is masked), the second _Masking Solution hypothesis_ (H3b) will be confirmed.
 
-```{r h2b, eval = FALSE, echo = TRUE}
-source(here::here("src/lib/functions.R"))
+```{r h3, eval = FALSE, echo = TRUE}
+h3a <- lmer(stance ~ masked + specification * ideological_distance_to_party +
+              political_knowledge + (1 | issue), data= d)
 
-df <- d %>% 
-  mutate(afd_vote == if_else(`D6-3`==1, 1, 0),
-         ethnic == if_else(populist=="Volk", "Ethnic", "Civic"))  %>%
-  select(`POST-2`, ethnic, afd_vote)
-h2a_1 <- lm(`POST-2` ~  ethnic * afd_vote,
-                   data = df)
-
-df <- d %>% 
-  mutate(afd_vote == if_else(`D6-3`==1, 1, 0),
-         ethnic == if_else(populist=="Volk", "Ethnic", "Civic"))  %>%
-  select(`POST-3`, ethnic, afd_vote)
-h2a_2 <- lm(`POST-3` ~  ethnic * afd_vote,
-                   data = df)
-
-df <- d %>% 
-  mutate(afd_vote == if_else(`D6-3`==1, 1, 0),
-         ethnic == if_else(populist=="Volk", "Ethnic", "Civic"))  %>%
-  select(`POST-4`, ethnic, afd_vote)
-h2a_3 <- lm(`POST-4` ~  ethnic * afd_vote,
-                   data = df)
+h3b <- lmer(stance ~ masked + 
+              specification * ypolitical_knowledge + 
+              ideological_distance_to_party + (1 | issue), data= d)
 ```

### 2022-06-02 src/analysis/balance-test.R

@@ -1,0 +1,57 @@
+covs <- d %>%
+  mutate(treat = if_else(actor == "masked" & sentence == "specified", 1,
+                 if_else(actor == "masked" & sentence == "underspecified", 2,
+                 if_else(actor == "party" & sentence == "specified", 3, 4)))) %>%
+  select(treat, D1:D7, PT1, PT7) %>%
+  drop_na()
+
+balanced <-bal.tab(treat ~ factor(D1) +  D2 + factor(D3) + 
+                     D5 + D7 + factor(PT1) + PT2 + 
+                     PT3 + PT4 + PT5 + PT6 + PT7,
+                   data = covs,
+                   thresholds = c(m = 0.05))[[1]] 
+
+df <- balanced %>%
+  mutate(variable = c("Male", "Age", "Region: Big Cities",
+                      "Region: West", "Region: North", "Region: East", 
+                      "Region: South", "Income",
+                      "High Levels of Education",
+                      "Low Levels of Education", "Medium Levels of Education",
+                      "50+", "BBB", "Bij1",  "Blanco", "CDA",
+                      "ChristenUnie", "D66", "Didn't vote", "Don't know",
+                      "Forum for Democracy","GroenLinks", "JA21", 
+                      "Not eligible", "Other Party","PvdA", "Animal Rights Party", "PVV", 
+                      "SGP", "SP", "VOLT","VVD",
+                      "Migration", "Climate", "Tax", "EU")) %>%
+  mutate(variable = factor(variable,
+                           levels = c("Male", "Age", "Region: Big Cities",
+                                      "Region: West", "Region: North", "Region: East", 
+                                      "Region: South", "Income",
+                                      "High Levels of Education",
+                                      "Low Levels of Education", "Medium Levels of Education",
+                                      "50+", "BBB", "Bij1",  "Blanco", "CDA",
+                                      "ChristenUnie", "D66", "Didn't vote", "Don't know",
+                                      "Forum for Democracy","GroenLinks", "JA21", 
+                                      "Not eligible", "Other Party","PvdA", "Animal Rights Party", "PVV", 
+                                      "SGP", "SP", "VOLT","VVD",
+                                      "Migration", "Climate", "Tax", "EU")),
+         difference = Corr.Un,2) %>%
+  select(variable, difference) %>%
+  drop_na() %>%
[…]

### 2023-05-26 docs/pap.Rmd
add analyses and draft paper

@@
 ## make sure to have the printr package installed: install.packages('printr')
-knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
+knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE,cache = TRUE)
 #library(printr)
@@
                               "[Labour Party/X] says tax rate should go up for highest earners.",
-                              "[Forum for Democracy/X] says that membership in the European Union has been especially bad for the Netherlands so far.".
+                              "[Forum for Democracy/X] says that membership in the European Union has been especially bad for the Netherlands so far.",
                               "[PVV/X] says many immigrants are coming this way.",
@@
                          "education (D7)"), 
-            `Wording ENG` = c("What is your gender?", "What is your year of birth?",
+            `Wording ENG` = c("What is your gender?", 
+                              "What is your year of birth?",
                              "In which region do you live?",
@@
                               "Can you indicate which income groups your net monthly household income falls into?",
-                             "What is your highest level of education?", ),
+                             "What is your highest level of education?"),
             `Wording NL` = c("Wat is uw geslacht?", "Wat is uw geboortejaar?", 
@@
                              "Kunt u aangeven in welke inkomensgroepen uw netto maandelijks huishoudinkomen valt?", 
-                             "Wat is uw hoogst GENOTEN opleiding",))
+                             "Wat is uw hoogst GENOTEN opleiding"))
 
@@
                                "The tax rate for the highest earners should go up",
-                               ""
+                               "The Netherlands should have a different role in the European Union.",
                                "In politics there is often talk about left and right. Where would you put yourself on this scale?",
@@
 If the groups are unbalanced on one of these variables -- i.e. standardized mean differences $<0.05$ -- we will add the covariates to the analyses.
-We will use the code below to conduct the balance tests (see [here](https://github.com/MarikenvdVelden/compromise-punish/tree/main/src/analysis/balance-test.R) for the R script).
+We will use the code below to conduct the balance tests (see [here](https://github.com/MarikenvdVelden/bias-experiment/blob/master/src/analysis/balance-test.R) for the R script).
 
@@
 Using the pooled data we will estimate a within groups fixed effects model. 
-$Y\hat{Y}_{r, i, t}$ in Equation \ref{eq:pooled} denotes the  evaluation  of  a stance by respondent $r$, during issue $i$ and at experimental round $t$ -- ranging from round 1 to round 4.
+$Y\hat{Y}_{r, i, t}$ in Equation \ref{eq:pooled1} denotes the  evaluation  of  a stance by respondent $r$, during issue $i$ and at experimental round $t$ -- ranging from round 1 to round 4.
 The standard errors are clustered at the individual level.\
 
-$$\begin{equation}\label{eq:pooled}
+\begin{equation}\label{eq:pooled1}
     \hat{stance-correct}_{r, i, t} =\beta_{0} + \beta_{1}masked_{r, i, t} +
@@
     \alpha_{i} + \gamma_{t} + \varepsilon_{r, i, t}
-\end{equation}$$
+\end{equation}
 
@@
 Using the pooled data we will estimate a within groups fixed effects model. 
-$Y\hat{Y}_{r, i, t}$ in Equation \ref{eq:pooled} denotes the  evaluation  of  a stance by respondent $r$, during issue $i$ and at experimental round $t$ -- ranging from round 1 to round 4.
+$Y\hat{Y}_{r, i, t}$ in Equation \ref{eq:pooled2} denotes the  evaluation  of  a stance by respondent $r$, during issue $i$ and at experimental round $t$ -- ranging from round 1 to round 4.
 The standard errors are clustered at the individual level
 
-$$\begin{equation}\label{eq:pooled}
+\begin{equation}\label{eq:pooled2}
     \hat{stance-correct}_{r, i, t} =\beta_{0} + \beta_{1}masked_{r, i, t} +
@@
 Using the pooled data we will estimate a within groups fixed effects model. 
-$Y\hat{Y}_{r, i, t}$ in Equation \ref{eq:pooled} denotes the  evaluation  of  a stance by respondent $r$, during issue $i$ and at experimental round $t$ -- ranging from round 1 to round 4.
+$Y\hat{Y}_{r, i, t}$ in Equation \ref{eq:pooled3} denotes the  evaluation  of  a stance by respondent $r$, during issue $i$ and at experimental round $t$ -- ranging from round 1 to round 4.
 The standard errors are clustered at the individual level
 
-$$\begin{equation}\label{eq:pooled}
+\begin{equation}\label{eq:pooled3}
     \hat{stance-correct}_{r, i, t} =\beta_{0} + \beta_{1}masked_{r, i, t} +
@@
 All the hypotheses are directional, and therefore all of the tests will be one-tailed.
-We will use an $\alpha$-value of $0.05$ as the value for statistical significance in all models above.
+We will use an $\alpha$-value of ``0.05` as the value for statistical significance in all models above.
 

### 2023-05-26 report/draft.Rmd
add analyses and draft paper

---
output: 
  stevetemplates::article:
    fig_caption: true
    extra_dependencies: ["float"]
bibliography: references.bib
urlcolor: black
linkcolor: black
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \floatsetup[table]{capposition=top}
   - \usepackage{booktabs}
   - \usepackage{longtable}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage[para,online,flushleft]{threeparttable}
   - \usepackage{threeparttablex}
   - \usepackage[normalem]{ulem}
   - \usepackage{makecell}
   - \usepackage{xcolor}
   - \usepackage{hyperref}
   - \usepackage{array}   
   - \usepackage{caption}
   - \usepackage{graphicx}
   - \usepackage{epstopdf}
   - \usepackage{siunitx}
   - \usepackage{hhline}
   - \usepackage{calc}
   - \usepackage{tabularx}
   - \usepackage{fontawesome}
   - \usepackage{amsthm}
   - \newtheorem{hypo}{Hypothesis}
biblio-style: apsr
title: "Whose Truth is it Anyway? An Experiment on Annotation Bias"
author:
- name: Mariken A.C.G. van der Velden*
- affiliation: Dep. of Communciation Science, Vrije Universiteit Amsterdam
- name: Wouter van Atteveldt
- affiliation: Dep. of Communciation Science, Vrije Universiteit Amsterdam
- name: Antse Fokkens
- affiliation: Computational Linguistics & Text Mining Lab, Vrije Universiteit Amsterdam
- name: Felicia Loecherbach
- affiliation: Center for Social Media & Politics, New York University
- name: Myrthe Reuver
- affiliation: Computational Linguistics & Text Mining Lab, Vrije Universiteit Amsterdam
- name: Kasper Welbers
- affiliation: Dep. of Communciation Science, Vrije Universiteit Amsterdam
thanks: "* = Corresponding author, Replication files are available on the author's Github account (https://github.com/MarikenvdVelden/bias-experiment); Author contributions: designed the study: MACGvdV, WvA, AF, FL, MR, & KW; conducted the study: MACGvdV, FL & MR; data cleaning & analysis: MACGvdV; writing of the paper: MACGvdV"
anonymous: TRUE
abstract: "Information is key to inform the behavior of citizens, and thereby for the social scientists studying them. The democratization of data has lead to numerous possibilities to gather and analyze textual data. These enourmous amounts of data are typically handled by machine learning techniques to classify into meaningful variables. The performance of these models are compared to a so-called golden standards, created by human annotators. Having a high level of agreement between these annotators is key, but some suggest personal characteristics of annotators, like political ideology or knowledge, interfere. We show in two pre-registered experiments that XXX. Thereby [contribution]."
keywords: "Experiment, Annotation Bias, Ideology, Measurinig Political Position, Text-as-Data, Political Knowledge"
geometry: margin=1in
mainfont: cochineal
fontsize: 11pt
params:
  anonymous: ""
doublespacing: TRUE
endnote: no
pandocparas: TRUE
sansitup: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message=FALSE, warning=FALSE,
                      fig.pos = "h", out.extra = "")
options(kableExtra.latex.load_packages = FALSE)
options(scipen = 1, digits = 2)
source(here::here("src/lib/functions.R"))
# Uncomment below if you want code captions
# oldSource <- knitr::knit_hooks$get("source")
# knitr::knit_hooks$set(source = function(x, options) {
#   x <- oldSource(x, options)
#   x <- ifelse(!is.null(options$code.cap), paste0(x, "\\captionof{chunk}{", options$code.cap,"}"), x)
#   ifelse(!is.null(options$ref), paste0(x, "\\label{", options$ref,"}"), x)
# })
# Add `chunkcaption: TRUE` to YAML as well.
```

# Introduction
Information is key to inform the behavior of citizens, and thereby for the social scientists studying them.
Classical theories of political communication, such as agenda setting or framing [e.g. @vanaelst2016political; @lecheler2019news], formulate that political information drives opinion formation and participation in politics -- from voting to protests [for overviews hereof across countries, see @pfetsch2013comparing].
The democratization of data has lead to numerous possibilities to gather and analyze textual data [for a recent overview, see @vanatteveldt2022computational].
In addition, advances in Natural Language Processing (NLP) have made it possible to automatically analyze large quantities of data using machine learning [e.g., see @bender2016linguistic; @wei2023overview].
The performance of such machine learning techniques are compared to so-called "golden standards", created by human annotators. 
Most data collection efforts to create such standards assume that there is only one correct interpretation for every input example, and that disagreement between the annotators is something that needs to be dealt with at all costs [@aroyo2015truth]. 
A recent study [@van2021validity] demonstrates that using crowd-coding platforms is a good way to collect such golden standards without too much disagreement.
These platforms have also been used for experiments, and the quality of the respondents has been a center of attention  [@coppock2019generalizing;@clifford2015samples;@huff2015these;@berinsky2014separating]. 
While some say that this data is of similar quality to data from a random sample of the population [@coppock2019generalizing], others have demonstrated that these online platforms are populated by people that are "unlike" the general public, being younger and holding more liberal values [@clifford2015samples;@huff2015these].
If the latter is true, this might cause a problem for the coding of political texts:
@ennser2018impact report that coders of political texts incorporate prior beliefs about parties' issue stances into their coding decisions.
The authors find that party labels cue coders to a stance. 
For example, coders are more likely to report a left-wing party to be pro-immigration and a populist right-wing party to be against based on the exact same sentence.
_Is this actually bias or a diversity of view points? And how big of a problem does this bias/diversity of view points generate for scholars relying on golden standard data?_

There is a long history of text annotation in studies analyzing political text.
While this yields lots of experiences as to how to train coders so that we get reliable hand-coded data, the procedure is expensive, protracted, and sometimes does not even get us the quality of data we need [@weber18].
While the crowd offers a solution to some of these issues [@van2021validity], the main underlying assumption that there is one correct interpretation for every input example remains untouched.
At the same time, trends of polarization have shown that people do interpret information according to their ideological position. 
_Does this mean that some are right and others are wrong? Or is there an ideological difference in the ground truth?_
This question presents a fundamental challenge to the main way of working when collecting gold standard data, as we operate from the baseline assumption that disagreement among the annotators should be avoided or reduced.
Typically, when specific cases continuously cause disagreement, more instructions are added to limit interpretations [for recent innovations to improve annotation, see for example: @barbera2021automated; @struthers2020bridging; @winter2020online; @debell2013harder; @benoit2016crowd, @ying2022topics].
However, work in computational linguistics has shown that increased annotation instructions do not increase quality [@parmar2022don].
This leaves us between a rock and a hard place. Is there a potential bias in annotators that we should account for?

In this paper, we build upon the NLP literature on disagreement -- or bias -- in annotation [e.g., see @shen2021sounds; @geva2019we; @sommerauer2020penguin; @plank2014learning] and so-called perspectivism [@basile2021toward; @havens2022beyond] -- i.e. the adoption of methods that integrate the opinions and perspectives of the human subjects involved in the knowledge representation step of the machine learning processes [@basile2021toward].
This literature puts forward that disagreement can occur because of differences in ideological position or political knowledge [@shen2021sounds; @alkiek2022classification; @joseph2021mis].
This allows us to test the extent to which disagreement takes place, for what type of stances, as well as gives us directions on how to deal with the diversity in conceptions and the political heterogeneity that nowadays potentially occurs in our sample of annotators.
To do so, we have fielded two high-powered pre-registered experiments (see [here](link) and [here](link)) in the Netherlands -- a low-level polarized country -- and in the U.S. -- a high-level polarized country testing the effect of ideological distance between the annotator and the political actor in the text (H1), the effect of overinterpretation based on political knowledge or ideological engagement (H2), and a offered solution of masking the political actor to mitigate the efffects of ideology and knowledge (H3).[^1]
In the experiments, we vary the level of specification with which a political actor takes a stance -- a declarative sentence versus a sentence where with some knowledge on politics, the stance might be inferred -- as well as whether the political actor is shown or masked with putting `[ACTOR]` instead of the political party. 
We do this for four different political issues: _Environment_, _Immigration_, _Tax Policy_, and _EU_ (for the Dutch case) or _Foreign Policy_ (for the American case).
This country selection does not only allow us to showcase the scope conditions of disagreement due to different levels of political heterogeneity, but also differentiates between languages.
English is not only the most dominant language for computational text analysis in the social science [@baden2022three; Dolinsky et al. 2023], crowd-coders do not need to be native speakers, given the dominance of English in our daily lives.
This is different for Dutch, it is a language spoken by a smaller community, typically native speakers, yet still an often-enough researched case in computational text analysis in the social science [@baden2022three; Dolinsky et al. 2023].

Our results demonstrate that overall sentence where with some knowledge on politics the stance might be inferred are really difficult for the crowd to annotate -- people overinterpret the position using their own knowledge of the world. 
This is problematic as these sentence are very common in political text -- like legislative debates or speeches -- as well as media reports. 
Moreover, our results also demonstrate that for these disagreements in the crowd to occur, the level of polarization needs to be high.
We do find support for our hypotheses in the American context, but not in the Dutch context -- except for the situation where masking overcomes political knowledge, we do find support for that in the Dutch case, but not in the American one.
Our findings thus underline the importance of taking disagreement seriously for the creation of gold standard text -- the bread-and-butter of all machine learning endaveours. 
We should look beyond the majority vote and modeling it in the data, because if an algorithm is trained on biased data from disagreeing annotators, it will reproduce and often exacerbate that bias when it is applied to new data [e.g., see @prost2019debiasing]. 
To be able to model these characteristic of annotators, we should survey the characteristics of annotators when using the crowd (see Webb-Williams et al. 2023 for a similar argument, yet different annotator characteristics).

[^1]: The data and research compendium is published on the [main author's github page](link) -- annonymized for the review process.

# Whose Truth is it Anyway? Disagreement & Perspectivism in Creating Gold Standard Data
To test whether societal trends of polarization have an effect on our crowd-coders, and by extension on the quality of our annotated gold standard data. To that end, we have formulated the following hypotheses:

> **Ideological Bias hypothesis** (_H1a_): The larger the ideological distance between respondent and the party, the less likely respondents code stances according to the party’s position.

> **Ideological Bias hypothesis** (_H1b_): The effect of H1a is stronger for the underspecified condition.

> **Ideological Overinterpretation hypothesis** (_H2a_):  The larger the ideological distance between respondent and the party, the more likely respondents interpret underspecified sentences as stance.

> **Political Knowledge Overinterpretation hypothesis** (_H2b_): The more political knowledge, the more likely people interpret underspecified sentences as stance.

> **Masking Solution hypothesis** (_H3a_):  Masking reduces the effect of respondents’ ideological position for coding stances according to the party’s position.

> **Masking Solution hypothesis** (_H3b_): Masking reduces the effect of respondents’ level of political knowledge for coding stances according to the party’s position.

# Data, Methods & Measurement
## Data
We  will conduct this survey experiment in the Netherlands in May 2022.
The sample, recruited through [KiesKompas](https://www.kieskompas.nl/en/),  consists of 3,000 participants (based on the power analysis presented in our [online compendium](LINK)) of 18 years and older.
Kieskompas works with non-random opt-in respondents.
Therefore, we measure many demographic background variables, and balance checks have been conducted to demonstrate whether certain categories are over represented in a certain experimental group.
The study has been approved by the [Research Ethics Review Committee](https://fsw.vu.nl/nl/onderzoek/research-ethics-review/index.aspx) of the _Vrije Universiteit Amsterdam_ (see the approval [here](https://github.com/MarikenvdVelden/bias-experiment/blob/master/docs/2022-3-30-59.pdf)).
To ensure good quality of our data, one attention check (discussed in more detail in [Section 3.3](#attention-checks)) is included [@berinsky2014separating].

## Measurement
_**Experimental Conditions.**_ Respondents are randomly assigned to either view a political party as an actor, or a masked condition, where they see `X` as an actor; simultaneously, respondents see either a fully specified sentence or a underspecified sentence, in which one needs additional information to interpret the positon on an actor.
Table \ref{tab:conditions} gives an overview of the variations in treatment  in the survey as well as their English translations.

```{r conditions, topcaption=TRUE}
d <- tibble(Condition = c(rep("Specified", 4),
                          rep("Underspecified", 4)), 
            `Wording ENG` = c("[PVV/X] says immigration should be made harder.",
                              "[GreenLeft/X] says nitrogen emissions need to be reduced.",
                              "[Labour Party/X] says tax rate should go up for highest earners.",
                              "[Forum for Democracy/X] says that membership in the European Union has been especially bad for the Netherlands so far.",
                              "[PVV/X] says many immigrants are coming this way.",
                              "[GreenLeft/X] says nitrogen policy must be different.",
                              "[Labour Party/X] says tax system must be changed.",
                              "[Forum for Democracy/X] says the Netherlands should have a different role in the European Union."),
            `Wording NL` = c("[PVV/X] zegt dat immigratie moeilijker gemaakt moet worden.",
                             "[GroenLinks/X] zegt dat stikstofuitstoot meer tegengegaan moet worden.",
                             "[PvdA/X] zegt dat het belastingtarief voor de hoogste inkomens omhoog moet.",
                             "[Forum voor Democratie/X] zegt dat het lidmaatschap van de Europese Unie tot nu toe vooral slecht geweest voor Nederland is.",
                             "[PVV/X] zegt dat veel immigranten deze kant op komen.",
                             "[GroenLinks/X] zegt dat het stikstofbeleid anders moet.",
                             "[PvdA/X] zegt dat het belastingstelsel moet worden aangepast.",
                             "[Forum voor Democratie/X] zegt dat Nederland een andere rol in de Europese Unie moet hebben."))

kbl(d, booktabs =T, caption = "\\label{tab:conditions}Survey Questions - Experimental Conditions") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "7cm") %>%
  column_spec(3, width = "7cm")
```

_**Dependent Variable.**_  We rely on whether or not a party's (implied) stance is coded according to the party’s position (H1 and H3) as well as whether or not the statement is coded as a stance at all (H3). For each issue, we ask the respondent `what is according to the sentence above the position of [ACTOR]?`, with the answer categories: `in favor`, `against`, `no stance`, `don't know`.

_**Moderating Covariates.**_ _Ideological position_ is measured using an 11-point scale ranging from left (`0`) to right (`10`). - _Political knowledge_ is measured with six items from the DPES.

_**Controle Variables.**_ In our analysis, we control for demographic information (gender, age, education, income, religion, job) as well as political background variables (trust in politics, ideological position on economic left-right scale and cultural progressive-conservative scale, and evaluations and prospects of the economy). 
Tables A.5 till A.17 in the OA demonstrate the descriptive information per country.

## Method
To test our hypotheses, we will conduct a multilevel model, with respondents clustered in issues, see Equation \ref{eq:pooled1}.
Using the pooled data we will estimate a within groups fixed effects model.
We have conducted a balance test based on demographics (age, gender, education, geographical region, level of urbanness,employment, and income), vote choice in the 2021 parliamentary elections, ideological self-placement, political knowledge, and positions on the issues, using the `cobalt` R package [@greifer2021]. 
This balance test indicated that none of the variables are unbalanced over the experimental groups, and therefore, as pre-registered, will not be added to the regression formula.
$Y\hat{Y}_{r, i, t}$ in Equation \ref{eq:pooled1} denotes the  evaluation  of  a stance by respondent $r$, during issue $i$ and at experimental round $t$ -- ranging from round 1 to round 4.
The standard errors are clustered at the individual level.

\begin{multline}\label{eq:pooled1}
    \hat{stance correct}_{r, i, t} =\beta_{0} + \beta_{1}masked_{r, i, t} +
    \beta_{2}specification_{r, i, t} + 
    \beta_{3}ideological distance to party_{r, i, t} +\\
    \beta_{4}political knowledge_{r, i, t} +
    \alpha_{i} + \gamma_{t} + \varepsilon_{r, i, t}
\end{multline}

# Results
To answer whether there is an ideological or knowledge-basd annotation bias, we have conducted a two-by-two experiment. Table \ref{tab:profile} demonstrates the average profile of respondents who annotate correctly and incorrectly. In terms of demographics, there is not much of a difference. Yet, people who are incorrectly identifying stances are more left-wing oriented compared to those who are correct -- i.e. an average score of 4 for those who are incorrect vs. an average score of 5 for those who are correct. For other positions on issues or political knowledge, we do not see a difference in averages between those who are correctly and incorrectly identifying stances.

```{r "profile"}
load(here("data/intermediate/cleaned_data_nl.RData"))
load(here("data/intermediate/cleaned_data_us.RData"))
source(here("src/analysis/profile_resp.R"))
kbl(e1, booktabs =T, caption = "\\label{tab:conditions}Profile Dutch Stance Annotators") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = "7cm") %>%
  column_spec(2, width = "7cm")

kbl(e1_us, booktabs =T, caption = "\\label{tab:conditions_us}Profile American Stance Annotators") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = "7cm") %>%
  column_spec(2, width = "7cm")
```

Looking at the effect of the experimental conditions on the two dependent variables -- 1) correctly identifying a stance; and 2) over-interpreting a stance -- Figure \ref{fig:results-base} visualizes the baseline. 
The left-hand panel demonstrates the effect of the two experimental conditions for correctly identifying the stance. 
On average, approximately half of the respondents where correct -- as indicated by the intercept. 
When we mask the political actor -- i.e. instead of mentioning the party, we put "X" -- we see that this improves correctly interpreting the stance significantly. However, an effect of `0.06` (i.e. `6%`) is not a big improvement.
We do see that the level of specification of a sentence has a significant effect. 
If a sentence is not fully specified, it has a substantive negative effect on the likelihood to correctly interpret the sentence.
A coefficient of `-0.38` indicates that compared to a fully specified sentence, `38%` of the respondents are more likely to be incorrect when the sentence is under-specified -- that is when the sentence does not state a clear position, but mentions the issue.
For the left-hand panel, we look at a different dependend variable, so instead of seeing if respondents where correct or not, we look at whether they interpreted the sentence as a stance or not. 
On average, almost nobody overinterprets a stance -- as indicated by an intercept of `-0.01`. 
However, if people see an X compared to a political actor, they are statistically significantly more likely to interpret the sentence as a stance. 
Yet, a coefficient of `0.02` (i.e. `2%`) is a very small effect.
For the condition of specification level, however, we see that compared to a fully specified sentence, people seeing an under-specified sentence are much more likely to interpret the sentence as a stance: an increase of `0.83`.
This indicates that people do not excel in this task without any instruction.
In the pre-registered section, we demonstrate the tests of the hypotheses, and afterwards, we visualize some explorations of the data to show the robustness of our findings.

```{r results-baseline, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-base} Baseline Results of Experimental Conditions"}
knitr::include_graphics(here("report/figures", "baseline-1.png"))
```

## Pre-registered Results
First, we test whether there is an ideological bias in interpreting stances (H1a), and if this bias increases for those that are further away from the ideological position of the political actor in the under-specified condition (H1b).
Figure \ref{fig:results-h1} demonstrates on the left-hand panel the regression coefficients and on the right-hand panel the average marginal effects of the interaction between ideological distance and under-specified sentences.
The upper-left panel of Figure \ref{fig:results-h1} demonstrates the coefficient of ideological distances for the likelihood of interpreting the stance correctly.
There is a negligible positive effect -- a coefficient of `0.002` -- that is borderline significant. Substantially, this means that there is no effect of ideological distance for correctly interpreting the stance.
Hence, no ideological bias found, thus no support for our H1a.
Looking at the lower-left, and right-hand panel of Figure \ref{fig:results-h1}, we see a small but significant effect of the interaction between ideological distance and the under-specified condition.
It is, however, going in the other direction than hypothesized. 
For those who are ideologically further away from the party, they are less likely to be wrong than those who are close too the party. 
Yet, the difference is about `3%` -- from a coefficient of `-0.38` to `-0.35`.
We thus find no evidence for H1b either. 
This demonstrates that even in times of hightened polarization, ideological annotation bias is not a huge concern.

```{r results-h1, out.width= "95%", fig.align = 'center', fig.cap = "\\label{fig:results-h1} Ideological Distance"}
knitr::include_graphics(here("report/figures", "h1-3.png"))
```

Secondly, we hypothesized that there is a risk of over-interpretation from those that are ideologically distant to the political actor (H2a) as well as those who have high levels of political knowledge (H2b). 
We measure this with an interaction between the experimental condition of specification and the variable of interest.
Following @brambor2006understanding, we visualize the average marginal effects for both interactions to enhance interpretation. 
Figure \ref{fig:results-h2} demonstrates the average marginal effects for both interaction effects: In the left-hand panel the results for H2a, and in the right-hand panel, the results for H2b.
The left-hand panel of Figure \ref{fig:results-h2} shows that those who are more distant from the position of the political actor are more likely to over-interpret the sentence as a stance.
The effect increases from about `80%` to those that are closest to the party to `91%` for those who are furthest away from the party.
This supports our H2a.
The right-hand panel of Figure \ref{fig:results-h2} shows that political knowledge does statistically significantly effect over-interpretation of a sentence as a stance.
The substantial effect, however, is small: From the least knowledgeable respondents to the most knowledgeable ones, the likelihood of over-interpretation increases with `2%`.
So, while we do find support for our H2b, we will in our recommendations section focus less on the political knowledge as a possible interfering feature, but focus on what to do with under-specified sentences when using crowd-coding.

```{r results-h2, out.width = "95%", out.height = "75%", fig.align = 'center', fig.cap = "\\label{fig:results-h2} Results Level of Sentence Specification"}
knitr::include_graphics(here("report/figures", "h2-1.png"))
```

Thirdly, we test whether masking of the political actor is a solution for potentially misinterpreting the stance. 
We hypothesized that masking should reduce the ideological bias (H3a) as well as the bias resulting from political knowledge (H3b).
We test these hypotheses with an interaction between the condition masking and the variables of interest and Figure \ref{fig:results-h3} demonstrates the average marginal effects.
The left-hand panel of Figure \ref{fig:results-h3} shows, against expectation, a slight increase: Those that are further away from the masked political actor are more likely to incorrectly interpret the stance. The decrease in slope is however negligible: From approximately `6%` to `7%` of the respondents being incorrect.
The slope for political knowledge, on the right-hand panel of Figure \ref{fig:results-h3}, is so-possible even flatter.
This means that masking of political actors does not help to correctly interpret a sentence as a stance -- i.e. no support for H3a and H3b.

```{r results-h3, out.width = "95%",  out.height = "75%", fig.align = 'center', fig.cap = "\\label{fig:results-h3} Results Masking Solution"}
knitr::include_graphics(here("report/figures", "h3-1.png"))
```

## Exploratory Results
To check the robustness of our findings, Figure \ref{fig:results-exp1} demonstrates the analyses for each issue separately. 
On the left-hand panel of Figure \ref{fig:results-exp1}, we show the results for correctly interpreting the stance.
We see some variation between issues. 
For the issue _Tax_ (lowest row), we see that almost everyone interprets the sentence correct -- an intercept of `0.96`.
This increases statistically significantly still with about `1%` when masking the political actor, and decreases statistically significantly with `68%` when the sentence is under-specified.
The regression results (in Online Appendix) show a small effect of ideological distance: the further away, the more likely to correctly interpret the stance -- i.e. against H1a.
Respondents found it most difficult to correctly interpret the issue _Environment_ (top row of Figure \ref{fig:results-exp1}), only `12%` has this correct initially, as indicated by the intercept. 
This increases statistically significantly still with about `8%` when masking the political actor, and decreases statistically significantly with `15%` when the sentence is under-specified.
The other two issues -- _EU_ and _Immigration_, in the middle rows of Figure \ref{fig:results-exp1} show a similar effect as the main base-line presented in Figure \ref{fig:results-base}.
The same pattern holds for the other dependent variable -- overinterpretation of a stance -- in the right-hand panel of Figure \ref{fig:results-exp1}.
Thus, while there are some differences in effect sizes between the issues, the overall findings are not driven by a single issue.

```{r results-exploration1, out.height = "85%",out.width = "95%",, fig.align = 'center', fig.cap = "\\label{fig:results-exp1} Exploration: Issue Speficific Analyses"}
knitr::include_graphics(here("report/figures", "issues-1.png"))
```

In addition to issue-specific analyses, we also explore an interaction between treatments, visualized in Figure \ref{fig:results-exp2} for both dependent variables.
Figure \ref{fig:results-exp2} shows that masking is of help when sentences are under-specified.
In the left-hand panel of Figure \ref{fig:results-exp2}, it demonstrates that for under-specifief sentences, people are less likely to incorrectly identify a sentence as a stance when the actor is masked (coefficient of `-0.30`) than when an actor is revealed (coefficient of `-0.45`). That means there is a 15% increase in having it correct.
The difference for over-interpreting is smaller between revealed and masked political actor -- shown in the right-hand panel of Figure \ref{fig:results-exp2} -- yet also statistically significant.
Compared to `85%` over-interpreting the sentence as a stance, in the masking solution "only" `80%` over-interprets the sentence as a stance. 
In the recommendation section, we will reflect on the masking solution for under-specified sentences.

```{r results-exploration2, out.width = "95%", out.height= "75%", fig.align = 'center', fig.cap = "\\label{fig:results-exp2} Exploration: Interaction with Treatments"}
knitr::include_graphics(here("report/figures", "exploration-3.png"))
knitr::include_graphics(here("report/figures", "exploration-4.png"))
```

Lastly, we explore three different ways of measuring ideological distance.
First, we measured ideological bias by looking at whether the respondent is congruent or not with the issue position in the sentence, visualized in Figure \ref{fig:results-exp3}.
Second, we measured ideological bias by looking at whether the person voted for the party displayed in the sentence, visualized in Figure \ref{fig:results-exp4}.
And thirdly, we measured ideological bias by looking at the ideology of the respondents -- not in relation to the political actor revealed, visualized in Figure \ref{fig:results-exp5}.
Figures \ref{fig:results-exp3}, \ref{fig:results-exp4}, and \ref{fig:results-exp5} show that our null-finding regarding ideological bias is not conditional upon the measure we used. In none of the analyses, we find evidence for ideological bias.

```{r results-exploration3, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-exp3} Exploration: Different Indicators of Ideological Distance (1)"}
knitr::include_graphics(here("report/figures", "exploration-1.png"))
```

```{r results-exploration4, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-exp4} Exploration: Different Indicators of Ideological Distance (2)"}
knitr::include_graphics(here("report/figures", "exploration-2.png"))
```

```{r results-exploration5, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-exp5} Exploration: Different Indicators of Ideological Distance (3)"}
knitr::include_graphics(here("report/figures", "exploration-3.png"))
```

```{r results-h2_expl, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-h2-expl} Results Level of Sentence Specification"}
knitr::include_graphics(here("report/figures", "h2-3.png"))
```

# Discussion
TBA

## Recommendations
TBA

\newpage
# References

### 2023-05-26 src/analysis/Analysis.Rmd
add analyses and draft paper

---
title: 'Analyses'
author: 
output: 
  github_document:
always_allow_html: true
editor_options: 
  chunk_output_type: console
---

```{r "setup", include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.align = "center",
                      fig.path=('../../report/figures/'))
library(rmarkdown)
```

# Required Packages &amp; Reproducibility
```{r "env", message=FALSE, warning=F}
rm(list=ls())
source(here::here("src/lib/functions.R"))
```

# Analyses
```{r "data for analyses"}
load(here("data/intermediate/cleaned_data_nl.RData"))
load(here("data/intermediate/cleaned_data_us.RData"))
```

## Balance Checks

The figure below shows that the data is not unbalanced for any of the variables: As described in the Pre-Analysis Plan (p.10), we will not add any covariates to the analyses as controls.

```{r "balance-checks", fig.height=12, fig.width=8}
source(here("src/analysis/balance-test.R"))
df
```

## Baseline Model
```{r "baseline", fig.align="center", fig.height=4, fig.width=8}
source(here("src/analysis/baseline.R"))
b

b_b
```

```{r "issues", fig.align="center", fig.height=16, fig.width=8}
source(here("src/analysis/issue_specific.R"))
b_i

b_ib
```

## Ideological Bias hypothesis
```{r "h1", fig.align="center", fig.height=8, fig.width=8}
source(here("src/analysis/h1.R"))
p1

p1b

p1 / p1b +
  plot_layout(heights = c(1, 1.3)) + 
  plot_layout(guides = "collect") & theme(legend.position = 'bottom')

p1_2 / p1b2 +
  plot_layout(heights = c(1, 1.3)) 
```

## Overinterpretation Hypotheses
```{r "h2", fig.align="center", fig.height=8, fig.width=8}
source(here("src/analysis/h2.R"))
p2

p2b

p2c
```

## Masking Hypothesis
```{r "h3", fig.align="center", fig.height=8, fig.width=8}
source(here("src/analysis/h3.R"))
p3

p3b

p3c
```

## Exploration
```{r "exploration", fig.height=7, fig.width=8}
source(here("src/analysis/exploration.R"))
p_e1
p_e2
p_e3_1
p_e3_2
p_b
p_b2

source(here("src/analysis/profile_resp.R"))
kbl(e1, booktabs =T, caption = "\\label{tab:conditions}Profile Dutch Stance Annotators") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = "7cm") %>%
  column_spec(2, width = "7cm")

kbl(e1_us, booktabs =T, caption = "\\label{tab:conditions_us}Profile American Stance Annotators") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = "7cm") %>%
  column_spec(2, width = "7cm")
```

### 2023-05-26 src/data-processing/prepare-data.Rmd
add analyses and draft paper

---
title: 'Prepare Data'
author: 
output: 
  github_document:
    toc: yes
always_allow_html: true
editor_options:

  chunk_output_type: console
---

```{r "setup", include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.align = "center",
                      fig.path=('../../report/figures/'))
library(printr)
options(digits = 2)
library(rmarkdown)
```

## Required Packages &amp; Reproducibility
```{r "env", message=FALSE, warning=F}
rm(list=ls())
source(here::here("src/lib/functions.R"))
#renv::snapshot()
```

## Tidy Data
This code chuck downloads the data from Qualtrics using the API and cleans the raw data.

```{r "get & clean data", results="hide"}
#Clean Dutch Data
d_nl <- read_sav(here("data/raw-private/NLOZ2206_Mariken_Bias-experiment.sav"))
source(here("src/data-processing/clean_data_nl.R"))

#Clean US Data
d_us <- fetch_survey(surveyID = "SV_86XBcsFUJka9eke", 
                    verbose = TRUE, force_request = T,
                    label = FALSE, convert = FALSE)
source(here("src/data-processing/clean_data_us.R"))
```

## Save Data for Analysis
```{r "save data"}
save(d_nl, file = here("data/intermediate/cleaned_data_nl.RData"))
save(d_us, file = here("data/intermediate/cleaned_data_us.RData"))
```

## Visualization of Data

### Dependent Variable by Issue
```{r "Dependent Variable", echo=F, fig.width=15, fig.height=8}
p1 <- d_nl %>%
  group_by(stance,issue,specification, masking) %>% 
  count() %>% 
  mutate(stance = ifelse(stance==0, "Not Correctly identified stance",
                         "Correctly identified stance"),
         specification = ifelse(specification == "Specified", 
                                "Condition: Specified Sentence", "Underspecified Sentence"),
         masking = ifelse(masking == "Masked", "Actor Masked", "Party Revealed")) %>% 
  ggplot(aes(y = issue, x = n, fill = stance)) +
  geom_col() +
  facet_grid(masking~specification) +
  labs(x = "", y = "", 
       title = "Dutch Data") +
  #scale_x_continuous(labels = scales::percent) +
  theme_ipsum() +
  scale_fill_manual(values = fig_cols[c(3, 2)]) +
  theme(legend.position = "bottom",
        legend.title = element_blank())

p2 <- d_us %>%
  group_by(stance,issue,specification, masking) %>% 
  count() %>% 
  mutate(stance = ifelse(stance==0, "Not Correctly identified stance",
                         "Correctly identified stance"),
         specification = ifelse(specification == "Specified", 
                                "Condition: Specified Sentence", "Underspecified Sentence"),
         masking = ifelse(masking == "Masked", "Actor Masked", "Party Revealed")) %>% 
  ggplot(aes(y = issue, x = n, fill = stance)) +
  geom_col() +
  facet_grid(masking~specification) +
  labs(x = "", y = "", 
       title = "US Data") +
  #scale_x_continuous(labels = scales::percent) +
  theme_ipsum() +
  scale_fill_manual(values = fig_cols[c(3, 2)]) +
  theme(legend.position = "bottom",
        legend.title = element_blank())

p1 + p2 + 
  plot_layout(guides = "collect") & theme(legend.position = 'bottom')

```

### Covariates
```{r "Independent Variables", echo=F, message=FALSE, warning=FALSE,fig.width=10, fig.height=10}
rm(p1,p2)
p <- d_nl %>% select(distance, PT7, PT6) %>% 
  pivot_longer(cols = everything(),
               values_to = "score") %>% 
  group_by(name) %>% 
  summarise(m = mean(score),
            stdev = sd(score)) %>% 
  mutate(name = recode(name,
                       `distance` = "Ideological Distance \n(> = More Distance)",
                       `PT6` = "Ideological Self Placement \n(0 = Left, 10 = Right)",
                       `PT7` = "Political Knowledge \n(0 = Low, 6 = High)"),
         country = "The Netherlands")

p2 <- d_us %>% 
  select(distance, PT7, PT7b, PT6) %>% 
  pivot_longer(cols = everything(),
               values_to = "score") %>% 
  group_by(name) %>% 
  summarise(m = mean(score),
            stdev = sd(score)) %>% 
  mutate(name = recode(name,
                       `distance` = "Ideological Distance \n(> = More Distance)",
                       `PT6` = "Ideological Self Placement \n(0 = Extremely Liberal, 6 = Extremely Conservative)",
                       `PT7` = "Political Knowledge \n(0 = Low, 3 = High)",
                       `PT7b` = "Bullshit Receptivity \n(0 = Low, 5 = High)"),
         country = "United States")

p %>% 
  add_case(p2) %>% 
  ggplot(aes(y = name, x = m, 
             xmin = (m - (1.96*stdev)),
             xmax = (m + (1.96*stdev)),
             color = country)) +
    geom_point(position = position_dodge(.2)) +
    geom_errorbar(width = 0, position = position_dodge(.2)) +
    theme_ipsum() +
    scale_colour_manual(values = fig_cols) +
    labs(x = "Value", y = "") +
  theme(legend.position = "bottom",
        legend.title = element_blank())
```

### Correlations Matrix
```{r "Correlations Matrix", echo=F, message=F, warning=F, fig.width=15, fig.height=15}
d_nl %>%
  drop_na() %>% 
  select(`Age` = D1,
         `Gender` = D2,
         `Education` = D3,
         `Region`  = D4,
         `Income`  = D5,
         `Job`  = D6,
         `Vote Recall`  = PT1,
         `Position: Immigration`  = PT2,
         `Position: Climate`  = PT3,
         `Position: Tax`  = PT4,
         `Position: EU`  = PT5,
         `Ideology`  = PT6,
         `Political Knowledge` = PT7,
         `Distance to Political Party` = distance,
         `Stance` = stance) %>%
  ggcorrmat(
  type = "robust", # correlation method
  cor.vars = c(everything()), # a range of variables can be selected
  colors = c(fig_cols[1], fig_cols[4], fig_cols[8]),
  title = "Correlalogram for Variables under Study",
  subtitle = "Dutch Data"
)

d_us %>%
  drop_na() %>% 
  select(`Age` = D1,
         `Gender` = D2,
         `Education` = D3,
         `Region`  = D4,
         `Income`  = D5,
         `Job`  = D6,
         `PID`  = PT1,
         `PID Strength` = PT1b,
         `Position: Immigration`  = PT2,
         `Position: Climate`  = PT3,
         `Position: Tax`  = PT4,
         `Position: EU`  = PT5,
         `Ideology`  = PT6,
         `Political Knowledge` = PT7,
         `Bullshit Receptivity` = PT7b,
         `Distance to Political Party` = distance,
         `Stance` = stance) %>%
  ggcorrmat(
  type = "robust", # correlation method
  cor.vars = c(everything()), # a range of variables can be selected
  colors = c(fig_cols[1], fig_cols[4], fig_cols[8]),
  title = "Correlalogram for Variables under Study",
  subtitle = "US Data"
)
```

### 2023-05-26 src/analysis/balance-test.R

@@ -1,45 +1,117 @@
-covs <- d %>%
-  mutate(treat = if_else(actor == "masked" & sentence == "specified", 1,
-                 if_else(actor == "masked" & sentence == "underspecified", 2,
-                 if_else(actor == "party" & sentence == "specified", 3, 4)))) %>%
-  select(treat, D1:D7, PT1, PT7) %>%
-  drop_na()
+## NL Data
+covs <- d_nl %>%
+  mutate(treat = if_else(masking == "Masked" & specification == "Specified", 1,
+                 if_else(masking == "Masked" & specification == "Underspecified", 2,
+                 if_else(masking == "Party" & specification == "Specified", 3, 4)))) %>%
+  select(treat, D1:D6, PT1:PT7, attention) %>%
+  filter(PT1 != "NA", D6 != "NA")
 
-balanced <-bal.tab(treat ~ factor(D1) +  D2 + factor(D3) + 
-                     D5 + D7 + factor(PT1) + PT2 + 
-                     PT3 + PT4 + PT5 + PT6 + PT7,
+balanced <-bal.tab(treat ~ D1 + factor(D2) + factor(D3) + 
+                     factor(D4) +D5 + factor(D6) + factor(PT1) + PT2 + 
+                     PT3 + PT4 + PT5 + PT6 + PT7 + attention,
                    data = covs,
                    thresholds = c(m = 0.05))[[1]] 
 
-df <- balanced %>%
-  mutate(variable = c("Male", "Age", "Region: Big Cities",
-                      "Region: West", "Region: North", "Region: East", 
-                      "Region: South", "Income",
-                      "High Levels of Education",
-                      "Low Levels of Education", "Medium Levels of Education",
-                      "50+", "BBB", "Bij1",  "Blanco", "CDA",
-                      "ChristenUnie", "D66", "Didn't vote", "Don't know",
+df_nl <- balanced %>%
+  mutate(variable = c( "Age", "Male",
+                       "High Levels of Education",
+                       "Low Levels of Education", "Medium Levels of Education",
+                       "Region: Big Cities",
+                       "Region: East", "Region: North",
+                       "Region: South", "Region: West",  
+                       "Income", "Job: Fulltime employed",
+                      "Job: Retired", "Job: Stay at home", "Job: Student",
+                      "Job: Unemployed", "Job: Unemployed and searching",
+                      "50+","Bij1", "BBB","CDA",
[…]

### 2023-05-26 src/analysis/baseline.R

@@ -1,0 +1,98 @@
+df_us <- d_us %>% 
+  select(stance, masking, specification, interpretation, country, issue)
+df_nl <- d_nl %>% 
+  select(stance, masking, specification, interpretation, country, issue)
+
+d <- df_nl %>% 
+  add_case(df_us)
+
+b1 <- broom.mixed::tidy(lmer(stance ~ masking + specification + 
+                         (country | issue), data= d)) %>%
+  mutate(y = "Y: Correctly Interpreting Stance",
+         type = "Pooled Analysis")
+
+tmp <- broom.mixed::tidy(lmer(stance ~ masking + specification + 
+                                (1 | issue), data= df_us)) %>%
+  mutate(y = "Y: Correctly Interpreting Stance",
+         type = "United States")
+
+b1 <- b1 %>% 
+  add_case(tmp)
+
+tmp <- broom.mixed::tidy(lmer(stance ~ masking + specification + 
+                                (1 | issue), data= df_nl)) %>%
+  mutate(y = "Y: Correctly Interpreting Stance",
+         type = "The Netherlands")
+
+b1 <- b1 %>% 
+  add_case(tmp)
+
+b2 <- broom.mixed::tidy(lmer(interpretation ~ masking + specification + 
+                               (country | issue), data= d)) %>%
+  mutate(y = "Y: Overinterpreting Stance",
+         type = "Pooled Analysis")
+
+tmp <- broom.mixed::tidy(lmer(interpretation ~ masking + specification + 
+                                (1 | issue), data= df_us)) %>%
+  mutate(y = "Y: Overinterpreting Stance",
+         type = "United States")
+
+b2 <- b2 %>% 
+  add_case(tmp)
+
+tmp <- broom.mixed::tidy(lmer(interpretation ~ masking + specification + 
+                                (1 | issue), data= df_nl)) %>%
+  mutate(y = "Y: Overinterpreting Stance",
+         type = "The Netherlands")
+
+b2 <- b2 %>% 
+  add_case(tmp)
+  
+b <- b1 %>% 
+  add_case(b2) %>% 
+  filter(term %in% c("(Intercept)",
+                     "maskingParty",
+                     "specificationUnderspecified"),
+         type != "Pooled Analysis") %>% 
+  mutate(term = recode(term,
+                       `(Intercept)` = "Intercept",
+                       `maskingParty` = "Condition:Masked Political Actor",
[…]

### 2023-05-26 src/analysis/exploration.R

@@ -1,0 +1,607 @@
+df_us <- d_us %>% 
+  select(stance, interpretation, masking, specification, distance, PT7, country, issue)
+df_nl <- d_nl %>% 
+  select(stance, interpretation, masking, specification, distance, PT7, country, issue)
+
+d <- df_nl %>% 
+  add_case(df_us)
+
+df <- d %>% 
+  mutate(b = specification,
+         a = recode(masking,
+                    `Masked` = 1,
+                    `Party` = 0))
+b3_a <- lmer(stance ~ b * a +
+                 (country | issue), data= df)
+b3_a <- tidy(margins::margins(b3_a, variables = "b",
+                               at = list("a" = 0:1))) %>% 
+  mutate(condition = "Condition: Underspecified Sentence",
+         y = "Y: Correctly Identifying Stance",
+         type = "Pooled Analysis")
+
+df <- df_us %>% 
+  mutate(b = specification,
+         a = recode(masking,
+                    `Masked` = 1,
+                    `Party` = 0))
+tmp <- lmer(stance ~ b * a + (1 | issue), data= df)
+tmp <- tidy(margins::margins(tmp, variables = "b",
+                             at = list("a" = 0:1))) %>% 
+  mutate(condition = "Condition: Underspecified Sentence",
+         y = "Y: Correctly Identifying Stance",
+         type = "United States")
+
+b3_a <- b3_a %>% 
+  add_case(tmp)
+
+df <- df_nl %>% 
+  mutate(b = specification,
+         a = recode(masking,
+                    `Masked` = 1,
+                    `Party` = 0))
+tmp <- lmer(stance ~ b * a + (1 | issue), data= df)
+tmp <- tidy(margins::margins(tmp, variables = "b",
+                             at = list("a" = 0:1))) %>% 
+  mutate(condition = "Condition: Underspecified Sentence",
+         y = "Y: Correctly Identifying Stance",
+         type = "The Netherlands")
+
+b3_a <- b3_a %>% 
+  add_case(tmp)
+
+df <- d %>% 
+  mutate(b = specification,
+         a = recode(masking,
+                    `Masked` = 1,
+                    `Party` = 0))
+b3_b <- lmer(interpretation ~ b * a +
+               (country | issue), data= df)
+b3_b <- tidy(margins::margins(b3_b, variables = "b",
+                              at = list("a" = 0:1))) %>% 
+  mutate(condition = "Condition: Underspecified Sentence",
[…]

### 2023-05-26 src/analysis/h1.R

@@ -1,0 +1,207 @@
+df_us <- d_us %>% 
+  select(stance, masking, specification, distance, PT7, country, issue)
+df_nl <- d_nl %>% 
+  select(stance, masking, specification, distance, PT7, country, issue)
+
+d <- df_nl %>% 
+  add_case(df_us)
+
+h1a <- broom.mixed::tidy(lmer(stance ~ masking + specification + 
+              distance + PT7 +
+              (country | issue), data= d)) %>% 
+  mutate(hyp = "H1a",
+         type = "Pooled Analysis") %>% 
+  filter(term %in% c("distance"))
+
+tmp <- broom.mixed::tidy(lmer(stance ~ masking + specification + 
+                                distance + PT7 +
+                                (1 | issue), data= df_us)) %>% 
+  mutate(hyp = "H1a",
+         type = "United States") %>% 
+  filter(term %in% c("distance"))
+
+h1a <- h1a %>% 
+  add_case(tmp)
+
+tmp <- broom.mixed::tidy(lmer(stance ~ masking + specification + 
+                                distance + PT7 +
+                                (1 | issue), data= df_nl)) %>% 
+  mutate(hyp = "H1a",
+         type = "the Netherlands") %>% 
+  filter(term %in% c("distance"))
+
+h1a <- h1a %>% 
+  add_case(tmp)
+rm(tmp)
+
+h1b <- broom.mixed::tidy(lmer(stance ~ masking + 
+                                specification * distance +
+                                PT7 + (country | issue), data= d)) %>% 
+  mutate(hyp = "H1b",
+         type = "Pooled Analysis") %>% 
+  filter(term %in% c("specificationUnderspecified",
+                     "distance",
+                     "specificationUnderspecified:distance"))
+
+tmp <- broom.mixed::tidy(lmer(stance ~ masking + 
+                                specification * distance +
+                                PT7 + (1 | issue), data= df_us)) %>% 
+  mutate(hyp = "H1b",
+         type = "United States") %>% 
+  filter(term %in% c("specificationUnderspecified",
+                     "distance",
+                     "specificationUnderspecified:distance"))
+
+h1b <- h1b %>% 
+  add_case(tmp)
+
+tmp <- broom.mixed::tidy(lmer(stance ~ masking + 
+                                specification * distance +
+                                PT7 + (1 | issue), data= df_nl)) %>% 
+  mutate(hyp = "H1b",
+         type = "the Netherlands") %>% 
+  filter(term %in% c("specificationUnderspecified",
[…]

### 2023-05-26 src/analysis/h2.R

@@ -1,0 +1,146 @@
+df_us <- d_us %>% 
+  select(interpretation, masking, specification, distance, PT7, country, issue)
+df_nl <- d_nl %>% 
+  select(interpretation, masking, specification, distance, PT7, country, issue)
+
+d <- df_nl %>% 
+  add_case(df_us)
+
+df <- d %>% 
+  mutate(b = specification,
+         a = distance)
+h2a <- lmer(interpretation ~ masking + b * a +
+              PT7 + (country | issue), data= df)
+h2a <- broom.mixed::tidy(margins::margins(h2a, variables = "b",
+                             at = list("a" = 0:8.5))) %>% 
+  mutate(hyp = "H2a: Ideological Distance \n(0=Low, 10 = High)",
+         type = "Pooled Analysis")
+
+df <- df_us %>% 
+  mutate(b = specification,
+         a = distance)
+tmp <- lmer(interpretation ~ masking + b * a +
+              PT7 + (1 | issue), data= df)
+tmp <- tidy(margins::margins(tmp, variables = "b",
+                             at = list("a" = 0:5))) %>% 
+  mutate(hyp = "H2a: Ideological Distance \n(0=Low, 10 = High)",
+         type = "United States")
+
+h2a <- h2a %>% 
+  add_case(tmp)
+
+df <- df_nl %>% 
+  mutate(b = specification,
+         a = distance)
+tmp <- lmer(interpretation ~ masking + b * a +
+              PT7 + (1 | issue), data= df)
+tmp <- tidy(margins::margins(tmp, variables = "b",
+                             at = list("a" = 0:8.5))) %>% 
+  mutate(hyp = "H2a: Ideological Distance \n(0=Low, 10 = High)",
+         type = "The Netherlands")
+
+h2a <- h2a %>% 
+  add_case(tmp)
+
+df <- d %>% 
+  mutate(b = specification,
+         a = PT7)
+h2b <- lmer(interpretation ~ masking + 
+              b * a + 
+              distance + (country | issue), data= df)
+h2b <- tidy(margins::margins(h2b, variables = "b",
+                             at = list("a" = 0:4))) %>% 
+  mutate(hyp = "H2b: Political Knowledge \n (0=Low, 6 = High)",
+         type = "Pooled Analysis")
+
+df <- df_us %>% 
+  mutate(b = specification,
+         a = PT7)
+tmp <- lmer(interpretation ~ masking + b * a +
[…]

### 2023-05-26 src/analysis/h3.R

@@ -1,0 +1,166 @@
+df_us <- d_us %>% 
+  select(stance, masking, specification, distance, PT7, country, issue)
+df_nl <- d_nl %>% 
+  select(stance, masking, specification, distance, PT7, country, issue)
+
+d <- df_nl %>% 
+  add_case(df_us)
+
+df <- d %>% 
+  mutate(b = recode(masking,
+                    `Masked` = 1,
+                    `Party` = 0),
+         a = distance)
+h3a <- lmer(stance ~ specification + b * a +
+              PT7 + (country | issue), data= df)
+h3a <- broom.mixed::tidy(margins::margins(h3a, variables = "b",
+                             at = list("a" = 0:8.5))) %>% 
+  mutate(hyp = "H3a: Ideological Distance \n(0=Low, 10 = High)",
+         type = "Pooled Analysis")
+
+df <- df_us %>% 
+  mutate(b = recode(masking,
+                    `Masked` = 1,
+                    `Party` = 0),
+         a = distance)
+tmp <- lmer(stance ~ specification + b * a +
+              PT7 + (1 | issue), data= df)
+tmp <- tidy(margins::margins(tmp, variables = "b",
+                             at = list("a" = 0:4))) %>% 
+  mutate(hyp = "H3a: Ideological Distance \n(0=Low, 10 = High)",
+         type = "United States")
+
+h3a <- h3a %>% 
+  add_case(tmp)
+
+df <- df_nl %>% 
+  mutate(b = recode(masking,
+                    `Masked` = 1,
+                    `Party` = 0),
+         a = distance)
+tmp <- lmer(stance ~ specification + b * a +
+              PT7 + (1 | issue), data= df)
+tmp <- tidy(margins::margins(tmp, variables = "b",
+                             at = list("a" = 0:8.5))) %>% 
+  mutate(hyp = "H3a: Ideological Distance \n(0=Low, 10 = High)",
+         type = "The Netherlands")
+
+h3a <- h3a %>% 
+  add_case(tmp)
+
+
+df <- d %>% 
+  mutate(b = recode(masking,
+                    `Masked` = 1,
+                    `Party` = 0),
+         a = PT7)
+h3b <- lmer(stance ~ specification + b * a +
+              distance + (country | issue), data= df)
+h3b <- tidy(margins::margins(h3b, variables = "b",
+                             at = list("a" = 0:4))) %>% 
[…]

### 2023-05-26 src/analysis/issue_specific.R

@@ -1,0 +1,141 @@
+df_us <- d_us %>% 
+  select(stance, masking, specification, interpretation, country, issue)
+df_nl <- d_nl %>% 
+  select(stance, masking, specification, interpretation, country, issue) %>% 
+  mutate(issue =ifelse(issue=="EU", "Foreign Policy", issue))
+
+d <- df_nl %>% 
+  add_case(df_us)
+
+issues <- unique(d$issue)
+
+for(i in 1:length(issues)){
+  df <- d %>% filter(issue==issues[i])
+  if(i==1){
+    b1 <- broom.mixed::tidy(lm(stance ~ masking + specification + 
+                                   factor(country), data= df)) %>%
+      mutate(y = "Y: Correctly Interpreting Stance",
+             issue = issues[i],
+             type = "Pooled Analysis")
+  }
+  else{
+    tmp <- broom.mixed::tidy(lm(stance ~ masking + specification + 
+                                    factor(country), data= df)) %>%
+      mutate(y = "Y: Correctly Interpreting Stance",
+             issue = issues[i],
+             type = "Pooled Analysis")
+    b1 <- b1 %>% 
+      add_case(tmp)
+  }
+}
+
+for(i in 1:length(issues)){
+  df <- df_us %>% filter(issue==issues[i])
+  tmp <- broom::tidy(lm(stance ~ masking + specification , data= df)) %>%
+      mutate(y = "Y: Correctly Interpreting Stance",
+             issue = issues[i],
+             type = "United States")
+    b1 <- b1 %>% 
+      add_case(tmp)
+}
+
+for(i in 1:length(issues)){
+  df <- df_nl %>% filter(issue==issues[i])
+  tmp <- broom::tidy(lm(stance ~ masking + specification , data= df)) %>%
+    mutate(y = "Y: Correctly Interpreting Stance",
+           issue = issues[i],
+           type = "The Netherlands")
+  b1 <- b1 %>% 
+    add_case(tmp)
+}
+
+
+for(i in 1:length(issues)){
+  df <- d %>% filter(issue==issues[i])
+  if(i==1){
+    b2 <- broom::tidy(lm(interpretation ~ masking + specification, data= df)) %>%
+      mutate(y = "Y: Overinterpreting Stance",
+             issue = issues[i],
+             type = "Pooled Analysis")
+  }
+  else{
+    tmp <- broom::tidy(lm(interpretation ~ masking + 
+                            specification + factor(country), data= df)) %>%
[…]

### 2023-05-26 src/analysis/profile_resp.R

@@ -1,0 +1,148 @@
+d <- d_nl %>% 
+  mutate(congruence = 0,
+         congruence = ifelse(PT2<3 & issue=="Immigration", 1,
+                             congruence),
+         congruence = ifelse(PT3<3 & issue=="Environment", 1,
+                             congruence),
+         congruence = ifelse(PT4>3 & issue=="Tax", 1,
+                             congruence),
+         congruence = ifelse(PT5>3 & issue=="EU", 1,
+                             congruence),
+         pid = "Other party",
+         pid = ifelse(PT1=="Forum voor Democratie", 
+                      "Forum voor Democratie", pid),
+         pid = ifelse(PT1=="PVV", 
+                      "PVV", pid),
+         pid = ifelse(PT1=="GroenLinks", 
+                      "GroenLinks", pid),
+         pid = ifelse(PT1=="PvdA", 
+                      "PvdA", pid),
+         pid = factor(pid))
+
+e1 <- d %>% 
+  group_by(stance) %>% 
+  summarise(Age = round(mean(D1),0),
+            Gender= "Male",
+            Education = "High-levels of education",
+            Location = "West of Netherlands",
+            Income = round(mean(D5),0),
+            Job = "Fulltime Employed",
+            `Vote Recall` = "D66",
+            `Position on Immigration` = round(mean(PT2),0),
+            `Position on Environment` = round(mean(PT3),0),
+            `Position on Tax` = round(mean(PT4),0),
+            `Position on EU` =round(mean(PT5),0),
+            `Ideological Position` = round(mean(PT6),0),
+            `Ideological Distance` = round(mean(distance),0),
+            `Issue Congruence` = round(mean(congruence),0),
+            `Political Knowledge` = round(mean(PT7),0)
+            ) %>% 
+  ungroup() 
+
+e1_a <- e1 %>% 
+  select(stance, Gender:Location, Job:`Vote Recall`) %>% 
+  pivot_longer(cols = Gender:`Vote Recall`) %>% 
+  filter(stance==0) %>% 
+  select(`Incorrectly Identified Stance` = value)
+
+e1_b <- e1 %>% 
+  select(stance, Gender:Location, Job:`Vote Recall`) %>% 
+  pivot_longer(cols = Gender:`Vote Recall`) %>% 
+  filter(stance==1) %>% 
[…]

### 2023-05-26 src/data-processing/clean_data_nl.R

@@ -1,0 +1,177 @@
+d_nl <- d_nl %>% 
+  zap_labels() %>% 
+  filter(age_cont>17) %>% 
+  select(id = ResponseId,
+         D1 = age_cont, #age - continues
+         D2 = sex_rec, #gender (2 categories)
+         D3 = edu_3_rec, #education 3 categories
+         D4 = geo_nielsen5_rec, #location (Nielen V)
+         D5, #income: no income (1), <500 (2), 501-1000 (3), 1001-1500 (4); 1501-2000 (5); 2001-2500 (6); 2501-3000 (7); 3001-3500 (8); 3501-4000 (9); 4001-4500 (10); 4501-7500 (11); >7501 (12)
+         D6, #job
+         PT1, #vote recall
+         PT2, #issue position: migration: 1 = neg, 5 = pos
+         PT3, #issue position: climate: 1 = neg, 5 = pos
+         PT4, #issue position: tax: 1 = neg, 5 = pos
+         PT5, #issue position: EU: 1 = neg, 5 = pos
+         PT6, #rile -self: 0 = left, 10 = right
+         PT7_1_3, PT7_1_4, PT7_1_5, PT7_1_16, #coalition partners (polknow)
+         PT7_2, #minister finance (polknow)
+         PT7_3, #2nd largest party (polknow)
+         PT7_4, #electing 1st chamber (polknow)
+         PT7_5, #who makes laws (polknow)
+         PT7_6, #coonstitutional court (polknow)
+         actor1:actor4,
+         issue1:issue4,
+         specified1:specified4,
+         S1:S4,
+         Attention1_1:Attention1_9) %>% 
+  mutate_at(vars(Attention1_1:Attention1_9), 
+            funs(recode(., .missing = 0, .default = 1))) %>% 
+  rowwise() %>%
+  mutate(tmp = sum(c_across(starts_with("Attention1_")), na.rm = T)) %>% 
+  ungroup() %>% 
+  mutate(D2 = recode(D2,
+                         `1` = "Male",
+                         `2` = "Female"),
+         D3 = recode(D3,
+                            `1` = "Low-level of education",
+                            `2` = "Mid-levels of education",
+                            `3` = "High-levels of education"),
+         D4 = recode(D4,
+                           `1` = "Big cities",
+                           `2` = "West of Netherlands",
+                           `3` = "North of Netherlands",
+                           `4` = "East of Netherlands",
+                           `5` = "South of Netherlands"),
[…]

### 2023-05-26 src/data-processing/clean_data_us.R

@@ -1,0 +1,237 @@
+d_us <- d_us %>% 
+  filter(consent == 1,
+         screener == 2) %>% 
+  unite("PT1b", matches("PT1b"),na.rm = TRUE) %>% 
+  select(id = PID,
+         gender = D1, #gender (3 categories)
+         age = D2,  #age - continues
+         state = D3,#state, 
+         income = D5, #income: no income (1), <500 (2), 501-1000 (3), 1001-1500 (4); 1501-2000 (5); 2001-2500 (6); 2501-3000 (7); 3001-3500 (8); 3501-4000 (9); 4001-4500 (10); 4501-7500 (11); >7501 (12)
+         job = D6, #job
+         education = D7, #education
+         PT1 = PT1a, #pid
+         PT1b, #pid strength
+         PT2, #issue position: migration: 1 = neg, 5 = pos
+         PT3, #issue position: climate: 1 = neg, 5 = pos
+         PT4, #issue position: tax: 1 = neg, 5 = pos
+         PT5, #issue position: foreign policy: 1 = neg, 5 = pos
+         PT6, #libcon -self: 0 = lib, 14 = cons
+         PT7_1, #senator years (polknow)
+         PT7_2, #majority in house (polknow)
+         PT7_3, #majority in senate (polknow)
+         PT7_4_1, PT7_4_2,  PT7_4_3, PT7_4_4, #bullshit receptivity
+         actor1:actor4,
+         issue1:issue4,
+         specified1:specified4,
+         S1:S4,
+         Attention1_1:Attention1_9) %>% 
+  mutate_at(vars(Attention1_1:Attention1_9), 
+            funs(recode(., .missing = 0, .default = 1))) %>% 
+  rowwise() %>%
+  mutate(tmp = sum(c_across(starts_with("Attention1_")), na.rm = T)) %>% 
+  ungroup() %>% 
+  select(id,
+         D1 = age,
+         D2 = gender,
+         D3 = education,
+         D4 = state,
+         D5 = income,
+         D6 = job,
+         PT1:Attention1_9, tmp) %>% 
+  mutate(D2 = recode(D2,
+                     `1` = "Male",
+                     `2` = "Female",
+                     .missing = "NA", .default = "NA"),
+         D3 = recode(D3,
+                     `1` = "Low-level of education",
+                     `9` = "Low-level of education",
+                     `10` = "Low-level of education",
+                     `11` = "Low-level of education",
+                     `12` = "Low-level of education",
+                     `13` = "Low-level of education",
[…]

### 2023-05-26 src/lib/functions.R

@@ -1,0 +1,18 @@
+library(tidyverse)
+library(haven)
+library(here)
+library(hrbrthemes)
+library(ggstatsplot)
+library(lme4)
+library(broom.mixed)
+library(cobalt)
+library(patchwork)
+library(kableExtra)
+library(qualtRics)
+
+fig_cols <- yarrr::piratepal(palette = "basel", 
+                             trans = .2)
+fig_cols <- as.character(fig_cols[1:9])
+
+api_key_fn <- here("data/raw-private/qualtrics_api_key.txt")
+API <- read_file(api_key_fn) %>% trimws()

### 2023-05-26 report/draft.Rmd
add analyses and draft paper

@@
 - affiliation: Dep. of Communciation Science, Vrije Universiteit Amsterdam
-thanks: "* = Corresponding author, Replication files are available on the author's Github account (https://github.com/MarikenvdVelden/bias-experiment); Author contributions: designed the study: MACGvdV, WvA, AF, FL, MR, & KW; conducted the study: MACGvdV, FL & MR; data cleaning & analysis: MACGvdV; writing of the paper: MACGvdV"
-anonymous: TRUE
+thanks: "* = Corresponding author, Replication files are available on the author's Github account (https://github.com/MarikenvdVelden/bias-experiment); Author contributions: a) designed the study: MACGvdV, WvA, AF, FL, MR, & KW; b) conducted the study: MACGvdV, FL & MR; c) data cleaning & analysis: MACGvdV; d) writing of the paper: MACGvdV"
+anonymous: FALSE
 abstract: "Information is key to inform the behavior of citizens, and thereby for the social scientists studying them.
@@
 _Does this mean that some are right and others are wrong? Or is there an ideological difference in the ground truth?_
-This question presents a fundamental challenge to the main way of working when collecting gold standard data, as we operate from the baseline assumption that disagreement among the annotators should be avoided or reduced.
+These questions presents a fundamental challenge to the main way of working when collecting gold standard data, as we operate from the baseline assumption that disagreement among the annotators should be avoided or reduced.
 Typically, when specific cases continuously cause disagreement, more instructions are added to limit interpretations [for recent innovations to improve annotation, see for example: @barbera2021automated; @struthers2020bridging; @winter2020online; @debell2013harder; @benoit2016crowd, @ying2022topics].
@@
 
-[^1]: The data and research compendium is published on the [main author's github page](link) -- annonymized for the review process.
+[^1]: The data and research compendium is published on the [main author's github page](https://github.com/MarikenvdVelden/bias-experiment) -- annonymized for the review process.
 
 # Whose Truth is it Anyway? Disagreement & Perspectivism in Creating Gold Standard Data
-To test whether societal trends of polarization have an effect on our crowd-coders, and by extension on the quality of our annotated gold standard data.
-To that end, we have formulated the following hypotheses:
+Generating large data-sets has become one of the main drivers of progress in natural language understanding.
+In studies of political communication, the most familiar annotation tasks involve identifying basic concepts.
+This often involve noting the topic of the text, the position of the actor or the tone of the text. 
+A recent study [@van2021validity] demonstrates that using crowd-coding platforms is a good way to collect such large data-sets.
+However, having only a few workers annotate the majority of text of interest has raised concerns about data diversity and models' ability to generalize beyond the crowd-workers:
+In a series of experiments, @geva2019we show that often models do not generalize well to annotations from annotators that did not contribute to the training set, suggesting that annotator bias should be monitored during data-set creation.
+One such potential bias, especially in times of increasing polarization [@iyengar2019origins; @gidron2019toward; @boxell2022cross], is based on ideological position of the annotator. 
+Given that annotators on crows-coding platforms tend to be younger and hold more liberal values than the general public [@clifford2015samples;@huff2015these], this could potentially hamper the data diversity and generalizability of the model.
+An additional reason to monitor the annotators ideological position as a potential source of annotator bias is that a recent study in NLP showed that experiential factors influence the consistency of how political ideologies are perceived [@shen2021sounds].
+Their finding challenges the "ground-truth" assumption we as researcher make that a position for example is either left-leaning or right-wing leaning. 
+People with different ideological backgrounds might experience that position differently. 
+This challenges us.
+We are interested in the effect of e.g. elite communication. 
+To study this, we allow for heterogeneous treatment effects in experimental work.
+This indicates that we often do not assume that the treatment, often using text, has the same effect for different partisans.
+Yet, at the same time, we forget or ignore that knowledge when creating large data-sets for our machine learning models. 
+To test whether people with different ideological backgrounds as well as their politicall knowledge might experience that position differently, challenging our ground-truth assumption in data annotation, we propose the following hypotheses:
 
-> **Ideological Bias hypothesis** (_H1a_): The larger the ideological distance between respondent and the party, the less likely respondents code stances according to the party’s position.
+> **Ideological Bias hypothesis** (_H1a_): The larger the ideological distance between respondent and the party, the less likely respondents annotate statements according to the party’s uttered position.
 
-> **Ideological Bias hypothesis** (_H1b_): The effect of H1a is stronger for the underspecified condition.
+> **Ideological Bias hypothesis** (_H1b_): The effect of H1a is stronger for sentences in which the party's position can be infered (i.e. underspecified sentences).
 
@@
 
+If these biases exists, how can we alleviate them?
+For the Austrian National Elections, @ennser2018impact already demonstrated that showing party labels impact annotators' assessment of the party position.
+So, one solution is masking:
+
 > **Masking Solution hypothesis** (_H3a_):  Masking reduces the effect of respondents’ ideological position for coding stances according to the party’s position.
@@
 > **Masking Solution hypothesis** (_H3b_): Masking reduces the effect of respondents’ level of political knowledge for coding stances according to the party’s position.
+
+Think about perspectivism for another solution!
 

### 2023-06-08 report/draft.Rmd
update paper

@@
 Yet, at the same time, we forget or ignore that knowledge when creating large data-sets for our machine learning models.
-To test whether people with different ideological backgrounds as well as their politicall knowledge might experience that position differently, challenging our ground-truth assumption in data annotation, we propose the following hypotheses:
+
+[Myrthe over NLP]
+
+To test whether people with different ideological backgrounds as well as their political knowledge might experience that position differently, challenging our ground-truth assumption in data annotation, we propose the following hypotheses:
 
@@
 > **Ideological Bias hypothesis** (_H1b_): The effect of H1a is stronger for sentences in which the party's position can be infered (i.e. underspecified sentences).
+
+[Myrthe over NLP]
+
+To test whether people with different ideological backgrounds as well as their political knowledge might experience that position differently, challenging our ground-truth assumption in data annotation, we propose the following hypotheses:
 
@@
 If these biases exists, how can we alleviate them?
+
+[Myrthe over NLP]
+
 For the Austrian National Elections, @ennser2018impact already demonstrated that showing party labels impact annotators' assessment of the party position.

### 2023-06-11 report/draft.Rmd
Added NLP papers and paragraphs

@@
 # Introduction
+
 Information is key to inform the behavior of citizens, and thereby for the social scientists studying them.
@@
 A recent study [@van2021validity] demonstrates that using crowd-coding platforms is a good way to collect such golden standards without too much disagreement.
-These platforms have also been used for experiments, and the quality of the respondents has been a center of attention  [@coppock2019generalizing;@clifford2015samples;@huff2015these;@berinsky2014separating].
-While some say that this data is of similar quality to data from a random sample of the population [@coppock2019generalizing], others have demonstrated that these online platforms are populated by people that are "unlike" the general public, being younger and holding more liberal values [@clifford2015samples;@huff2015these].
-If the latter is true, this might cause a problem for the coding of political texts:
-@ennser2018impact report that coders of political texts incorporate prior beliefs about parties' issue stances into their coding decisions.
+These platforms have also been used for experiments, and the quality of the respondents has been a center of attention [@coppock2019generalizing; @clifford2015samples; @huff2015these; @berinsky2014separating].
+While some say that this data is of similar quality to data from a random sample of the population [@coppock2019generalizing], others have demonstrated that these online platforms are populated by people that are "unlike" the general public, being younger and holding more liberal values [@clifford2015samples; @huff2015these].
+If the latter is true, this might cause a problem for the coding of political texts: @ennser2018impact report that coders of political texts incorporate prior beliefs about parties' issue stances into their coding decisions.
 The authors find that party labels cue coders to a stance.
 For example, coders are more likely to report a left-wing party to be pro-immigration and a populist right-wing party to be against based on the exact same sentence.
-_Is this actually bias or a diversity of view points? And how big of a problem does this bias/diversity of view points generate for scholars relying on golden standard data?_
+*Is this actually bias or a diversity of view points?
+And how big of a problem does this bias/diversity of view points generate for scholars relying on golden standard data?*
 
@@
 At the same time, trends of polarization have shown that people do interpret information according to their ideological position.
-_Does this mean that some are right and others are wrong? Or is there an ideological difference in the ground truth?_
-These questions presents a fundamental challenge to the main way of working when collecting gold standard data, as we operate from the baseline assumption that disagreement among the annotators should be avoided or reduced.
+*Does this mean that some are right and others are wrong?
+Or is there an ideological difference in the ground truth?* These questions presents a fundamental challenge to the main way of working when collecting gold standard data, as we operate from the baseline assumption that disagreement among the annotators should be avoided or reduced.
 Typically, when specific cases continuously cause disagreement, more instructions are added to limit interpretations [for recent innovations to improve annotation, see for example: @barbera2021automated; @struthers2020bridging; @winter2020online; @debell2013harder; @benoit2016crowd, @ying2022topics].
 However, work in computational linguistics has shown that increased annotation instructions do not increase quality [@parmar2022don].
-This leaves us between a rock and a hard place. Is there a potential bias in annotators that we should account for?
+This leaves us between a rock and a hard place.
+Is there a potential bias in annotators that we should account for?
 
@@
 To do so, we have fielded two high-powered pre-registered experiments (see [here](link) and [here](link)) in the Netherlands -- a low-level polarized country -- and in the U.S.
--- a high-level polarized country testing the effect of ideological distance between the annotator and the political actor in the text (H1), the effect of overinterpretation based on political knowledge or ideological engagement (H2), and a offered solution of masking the political actor to mitigate the efffects of ideology and knowledge (H3).[^1]
-In the experiments, we vary the level of specification with which a political actor takes a stance -- a declarative sentence versus a sentence where with some knowledge on politics, the stance might be inferred -- as well as whether the political actor is shown or masked with putting `[ACTOR]` instead of the political party.
-We do this for four different political issues: _Environment_, _Immigration_, _Tax Policy_, and _EU_ (for the Dutch case) or _Foreign Policy_ (for the American case).
+-- a high-level polarized country testing the effect of ideological distance between the annotator and the political actor in the text (H1), the effect of overinterpretation based on political knowledge or ideological engagement (H2), and a offered solution of masking the political actor to mitigate the efffects of ideology and knowledge (H3).[^1] In the experiments, we vary the level of specification with which a political actor takes a stance -- a declarative sentence versus a sentence where with some knowledge on politics, the stance might be inferred -- as well as whether the political actor is shown or masked with putting `[ACTOR]` instead of the political party.
+We do this for four different political issues: *Environment*, *Immigration*, *Tax Policy*, and *EU* (for the Dutch case) or *Foreign Policy* (for the American case).
 This country selection does not only allow us to showcase the scope conditions of disagreement due to different levels of political heterogeneity, but also differentiates between languages.
@@
 2023].
+
+[^1]: The data and research compendium is published on the [main author's github page](https://github.com/MarikenvdVelden/bias-experiment) -- annonymized for the review process.
 
@@
 
-[^1]: The data and research compendium is published on the [main author's github page](https://github.com/MarikenvdVelden/bias-experiment) -- annonymized for the review process.
+# Whose Truth is it Anyway? Disagreement & Perspectivism in Creating Gold Standard Data
 
-# Whose Truth is it Anyway? Disagreement & Perspectivism in Creating Gold Standard Data
 Generating large data-sets has become one of the main drivers of progress in natural language understanding.
@@
 A recent study [@van2021validity] demonstrates that using crowd-coding platforms is a good way to collect such large data-sets.
-However, having only a few workers annotate the majority of text of interest has raised concerns about data diversity and models' ability to generalize beyond the crowd-workers:
-In a series of experiments, @geva2019we show that often models do not generalize well to annotations from annotators that did not contribute to the training set, suggesting that annotator bias should be monitored during data-set creation.
+However, having only a few workers annotate the majority of text of interest has raised concerns about data diversity and models' ability to generalize beyond the crowd-workers: In a series of experiments, @geva2019we show that often models do not generalize well to annotations from annotators that did not contribute to the training set, suggesting that annotator bias should be monitored during data-set creation.
 One such potential bias, especially in times of increasing polarization [@iyengar2019origins; @gidron2019toward; @boxell2022cross], is based on ideological position of the annotator.
-Given that annotators on crows-coding platforms tend to be younger and hold more liberal values than the general public [@clifford2015samples;@huff2015these], this could potentially hamper the data diversity and generalizability of the model.
+Given that annotators on crows-coding platforms tend to be younger and hold more liberal values than the general public [@clifford2015samples; @huff2015these], this could potentially hamper the data diversity and generalizability of the model.
 An additional reason to monitor the annotators ideological position as a potential source of annotator bias is that a recent study in NLP showed that experiential factors influence the consistency of how political ideologies are perceived [@shen2021sounds].
@@
 This challenges us.
-We are interested in the effect of e.g. elite communication. 
+We are interested in the effect of e.g.
+elite communication.
 To study this, we allow for heterogeneous treatment effects in experimental work.
@@
 
-[Myrthe over NLP]
+The field of Natural Language Processing often works on automatically classifying texts on labels of concepts such as stance, sentiment, and political orientation.
+These models are trained with training data, which is annotated by human annotators.
+Often, this process has a final step where disagreements and differences in annotations are levelled by aggregating, averaging, or in other ways coming to a consensus on one label for one example, which is then seen as "ground truth"[@aroyo2015].
+Differences from this ground truth label are seen as errors that need to be removed or sorted out.
+Models then learn to predict labels for new examples based on this ground truth.
+However, since several years there is some discussion on how realistic it is to have one label, especially for subjective or complex concepts and/or texts with multiple interpretations.
+@aroyo2015 describes succinctly how the predominant annotation procedures for classification models run into myths such as "disagreement is bad" and "one annotation is enough".
+@plank2014learning adds to this that underlying ambiguity and linguistic complexity should be considered for disagreement in annotations: not all linguistic examples are created equal.
+Disagreement even occurs in seemingly objective tasks such as Part of Speech tagging [@fornaciari2021, @plank2014learning].
+
+Another aspect to consider is that disagreement can be informative for the concept under measure - sometimes agreement can be used to validate hypotheses about how universal the perceptions of such concepts are [@sommerauer].
+Additional doubts on smoothing out disagreement in annotation have focussed on the lack of diversity when only annotating with one label or annotator, leading to a homogeneity especially in subjective and social tasks [@geva2019we] such as hatespeech detection or political affiliation classification.
+The question then is: whose perspective is being recorded in these datasets, and then later in the models trained on these datasets?
+Or, as [@geva2019we] asked: are we modelling the task, or the annotator?
+
+Several tasks are intrinsically societal, with answers that differ based on the make-up of the community of annotators.
+The answer of the annotators in turn influence how machine learning examples classify new models.
+For instance, hate speech detection and abuse detection is one NLP task where the race and gender of annotators influences annotators and model performance[@gordon2022, @Larimore, @waseem2016].
+Language is inherently connected to society and culture: @shenDarling analyze sentiment analysis, and find that human annotators lead to certain perspectives on sentiment being recorded, and that notably that African American English dialects are often misunderstood by such models.
+
+Most recently, new annotation paradigms have gone one step further: maybe we are annotating the annotator rather than the task, and maybe that is actually beneficial to the models.
+@pavlick2019 find that for the logical coherence task Natural Language Inference, annotators have several valid interpretations that are not reflected in one ground truth label.
+They call for new training paradigms that can reflect "the full range of possible human inferences"[@pavlick2019, p.
+688].
+Recent approaches in NLP have sought to explicitly incorporate disagreement and diversity in training data annotations.
+[@röttger2021] introduces the idea of an explicitly subjective annotation paradigm existing in addition to one focussed on one label and "ground truth".
+Such a subjective annotation paradigm can be used for purposes where the goal is finding diverse perspectives on the task or concepts, and for models to model more accurately how humans interpret a task or text.
+Additionally, "perspectivism" [@basile2021toward] is a paradigm and research agenda where different perspectives are explicitly incorporated in the training data, and used by models to provide more human-like classifications.
+Another paradigm is "jury learning" [@gordon2022], in which machine learning models do not learn to replicate one specific ground truth, but are trained with different annotator juries to reflect the judgement of different populations.
+In both approaches, demographic and other individual aspects of the annotator are explicitly mentioned and highlighted as having an influence on classification performance, but is used as an asset rather than as an error.
+
+@troiano2021[p.2] also note how for a complex concept such as the "emotion" of a text, the annotator can make several assumptions during the annotation process on what is wanted, that are all valid and may or may not be useful in different contexts: "It is possible to assess one\'s own emotion after reading the text, to reconstruct the affective state of the writers who produced it, to guess the reaction that they intended to elicit in the readers, and so on.\" @shen2021sounds find that political ideology is not an inherent concept in many texts, but rather dependent on who is asked to annotate and their perceptions and background.
+Extralinguistic factors, such as annotator's own political ideology and also knowledge, influenced annotation and in turn model performance.
+@thornjakobsen2022 specifically analyze a task related to stance detection, argumentative sentence detection, on these extralinguistic factors, and find an effect of gender and political leaning on annotations and also model performance.
+However, to our knowledge this phenomenon has not yet been tested in a controlled experimental setting with manipulations in the texts.
 
@@
 
-> **Ideological Bias hypothesis** (_H1a_): The larger the ideological distance between respondent and the party, the less likely respondents annotate statements according to the party’s uttered position.
+> **Ideological Bias hypothesis** (*H1a*): The larger the ideological distance between respondent and the party, the less likely respondents annotate statements according to the party's uttered position.
 
-> **Ideological Bias hypothesis** (_H1b_): The effect of H1a is stronger for sentences in which the party's position can be infered (i.e. underspecified sentences).
+> **Ideological Bias hypothesis** (*H1b*): The effect of H1a is stronger for sentences in which the party's position can be infered (i.e. underspecified sentences).
 
-[Myrthe over NLP]
+As noted by @plank2014, some linguistic examples are ambiguous, and open to multiple interpretations even for seemingly objective tasks such as part of speech tagging, where annotators have to distinguish parts of speech such as nouns and verbs.
+A complex concept such as political ideology is much more likely to lead to multiple interpretations.
+We call such sentences with more possible interpretations and less explicit standpoints "underspecified".
+A lack of explicitness in the annotated text is one of the main causes of the disagreements in earlier literature.
+@thornjakobsen2022 deduce that annotator bias coms from a process nown as the affect heuristic [@slovic2007 ]: making a secision based on the emotional response related to your own personal attitude towards the discussed topic, especially when the text is relatively ambiguous.
 
@@
 
-> **Ideological Overinterpretation hypothesis** (_H2a_):  The larger the ideological distance between respondent and the party, the more likely respondents interpret underspecified sentences as stance.
+> **Ideological Overinterpretation hypothesis** (*H2a*): The larger the ideological distance between respondent and the party, the more likely respondents interpret underspecified sentences as stance.
 
-> **Political Knowledge Overinterpretation hypothesis** (_H2b_): The more political knowledge, the more likely people interpret underspecified sentences as stance.
+> **Political Knowledge Overinterpretation hypothesis** (*H2b*): The more political knowledge, the more likely people interpret underspecified sentences as stance.
 
@@
 
-[Myrthe over NLP]
+There have been several previous approaches to solve biased annotations, especially where it concerns political or societal aspects.
+@geva2019we introduce an approach where training set annotators are separated from annotators annotating datasets that evaluate the models, to ensure the evaluation is not simply accurate at replicating the original annotators, but can generalize to new annotators' judgements.
+However, these approaches are not aimed at reducing biases during the training set annotation procedure.
+Other approaches are aimed at leveraging multiple perspectives - but this is not useful when one wants one label to learn from.
 
-For the Austrian National Elections, @ennser2018impact already demonstrated that showing party labels impact annotators' assessment of the party position.
-So, one solution is masking:
+Very woke hot take that I am unsure whether to include and also where: "\"The problem is that framing arbitrary representations in data as \"bias\" misses the political character of datasets: there is no neutral data and no apolitical standpoint from where we can call out bias [23].
+Datasets are always \"a worldview\" [26] and, as such, data always remains biased.\" (Miceli et al., 2022, p.
+5)"
 
-> **Masking Solution hypothesis** (_H3a_):  Masking reduces the effect of respondents’ ideological position for coding stances according to the party’s position.
+For the Austrian National Elections, @ennser2018impact already demonstrated that showing party labels impact annotators' assessment of the party position. So, one solution is masking:
 
-> **Masking Solution hypothesis** (_H3b_): Masking reduces the effect of respondents’ level of political knowledge for coding stances according to the party’s position.
+> **Masking Solution hypothesis** (*H3a*): Masking reduces the effect of respondents' ideological position for coding stances according to the party's position.
+
+> **Masking Solution hypothesis** (*H3b*): Masking reduces the effect of respondents' level of political knowledge for coding stances according to the party's position.
 
@@
 # Data, Methods & Measurement
+
 ## Data
+
 We will conduct this survey experiment in the Netherlands in May 2022.
@@
 Therefore, we measure many demographic background variables, and balance checks have been conducted to demonstrate whether certain categories are over represented in a certain experimental group.
-The study has been approved by the [Research Ethics Review Committee](https://fsw.vu.nl/nl/onderzoek/research-ethics-review/index.aspx) of the _Vrije Universiteit Amsterdam_ (see the approval [here](https://github.com/MarikenvdVelden/bias-experiment/blob/master/docs/2022-3-30-59.pdf)).
+The study has been approved by the [Research Ethics Review Committee](https://fsw.vu.nl/nl/onderzoek/research-ethics-review/index.aspx) of the *Vrije Universiteit Amsterdam* (see the approval [here](https://github.com/MarikenvdVelden/bias-experiment/blob/master/docs/2022-3-30-59.pdf)).
 To ensure good quality of our data, one attention check (discussed in more detail in [Section 3.3](#attention-checks)) is included [@berinsky2014separating].
@@
 ## Measurement
-_**Experimental Conditions.**_ Respondents are randomly assigned to either view a political party as an actor, or a masked condition, where they see `X` as an actor; simultaneously, respondents see either a fully specified sentence or a underspecified sentence, in which one needs additional information to interpret the positon on an actor.
+
+***Experimental Conditions.*** Respondents are randomly assigned to either view a political party as an actor, or a masked condition, where they see `X` as an actor; simultaneously, respondents see either a fully specified sentence or a underspecified sentence, in which one needs additional information to interpret the positon on an actor.
 Table \ref{tab:conditions} gives an overview of the variations in treatment in the survey as well as their English translations.
@@
 
-_**Dependent Variable.**_  We rely on whether or not a party's (implied) stance is coded according to the party’s position (H1 and H3) as well as whether or not the statement is coded as a stance at all (H3).
+***Dependent Variable.*** We rely on whether or not a party's (implied) stance is coded according to the party's position (H1 and H3) as well as whether or not the statement is coded as a stance at all (H3).
 For each issue, we ask the respondent `what is according to the sentence above the position of [ACTOR]?`, with the answer categories: `in favor`, `against`, `no stance`, `don't know`.
 
-_**Moderating Covariates.**_ _Ideological position_ is measured using an 11-point scale ranging from left (`0`) to right (`10`). - _Political knowledge_ is measured with six items from the DPES.
+***Moderating Covariates.*** *Ideological position* is measured using an 11-point scale ranging from left (`0`) to right (`10`). - *Political knowledge* is measured with six items from the DPES.
 
-_**Controle Variables.**_ In our analysis, we control for demographic information (gender, age, education, income, religion, job) as well as political background variables (trust in politics, ideological position on economic left-right scale and cultural progressive-conservative scale, and evaluations and prospects of the economy).
+***Controle Variables.*** In our analysis, we control for demographic information (gender, age, education, income, religion, job) as well as political background variables (trust in politics, ideological position on economic left-right scale and cultural progressive-conservative scale, and evaluations and prospects of the economy).
 Tables A.5 till A.17 in the OA demonstrate the descriptive information per country.
@@
 ## Method
+
 To test our hypotheses, we will conduct a multilevel model, with respondents clustered in issues, see Equation \ref{eq:pooled1}.
@@
 
+```{=tex}
 \begin{multline}\label{eq:pooled1}
@@
 \end{multline}
+```
+# Results
 
-# Results
 To answer whether there is an ideological or knowledge-basd annotation bias, we have conducted a two-by-two experiment.
@@
 However, if people see an X compared to a political actor, they are statistically significantly more likely to interpret the sentence as a stance.
-Yet, a coefficient of `0.02` (i.e. `2%`) is a very small effect.
+Yet, a coefficient of `0.02` (i.e.
+`2%`) is a very small effect.
 For the condition of specification level, however, we see that compared to a fully specified sentence, people seeing an under-specified sentence are much more likely to interpret the sentence as a stance: an increase of `0.83`.
@@
 ## Pre-registered Results
+
 First, we test whether there is an ideological bias in interpreting stances (H1a), and if this bias increases for those that are further away from the ideological position of the political actor in the under-specified condition (H1b).
@@
 The slope for political knowledge, on the right-hand panel of Figure \ref{fig:results-h3}, is so-possible even flatter.
-This means that masking of political actors does not help to correctly interpret a sentence as a stance -- i.e. no support for H3a and H3b.
+This means that masking of political actors does not help to correctly interpret a sentence as a stance -- i.e.
+no support for H3a and H3b.
 
@@
 ## Exploratory Results
+
 To check the robustness of our findings, Figure \ref{fig:results-exp1} demonstrates the analyses for each issue separately.
@@
 We see some variation between issues.
-For the issue _Tax_ (lowest row), we see that almost everyone interprets the sentence correct -- an intercept of `0.96`.
+For the issue *Tax* (lowest row), we see that almost everyone interprets the sentence correct -- an intercept of `0.96`.
 This increases statistically significantly still with about `1%` when masking the political actor, and decreases statistically significantly with `68%` when the sentence is under-specified.
-The regression results (in Online Appendix) show a small effect of ideological distance: the further away, the more likely to correctly interpret the stance -- i.e. against H1a.
-Respondents found it most difficult to correctly interpret the issue _Environment_ (top row of Figure \ref{fig:results-exp1}), only `12%` has this correct initially, as indicated by the intercept. 
+The regression results (in Online Appendix) show a small effect of ideological distance: the further away, the more likely to correctly interpret the stance -- i.e.
+against H1a.
+Respondents found it most difficult to correctly interpret the issue *Environment* (top row of Figure \ref{fig:results-exp1}), only `12%` has this correct initially, as indicated by the intercept.
 This increases statistically significantly still with about `8%` when masking the political actor, and decreases statistically significantly with `15%` when the sentence is under-specified.
-The other two issues -- _EU_ and _Immigration_, in the middle rows of Figure \ref{fig:results-exp1} show a similar effect as the main base-line presented in Figure \ref{fig:results-base}.
+The other two issues -- *EU* and *Immigration*, in the middle rows of Figure \ref{fig:results-exp1} show a similar effect as the main base-line presented in Figure \ref{fig:results-base}.
 The same pattern holds for the other dependent variable -- overinterpretation of a stance -- in the right-hand panel of Figure \ref{fig:results-exp1}.
@@
 # Discussion
+
 TBA
@@
 ## Recommendations
+
 TBA
@@
 \newpage
+
 # References

### 2023-06-12 report/draft.Rmd
integration NLP text

@@
 People with different ideological backgrounds might experience that position differently.
-This challenges us.
+This challenges our way of data collection.
 We are interested in the effect of e.g.
@@
 The field of Natural Language Processing often works on automatically classifying texts on labels of concepts such as stance, sentiment, and political orientation.
-These models are trained with training data, which is annotated by human annotators.
-Often, this process has a final step where disagreements and differences in annotations are levelled by aggregating, averaging, or in other ways coming to a consensus on one label for one example, which is then seen as "ground truth"[@aroyo2015].
+These models are trained on data created by human annotators.
+Often, this process has a final step where disagreements and differences in annotations are leveled by aggregating, averaging, or in other ways coming to a consensus on one label for one example, which is then seen as "ground truth" [@aroyo2015].
 Differences from this ground truth label are seen as errors that need to be removed or sorted out. 
@@
 
-Another aspect to consider is that disagreement can be informative for the concept under measure - sometimes agreement can be used to validate hypotheses about how universal the perceptions of such concepts are [@sommerauer].
+Another aspect to consider is that disagreement can be informative for the concept under measure - sometimes agreement can be used to validate hypotheses about how universal the perceptions of such concepts are [@sommerauer2020penguin].
 Additional doubts on smoothing out disagreement in annotation have focussed on the lack of diversity when only annotating with one label or annotator, leading to a homogeneity especially in subjective and social tasks [@geva2019we] such as hatespeech detection or political affiliation classification.
-The question then is: whose perspective is being recorded in these datasets, and then later in the models trained on these datasets?
-Or, as [@geva2019we] asked: are we modelling the task, or the annotator?
-
-Several tasks are intrinsically societal, with answers that differ based on the make-up of the community of annotators.
+The question then is: _Whose perspective is being recorded in these datasets, and then later in the models trained on these datasets?_ 
+Framing arbitrary representations in data as \"bias\" misses the political character of datasets: There is no neutral data and no apolitical standpoint from where we can call out bias.
+Datasets are always \"a worldview\" [26] and, as such, data always remains biased.\" (Miceli et al., 2022, p.
+5).
+This is key to social scientist in general, and those studying political text in particular, since several tasks of interest are intrinsically societal, with answers that differ based on the make-up of the worldview of annotators.
 The answer of the annotators in turn influence how machine learning examples classify new models. 
@@
 
-Most recently, new annotation paradigms have gone one step further: maybe we are annotating the annotator rather than the task, and maybe that is actually beneficial to the models.
+Most recently, new annotation paradigms have gone one step further by asking whether we  modelling the task, or the annotator [@geva2019we]?
 @pavlick2019 find that for the logical coherence task Natural Language Inference, annotators have several valid interpretations that are not reflected in one ground truth label.
@@
 @thornjakobsen2022 deduce that annotator bias coms from a process nown as the affect heuristic [@slovic2007 ]: making a secision based on the emotional response related to your own personal attitude towards the discussed topic, especially when the text is relatively ambiguous.
-
 To test whether people with different ideological backgrounds as well as their political knowledge might experience that position differently, challenging our ground-truth assumption in data annotation, we propose the following hypotheses:
@@
 
-If these biases exists, how can we alleviate them?
-
+_If these biases exists, how can we alleviate them?_
 There have been several previous approaches to solve biased annotations, especially where it concerns political or societal aspects.
@@
 Other approaches are aimed at leveraging multiple perspectives - but this is not useful when one wants one label to learn from.
-
-Very woke hot take that I am unsure whether to include and also where: "\"The problem is that framing arbitrary representations in data as \"bias\" misses the political character of datasets: there is no neutral data and no apolitical standpoint from where we can call out bias [23].
-Datasets are always \"a worldview\" [26] and, as such, data always remains biased.\" (Miceli et al., 2022, p.
-5)"
-
 For the Austrian National Elections, @ennser2018impact already demonstrated that showing party labels impact annotators' assessment of the party position. So, one solution is masking:
@@
 > **Masking Solution hypothesis** (*H3b*): Masking reduces the effect of respondents' level of political knowledge for coding stances according to the party's position.
-
-Think about perspectivism for another solution!
 

### 2023-06-12 report/draft.Rmd
integration NLP text

@@
 - affiliation: Dep. of Communciation Science, Vrije Universiteit Amsterdam
+- name: Myrthe Reuver
+- affiliation: Computational Linguistics & Text Mining Lab, Vrije Universiteit Amsterdam
 - name: Wouter van Atteveldt
@@
 - affiliation: Center for Social Media & Politics, New York University
-- name: Myrthe Reuver
-- affiliation: Computational Linguistics & Text Mining Lab, Vrije Universiteit Amsterdam
 - name: Kasper Welbers

### 2023-06-14 report/draft.Rmd

@@
 
-Information is key to inform the behavior of citizens, and thereby for the social scientists studying them.
+Information is key to inform the behavior of citizens, and thereby for the social scientists studthem.
 Classical theories of political communication, such as agenda setting or framing [e.g.
@@
 In addition, advances in Natural Language Processing (NLP) have made it possible to automatically analyze large quantities of data using machine learning [e.g., see @bender2016linguistic; @wei2023overview].
-The performance of such machine learning techniques are compared to so-called "golden standards", created by human annotators.
-Most data collection efforts to create such standards assume that there is only one correct interpretation for every input example, and that disagreement between the annotators is something that needs to be dealt with at all costs [@aroyo2015truth].
+The performance of such machine learning techniques are compared to so-called "gold standards", created by human annotators.
+Most data collection efforts to create such standards assume that there is only one correct interpretation for every input example, and that disagreement between the annotators is a problem that needs to be solved, e.g.
+by improving the codebook or coder training [@aroyo2015truth].
 A recent study [@van2021validity] demonstrates that using crowd-coding platforms is a good way to collect such golden standards without too much disagreement.
@@
 Or is there an ideological difference in the ground truth?* These questions presents a fundamental challenge to the main way of working when collecting gold standard data, as we operate from the baseline assumption that disagreement among the annotators should be avoided or reduced.
-Typically, when specific cases continuously cause disagreement, more instructions are added to limit interpretations [for recent innovations to improve annotation, see for example: @barbera2021automated; @struthers2020bridging; @winter2020online; @debell2013harder; @benoit2016crowd, @ying2022topics].
+Typically, when specific cases continuously cause disagreement, more instructions are added to limit interpretations (for recent innovations to improve annotation, see for example: @barbera2021automated; @struthers2020bridging; @winter2020online; @debell2013harder; @benoit2016crowd, @ying2022topics).
 However, work in computational linguistics has shown that increased annotation instructions do not increase quality [@parmar2022don].
@@
 2023].
+
+<!-- WvA: check referentie dolinsky hierboven. Ik zou trouwens de [link]s naar de prereg ook in de voetnoot zetten denk ik? -->
 
@@
 @plank2014learning adds to this that underlying ambiguity and linguistic complexity should be considered for disagreement in annotations: not all linguistic examples are created equal.
-Disagreement even occurs in seemingly objective tasks such as Part of Speech tagging [@fornaciari2021, @plank2014learning].
+Disagreement even occurs in seemingly objective tasks such as Part of Speech tagging [@fornaciari2021; @plank2014learning].
 
@@
 Framing arbitrary representations in data as \"bias\" misses the political character of datasets: There is no neutral data and no apolitical standpoint from where we can call out bias.
-Datasets are always \"a worldview\" [26] and, as such, data always remains biased.\" (Miceli et al., 2022, p.
+Datasets are always \"a worldview\" and, as such, data always remains biased.\" (Miceli et al., 2022, p.
 5).
@@
 
-Most recently, new annotation paradigms have gone one step further by asking whether we  modelling the task, or the annotator [@geva2019we]?
+<!-- WvA: ik heb de [26] weggehaald, maar dat was in Miceli ongetwijfeld een referentie, ik weet niet wat een nette manier is om dat te doen. Miceli moet nog in bibtex gezet worden trouwens -->
+
+Most recently, new annotation paradigms have gone one step further by asking whether we are modeling the task, or the annotator [@geva2019we]?
 @pavlick2019 find that for the logical coherence task Natural Language Inference, annotators have several valid interpretations that are not reflected in one ground truth label.
@@
 
-***Controle Variables.*** In our analysis, we control for demographic information (gender, age, education, income, religion, job) as well as political background variables (trust in politics, ideological position on economic left-right scale and cultural progressive-conservative scale, and evaluations and prospects of the economy).
+***Control Variables.*** In our analysis, we control for demographic information (gender, age, education, income, religion, job) as well as political background variables (trust in politics, ideological position on economic left-right scale and cultural progressive-conservative scale, and evaluations and prospects of the economy).
 Tables A.5 till A.17 in the OA demonstrate the descriptive information per country.
@@
 This balance test indicated that none of the variables are unbalanced over the experimental groups, and therefore, as pre-registered, will not be added to the regression formula.
-$Y\hat{Y}_{r, i, t}$ in Equation \ref{eq:pooled1} denotes the evaluation of a stance by respondent $r$, during issue $i$ and at experimental round $t$ -- ranging from round 1 to round 4.
+$\hat{Y}_{r, i, t}$ in Equation \ref{eq:pooled1} denotes the evaluation of a stance by respondent $r$, during issue $i$ and at experimental round $t$ -- ranging from round 1 to round 4.
 The standard errors are clustered at the individual level.
@@
 \begin{multline}\label{eq:pooled1}
-    \hat{stance correct}_{r, i, t} =\beta_{0} + \beta_{1}masked_{r, i, t} +
-    \beta_{2}specification_{r, i, t} + 
-    \beta_{3}ideological distance to party_{r, i, t} +\\
-    \beta_{4}political knowledge_{r, i, t} +
+    \hat{Y}_{r, i, t} =\beta_{0} + 
+    \beta_{1}\cdot masked_{r, i, t} +
+    \beta_{2}\cdot specification_{r, i, t} + 
+    \beta_{3}\cdot ideological distance to party_{r, i, t} +\\
+    \beta_{4}\cdot political knowledge_{r, i, t} +
     \alpha_{i} + \gamma_{t} + \varepsilon_{r, i, t}
@@
 ```
+
+<!-- WvA:
+
 # Results
@@
 
-## Pre-registered Results
+## Results on pre-registered hypotheses
 

### 2023-06-14 report/draft.Rmd

--- 2023-06-12 report/draft.Rmd
@@
 To answer whether there is an ideological or knowledge-basd annotation bias, we have conducted a two-by-two experiment.
-Table \ref{tab:profile} demonstrates the average profile of respondents who annotate correctly and incorrectly.
+Tables \ref{tab:profilenl} and \ref{tab:profileus} demonstrate the average profile of respondents who annotate correctly and incorrectly (where respondents who annotated some stances correctly and some incorrectly are weighted by proportion (in)correct).
 In terms of demographics, there is not much of a difference.
@@
 
-```{r "profile"}
+```{r "profilenl"}
 load(here("data/intermediate/cleaned_data_nl.RData"))
@@
   column_spec(2, width = "7cm")
-
+```
+```{r "profileus"}
 kbl(e1_us, booktabs =T, caption = "\\label{tab:conditions_us}Profile American Stance Annotators") %>%

### 2023-06-14 report/draft.Rmd
some editing and adding missing bib entries

@@
 These enourmous amounts of data are typically handled by machine learning techniques to classify into meaningful variables.
-The performance of these models are compared to a so-called golden standards, created by human annotators.
+The performance of these models is evaluated based on a gold standard, created by human annotators.
 Having a high level of agreement between these annotators is key, but some suggest personal characteristics of annotators, like political ideology or knowledge, interfere.
@@
 @vanaelst2016political; @lecheler2019news], formulate that political information drives opinion formation and participation in politics -- from voting to protests [for overviews hereof across countries, see @pfetsch2013comparing].
-The democratization of data has lead to numerous possibilities to gather and analyze textual data [for a recent overview, see @vanatteveldt2022computational].
-In addition, advances in Natural Language Processing (NLP) have made it possible to automatically analyze large quantities of data using machine learning [e.g., see @bender2016linguistic; @wei2023overview].
-The performance of such machine learning techniques are compared to so-called "golden standards", created by human annotators.
-Most data collection efforts to create such standards assume that there is only one correct interpretation for every input example, and that disagreement between the annotators is something that needs to be dealt with at all costs [@aroyo2015truth].
+The democratization of data and advent of computational social science has paved the way for new possibilities to gather and analyze textual data [for a recent overview, see @vanatteveldt2022computational].
+In particular, advances in Natural Language Processing (NLP) have made it possible to automatically analyze large quantities of data using machine learning [e.g., see @bender2016linguistic; @wei2023overview].
+To determine the validity of such large scale, computational analyses, we rely on "gold standard" data, created by human annotators. 
+If follows that the validity of these analyses hinges on the quality of these golden standards.
+If a gold standard contains biases, i.e., systematic errors, it can foster bias in any downstream analysis.
+
+Most data collection efforts to create gold standards assume that there is only one correct interpretation for every input example, and that disagreement between the annotators is something that needs to be dealt with at all costs [@aroyo2015truth].
 A recent study [@van2021validity] demonstrates that using crowd-coding platforms is a good way to collect such golden standards without too much disagreement.
 These platforms have also been used for experiments, and the quality of the respondents has been a center of attention [@coppock2019generalizing; @clifford2015samples; @huff2015these; @berinsky2014separating].
-While some say that this data is of similar quality to data from a random sample of the population [@coppock2019generalizing], others have demonstrated that these online platforms are populated by people that are "unlike" the general public, being younger and holding more liberal values [@clifford2015samples; @huff2015these].
+While it has been argued that this data is of similar quality to data from a random sample of the population [@coppock2019generalizing], others have demonstrated that these online platforms are populated by people that are "unlike" the general public, being younger and holding more liberal values [@clifford2015samples; @huff2015these].
 If the latter is true, this might cause a problem for the coding of political texts: @ennser2018impact report that coders of political texts incorporate prior beliefs about parties' issue stances into their coding decisions.
@@
 *Does this mean that some are right and others are wrong?
-Or is there an ideological difference in the ground truth?* These questions presents a fundamental challenge to the main way of working when collecting gold standard data, as we operate from the baseline assumption that disagreement among the annotators should be avoided or reduced.
+Or is there an ideological difference in the ground truth?* These questions present a fundamental challenge to the main way of working when collecting gold standard data, as we operate from the baseline assumption that disagreement among the annotators should be avoided or reduced.
 Typically, when specific cases continuously cause disagreement, more instructions are added to limit interpretations [for recent innovations to improve annotation, see for example: @barbera2021automated; @struthers2020bridging; @winter2020online; @debell2013harder; @benoit2016crowd, @ying2022topics].
@@
 To do so, we have fielded two high-powered pre-registered experiments (see [here](link) and [here](link)) in the Netherlands -- a low-level polarized country -- and in the U.S.
--- a high-level polarized country testing the effect of ideological distance between the annotator and the political actor in the text (H1), the effect of overinterpretation based on political knowledge or ideological engagement (H2), and a offered solution of masking the political actor to mitigate the efffects of ideology and knowledge (H3).[^1] In the experiments, we vary the level of specification with which a political actor takes a stance -- a declarative sentence versus a sentence where with some knowledge on politics, the stance might be inferred -- as well as whether the political actor is shown or masked with putting `[ACTOR]` instead of the political party.
+-- a high-level polarized country testing the effect of ideological distance between the annotator and the political actor in the text (H1), the effect of overinterpretation based on political knowledge or ideological engagement (H2), and a offered solution of masking the political actor to mitigate the effects of ideology and knowledge (H3).[^1] In the experiments, we vary the level of specification with which a political actor takes a stance -- a declarative sentence versus a sentence where with some knowledge on politics, the stance might be inferred -- as well as whether the political actor is shown or masked with putting `[ACTOR]` instead of the political party.
 We do this for four different political issues: *Environment*, *Immigration*, *Tax Policy*, and *EU* (for the Dutch case) or *Foreign Policy* (for the American case).
@@
 A lack of explicitness in the annotated text is one of the main causes of the disagreements in earlier literature.
-@thornjakobsen2022 deduce that annotator bias coms from a process nown as the affect heuristic [@slovic2007 ]: making a secision based on the emotional response related to your own personal attitude towards the discussed topic, especially when the text is relatively ambiguous.
+@thornjakobsen2022 deduce that annotator bias comes from a process known as the affect heuristic [@slovic2007affect ]: making a decision based on the emotional response related to your own personal attitude towards the discussed topic, especially when the text is relatively ambiguous.
 To test whether people with different ideological backgrounds as well as their political knowledge might experience that position differently, challenging our ground-truth assumption in data annotation, we propose the following hypotheses:
@@
 ```
+
 # Results

### 2023-06-15 src/data-processing/prepare-data.Rmd
add variable for strict and lenient stance

@@
 ### Dependent Variable by Issue
-```{r "Dependent Variable", echo=F, fig.width=15, fig.height=8}
+```{r "Dependent Variable", echo=F, fig.width=15, fig.height=8, eval = F}
 p1 <- d_nl %>%
-  group_by(stance,issue,specification, masking) %>% 
+  group_by(stance_nlp,issue,specification, masking) %>% 
   count() %>% 
@@
 ### Covariates
-```{r "Independent Variables", echo=F, message=FALSE, warning=FALSE,fig.width=10, fig.height=10}
+```{r "Independent Variables", echo=F, message=FALSE, warning=FALSE,fig.width=10, fig.height=10, eval = F}
 rm(p1,p2)
@@
 ### Correlations Matrix
-```{r "Correlations Matrix", echo=F, message=F, warning=F, fig.width=15, fig.height=15}
+```{r "Correlations Matrix", echo=F, message=F, warning=F, fig.width=15, fig.height=15, eval = F}
 d_nl %>%

### 2023-06-15 src/data-processing/clean_data_nl.R

@@ -101,22 +101,44 @@ d_nl <- d_nl %>%
          specified4 = recode(specified4,
                              `zegt dat het lidmaatschap van de Europese Unie tot nu toe vooral slecht geweest voor Nederland is.` = "Specified",
                              `zegt dat Nederland een andere rol in de Europese Unie moet hebben.` = "Underspecified"),
-         stance1 = ifelse(specified1 == "Specified" & S1 == 2, 
+         stance1_nlp = ifelse(specified1 == "Specified" & S1 == 2, 
                           "correct", "not correct"),
-         stance1 = ifelse(specified1 == "Underspecified" & S1 == 3, 
-                          "correct", stance1),
-         stance2 = ifelse(specified2 == "Specified" & S2 == 1, 
+         stance1_nlp = ifelse(specified1 == "Underspecified" & S1 == 3, 
+                          "correct", stance1_nlp),
+         stance1_ss = ifelse(specified1 == "Specified" & S1 == 2, 
+                              "correct", "not correct"),
+         stance1_ss = ifelse(specified1 == "Underspecified" & S1 == 3, 
+                              "correct", stance1_ss),
+         stance2_nlp = ifelse(specified2 == "Specified" & S2 == 1, 
+                              "correct", "not correct"),
+         stance2_nlp = ifelse(specified2 == "Underspecified" & S2 == 3, 
+                              "correct", stance2_nlp),
+         stance2_ss = ifelse(specified2 == "Specified" & S2 == 1, 
                           "correct", "not correct"),
-         stance2 = ifelse(specified2 == "Underspecified" & S2 == 3,
-                          "correct", stance2),
-         stance3 = ifelse(specified3 == "Specified" & S3 == 1,
+         stance2_ss = ifelse(specified2 == "Underspecified" & S2 == 2, 
+                             "correct", stance2_ss),
+         stance2_ss= ifelse(specified2 == "Underspecified" & S2 == 3,
+                          "correct", stance2_ss),
+         stance3_nlp = ifelse(specified3 == "Specified" & S3 == 1,
                           "correct", "not correct"),
-         stance3 = ifelse(specified3 == "Underspecified" & S3 == 3, 
-                          "correct", stance3),
-         stance4 = ifelse(specified4 == "Specified" & S4 == 2, 
+         stance3_nlp = ifelse(specified3 == "Underspecified" & S3 == 3, 
[…]

### 2023-06-15 src/data-processing/clean_data_us.R

@@ -166,22 +166,44 @@ d_us <- d_us %>%
          specified4 = recode(specified4,
                              `say there should be a different military presence in the Pacific Ocean.` = "Underspecified",
                              `say the U.S. needs to consider military build-up in the Pacific Ocean` = "Specified"),
-         stance1 = ifelse(specified1 == "Specified" & S1 == 2, 
-                          "correct", "not correct"),
-         stance1 = ifelse(specified1 == "Underspecified" & S1 == 3, 
-                          "correct", stance1),
-         stance2 = ifelse(specified2 == "Specified" & S2 == 1, 
-                          "correct", "not correct"),
-         stance2 = ifelse(specified2 == "Underspecified" & S2 == 3,
-                          "correct", stance2),
-         stance3 = ifelse(specified3 == "Specified" & S3 == 1,
-                          "correct", "not correct"),
-         stance3 = ifelse(specified3 == "Underspecified" & S3 == 3, 
-                          "correct", stance3),
-         stance4 = ifelse(specified4 == "Specified" & S4 == 1, 
-                          "correct", "not correct"),
-         stance4 = ifelse(specified4 == "Underspecified" & S4 == 3,
-                          "correct", stance4),
+         stance1_nlp = ifelse(specified1 == "Specified" & S1 == 2, 
+                              "correct", "not correct"),
+         stance1_nlp = ifelse(specified1 == "Underspecified" & S1 == 3, 
+                              "correct", stance1_nlp),
+         stance1_ss = ifelse(specified1 == "Specified" & S1 == 2, 
+                             "correct", "not correct"),
+         stance1_ss = ifelse(specified1 == "Underspecified" & S1 == 3, 
+                             "correct", stance1_ss),
+         stance2_nlp = ifelse(specified2 == "Specified" & S2 == 1, 
+                              "correct", "not correct"),
+         stance2_nlp = ifelse(specified2 == "Underspecified" & S2 == 3, 
+                              "correct", stance2_nlp),
+         stance2_ss = ifelse(specified2 == "Specified" & S2 == 1, 
+                             "correct", "not correct"),
+         stance2_ss = ifelse(specified2 == "Underspecified" & S2 == 2, 
+                             "correct", stance2_ss),
[…]

### 2023-06-20 report/draft.Rmd
update analyses: lenient and strict stance perception

@@
 For other positions on issues or political knowledge, we do not see a difference in averages between those who are correctly and incorrectly identifying stances.
+This profile is quite similar for the lenient interpretation of what a stance is.
 
-```{r "profilenl"}
+```{r "profilenl-strict"}
 load(here("data/intermediate/cleaned_data_nl.RData"))
@@
 source(here("src/analysis/profile_resp.R"))
-kbl(e1, booktabs =T, caption = "\\label{tab:conditions}Profile Dutch Stance Annotators") %>%
+kbl(e1_nlp, booktabs =T, caption = "\\label{tab:profilenl}Profile Dutch Stance Annotators") %>%
   kable_styling(latex_options = c("striped", "hold_position"),
@@
 ```
+
 ```{r "profileus"}
-kbl(e1_us, booktabs =T, caption = "\\label{tab:conditions_us}Profile American Stance Annotators") %>%
+kbl(e1_us_nlp, booktabs =T, caption = "\\label{tab:conditions_us}Profile American Stance Annotators") %>%
   kable_styling(latex_options = c("striped", "hold_position"),
@@
 
-```{r results-h1, out.width= "95%", fig.align = 'center', fig.cap = "\\label{fig:results-h1} Ideological Distance"}
-knitr::include_graphics(here("report/figures", "h1-3.png"))
-```
-
 Secondly, we hypothesized that there is a risk of over-interpretation from those that are ideologically distant to the political actor (H2a) as well as those who have high levels of political knowledge (H2b).
@@
 
-```{r results-h2, out.width = "95%", out.height = "75%", fig.align = 'center', fig.cap = "\\label{fig:results-h2} Results Level of Sentence Specification"}
-knitr::include_graphics(here("report/figures", "h2-1.png"))
+\begin{landscape}
+```{r results-h1, fig.align = 'center', fig.cap = "\\label{fig:results-h1} Ideological Distance", fig.pos="H", fig.show="hold",out.width = "49%", out.height="40%"}
+knitr::include_graphics(c(here("report/figures", "h1-1.png"), here("report/figures", "h1-2.png")))
 ```
+```{r results-h2, fig.align = 'center', fig.cap = "\\label{fig:results-h2} Results Level of Sentence Specification", fig.pos="H", fig.show="hold",out.width = "49%", out.height="40%"}
+knitr::include_graphics(c(here("report/figures", "h2-1.png"), here("report/figures", "h2-2.png")))
+```
+\end{landscape}
 
@@
 
-```{r results-h3, out.width = "95%",  out.height = "75%", fig.align = 'center', fig.cap = "\\label{fig:results-h3} Results Masking Solution"}
+```{r results-h3, fig.align = 'center', fig.cap = "\\label{fig:results-h3} Results Masking Solution", fig.pos="H", fig.show="hold",out.width = "59%", out.height="50%"}
 knitr::include_graphics(here("report/figures", "h3-1.png"))
@@
 
-```{r results-exploration1, out.height = "85%",out.width = "95%",, fig.align = 'center', fig.cap = "\\label{fig:results-exp1} Exploration: Issue Speficific Analyses"}
-knitr::include_graphics(here("report/figures", "issues-1.png"))
+```{r results-exploration1, out.height = "85%",out.width = "95%",, fig.align = 'center', fig.cap = "\\label{fig:results-exp1} Exploration: Issue Speficific Analyses", eval = F}
+knitr::include_graphics(c(here("report/figures", "issues-1.png")))
 ```
@@
 
-```{r results-exploration2, out.width = "95%", out.height= "75%", fig.align = 'center', fig.cap = "\\label{fig:results-exp2} Exploration: Interaction with Treatments"}
+```{r results-exploration2, out.width = "95%", out.height= "75%", fig.align = 'center', fig.cap = "\\label{fig:results-exp2} Exploration: Interaction with Treatments", eval = F}
 knitr::include_graphics(here("report/figures", "exploration-3.png"))
@@
 
-```{r results-exploration3, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-exp3} Exploration: Different Indicators of Ideological Distance (1)"}
+```{r results-exploration3, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-exp3} Exploration: Different Indicators of Ideological Distance (1)", eval = F}
 knitr::include_graphics(here("report/figures", "exploration-1.png"))
@@
 
-```{r results-exploration4, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-exp4} Exploration: Different Indicators of Ideological Distance (2)"}
+```{r results-exploration4, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-exp4} Exploration: Different Indicators of Ideological Distance (2)", eval = F}
 knitr::include_graphics(here("report/figures", "exploration-2.png"))
@@
 
-```{r results-exploration5, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-exp5} Exploration: Different Indicators of Ideological Distance (3)"}
+```{r results-exploration5, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-exp5} Exploration: Different Indicators of Ideological Distance (3)", eval = F}
 knitr::include_graphics(here("report/figures", "exploration-3.png"))
@@
 
-```{r results-h2_expl, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-h2-expl} Results Level of Sentence Specification"}
+```{r results-h2_expl, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-h2-expl} Results Level of Sentence Specification", eval = F}
 knitr::include_graphics(here("report/figures", "h2-3.png"))

### 2023-06-20 src/analysis/Analysis.Rmd
update analyses: lenient and strict stance perception

@@
 b
-
-b_b
+rm(b_b, b1, b2, d, df, df_nl, df_us, tmp)
 ```
@@
 b_i
-
-b_ib
+rm(b1, b2, d, df, df_nl, df_us, tmp)
 ```
@@
 source(here("src/analysis/h1.R"))
-p1
-
+p1a
 p1b
-
-p1 / p1b +
-  plot_layout(heights = c(1, 1.3)) + 
-  plot_layout(guides = "collect") & theme(legend.position = 'bottom')
-
-p1_2 / p1b2 +
-  plot_layout(heights = c(1, 1.3)) 
+rm(df, df_nl, df_us, h1a, h1b, tmp)
 ```
@@
 source(here("src/analysis/h2.R"))
-p2
-
+p2a
 p2b
-
-p2c
+p2c1
+p2c2
+rm(df, df_nl, df_us, h2a, h2a_2, h2b, h2b_2, tmp, tmpp)
 ```
@@
 source(here("src/analysis/h3.R"))
-p3
-
+p3a
 p3b
-
-p3c
+p3c1
+p3c2
+rm(df, df_nl, df_us, h3a, h3a_2, h3b, h3b_2,h3b_bs, h3b_2bs, tmp, tmpp)
 ```
@@
 source(here("src/analysis/exploration.R"))
-p_e1
-p_e2
-p_e3_1
-p_e3_2
-p_b
+p_b1
 p_b2
+p_e1a
+p_e1b
 
 source(here("src/analysis/profile_resp.R"))
-kbl(e1, booktabs =T, caption = "\\label{tab:conditions}Profile Dutch Stance Annotators") %>%
+kbl(e1_nlp, booktabs =T, caption = "\\label{tab:conditions}Profile Dutch Stance Annotators") %>%
   kable_styling(latex_options = c("striped", "hold_position"),
@@
 
-kbl(e1_us, booktabs =T, caption = "\\label{tab:conditions_us}Profile American Stance Annotators") %>%
+kbl(e1_ss, booktabs =T, caption = "\\label{tab:conditions2}Profile Dutch Stance Annotators") %>%
+  kable_styling(latex_options = c("striped", "hold_position"),
+                full_width = F, fixed_thead = T, position = "center") %>%
+  column_spec(1, width = "7cm") %>%
+  column_spec(2, width = "7cm")
+
+kbl(e1_us_nlp, booktabs =T, caption = "\\label{tab:conditions_us}Profile American Stance Annotators") %>%
+  kable_styling(latex_options = c("striped", "hold_position"),
+                full_width = F, fixed_thead = T, position = "center") %>%
+  column_spec(1, width = "7cm") %>%
+  column_spec(2, width = "7cm")
+
+kbl(e1_us_ss, booktabs =T, caption = "\\label{tab:conditions_us2}Profile American Stance Annotators") %>%
   kable_styling(latex_options = c("striped", "hold_position"),

### 2023-06-20 src/analysis/baseline.R

@@ -1,48 +1,98 @@
 df_us <- d_us %>% 
-  select(stance, masking, specification, interpretation, country, issue)
+  mutate(country = "United States") %>% 
+  select(stance_nlp, stance_ss, masking, specification, interpretation_nlp, interpretation_ss, country, issue)
 df_nl <- d_nl %>% 
-  select(stance, masking, specification, interpretation, country, issue)
+  mutate(country = "The Netherlands") %>% 
+  select(stance_nlp, stance_ss, masking, specification, interpretation_nlp, interpretation_ss, country, issue)
 
 d <- df_nl %>% 
   add_case(df_us)
 
-b1 <- broom.mixed::tidy(lmer(stance ~ masking + specification + 
+b1 <- broom.mixed::tidy(lmer(stance_nlp ~ masking + specification + 
                          (country | issue), data= d)) %>%
-  mutate(y = "Y: Correctly Interpreting Stance",
+  mutate(y = "Y: Correctly Interpreting Stance (Strict Interpretation)",
          type = "Pooled Analysis")
 
-tmp <- broom.mixed::tidy(lmer(stance ~ masking + specification + 
+tmp <- broom.mixed::tidy(lmer(stance_ss ~ masking + specification + 
+                               (country | issue), data= d)) %>%
+  mutate(y = "Y: Correctly Interpreting Stance (Lenient Interpretation)",
+         type = "Pooled Analysis")
+
+b1 <- b1 %>% 
+  add_case(tmp)
+
+tmp <- broom.mixed::tidy(lmer(stance_nlp ~ masking + specification + 
                                 (1 | issue), data= df_us)) %>%
-  mutate(y = "Y: Correctly Interpreting Stance",
+  mutate(y = "Y: Correctly Interpreting Stance (Strict Interpretation)",
          type = "United States")
 
 b1 <- b1 %>% 
   add_case(tmp)
 
-tmp <- broom.mixed::tidy(lmer(stance ~ masking + specification + 
+tmp <- broom.mixed::tidy(lmer(stance_ss ~ masking + specification + 
+                                (1 | issue), data= df_us)) %>%
+  mutate(y = "Y: Correctly Interpreting Stance (Lenient Interpretation)",
+         type = "United States")
+
+b1 <- b1 %>% 
+  add_case(tmp)
+
+tmp <- broom.mixed::tidy(lmer(stance_nlp ~ masking + specification + 
[…]

### 2023-06-20 src/analysis/exploration.R

@@ -1,34 +1,36 @@
 df_us <- d_us %>% 
-  select(stance, interpretation, masking, specification, distance, PT7, country, issue)
+  select(stance_nlp, stance_ss, interpretation_nlp, interpretation_ss, masking, specification, distance, PT7, issue)
 df_nl <- d_nl %>% 
-  select(stance, interpretation, masking, specification, distance, PT7, country, issue)
+  select(stance_nlp, stance_ss, interpretation_nlp, interpretation_ss, masking, specification, distance, PT7, issue)
 
-d <- df_nl %>% 
-  add_case(df_us)
-
-df <- d %>% 
+df <- d_us %>% 
   mutate(b = specification,
          a = recode(masking,
                     `Masked` = 1,
                     `Party` = 0))
-b3_a <- lmer(stance ~ b * a +
-                 (country | issue), data= df)
+b3_a <- lmer(stance_nlp ~ b * a +
+                 (1 | issue), data= df)
+b3_a2 <- tidy(b3_a) %>% 
+  mutate(condition = "Condition: Underspecified Sentence",
+         y = "Y: Correctly Identifying Stance (Strict Interpretation)",
+         type = "United States")
 b3_a <- tidy(margins::margins(b3_a, variables = "b",
                                at = list("a" = 0:1))) %>% 
   mutate(condition = "Condition: Underspecified Sentence",
-         y = "Y: Correctly Identifying Stance",
-         type = "Pooled Analysis")
+         y = "Y: Correctly Identifying Stance (Strict Interpretation)",
+         type = "United States")
 
-df <- df_us %>% 
-  mutate(b = specification,
-         a = recode(masking,
-                    `Masked` = 1,
-                    `Party` = 0))
-tmp <- lmer(stance ~ b * a + (1 | issue), data= df)
+tmp <- lmer(stance_ss ~ b * a + (1 | issue), data= df)
+tmpp <- tidy(tmp) %>% 
+  mutate(condition = "Condition: Underspecified Sentence",
+         y = "Y: Correctly Identifying Stance (Lenient Interpretation)",
+         type = "United States")
+b3_a2 <- b3_a2 %>% 
+  add_case(tmpp)
 tmp <- tidy(margins::margins(tmp, variables = "b",
                              at = list("a" = 0:1))) %>% 
   mutate(condition = "Condition: Underspecified Sentence",
[…]

### 2023-06-20 src/analysis/h1.R

@@ -1,53 +1,66 @@
 df_us <- d_us %>% 
-  select(stance, masking, specification, distance, PT7, country, issue)
+  select(stance_nlp, stance_ss, masking, specification, distance, PT7, issue)
 df_nl <- d_nl %>% 
-  select(stance, masking, specification, distance, PT7, country, issue)
+  select(stance_nlp, stance_ss, masking, specification, distance, PT7, issue)
 
-d <- df_nl %>% 
-  add_case(df_us)
-
-h1a <- broom.mixed::tidy(lmer(stance ~ masking + specification + 
-              distance + PT7 +
-              (country | issue), data= d)) %>% 
-  mutate(hyp = "H1a",
-         type = "Pooled Analysis") %>% 
-  filter(term %in% c("distance"))
-
-tmp <- broom.mixed::tidy(lmer(stance ~ masking + specification + 
+h1a <- broom.mixed::tidy(lmer(stance_nlp ~ masking + specification + 
                                 distance + PT7 +
                                 (1 | issue), data= df_us)) %>% 
   mutate(hyp = "H1a",
-         type = "United States") %>% 
+         type = "United States",
+         y = "Interpreting Stance Correctly (Strict Interpretation)") %>% 
+  filter(term %in% c("distance"))
+
+tmp <- broom.mixed::tidy(lmer(stance_ss ~ masking + specification + 
+                                distance + PT7 +
+                                (1 | issue), data= df_us)) %>% 
+  mutate(hyp = "H1a",
+         type = "United States",
+         y = "Interpreting Stance Correctly (Lenient Interpretation)") %>% 
   filter(term %in% c("distance"))
 
 h1a <- h1a %>% 
   add_case(tmp)
 
-tmp <- broom.mixed::tidy(lmer(stance ~ masking + specification + 
+tmp <- broom.mixed::tidy(lmer(stance_nlp ~ masking + specification + 
                                 distance + PT7 +
                                 (1 | issue), data= df_nl)) %>% 
   mutate(hyp = "H1a",
-         type = "the Netherlands") %>% 
+         type = "The Netherlands",
+         y = "Interpreting Stance Correctly (Strict Interpretation)") %>% 
+  filter(term %in% c("distance"))
+
+h1a <- h1a %>% 
+  add_case(tmp)
+
+tmp <- broom.mixed::tidy(lmer(stance_ss ~ masking + specification + 
+                                distance + PT7 +
[…]

### 2023-06-20 src/analysis/h2.R

@@ -1,30 +1,38 @@
 df_us <- d_us %>% 
-  select(interpretation, masking, specification, distance, PT7, country, issue)
+  select(stance_nlp, stance_ss, interpretation_nlp, interpretation_ss,
+         masking, specification, distance, PT7, PT7b, issue)
 df_nl <- d_nl %>% 
-  select(interpretation, masking, specification, distance, PT7, country, issue)
-
-d <- df_nl %>% 
-  add_case(df_us)
-
-df <- d %>% 
-  mutate(b = specification,
-         a = distance)
-h2a <- lmer(interpretation ~ masking + b * a +
-              PT7 + (country | issue), data= df)
-h2a <- broom.mixed::tidy(margins::margins(h2a, variables = "b",
-                             at = list("a" = 0:8.5))) %>% 
-  mutate(hyp = "H2a: Ideological Distance \n(0=Low, 10 = High)",
-         type = "Pooled Analysis")
+  select(stance_nlp, stance_ss, interpretation_nlp, interpretation_ss,
+         masking, specification, distance, PT7, issue)
 
 df <- df_us %>% 
   mutate(b = specification,
          a = distance)
-tmp <- lmer(interpretation ~ masking + b * a +
+h2a <- lmer(interpretation_nlp ~ masking + b * a +
               PT7 + (1 | issue), data= df)
+h2a_2 <- tidy(h2a) %>% 
+  mutate(hyp = "H2a: Ideological Distance",
+         type = "United States",
+         y = "Overinterpreting Stance (Strict Interpretation)")
+h2a <- tidy(margins::margins(h2a, variables = "b",
+                             at = list("a" = 0:7))) %>% 
+  mutate(hyp = "H2a: Ideological Distance",
+         type = "United States",
+         y = "Overinterpreting Stance (Strict Interpretation)")
+
+tmp <- lmer(interpretation_ss ~ masking + b * a +
+              PT7 + (1 | issue), data= df)
+tmpp <- tidy(tmp) %>% 
+  mutate(hyp = "H2a: Ideological Distance",
+         type = "United States",
+         y = "Overinterpreting Stance (Lenient Interpretation)")
+h2a_2 <- h2a_2 %>% 
+  add_case(tmpp) 
 tmp <- tidy(margins::margins(tmp, variables = "b",
-                             at = list("a" = 0:5))) %>% 
[…]

### 2023-06-20 src/analysis/h3.R

@@ -1,34 +1,39 @@
 df_us <- d_us %>% 
-  select(stance, masking, specification, distance, PT7, country, issue)
+  select(stance_nlp, stance_ss, masking, specification, distance, PT7, PT7b, issue)
 df_nl <- d_nl %>% 
-  select(stance, masking, specification, distance, PT7, country, issue)
-
-d <- df_nl %>% 
-  add_case(df_us)
-
-df <- d %>% 
-  mutate(b = recode(masking,
-                    `Masked` = 1,
-                    `Party` = 0),
-         a = distance)
-h3a <- lmer(stance ~ specification + b * a +
-              PT7 + (country | issue), data= df)
-h3a <- broom.mixed::tidy(margins::margins(h3a, variables = "b",
-                             at = list("a" = 0:8.5))) %>% 
-  mutate(hyp = "H3a: Ideological Distance \n(0=Low, 10 = High)",
-         type = "Pooled Analysis")
+  select(stance_nlp, stance_ss, masking, specification, distance, PT7, issue)
 
 df <- df_us %>% 
   mutate(b = recode(masking,
                     `Masked` = 1,
                     `Party` = 0),
+         b= factor(b),
          a = distance)
-tmp <- lmer(stance ~ specification + b * a +
+h3a <- lmer(stance_nlp ~ specification + b * a +
               PT7 + (1 | issue), data= df)
+h3a_2 <- tidy(h3a) %>% 
+  mutate(hyp = "H3a: Ideological Distance",
+         type = "United States",
+         y = "Overinterpreting Stance (Strict Interpretation)")
+h3a <- broom.mixed::tidy(margins::margins(h3a, variables = "b",
+                             at = list("a" = 0:7))) %>% 
+  mutate(hyp = "H3a: Ideological Distance",
+         type = "United States",
+         y = "Overinterpreting Stance (Strict Interpretation)")
+
+tmp <- lmer(stance_ss ~ specification + b * a +
+              PT7 + (1 | issue), data= df)
+tmpp <- tidy(tmp) %>% 
+  mutate(hyp = "H3a: Ideological Distance",
+         type = "United States",
+         y = "Overinterpreting Stance (Lenient Interpretation)")
+h3a_2 <- h3a_2 %>% 
+  add_case(tmpp)
 tmp <- tidy(margins::margins(tmp, variables = "b",
-                             at = list("a" = 0:4))) %>% 
[…]

### 2023-06-20 src/analysis/issue_specific.R

@@ -1,89 +1,104 @@
 df_us <- d_us %>% 
-  select(stance, masking, specification, interpretation, country, issue)
+  mutate(country = "United States") %>% 
+  select(stance_nlp, stance_ss, masking, specification, interpretation_nlp, interpretation_ss, country, issue)
 df_nl <- d_nl %>% 
-  select(stance, masking, specification, interpretation, country, issue) %>% 
-  mutate(issue =ifelse(issue=="EU", "Foreign Policy", issue))
+  mutate(country = "The Netherlands") %>% 
+  select(stance_nlp, stance_ss, masking, specification, interpretation_nlp, interpretation_ss, country, issue)
 
-d <- df_nl %>% 
-  add_case(df_us)
+issues_us <- unique(df_us$issue)
+issues_nl <- unique(df_nl$issue)
 
-issues <- unique(d$issue)
-
-for(i in 1:length(issues)){
-  df <- d %>% filter(issue==issues[i])
+for(i in 1:length(issues_us)){
+  df <- df_us %>% filter(issue==issues_us[i])
   if(i==1){
-    b1 <- broom.mixed::tidy(lm(stance ~ masking + specification + 
-                                   factor(country), data= df)) %>%
-      mutate(y = "Y: Correctly Interpreting Stance",
-             issue = issues[i],
-             type = "Pooled Analysis")
-  }
+    b1 <- broom::tidy(lm(stance_nlp ~ masking + specification , data= df)) %>%
+      mutate(y = "Y: Correctly Interpreting Stance (Strict Interpretation)",
+             issue = issues_us[i],
+             type = "United States")}
   else{
-    tmp <- broom.mixed::tidy(lm(stance ~ masking + specification + 
-                                    factor(country), data= df)) %>%
-      mutate(y = "Y: Correctly Interpreting Stance",
-             issue = issues[i],
-             type = "Pooled Analysis")
-    b1 <- b1 %>% 
-      add_case(tmp)
-  }
-}
-
-for(i in 1:length(issues)){
-  df <- df_us %>% filter(issue==issues[i])
-  tmp <- broom::tidy(lm(stance ~ masking + specification , data= df)) %>%
-      mutate(y = "Y: Correctly Interpreting Stance",
-             issue = issues[i],
+    tmp <- broom::tidy(lm(stance_nlp ~ masking + specification , data= df)) %>%
[…]

### 2023-06-20 src/analysis/profile_resp.R

@@ -20,7 +20,7 @@ d <- d_nl %>%
          pid = factor(pid))
 
 e1 <- d %>% 
-  group_by(stance) %>% 
+  group_by(stance_nlp) %>% 
   summarise(Age = round(mean(D1),0),
             Gender= "Male",
             Education = "High-levels of education",
@@ -39,35 +39,84 @@ e1 <- d %>%
             ) %>% 
   ungroup() 
 
+e1b <- d %>% 
+  group_by(stance_ss) %>% 
+  summarise(Age = round(mean(D1),0),
+            Gender= "Male",
+            Education = "High-levels of education",
+            Location = "West of Netherlands",
+            Income = round(mean(D5),0),
+            Job = "Fulltime Employed",
+            `Vote Recall` = "D66",
+            `Position on Immigration` = round(mean(PT2),0),
+            `Position on Environment` = round(mean(PT3),0),
+            `Position on Tax` = round(mean(PT4),0),
+            `Position on EU` =round(mean(PT5),0),
+            `Ideological Position` = round(mean(PT6),0),
+            `Ideological Distance` = round(mean(distance),0),
+            `Issue Congruence` = round(mean(congruence),0),
+            `Political Knowledge` = round(mean(PT7),0)
+  ) %>% 
+  ungroup() 
+
 e1_a <- e1 %>% 
-  select(stance, Gender:Location, Job:`Vote Recall`) %>% 
+  select(stance_nlp, Gender:Location, Job:`Vote Recall`) %>% 
   pivot_longer(cols = Gender:`Vote Recall`) %>% 
-  filter(stance==0) %>% 
-  select(`Incorrectly Identified Stance` = value)
+  filter(stance_nlp==0) %>% 
+  select(`Incorrectly Identified Stance (Strict Interpretation)` = value)
 
 e1_b <- e1 %>% 
-  select(stance, Gender:Location, Job:`Vote Recall`) %>% 
+  select(stance_nlp, Gender:Location, Job:`Vote Recall`) %>% 
   pivot_longer(cols = Gender:`Vote Recall`) %>% 
-  filter(stance==1) %>% 
-  select(`Correctly Identified Stance` = value)
+  filter(stance_nlp==1) %>% 
[…]

### 2023-06-20 src/data-processing/clean_data_us.R

@@ -220,7 +220,7 @@ d_us <- d_us %>%
                              "Stance", "No Stance"),
          interpret4 = ifelse(specified4 == "Underspecified" & S4 == 2, 
                              "Stance", interpret4)) %>% 
-  select(id, D1:D6, PT1:PT6, PT7, attention,
+  select(id, D1:D6, PT1:PT6, PT7, PT7b, attention,
          actor1:actor4,
          issue1:issue4,
          specified1:specified4, 
@@ -235,7 +235,7 @@ d_us <- d_us %>%
                values_to = "condition",
                names_to = "round") %>% 
   separate(condition, c("actor", "issue", "specification", "stance_nlp", "stance_ss", "interpretation"), "-") %>% 
-  select(id, D1:PT7, attention,
+  select(id, D1:PT7b, attention,
          round, actor, issue, specification, stance_nlp, stance_ss, interpretation) %>% 
   mutate(masking = ifelse(actor == "X", "Masked", "Party"),
          stance_nlp = ifelse(stance_nlp =="correct", 1 ,0),

### 2023-06-21 report/draft.Rmd
update analyses in report

@@
 
-We will conduct this survey experiment in the Netherlands in May 2022.
-The sample, recruited through [KiesKompas](https://www.kieskompas.nl/en/), consists of 3,000 participants (based on the power analysis presented in our [online compendium](LINK)) of 18 years and older.
-Kieskompas works with non-random opt-in respondents.
-Therefore, we measure many demographic background variables, and balance checks have been conducted to demonstrate whether certain categories are over represented in a certain experimental group.
-The study has been approved by the [Research Ethics Review Committee](https://fsw.vu.nl/nl/onderzoek/research-ethics-review/index.aspx) of the *Vrije Universiteit Amsterdam* (see the approval [here](https://github.com/MarikenvdVelden/bias-experiment/blob/master/docs/2022-3-30-59.pdf)).
+We have conducted the survey experiments in the Netherlands in May 2022 and in the United States in January 2023. 
+Both samples, recruited through [KiesKompas](https://www.kieskompas.nl/en/) and [Prolific](https://www.prolific.co/) for respectively the Dutch and American case, consists of 3,000 participants (based on the power analysis presented in our [online compendium](LINK)) of 18 years and older.
+Both survey companies works with non-random opt-in respondents. 
+Therefore, we measured many demographic background variables, and balance checks have been conducted to demonstrate whether certain categories are over represented in a certain experimental group. 
+Our study has been approved by the [Research Ethics Review Committee](https://fsw.vu.nl/nl/onderzoek/research-ethics-review/index.aspx) of the *Vrije Universiteit Amsterdam* (see the approval [here](https://github.com/MarikenvdVelden/bias-experiment/blob/master/docs/2022-3-30-59.pdf)).
 To ensure good quality of our data, one attention check (discussed in more detail in [Section 3.3](#attention-checks)) is included [@berinsky2014separating].
@@
 
-***Experimental Conditions.*** Respondents are randomly assigned to either view a political party as an actor, or a masked condition, where they see `X` as an actor; simultaneously, respondents see either a fully specified sentence or a underspecified sentence, in which one needs additional information to interpret the positon on an actor.
-Table \ref{tab:conditions} gives an overview of the variations in treatment in the survey as well as their English translations.
+***Experimental Conditions.*** Respondents are randomly assigned to either view a political party as an actor, or a masked condition, where they see `X` as an actor; simultaneously, respondents see either a fully specified sentence or a underspecified sentence, in which one needs additional information to interpret the position on an actor.
+Table \ref{tab:conditions} gives an overview of the variations in treatment in the surveys.
 
 ```{r conditions, topcaption=TRUE}
-d <- tibble(Condition = c(rep("Specified", 4),
+d <- tibble(`Condition` = c(rep("Specified", 4),
                           rep("Underspecified", 4)), 
-            `Wording ENG` = c("[PVV/X] says immigration should be made harder.",
+            `US Experiment` = c("[Republicans/X] say immigration should be made more difficult.",
+                              "[Democrats/X] say we need to put a tax on carbon emissions",
+                              "[Democrats/X] say we should implement a wealth tax for the richest Americans.",
+                              "[Republicans/X] say the U.S. needs to consider military build-up in the Pacific Ocean",
+                              "[Republicans/ X] say many immigrants are crossing our borders.",
+                              "[Democrats/X] say carbon emissions policy should be implemented differently.",
+                              "[Democrats/X] say the tax system should be implemented differently.",
+                              "[Republicans/ X] say there should be a different military presence in the Pacific Ocean."),
+            `NL Experiment` = c("[PVV/X] says immigration should be made harder.",
                              "[GreenLeft/X] says nitrogen emissions need to be reduced.",
@@
                              "[Labour Party/X] says tax system must be changed.",
-                              "[Forum for Democracy/X] says the Netherlands should have a different role in the European Union."),
-            `Wording NL` = c("[PVV/X] zegt dat immigratie moeilijker gemaakt moet worden.",
-                             "[GroenLinks/X] zegt dat stikstofuitstoot meer tegengegaan moet worden.",
-                             "[PvdA/X] zegt dat het belastingtarief voor de hoogste inkomens omhoog moet.",
-                             "[Forum voor Democratie/X] zegt dat het lidmaatschap van de Europese Unie tot nu toe vooral slecht geweest voor Nederland is.",
-                             "[PVV/X] zegt dat veel immigranten deze kant op komen.",
-                             "[GroenLinks/X] zegt dat het stikstofbeleid anders moet.",
-                             "[PvdA/X] zegt dat het belastingstelsel moet worden aangepast.",
-                             "[Forum voor Democratie/X] zegt dat Nederland een andere rol in de Europese Unie moet hebben."))
+                             "[Forum for Democracy/X] says the Netherlands should have a different role in the European Union."))
 
@@
 For each issue, we ask the respondent `what is according to the sentence above the position of [ACTOR]?`, with the answer categories: `in favor`, `against`, `no stance`, `don't know`.
+We use both a very strict interpretation of stance -- specification of change and direction -- and a lenient interpretation -- specification of change.
+Using the strict interpretation, respondents are correct if they say `no stance` for the underspecified sentences and `against`, `in favor`, `in favor`, and `against` to the specified sentences one to four.
+Using a more lenient interpretation, respondents could say either `in favor` or `against` as well for underspecified sentence two to four.
 
-***Moderating Covariates.*** *Ideological position* is measured using an 11-point scale ranging from left (`0`) to right (`10`). - *Political knowledge* is measured with six items from the DPES.
+***Moderating Covariates.*** *Ideological position* is measured using an 11-point scale ranging from left (`0`) to right (`10`).
+*Political knowledge* is measured with six items from the Dutch Parliamentary Election Studies for the Dutch sample, and the three items from the American National Election Studies.[^]:The questionaire can be found in OA **XX**.
 
@@
 
-Looking at the effect of the experimental conditions on the two dependent variables -- 1) correctly identifying a stance; and 2) over-interpreting a stance -- Figure \ref{fig:results-base} visualizes the baseline.
-The left-hand panel demonstrates the effect of the two experimental conditions for correctly identifying the stance.
-On average, approximately half of the respondents where correct -- as indicated by the intercept.
+Looking at the effect of the experimental conditions on the four dependent variables -- 1) correctly identifying a stance; and 2) over-interpreting a stance for both a strict and a lenient interpretation of stance -- Figure \ref{fig:results-base} visualizes the baseline.
+The left-hand panel demonstrates the effect of the two experimental conditions for correctly identifying the stance in the Dutch case.
+The right-hand panel does so for the American case.
+On average, many respondents in both cases (respectively `85%` in the Dutch case and `65%` in the American case) correctly interpreted the stance using either the lenient or strict interpretation (respectively in blue and red) -- as indicated by the intercept.
 When we mask the political actor -- i.e.
-instead of mentioning the party, we put "X" -- we see that this improves correctly interpreting the stance significantly.
-However, an effect of `0.06` (i.e.
-`6%`) is not a big improvement.
-We do see that the level of specification of a sentence has a significant effect.
-If a sentence is not fully specified, it has a substantive negative effect on the likelihood to correctly interpret the sentence.
-A coefficient of `-0.38` indicates that compared to a fully specified sentence, `38%` of the respondents are more likely to be incorrect when the sentence is under-specified -- that is when the sentence does not state a clear position, but mentions the issue.
-For the left-hand panel, we look at a different dependend variable, so instead of seeing if respondents where correct or not, we look at whether they interpreted the sentence as a stance or not.
-On average, almost nobody overinterprets a stance -- as indicated by an intercept of `-0.01`.
-However, if people see an X compared to a political actor, they are statistically significantly more likely to interpret the sentence as a stance.
+instead of mentioning the party, we put "X" -- we see that this does on average not improves correctly interpreting the stance significantly in neither the Dutch or the American case.
+Additionally, we do see that the level of specification of a sentence has a significant effect. 
+If a sentence is not fully specified, it has a substantive negative effect on the likelihood to correctly interpret the stance in both the lenient and strict interpretation of a stance.
+These effects are substantial in both the American and Dutch case, with coefficients varying between `-0.2` and `-0.5`.
+This indicates that compared to a fully specified sentence, between `20%` and `50%` of the respondents are more likely to be incorrect when the sentence is under-specified -- that is when the sentence does not state a clear position, but mentions the issue.
+Looking at the other dependent variable, whether they interpreted the sentence as a stance or not, we see that almost nobody overinterprets a stance in the strict interpretation in either the Dutch or American case.
+Yet, they do overinterpret a stance in the lenient interpretation.
+Moreover, if people see an X compared to a political actor, they are statistically significantly more likely to interpret the sentence as a stance in its strict interpretation.
 Yet, a coefficient of `0.02` (i.e.
 `2%`) is a very small effect.
-For the condition of specification level, however, we see that compared to a fully specified sentence, people seeing an under-specified sentence are much more likely to interpret the sentence as a stance: an increase of `0.83`.
+For the condition of specification level, however, we see that compared to a fully specified sentence, people seeing an under-specified sentence are much more likely to interpret the sentence as a stance in the strict interpretation: an increase of `0.83`.
 This indicates that people do not excel in this task without any instruction.
-In the pre-registered section, we demonstrate the tests of the hypotheses, and afterwards, we visualize some explorations of the data to show the robustness of our findings.
+Using the lenient interpretation, however, people seem less likely to annotate the sentence as a stance.
+In the pre-registered section, we demonstrate the tests of the hypotheses, and afterwards, we discuss some explorations of the data to show the robustness of our findings, the visualizations thereof are displayed in OA **XX**.
 
@@
 There is a negligible positive effect -- a coefficient of `0.002` -- that is borderline significant.
-Substantially, this means that there is no effect of ideological distance for correctly interpreting the stance.
+Substantially, this means that there is no effect of ideological distance for correctly interpreting the stance in either the American or Dutch case.
 Hence, no ideological bias found, thus no support for our H1a.
-Looking at the lower-left, and right-hand panel of Figure \ref{fig:results-h1}, we see a small but significant effect of the interaction between ideological distance and the under-specified condition.
-It is, however, going in the other direction than hypothesized.
-For those who are ideologically further away from the party, they are less likely to be wrong than those who are close too the party.
-Yet, the difference is about `3%` -- from a coefficient of `-0.38` to `-0.35`.
-We thus find no evidence for H1b either.
-This demonstrates that even in times of hightened polarization, ideological annotation bias is not a huge concern.
-
-Secondly, we hypothesized that there is a risk of over-interpretation from those that are ideologically distant to the political actor (H2a) as well as those who have high levels of political knowledge (H2b).
-We measure this with an interaction between the experimental condition of specification and the variable of interest.
-Following @brambor2006understanding, we visualize the average marginal effects for both interactions to enhance interpretation.
-Figure \ref{fig:results-h2} demonstrates the average marginal effects for both interaction effects: In the left-hand panel the results for H2a, and in the right-hand panel, the results for H2b.
-The left-hand panel of Figure \ref{fig:results-h2} shows that those who are more distant from the position of the political actor are more likely to over-interpret the sentence as a stance.
-The effect increases from about `80%` to those that are closest to the party to `91%` for those who are furthest away from the party.
-This supports our H2a.
-The right-hand panel of Figure \ref{fig:results-h2} shows that political knowledge does statistically significantly effect over-interpretation of a sentence as a stance.
-The substantial effect, however, is small: From the least knowledgeable respondents to the most knowledgeable ones, the likelihood of over-interpretation increases with `2%`.
-So, while we do find support for our H2b, we will in our recommendations section focus less on the political knowledge as a possible interfering feature, but focus on what to do with under-specified sentences when using crowd-coding.
+Looking at the lower-left, and right-hand panel of Figure \ref{fig:results-h1}, we see a small but significant effect of the interaction between ideological distance and the under-specified condition for both interpretations in the American case, and only for the strict interpretation of a stance in the Dutch case.
+These effects are, however, going in the other direction than hypothesized. 
+For those who are ideologically further away from the party, they are less likely to be wrong than those who are close too the party, as shown in the right-hand panel of Figure \ref{fig:results-h1}. 
+For the Dutch case, using the strict interpretation, the difference is about `20%` -- from a coefficient of `-0.6` to `-0.4`.
+In the American case, the slope is even steeper, with a difference of about `40%`.
+We thus find evidence for H1b.
+This demonstrates that even in times of heightened polarization, ideological annotation bias can be a concern.
 
@@
 
+Secondly, we hypothesized that there is a risk of over-interpretation from those that are ideologically distant to the political actor (H2a) as well as those who have high levels of political knowledge (H2b).
+We measure this with an interaction between the experimental condition of specification and the variable of interest. 
+Following @brambor2006understanding, we visualize the average marginal effects for both interactions to enhance interpretation in the right-hand panel of Figure \ref{fig:results-h2}. 
+Figure \ref{fig:results-h2} demonstrates the average marginal effects for both interaction effects: In the upper-panel the results for H2a, and in the bottom-panel, the results for H2b.
+In both countries, we see that the different interpretations of stance have opposite effects: A strict interpretation increases the chance of overinterpreting, but this is not exacerbated by ideological distance or political knowledge.
+Using a more lenient interpretation, we see that this decreasess the chance of overinterpreting.
+While this is not further diminished by political knowledge, ideological disctance does further diminish the change in the American case -- i.e. against expectation H2b.
+In the Dutch case, however, the further you are from the party the more likely you are to overinterpret the sentence as a stance.
+So, while we do find some support for our H2a, we will reflect on the interpretation of a stance in combination with the context, as this effects how crowds interpret underspecified sentences.
+
 Thirdly, we test whether masking of the political actor is a solution for potentially misinterpreting the stance.
 We hypothesized that masking should reduce the ideological bias (H3a) as well as the bias resulting from political knowledge (H3b).
-We test these hypotheses with an interaction between the condition masking and the variables of interest and Figure \ref{fig:results-h3} demonstrates the average marginal effects.
-The left-hand panel of Figure \ref{fig:results-h3} shows, against expectation, a slight increase: Those that are further away from the masked political actor are more likely to incorrectly interpret the stance.
-The decrease in slope is however negligible: From approximately `6%` to `7%` of the respondents being incorrect.
-The slope for political knowledge, on the right-hand panel of Figure \ref{fig:results-h3}, is so-possible even flatter.
-This means that masking of political actors does not help to correctly interpret a sentence as a stance -- i.e.
-no support for H3a and H3b.
+We test these hypotheses with an interaction between the condition masking and the variables of interest and Figure \ref{fig:results-h3} demonstrates the regression coefficients.
+The upper-left panel of Figure \ref{fig:results-h3} shows, against expectation, a slight increase in the Dutch case: Those that are further away from the masked political actor are more likely to incorrectly interpret the stance.
+The decrease in slope is however negligible: From approximately `0.1%` to `0.2%` of the respondents being incorrect.
+There is no effect found in the American case.
+The same goes for the interaction between masking and political knowledge, on the bottom-panel of Figure \ref{fig:results-h3}.
+There is no significant effect found in the US, but a very small negative effect in the Dutch case.
+This means that masking of political actors does not help to correctly interpret a sentence as a stance -- i.e. no support for H3a and H3b.
 
@@
 
-To check the robustness of our findings, Figure \ref{fig:results-exp1} demonstrates the analyses for each issue separately.
-On the left-hand panel of Figure \ref{fig:results-exp1}, we show the results for correctly interpreting the stance.
-We see some variation between issues.
-For the issue *Tax* (lowest row), we see that almost everyone interprets the sentence correct -- an intercept of `0.96`.
-This increases statistically significantly still with about `1%` when masking the political actor, and decreases statistically significantly with `68%` when the sentence is under-specified.
-The regression results (in Online Appendix) show a small effect of ideological distance: the further away, the more likely to correctly interpret the stance -- i.e.
-against H1a.
-Respondents found it most difficult to correctly interpret the issue *Environment* (top row of Figure \ref{fig:results-exp1}), only `12%` has this correct initially, as indicated by the intercept.
-This increases statistically significantly still with about `8%` when masking the political actor, and decreases statistically significantly with `15%` when the sentence is under-specified.
-The other two issues -- *EU* and *Immigration*, in the middle rows of Figure \ref{fig:results-exp1} show a similar effect as the main base-line presented in Figure \ref{fig:results-base}.
-The same pattern holds for the other dependent variable -- overinterpretation of a stance -- in the right-hand panel of Figure \ref{fig:results-exp1}.
-Thus, while there are some differences in effect sizes between the issues, the overall findings are not driven by a single issue.
+To check the robustness of our findings, Figure \ref{fig:results-exp1} demonstrates the analyses for each issue separately. The different colors visualize the different dependent variables.
+We do not see much variation between issues *Tax*, *EU/Foreign Policy*, and *Environment*. For those issues, we see that almost everyone interprets the sentence correct (in blue and red).
+We also see that for a lenient interpretation of stances, people are quite likely to over interpret a position as a stance.
+Being correct about the stance does not when masking the political actor in both cases.
+Yet, the chance of being correct decreases statistically significantly when the sentence is underspecified.
+The same holds for overinterpreting for the lenient interpretation, but the opposite is true for the strict interpretation; there overinterpretation is more likely with underspecified sentences. 
+Looking at *Immigration*, we see a different pattern.
+We see that masking does not increase the likelihood of being correct, but does increase the likelihood of overinterpretation regardless of how one defines a stance.
+Underspecified sentence are less likely to be correctly identified and more likely to be overinterpreted regardless of the definition of a stance.
+So, while there are some differences in effect sizes between the issues, the overall findings are not driven by a single issue.
 
-```{r results-exploration1, out.height = "85%",out.width = "95%",, fig.align = 'center', fig.cap = "\\label{fig:results-exp1} Exploration: Issue Speficific Analyses", eval = F}
+```{r results-exploration1, out.height = "85%",out.width = "95%",, fig.align = 'center', fig.cap = "\\label{fig:results-exp1} Exploration: Issue Speficific Analyses", fig.pos="H", fig.show="hold"}
 knitr::include_graphics(c(here("report/figures", "issues-1.png")))
 ```
-
-In addition to issue-specific analyses, we also explore an interaction between treatments, visualized in Figure \ref{fig:results-exp2} for both dependent variables.
-Figure \ref{fig:results-exp2} shows that masking is of help when sentences are under-specified.
-In the left-hand panel of Figure \ref{fig:results-exp2}, it demonstrates that for under-specifief sentences, people are less likely to incorrectly identify a sentence as a stance when the actor is masked (coefficient of `-0.30`) than when an actor is revealed (coefficient of `-0.45`).
+In addition to issue-specific analyses, we also explore an interaction between treatments, visualized in Figure **OA.XX** for both dependent variables.
+This shoows that masking is of help when sentences are under-specified.
+In the left-hand panel of Figure **OA.XX**, it demonstrates that for under-specifief sentences, people are less likely to incorrectly identify a sentence as a stance when the actor is masked (coefficient of `-0.30`) than when an actor is revealed (coefficient of `-0.45`).
 That means there is a 15% increase in having it correct.
-The difference for over-interpreting is smaller between revealed and masked political actor -- shown in the right-hand panel of Figure \ref{fig:results-exp2} -- yet also statistically significant.
+The difference for over-interpreting is smaller between revealed and masked political actor -- shown in the right-hand panel of Figure **OA.XX** -- yet also statistically significant.
 Compared to `85%` over-interpreting the sentence as a stance, in the masking solution "only" `80%` over-interprets the sentence as a stance.
@@
 
-```{r results-exploration2, out.width = "95%", out.height= "75%", fig.align = 'center', fig.cap = "\\label{fig:results-exp2} Exploration: Interaction with Treatments", eval = F}
-knitr::include_graphics(here("report/figures", "exploration-3.png"))
-knitr::include_graphics(here("report/figures", "exploration-4.png"))
-```
-
-Lastly, we explore three different ways of measuring ideological distance.
-First, we measured ideological bias by looking at whether the respondent is congruent or not with the issue position in the sentence, visualized in Figure \ref{fig:results-exp3}.
-Second, we measured ideological bias by looking at whether the person voted for the party displayed in the sentence, visualized in Figure \ref{fig:results-exp4}.
-And thirdly, we measured ideological bias by looking at the ideology of the respondents -- not in relation to the political actor revealed, visualized in Figure \ref{fig:results-exp5}.
-Figures \ref{fig:results-exp3}, \ref{fig:results-exp4}, and \ref{fig:results-exp5} show that our null-finding regarding ideological bias is not conditional upon the measure we used.
-In none of the analyses, we find evidence for ideological bias.
-
-```{r results-exploration3, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-exp3} Exploration: Different Indicators of Ideological Distance (1)", eval = F}
-knitr::include_graphics(here("report/figures", "exploration-1.png"))
-```
-
-```{r results-exploration4, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-exp4} Exploration: Different Indicators of Ideological Distance (2)", eval = F}
-knitr::include_graphics(here("report/figures", "exploration-2.png"))
-```
-
-```{r results-exploration5, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-exp5} Exploration: Different Indicators of Ideological Distance (3)", eval = F}
-knitr::include_graphics(here("report/figures", "exploration-3.png"))
-```
-
-```{r results-h2_expl, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-h2-expl} Results Level of Sentence Specification", eval = F}
-knitr::include_graphics(here("report/figures", "h2-3.png"))
-```
+Lastly, we explore three different ways of measuring ideological distance and an alternative for political knowledge in the American case.
+First, we measured ideological bias by looking at whether the respondent is congruent or not with the issue position in the sentence, visualized in Figure **OA.XX** .
+Second, we measured ideological bias by looking at whether the person voted for the party displayed in the sentence, visualized in Figure **OA.XX**.
+And thirdly, we measured ideological bias by looking at the ideology of the respondents -- not in relation to the political actor revealed, visualized in Figure **OA.XX**.
+These figures show that our null-finding regarding ideological bias is not conditional upon the measure we used. In none of the analyses, we find evidence for ideological bias.
+Also for the alternative measurement of political knowledge in the US, we find the same results as reported in the main analyses.
 

### 2023-06-22 report/draft.Rmd
update analyses in report
Last additions
Some copyediting, adding a sentence on ChatGPT and two sentences on strict/lenient stances

@@
 sansitup: FALSE
+editor_options: 
+  markdown: 
+    wrap: sentence
 ---
@@
 To determine the validity of such large scale, computational analyses, we rely on "gold standard" data, created by human annotators.
-If follows that the validity of these analyses hinges on the quality of these golden standards.
+It follows that the validity of these analyses hinges on the quality of these golden standards.
 If a gold standard contains biases, i.e., systematic errors, it can foster bias in any downstream analysis.
@@
 *Is this actually bias or a diversity of view points?
-And how big of a problem does this bias/diversity of view points generate for scholars relying on golden standard data?*
+And how big of a problem does this bias/diversity of view points generate for scholars relying on golden standard data?* This question also taps into newer developments related to data annotations: Increasingly, Large Language Models such as ChatGPT are being used by researchers [@gilardi2023chatgpt] as well as annotators themselves [@veselovsky2023artificial] to "simulate" human respondents for annotation tasks.
+These models draw among other things from the gold standard data sets that have been created by researchers, perpetuating the specific biases that are present in them and not allowing for any further disagreement or diversity of viewpoints.
 
@@
 While the crowd offers a solution to some of these issues [@van2021validity], the main underlying assumption that there is one correct interpretation for every input example remains untouched.
-At the same time, trends of polarization have shown that people do interpret information according to their ideological position.
+However, especially for more complex coding tasks different research fields have different interpretations of what counts as *correct* solution depending on the main goal of annotation: Do we want to create a clean benchmark data set that follows strict linguistic rules or one that reflects how statements are being interpreted by "real" readers that do not necessarily follow theoretical definitions on which elements are needed for a specified stance?
+Trends of polarization have shown that people do interpret information according to their ideological position.
 *Does this mean that some are right and others are wrong?
@@
 To do so, we have fielded two high-powered pre-registered experiments (see [here](link) and [here](link)) in the Netherlands -- a low-level polarized country -- and in the U.S.
--- a high-level polarized country testing the effect of ideological distance between the annotator and the political actor in the text (H1), the effect of overinterpretation based on political knowledge or ideological engagement (H2), and a offered solution of masking the political actor to mitigate the effects of ideology and knowledge (H3).[^1] In the experiments, we vary the level of specification with which a political actor takes a stance -- a declarative sentence versus a sentence where with some knowledge on politics, the stance might be inferred -- as well as whether the political actor is shown or masked with putting `[ACTOR]` instead of the political party.
+-- a high-level polarized country testing the effect of ideological distance between the annotator and the political actor in the text (H1), the effect of overinterpretation based on political knowledge or ideological engagement (H2), and an offered solution of masking the political actor to mitigate the effects of ideology and knowledge (H3).[^1]
+In the experiments, we vary the level of specification with which a political actor takes a stance -- a declarative sentence versus a sentence where with some knowledge on politics, the stance might be inferred -- as well as whether the political actor is shown or masked with putting `[ACTOR]` instead of the political party.
 We do this for four different political issues: *Environment*, *Immigration*, *Tax Policy*, and *EU* (for the Dutch case) or *Foreign Policy* (for the American case).
@@
 
-Our results demonstrate that overall sentence where with some knowledge on politics the stance might be inferred are really difficult for the crowd to annotate -- people overinterpret the position using their own knowledge of the world.
-This is problematic as these sentence are very common in political text -- like legislative debates or speeches -- as well as media reports.
+Our results demonstrate that overall sentences where with some knowledge on politics the stance might be inferred are really difficult for the crowd to annotate -- people overinterpret the position using their own knowledge of the world.
+This is problematic as these sentences are very common in political text -- like legislative debates or speeches -- as well as media reports.
 Moreover, our results also demonstrate that for these disagreements in the crowd to occur, the level of polarization needs to be high.
-We do find support for our hypotheses in the American context, but not in the Dutch context -- except for the situation where masking overcomes political knowledge, we do find support for that in the Dutch case, but not in the American one.
-Our findings thus underline the importance of taking disagreement seriously for the creation of gold standard text -- the bread-and-butter of all machine learning endaveours.
-We should look beyond the majority vote and modeling it in the data, because if an algorithm is trained on biased data from disagreeing annotators, it will reproduce and often exacerbate that bias when it is applied to new data [e.g., see @prost2019debiasing].
+We do find support for our hypotheses in the American context, but not in the Dutch context -- except for the situation where masking overcomes political knowledge.
+In this case, we do find support in the Dutch experiment, but not in the American one.
+Our findings thus underline the importance of taking disagreement seriously for the creation of gold standard text -- the bread-and-butter of all machine learning endeavours.
+We should look beyond the majority vote and model it in the data, because if an algorithm is trained on biased data from disagreeing annotators, it will reproduce and often exacerbate that bias when it is applied to new data [e.g., see @prost2019debiasing].
 To be able to model these characteristic of annotators, we should survey the characteristics of annotators when using the crowd (see Webb-Williams et al.
@@
 In studies of political communication, the most familiar annotation tasks involve identifying basic concepts.
-This often involve noting the topic of the text, the position of the actor or the tone of the text.
+This includes noting the topic of the text, the position of the actor or the tone of the text.
 A recent study [@van2021validity] demonstrates that using crowd-coding platforms is a good way to collect such large data-sets.
@@
 One such potential bias, especially in times of increasing polarization [@iyengar2019origins; @gidron2019toward; @boxell2022cross], is based on ideological position of the annotator.
-Given that annotators on crows-coding platforms tend to be younger and hold more liberal values than the general public [@clifford2015samples; @huff2015these], this could potentially hamper the data diversity and generalizability of the model.
-An additional reason to monitor the annotators ideological position as a potential source of annotator bias is that a recent study in NLP showed that experiential factors influence the consistency of how political ideologies are perceived [@shen2021sounds].
+Given that annotators on crowd-coding platforms tend to be younger and hold more liberal values than the general public [@clifford2015samples; @huff2015these], this could potentially hamper the data diversity and generalizability of the model.
+An additional reason to monitor the annotators' ideological position as a potential source of annotator bias is that a recent study in NLP showed that experiential factors influence the consistency of how political ideologies are perceived [@shen2021sounds].
 Their finding challenges the "ground-truth" assumption we as researcher make that a position for example is either left-leaning or right-wing leaning.
@@
 This challenges our way of data collection.
-We are interested in the effect of e.g.
-elite communication.
+We are interested in the effect of e.g. elite communication.
 To study this, we allow for heterogeneous treatment effects in experimental work.
@@
 Additional doubts on smoothing out disagreement in annotation have focussed on the lack of diversity when only annotating with one label or annotator, leading to a homogeneity especially in subjective and social tasks [@geva2019we] such as hatespeech detection or political affiliation classification.
-The question then is: _Whose perspective is being recorded in these datasets, and then later in the models trained on these datasets?_ 
-Framing arbitrary representations in data as \"bias\" misses the political character of datasets: There is no neutral data and no apolitical standpoint from where we can call out bias.
-Datasets are always \"a worldview\" [26] and, as such, data always remains biased.\" (Miceli et al., 2022, p.
+The question then is: *Whose perspective is being recorded in these datasets, and then later in the models trained on these datasets?* Framing arbitrary representations in data as "bias" misses the political character of datasets: There is no neutral data and no apolitical standpoint from where we can call out bias.
+Datasets are always "a worldview" [26] and, as such, data always remains biased." (Miceli et al., 2022, p.
 5).
@@
 
-Most recently, new annotation paradigms have gone one step further by asking whether we  modelling the task, or the annotator [@geva2019we]?
+Most recently, new annotation paradigms have gone one step further by asking whether we are modelling the task, or the annotator [@geva2019we].
 @pavlick2019 find that for the logical coherence task Natural Language Inference, annotators have several valid interpretations that are not reflected in one ground truth label.
-They call for new training paradigms that can reflect "the full range of possible human inferences"[@pavlick2019, p.
-688].
+They call for new training paradigms that can reflect "the full range of possible human inferences"[@pavlick2019, p. 688].
 Recent approaches in NLP have sought to explicitly incorporate disagreement and diversity in training data annotations.
@@
 
-@troiano2021[p.2] also note how for a complex concept such as the "emotion" of a text, the annotator can make several assumptions during the annotation process on what is wanted, that are all valid and may or may not be useful in different contexts: "It is possible to assess one\'s own emotion after reading the text, to reconstruct the affective state of the writers who produced it, to guess the reaction that they intended to elicit in the readers, and so on.\" @shen2021sounds find that political ideology is not an inherent concept in many texts, but rather dependent on who is asked to annotate and their perceptions and background.
+@troiano2021 [p.2] also note how for a complex concept such as the "emotion" of a text, the annotator can make several assumptions during the annotation process on what is wanted, that are all valid and may or may not be useful in different contexts: "It is possible to assess one's own emotion after reading the text, to reconstruct the affective state of the writers who produced it, to guess the reaction that they intended to elicit in the readers, and so on." @shen2021sounds find that political ideology is not an inherent concept in many texts, but rather dependent on who is asked to annotate and their perceptions and background.
 Extralinguistic factors, such as annotator's own political ideology and also knowledge, influenced annotation and in turn model performance.
@@
 
-> **Ideological Bias hypothesis** (*H1b*): The effect of H1a is stronger for sentences in which the party's position can be infered (i.e. underspecified sentences).
+> **Ideological Bias hypothesis** (*H1b*): The effect of H1a is stronger for sentences in which the party's position can be inferred (i.e. underspecified sentences).
 
@@
 @thornjakobsen2022 deduce that annotator bias comes from a process known as the affect heuristic [@slovic2007affect ]: making a decision based on the emotional response related to your own personal attitude towards the discussed topic, especially when the text is relatively ambiguous.
-To test whether people with different ideological backgrounds as well as their political knowledge might experience that position differently, challenging our ground-truth assumption in data annotation, we propose the following hypotheses:
+In the tradition of more strict interpretations of what constitutes as a stance especially in (computational) linguistics, sentences that are underspecified should always be annotated as "not a stance".
+However, this might not be the desired annotation for other research purposes: When for example the main goal is to understand how readers are affected by statements shown in e.g., a newspaper article, a more lenient definition of stance that allows annotators to infer the direction of a political stance from context and prior knowledge even though a statement is strictly speaking underspecified might be more useful.
+In everyday life, people will not follow strict linguistic definitions, for understanding media effects we might thus be more interested in whether stances, giving context information, are "correctly overinterpreted".
+To test whether people with different ideological backgrounds as well as their political knowledge might experience underidentified position differently, challenging our ground-truth assumption in data annotation, we propose the following hypotheses:
 
@@
 
-_If these biases exists, how can we alleviate them?_
-There have been several previous approaches to solve biased annotations, especially where it concerns political or societal aspects.
-@geva2019we introduce an approach where training set annotators are separated from annotators annotating datasets that evaluate the models, to ensure the evaluation is not simply accurate at replicating the original annotators, but can generalize to new annotators' judgements.
+We will test these hypotheses both with a strict and a lenient interpretation of stance to illustrate whether decisions made in the research process on what constitutes a stance influence the results.
+In a next step, we ask: *If these biases exists, how can we alleviate them?* There have been several previous approaches to solve biased annotations, especially where it concerns political or societal aspects.
+@geva2019we introduce an approach where training set annotators are separated from annotators annotating data sets that evaluate the models, to ensure the evaluation is not simply accurate at replicating the original annotators, but can generalize to new annotators' judgements.
 However, these approaches are not aimed at reducing biases during the training set annotation procedure.
 Other approaches are aimed at leveraging multiple perspectives - but this is not useful when one wants one label to learn from.
-For the Austrian National Elections, @ennser2018impact already demonstrated that showing party labels impact annotators' assessment of the party position. So, one solution is masking:
+For the Austrian National Elections, @ennser2018impact already demonstrated that showing party labels impact annotators' assessment of the party position.
+So, one solution is masking:
 
@@
 We have conducted the survey experiments in the Netherlands in May 2022 and in the United States in January 2023.
-Both samples, recruited through [KiesKompas](https://www.kieskompas.nl/en/) and [Prolific](https://www.prolific.co/) for respectively the Dutch and American case, consists of 3,000 participants (based on the power analysis presented in our [online compendium](LINK)) of 18 years and older.
+Both samples, recruited through [KiesKompas](https://www.kieskompas.nl/en/) and [Prolific](https://www.prolific.co/) for respectively the Dutch and American case, consist of 3,000 participants (based on the power analysis presented in our [online compendium](LINK)) of 18 years and older.
 Both survey companies works with non-random opt-in respondents.
@@
 ***Moderating Covariates.*** *Ideological position* is measured using an 11-point scale ranging from left (`0`) to right (`10`).
-*Political knowledge* is measured with six items from the Dutch Parliamentary Election Studies for the Dutch sample, and the three items from the American National Election Studies.[^]:The questionaire can be found in OA **XX**.
+*Political knowledge* is measured with six items from the Dutch Parliamentary Election Studies for the Dutch sample, and the three items from the American National Election Studies.[\^]:The questionnaire can be found in OA **XX**.
 
-***Controle Variables.*** In our analysis, we control for demographic information (gender, age, education, income, religion, job) as well as political background variables (trust in politics, ideological position on economic left-right scale and cultural progressive-conservative scale, and evaluations and prospects of the economy).
+***Control Variables.*** In our analysis, we control for demographic information (gender, age, education, income, religion, job) as well as political background variables (trust in politics, ideological position on economic left-right scale and cultural progressive-conservative scale, and evaluations and prospects of the economy).
 Tables A.5 till A.17 in the OA demonstrate the descriptive information per country.
@@
 ```
-
 # Results
 
-To answer whether there is an ideological or knowledge-basd annotation bias, we have conducted a two-by-two experiment.
+To answer whether there is an ideological or knowledge-based annotation bias, we have conducted a two-by-two experiment.
 Tables \ref{tab:profilenl} and \ref{tab:profileus} demonstrate the average profile of respondents who annotate correctly and incorrectly (where respondents who annotated some stances correctly and some incorrectly are weighted by proportion (in)correct).
@@
 When we mask the political actor -- i.e.
-instead of mentioning the party, we put "X" -- we see that this does on average not improves correctly interpreting the stance significantly in neither the Dutch or the American case.
+instead of mentioning the party, we put "X" -- we see that this does on average not improve correctly interpreting the stance significantly in neither the Dutch or the American case.
 Additionally, we do see that the level of specification of a sentence has a significant effect.
@@
 Moreover, if people see an X compared to a political actor, they are statistically significantly more likely to interpret the sentence as a stance in its strict interpretation.
-Yet, a coefficient of `0.02` (i.e.
-`2%`) is a very small effect.
+Yet, a coefficient of `0.02` (i.e. `2%`) is a very small effect.
 For the condition of specification level, however, we see that compared to a fully specified sentence, people seeing an under-specified sentence are much more likely to interpret the sentence as a stance in the strict interpretation: an increase of `0.83`.
@@
 
+````{=tex}
 \begin{landscape}
@@
 \end{landscape}
+````
 
@@
 In both countries, we see that the different interpretations of stance have opposite effects: A strict interpretation increases the chance of overinterpreting, but this is not exacerbated by ideological distance or political knowledge.
-Using a more lenient interpretation, we see that this decreasess the chance of overinterpreting.
-While this is not further diminished by political knowledge, ideological disctance does further diminish the change in the American case -- i.e. against expectation H2b.
+Using a more lenient interpretation, we see that this decreases the chance of overinterpreting.
+While this is not further diminished by political knowledge, ideological distance does further diminish the change in the American case -- i.e. against expectation H2b.
 In the Dutch case, however, the further you are from the party the more likely you are to overinterpret the sentence as a stance.
@@
 
-To check the robustness of our findings, Figure \ref{fig:results-exp1} demonstrates the analyses for each issue separately. The different colors visualize the different dependent variables.
-We do not see much variation between issues *Tax*, *EU/Foreign Policy*, and *Environment*. For those issues, we see that almost everyone interprets the sentence correct (in blue and red).
+To check the robustness of our findings, Figure \ref{fig:results-exp1} demonstrates the analyses for each issue separately.
+The different colors visualize the different dependent variables.
+We do not see much variation between issues *Tax*, *EU/Foreign Policy*, and *Environment*.
+For those issues, we see that almost everyone interprets the sentence correct (in blue and red).
 We also see that for a lenient interpretation of stances, people are quite likely to over interpret a position as a stance.
-Being correct about the stance does not when masking the political actor in both cases.
+**Being correct about the stance does not when masking the political actor in both cases**.
 Yet, the chance of being correct decreases statistically significantly when the sentence is underspecified.
@@
 We see that masking does not increase the likelihood of being correct, but does increase the likelihood of overinterpretation regardless of how one defines a stance.
-Underspecified sentence are less likely to be correctly identified and more likely to be overinterpreted regardless of the definition of a stance.
+Underspecified sentences are less likely to be correctly identified and more likely to be overinterpreted regardless of the definition of a stance.
 So, while there are some differences in effect sizes between the issues, the overall findings are not driven by a single issue.
@@
 ```
+
 In addition to issue-specific analyses, we also explore an interaction between treatments, visualized in Figure **OA.XX** for both dependent variables.
-This shoows that masking is of help when sentences are under-specified.
-In the left-hand panel of Figure **OA.XX**, it demonstrates that for under-specifief sentences, people are less likely to incorrectly identify a sentence as a stance when the actor is masked (coefficient of `-0.30`) than when an actor is revealed (coefficient of `-0.45`).
+This shows that masking is of help when sentences are under-specified.
+In the left-hand panel of Figure **OA.XX**, it demonstrates that for under-specified sentences, people are less likely to incorrectly identify a sentence as a stance when the actor is masked (coefficient of `-0.30`) than when an actor is revealed (coefficient of `-0.45`).
 That means there is a 15% increase in having it correct.
@@
 And thirdly, we measured ideological bias by looking at the ideology of the respondents -- not in relation to the political actor revealed, visualized in Figure **OA.XX**.
-These figures show that our null-finding regarding ideological bias is not conditional upon the measure we used. In none of the analyses, we find evidence for ideological bias.
+These figures show that our null-finding regarding ideological bias is not conditional upon the measure we used.
+In none of the analyses, we find evidence for ideological bias.
 Also for the alternative measurement of political knowledge in the US, we find the same results as reported in the main analyses.

### 2023-06-22 report/draft.Rmd
update references

@@
 The answer of the annotators in turn influence how machine learning examples classify new models.
-For instance, hate speech detection and abuse detection is one NLP task where the race and gender of annotators influences annotators and model performance[@gordon2022, @Larimore, @waseem2016].
+For instance, hate speech detection and abuse detection is one NLP task where the race and gender of annotators influences annotators and model performance[@gordon2022, @larimore2021reconsidering, @waseem2016].
 Language is inherently connected to society and culture: @shenDarling analyze sentiment analysis, and find that human annotators lead to certain perspectives on sentiment being recorded, and that notably that African American English dialects are often misunderstood by such models.
@@
 
+Secondly, we hypothesized that there is a risk of over-interpretation from those that are ideologically distant to the political actor (H2a) as well as those who have high levels of political knowledge (H2b).
+We measure this with an interaction between the experimental condition of specification and the variable of interest.
+Following @brambor2006understanding, we visualize the average marginal effects for both interactions to enhance interpretation in the right-hand panel of Figure \ref{fig:results-h2}.
+Figure \ref{fig:results-h2} demonstrates the average marginal effects for both interaction effects: In the upper-panel the results for H2a, and in the bottom-panel, the results for H2b.
+In both countries, we see that the different interpretations of stance have opposite effects: A strict interpretation increases the chance of overinterpreting, but this is not exacerbated by ideological distance or political knowledge.
+Using a more lenient interpretation, we see that this decreases the chance of overinterpreting.
+While this is not further diminished by political knowledge, ideological distance does further diminish the change in the American case -- i.e. against expectation H2b.
+In the Dutch case, however, the further you are from the party the more likely you are to overinterpret the sentence as a stance.
+So, while we do find some support for our H2a, we will reflect on the interpretation of a stance in combination with the context, as this effects how crowds interpret underspecified sentences.
+
 ````{=tex}
@@
 ````
-
-Secondly, we hypothesized that there is a risk of over-interpretation from those that are ideologically distant to the political actor (H2a) as well as those who have high levels of political knowledge (H2b).
-We measure this with an interaction between the experimental condition of specification and the variable of interest.
-Following @brambor2006understanding, we visualize the average marginal effects for both interactions to enhance interpretation in the right-hand panel of Figure \ref{fig:results-h2}.
-Figure \ref{fig:results-h2} demonstrates the average marginal effects for both interaction effects: In the upper-panel the results for H2a, and in the bottom-panel, the results for H2b.
-In both countries, we see that the different interpretations of stance have opposite effects: A strict interpretation increases the chance of overinterpreting, but this is not exacerbated by ideological distance or political knowledge.
-Using a more lenient interpretation, we see that this decreases the chance of overinterpreting.
-While this is not further diminished by political knowledge, ideological distance does further diminish the change in the American case -- i.e. against expectation H2b.
-In the Dutch case, however, the further you are from the party the more likely you are to overinterpret the sentence as a stance.
-So, while we do find some support for our H2a, we will reflect on the interpretation of a stance in combination with the context, as this effects how crowds interpret underspecified sentences.
 

### 2023-06-22 report/draft.Rmd
update references

@@
 This country selection does not only allow us to showcase the scope conditions of disagreement due to different levels of political heterogeneity, but also differentiates between languages.
-English is not only the most dominant language for computational text analysis in the social science [@baden2022three; Dolinsky et al.
-2023], crowd-coders do not need to be native speakers, given the dominance of English in our daily lives.
+English is not only the most dominant language for computational text analysis in the social science [@baden2022three; @dolinsky2023threegaps], crowd-coders do not need to be native speakers, given the dominance of English in our daily lives.
 This is different for Dutch, it is a language spoken by a smaller community, typically native speakers, yet still an often-enough researched case in computational text analysis in the social science [@baden2022three; Dolinsky et al.

### 2023-07-06 report/draft.Rmd
start of discussion (in progress)

@@
 
-To test whether people with different ideological backgrounds as well as their political knowledge might experience that position differently, challenging our ground-truth assumption in data annotation, we propose the following hypotheses:
+To test whether people with different ideological backgrounds as well as their political knowledge might experience that position differently, challenging our ground-truth assumption in data annotation, we propose the following hypothesis:
 
 > **Ideological Bias hypothesis** (*H1a*): The larger the ideological distance between respondent and the party, the less likely respondents annotate statements according to the party's uttered position.
-
-> **Ideological Bias hypothesis** (*H1b*): The effect of H1a is stronger for sentences in which the party's position can be inferred (i.e. underspecified sentences).
 
@@
 @thornjakobsen2022 deduce that annotator bias comes from a process known as the affect heuristic [@slovic2007affect ]: making a decision based on the emotional response related to your own personal attitude towards the discussed topic, especially when the text is relatively ambiguous.
+We therefore expect that the strength of H1a is conditional on this ambiguity:
+
+> **Ideological Bias hypothesis** (*H1b*): The effect of H1a is stronger for underspecified sentences.
+
 In the tradition of more strict interpretations of what constitutes as a stance especially in (computational) linguistics, sentences that are underspecified should always be annotated as "not a stance".
@@
 
-TBA
+(in progress)
+
+In this paper we set out to critically examine annotation bias in the classification of political stances. 
+In recent years there have been groundbreaking advances in the field of natural language processing that make it increasingly possible to automatically classify stances in texts on a large scale, which holds great potential for communication research.
+However, research in the field of communication and political science gives reason for concern that the human annotated gold standard data that is used to train and validate these stance classification models suffers from systematic biases.
+This appears especially likely in crowd-coded data, where disagreement is mostly resolved with majority voting rather then calibrating coders through more explicit coding rules.
+We show in two pre-registered crowd-coding experiments that ideological differences can systematically affect coding decisions, and reflect on how the field of communication should deal with this form of annotation bias.
+
+First, we tested whether ideological bias affects the interpretation of a party's stance by analyzing whether a person's ideological distance from a party affects how they classify their stance. 
+Although we did not find a main effect of ideological distance (H1a), we did find evidence that the effect of ideological distance is greater for underspecified sentences.
+That is, for sentences with clear political stances on an issue (e.g., immigration should be made more difficult) we did not find an effect, but if the sentence refers to the issue without making a clear stance (e.g., many immigrants are crossing our borders) we do see a clear effect of ideological distance.
+
+Second, we looked at differences in terms of overinterpretation, which we define as coding stance in an underspecified sentence (i.e. in which the stance is not explicitly made clear).
+In other words, when coders inferred the stance of the party on the issue based on prior knowledge or beliefs about the party.
+We expected that overinterpretation is more likely if the ideological distance of the coder to the party is greater (H2a) and if the coder has more political knowledge (H2b). 
+Overall, our findings did not support these hypotheses.
+With the more lenient interpretation of overinterpretation, we did find support for H2a for the Dutch context, but we found an opposite effect for the US. 
+While this does not support the hypothesis in terms of the direction of the biase, it does imply that ideological distance can affect how stance is processed.
+
+Finally, we investigated whether annotation biases due to ideological distance from the party (H3a) and political knowledge (H3b) could be alleviated by masking the political party.
+Here we again found no convincing evidence.
+
+Although most of our hypotheses were not supported, we believe that the confirmation of H1b could pose important implications for stance analysis. 
+something something
+
+(the following is just some notes)
+
+## implications:
+- improving stance detection models also requires improving human annotations
+- if the perception of political stance is contingent on ideological differences, then it can be inherently problematic to conceptualize the political stance, as expressed in a text, as a ground truth.
+
+There are two general solutions for dealing with this bias.
+The first and more classic solution would be to aim to improve intercoder reliability. 
+This could be achieved both through stricter training and by design of the coding task.
+If prior knowledge or beliefs about a political actor affects interpretation, the actor could be masked.
+
+The second solution is to conceptualize political stance as having an inherently subjective quality, and measure it accordingly. 
+This aligns more with the literature on perspectivism (refs), which still appears under-lighted in the field of communication. 
+Perspectivism seems particularly relevant for media effect research, where political stances in texts are measured to make inferences about the audience.
+Our results show that XXX.
+As such, to make accurate inferences about how a citizen perceived political stance, and consequently how this might have affected behavior such as voting or participating in protests, it might be necessary to take personal characteristics
+Note that this does not necessarily imply that intercoder reliability is no longer important.
+Rather, by
 

### 2023-07-07 report/draft.Rmd
first minor changes (mainly typos and reference correction

@@
 abstract: "Information is key to inform the behavior of citizens, and thereby for the social scientists studying them. 
-The democratization of data has lead to numerous possibilities to gather and analyze textual data.
+#This first sentence is a bit strange: information to inform behavior? I am not sure what it means, so also not sure how to rephrase..
+The democratization of data has lead to numerous possibilities for gathering and analyzing textual data.
 These enourmous amounts of data are typically handled by machine learning techniques to classify into meaningful variables.
 The performance of these models is evaluated based on a gold standard, created by human annotators.
-Having a high level of agreement between these annotators is key, but some suggest personal characteristics of annotators, like political ideology or knowledge, interfere.
-We show in two pre-registered experiments that XXX.
-Thereby [contribution]."
+Having a high level of agreement between these annotators is key, but some
+#some what?
+suggest personal characteristics of annotators, like political ideology or knowledge, interfere. We show in two pre-registered experiments that XXX. Thereby [contribution]."
 keywords: "Experiment, Annotation Bias, Ideology, Measurinig Political Position, Text-as-Data, Political Knowledge"
@@
 Information is key to inform the behavior of citizens, and thereby for the social scientists studying them. 
+#See comment on abstract...
 Classical theories of political communication, such as agenda setting or framing [e.g.
 @vanaelst2016political; @lecheler2019news], formulate that political information drives opinion formation and participation in politics -- from voting to protests [for overviews hereof across countries, see @pfetsch2013comparing].
-The democratization of data and advent of computational social science has paved the way for new possibilities to gather and analyze textual data [for a recent overview, see @vanatteveldt2022computational].
-In particular, advances in Natural Language Processing (NLP) have made it possible to automatically analyze large quantities of data using machine learning [e.g., see @bender2016linguistic; @wei2023overview].
-To determine the validity of such large scale, computational analyses, we rely on "gold standard" data, created by human annotators.
+The democratization of data and advent of computational social science has paved the way for new possibilities of gathering and analyzing textual data [for a recent overview, see @vanatteveldt2022computational].
+#possibility to is ungrammatical
+In particular, advances in Natural Language Processing (NLP) have made it possible to automatically analyze large quantities of data using machine learning [e.g., see [ADD REF, Bender does not make sense here] @wei2023overview].
+To determine the validity of such large scale computational analyses, we rely on "gold standard" data, created by human annotators.
 It follows that the validity of these analyses hinges on the quality of these golden standards.
@@
 Most data collection efforts to create gold standards assume that there is only one correct interpretation for every input example, and that disagreement between the annotators is something that needs to be dealt with at all costs [@aroyo2015truth].
-A recent study [@van2021validity] demonstrates that using crowd-coding platforms is a good way to collect such golden standards without too much disagreement.
+A recent study     demonstrates that using crowd-coding platforms is a good way to collect such golden standards without too much disagreement.
 These platforms have also been used for experiments, and the quality of the respondents has been a center of attention [@coppock2019generalizing; @clifford2015samples; @huff2015these; @berinsky2014separating].
@@
 *Is this actually bias or a diversity of view points?
-And how big of a problem does this bias/diversity of view points generate for scholars relying on golden standard data?* This question also taps into newer developments related to data annotations: Increasingly, Large Language Models such as ChatGPT are being used by researchers [@gilardi2023chatgpt] as well as annotators themselves [@veselovsky2023artificial] to "simulate" human respondents for annotation tasks.
+And how big of a problem does this bias/diversity of view points generate for researchers relying on golden standard data?* This question also taps into newer developments related to data annotations: Increasingly, Large Language Models such as ChatGPT are being used by researchers [@gilardi2023chatgpt] as well as annotators themselves [@veselovsky2023artificial] to "simulate" human respondents for annotation tasks.
 These models draw among other things from the gold standard data sets that have been created by researchers, perpetuating the specific biases that are present in them and not allowing for any further disagreement or diversity of viewpoints.
@@
 
-In this paper, we build upon the NLP literature on disagreement -- or bias -- in annotation [e.g., see @shen2021sounds; @geva2019we; @sommerauer2020penguin; @plank2014learning] and so-called perspectivism [@basile2021toward; @havens2022beyond] -- i.e.
-the adoption of methods that integrate the opinions and perspectives of the human subjects involved in the knowledge representation step of the machine learning processes [@basile2021toward].
+In this paper, we build upon the NLP literature on disagreement -- or bias -- in annotation [e.g., see @shen2021sounds; @geva2019we; @sommerauer2020penguin; @plank2014learning] and so-called perspectivism [@basile2021toward; @havens2022beyond] -- i.e.\ the adoption of methods that integrate the opinions and perspectives of the human subjects involved in the knowledge representation step of the machine learning processes [@basile2021toward].
 This literature puts forward that disagreement can occur because of differences in ideological position or political knowledge [@shen2021sounds; @alkiek2022classification; @joseph2021mis].
@@
 This country selection does not only allow us to showcase the scope conditions of disagreement due to different levels of political heterogeneity, but also differentiates between languages.
-English is not only the most dominant language for computational text analysis in the social science [@baden2022three; @dolinsky2023threegaps], crowd-coders do not need to be native speakers, given the dominance of English in our daily lives.
-This is different for Dutch, it is a language spoken by a smaller community, typically native speakers, yet still an often-enough researched case in computational text analysis in the social science [@baden2022three; Dolinsky et al.
+English is not only the most dominant language for computational text analysis in the social science [@baden2022three; @dolinsky2023threegaps], crowd-coders do not need to be first language speakers, given the dominance of English in our daily lives.
+This is different for Dutch, it is a language spoken by a smaller community, typically first language speakers, yet still an often-enough researched case in computational text analysis in the social science [@baden2022three; Dolinsky et al.
 2023].
@@
 This challenges our way of data collection.
-We are interested in the effect of e.g. elite communication.
+We are interested in the effect of e.g.\ elite communication.
 To study this, we allow for heterogeneous treatment effects in experimental work.
@@
 
-Another aspect to consider is that disagreement can be informative for the concept under measure - sometimes agreement can be used to validate hypotheses about how universal the perceptions of such concepts are [@sommerauer2020penguin].
+Another aspect to consider is that disagreement can be informative for the concept under measure - sometimes agreement can be used to validate hypotheses about how universal the perceptions of such concepts are [@sommerauer2020would].
 Additional doubts on smoothing out disagreement in annotation have focussed on the lack of diversity when only annotating with one label or annotator, leading to a homogeneity especially in subjective and social tasks [@geva2019we] such as hatespeech detection or political affiliation classification.
@@
 The answer of the annotators in turn influence how machine learning examples classify new models.
-For instance, hate speech detection and abuse detection is one NLP task where the race and gender of annotators influences annotators and model performance[@gordon2022, @larimore2021reconsidering, @waseem2016].
+For instance, hate speech detection and abuse detection is one NLP task where the race and gender of annotators influences annotators and model performance [@gordon2022, @larimore2021reconsidering, @waseem2016].
 Language is inherently connected to society and culture: @shenDarling analyze sentiment analysis, and find that human annotators lead to certain perspectives on sentiment being recorded, and that notably that African American English dialects are often misunderstood by such models.
@@
 Recent approaches in NLP have sought to explicitly incorporate disagreement and diversity in training data annotations.
-[@röttger2021] introduces the idea of an explicitly subjective annotation paradigm existing in addition to one focussed on one label and "ground truth".
+@röttger2021 introduce the idea of an explicitly subjective annotation paradigm existing in addition to one focussed on one label and "ground truth".
 Such a subjective annotation paradigm can be used for purposes where the goal is finding diverse perspectives on the task or concepts, and for models to model more accurately how humans interpret a task or text.
@@
 Using the pooled data we will estimate a within groups fixed effects model.
-We have conducted a balance test based on demographics (age, gender, education, geographical region, level of urbanness,employment, and income), vote choice in the 2021 parliamentary elections, ideological self-placement, political knowledge, and positions on the issues, using the `cobalt` R package [@greifer2021].
+We have conducted a balance test based on demographics (age, gender, education, geographical region, level of urbanness, employment, and income), vote choice in the 2021 parliamentary elections, ideological self-placement, political knowledge, and positions on the issues, using the `cobalt` R package [@greifer2021].
 This balance test indicated that none of the variables are unbalanced over the experimental groups, and therefore, as pre-registered, will not be added to the regression formula.

### 2023-07-07 report/draft.Rmd
draft draft discussion

@@
 Or is there an ideological difference in the ground truth?* These questions present a fundamental challenge to the main way of working when collecting gold standard data, as we operate from the baseline assumption that disagreement among the annotators should be avoided or reduced.
-Typically, when specific cases continuously cause disagreement, more instructions are added to limit interpretations [for recent innovations to improve annotation, see for example: @barbera2021automated; @struthers2020bridging; @winter2020online; @debell2013harder; @benoit2016crowd, @ying2022topics].
+Typically, when specific cases continuously cause disagreement, more instructions are added to limit interpretations (for recent innovations to improve annotation, see for example: @barbera2021automated; @struthers2020bridging; @winter2020online; @debell2013harder; @benoit2016crowd, @ying2022topics).
 However, work in computational linguistics has shown that increased annotation instructions do not increase quality [@parmar2022don].
@@
 
-(in progress)
-
 In this paper we set out to critically examine annotation bias in the classification of political stances. 
@@
 
-First, we tested whether ideological bias affects the interpretation of a party's stance by analyzing whether a person's ideological distance from a party affects how they classify their stance. 
+First, we tested whether ideological bias affects the interpretation of a party's stance by analyzing whether a person's ideological distance to a party affects how they classify the party's stance. 
 Although we did not find a main effect of ideological distance (H1a), we did find evidence that the effect of ideological distance is greater for underspecified sentences.
-That is, for sentences with clear political stances on an issue (e.g., immigration should be made more difficult) we did not find an effect, but if the sentence refers to the issue without making a clear stance (e.g., many immigrants are crossing our borders) we do see a clear effect of ideological distance.
-
+This implies that ideological distance can indeed influence how annotators infer stance, but only if the sentence leaves enough room for interpretation.
 Second, we looked at differences in terms of overinterpretation, which we define as coding stance in an underspecified sentence (i.e. in which the stance is not explicitly made clear).
-In other words, when coders inferred the stance of the party on the issue based on prior knowledge or beliefs about the party.
+In other words, when annotators inferred the stance of the party on the issue based on prior knowledge or beliefs about the party.
 We expected that overinterpretation is more likely if the ideological distance of the coder to the party is greater (H2a) and if the coder has more political knowledge (H2b). 
@@
 With the more lenient interpretation of overinterpretation, we did find support for H2a for the Dutch context, but we found an opposite effect for the US. 
-While this does not support the hypothesis in terms of the direction of the biase, it does imply that ideological distance can affect how stance is processed.
-
+While this does not support the hypothesis in terms of the direction of the bias, it does imply that ideological distance can affect how stance is processed.
 Finally, we investigated whether annotation biases due to ideological distance from the party (H3a) and political knowledge (H3b) could be alleviated by masking the political party.
@@
 
-Although most of our hypotheses were not supported, we believe that the confirmation of H1b could pose important implications for stance analysis. 
-something something
+Although most of our hypotheses were not supported, our findings for H1a and H2b in particular pose important implications.
+For sentences with clear political stances on an issue (e.g., immigration should be made more difficult) annotators generally agreed on the stance that is expressed.
+But if the sentence refers to the issue without making a clear stance (e.g., many immigrants are crossing our borders) we do see a clear effect of ideological distance.
+This indicates that for annotation bias to occur, there has to be sufficient room for different interpretations. 
+If the stance is clearly specified, then there is more of a ground truth, and the opinions and perspectives of the annotator therefore matter less.
+It is when the stance is underspecified that perspectivism [@basile2021toward; @havens2022beyond] comes into play.
+When a party talks about a political issue without taking an explicit stance, annotators need to fill in the gaps and make their own inferences about the party's stance on this issue.
+Our results imply that political ideology is indeed a relevant dimension in shaping how an annotator perceives political stances.
 
-(the following is just some notes)
+This means that to improve stance detection models, we need to critically evaluate how we create our gold standards.
+In the best case scenario, ignoring personal differences on dimensions such as political ideology only adds noise to the data.
+However, this is only the case if a sufficiently large sample of annotators is randomly drawn from the population.
+In reality, gold standards are often the product of a handful of expert coders, that typically study or work at universities.
+Even if a good amount of crowd coders is used, it cannot blindly be assumed that this provides a diverse and balanced reflection of political ideologies in society, as we also observed in our samples (Table 2).
+Our results indicate that an annotation team that is skewed towards one end of the political spectrum could indeed result in a biased gold standard.
 
-## implications:
-- improving stance detection models also requires improving human annotations
-- if the perception of political stance is contingent on ideological differences, then it can be inherently problematic to conceptualize the political stance, as expressed in a text, as a ground truth.
-
-There are two general solutions for dealing with this bias.
-The first and more classic solution would be to aim to improve intercoder reliability. 
-This could be achieved both through stricter training and by design of the coding task.
-If prior knowledge or beliefs about a political actor affects interpretation, the actor could be masked.
-
-The second solution is to conceptualize political stance as having an inherently subjective quality, and measure it accordingly. 
-This aligns more with the literature on perspectivism (refs), which still appears under-lighted in the field of communication. 
-Perspectivism seems particularly relevant for media effect research, where political stances in texts are measured to make inferences about the audience.
-Our results show that XXX.
-As such, to make accurate inferences about how a citizen perceived political stance, and consequently how this might have affected behavior such as voting or participating in protests, it might be necessary to take personal characteristics
-Note that this does not necessarily imply that intercoder reliability is no longer important.
-Rather, by
+Furthermore, if the perception of political stance is contingent on ideological differences, then it can be inherently problematic to conceptualize political stance, as expressed in a text, as a ground truth.
+Depending on the goal of a study, it might be more appropriate to conceptualize political stance as having an inherently subjective quality, and strive to measure this subjective space.
+As a general guideline, we propose making a distinction between content analysis research that makes inferences about the author of a text, and research that makes inferences about the audience.
+For example, a study that compares political stances between news outlets to analyze political bias in the news does require a consistent and reliable measurement of political stances.
+This could indeed require a ground-truth definition of political stance, that is drilled into annotators through well-defined and specific instructions and extensive training.
+But if the study involves making inferences about how the audience interprets political stances, for example in media effects research, then a ground-truth definition of political stance is problematic.
+To make accurate inferences about how a citizen subjectively perceived political content, and consequently how this might have affected behavior such as voting, we need to learn more about how shared personal characteristics relate to shared modes of perspective.
 

### 2023-07-07 report/draft.Rmd
fixed references

--- 2023-07-07 report/draft.Rmd
@@
 
-In this paper, we build upon the NLP literature on disagreement -- or bias -- in annotation [e.g., see @shen2021sounds; @geva2019we; @sommerauer2020penguin; @plank2014learning] and so-called perspectivism [@basile2021toward; @havens2022beyond] -- i.e.\ the adoption of methods that integrate the opinions and perspectives of the human subjects involved in the knowledge representation step of the machine learning processes [@basile2021toward].
+In this paper, we build upon the NLP literature on disagreement -- or bias -- in annotation [e.g., see @shen2021sounds; @geva2019we; @sommerauer2020penguin; @plank2014learning] and so-called perspectivism [@cabitza2023toward; @havens2022beyond] -- i.e.\ the adoption of methods that integrate the opinions and perspectives of the human subjects involved in the knowledge representation step of the machine learning processes [@cabitza2023toward].
 This literature puts forward that disagreement can occur because of differences in ideological position or political knowledge [@shen2021sounds; @alkiek2022classification; @joseph2021mis].
@@
 Recent approaches in NLP have sought to explicitly incorporate disagreement and diversity in training data annotations.
-@röttger2021 introduce the idea of an explicitly subjective annotation paradigm existing in addition to one focussed on one label and "ground truth".
+@rottger2022two introduce the idea of an explicitly subjective annotation paradigm existing in addition to one focussed on one label and "ground truth".
 Such a subjective annotation paradigm can be used for purposes where the goal is finding diverse perspectives on the task or concepts, and for models to model more accurately how humans interpret a task or text.
-Additionally, "perspectivism" [@basile2021toward] is a paradigm and research agenda where different perspectives are explicitly incorporated in the training data, and used by models to provide more human-like classifications.
+Additionally, "perspectivism" [@cabitza2023toward] is a paradigm and research agenda where different perspectives are explicitly incorporated in the training data, and used by models to provide more human-like classifications.
 Another paradigm is "jury learning" [@gordon2022], in which machine learning models do not learn to replicate one specific ground truth, but are trained with different annotator juries to reflect the judgement of different populations.
@@
 
-@troiano2021 [p.2] also note how for a complex concept such as the "emotion" of a text, the annotator can make several assumptions during the annotation process on what is wanted, that are all valid and may or may not be useful in different contexts: "It is possible to assess one's own emotion after reading the text, to reconstruct the affective state of the writers who produced it, to guess the reaction that they intended to elicit in the readers, and so on." @shen2021sounds find that political ideology is not an inherent concept in many texts, but rather dependent on who is asked to annotate and their perceptions and background.
+@troiano2021emotion [p.2] also note how for a complex concept such as the "emotion" of a text, the annotator can make several assumptions during the annotation process on what is wanted, that are all valid and may or may not be useful in different contexts: "It is possible to assess one's own emotion after reading the text, to reconstruct the affective state of the writers who produced it, to guess the reaction that they intended to elicit in the readers, and so on." @shen2021sounds find that political ideology is not an inherent concept in many texts, but rather dependent on who is asked to annotate and their perceptions and background.
 Extralinguistic factors, such as annotator's own political ideology and also knowledge, influenced annotation and in turn model performance.

### 2023-07-11 report/draft.Rmd
updates in manuscript:solving comments

@@
 anonymous: FALSE
-abstract: "Information is key to inform the behavior of citizens, and thereby for the social scientists studying them. 
-#This first sentence is a bit strange: information to inform behavior? I am not sure what it means, so also not sure how to rephrase..
+abstract: "The behavior of citizens and the study of them by social scientists rely heavily on key information.
 The democratization of data has lead to numerous possibilities for gathering and analyzing textual data.
@@
 The performance of these models is evaluated based on a gold standard, created by human annotators.
-Having a high level of agreement between these annotators is key, but some
-#some what?
-suggest personal characteristics of annotators, like political ideology or knowledge, interfere. We show in two pre-registered experiments that XXX. Thereby [contribution]."
+Having a high level of agreement between these annotators is key, but recent research suggests personal characteristics of annotators, like political ideology or knowledge, interfere.
+We show in two pre-registered experiments that XXX.
+Thereby [contribution]."
 keywords: "Experiment, Annotation Bias, Ideology, Measurinig Political Position, Text-as-Data, Political Knowledge"
@@
 
-Information is key to inform the behavior of citizens, and thereby for the social scientists studying them. 
-#See comment on abstract...
+The behavior of citizens and the study of them by social scientists rely heavily on key information.
 Classical theories of political communication, such as agenda setting or framing [e.g.
@@
 The democratization of data and advent of computational social science has paved the way for new possibilities of gathering and analyzing textual data [for a recent overview, see @vanatteveldt2022computational].
-#possibility to is ungrammatical
 In particular, advances in Natural Language Processing (NLP) have made it possible to automatically analyze large quantities of data using machine learning [e.g., see [ADD REF, Bender does not make sense here] @wei2023overview].
@@
 Most data collection efforts to create gold standards assume that there is only one correct interpretation for every input example, and that disagreement between the annotators is something that needs to be dealt with at all costs [@aroyo2015truth].
-A recent study     demonstrates that using crowd-coding platforms is a good way to collect such golden standards without too much disagreement.
+A recent study  by @van2021validity demonstrates that using crowd-coding platforms is a good way to collect such golden standards without too much disagreement.
 These platforms have also been used for experiments, and the quality of the respondents has been a center of attention [@coppock2019generalizing; @clifford2015samples; @huff2015these; @berinsky2014separating].
@@
 However, especially for more complex coding tasks different research fields have different interpretations of what counts as *correct* solution depending on the main goal of annotation: Do we want to create a clean benchmark data set that follows strict linguistic rules or one that reflects how statements are being interpreted by "real" readers that do not necessarily follow theoretical definitions on which elements are needed for a specified stance?
-Trends of polarization have shown that people do interpret information according to their ideological position.
+Trends of polarization have shown that people do interpret information according to their ideological position [e.g., @rekker2022understanding; @lee2021more].
 *Does this mean that some are right and others are wrong?

### 2023-07-11 report/draft.Rmd
updates in manuscript:solving comments

@@
 biblio-style: apsr
-title: "Whose Truth is it Anyway? An Experiment on Annotation Bias"
+title: "Whose Truth is it Anyway? An Experiment on Annotation Bias in Times of Factual Opinion Polarization"
 author:
-- name: Mariken A.C.G. van der Velden*
+- name: Mariken A.C.G. van der Velden
 - affiliation: Dep. of Communciation Science, Vrije Universiteit Amsterdam
@@
 - affiliation: Dep. of Communciation Science, Vrije Universiteit Amsterdam
-thanks: "* = Corresponding author, Replication files are available on the author's Github account (https://github.com/MarikenvdVelden/bias-experiment); Author contributions: a) designed the study: MACGvdV, WvA, AF, FL, MR, & KW; b) conducted the study: MACGvdV, FL & MR; c) data cleaning & analysis: MACGvdV; d) writing of the paper: MACGvdV"
+thanks: "Corresponding author: MACGvdV, Replication files are available on the author's Github account (https://github.com/MarikenvdVelden/bias-experiment); Author contributions: a) designed the study: MACGvdV, MR, WvA, AF, FL, & KW; b) conducted the study: MACGvdV, FL & MR; c) data cleaning & analysis: MACGvdV; d) writing of the paper: MACGvdV, MR, WvA, AF, FL, & KW"
 anonymous: FALSE
@@
 
-```{r results-h3, fig.align = 'center', fig.cap = "\\label{fig:results-h3} Results Masking Solution", fig.pos="H", fig.show="hold",out.width = "59%", out.height="50%"}
+```{r results-h3, fig.align = 'center', fig.cap = "\\label{fig:results-h3} Results Masking Solution", fig.pos="H", fig.show="hold",out.width = "80%", out.height="80%"}
 knitr::include_graphics(here("report/figures", "h3-1.png"))
@@
 
-```{r results-exploration1, out.height = "85%",out.width = "95%",, fig.align = 'center', fig.cap = "\\label{fig:results-exp1} Exploration: Issue Speficific Analyses", fig.pos="H", fig.show="hold"}
-knitr::include_graphics(c(here("report/figures", "issues-1.png")))
-```
-
 In addition to issue-specific analyses, we also explore an interaction between treatments, visualized in Figure **OA.XX** for both dependent variables.
@@
 Also for the alternative measurement of political knowledge in the US, we find the same results as reported in the main analyses.
+
+```{r results-exploration1, out.height = "85%",out.width = "95%",, fig.align = 'center', fig.cap = "\\label{fig:results-exp1} Exploration: Issue Speficific Analyses", fig.pos="H", fig.show="hold"}
+knitr::include_graphics(c(here("report/figures", "issues-1.png")))
+```
 

### 2023-07-13 report/draft.Rmd
update draft (conclusion and refs)

@@
    - \newtheorem{hypo}{Hypothesis}
+   - \usepackage{setspace}
 biblio-style: apsr
@@
 thanks: "Corresponding author: MACGvdV, Replication files are available on the author's Github account (https://github.com/MarikenvdVelden/bias-experiment); Author contributions: a) designed the study: MACGvdV, MR, WvA, AF, FL, & KW; b) conducted the study: MACGvdV, FL & MR; c) data cleaning & analysis: MACGvdV; d) writing of the paper: MACGvdV, MR, WvA, AF, FL, & KW"
-anonymous: FALSE
+anonymous: TRUE
 abstract: "The behavior of citizens and the study of them by social scientists rely heavily on key information.
-The democratization of data has lead to numerous possibilities for gathering and analyzing textual data.
-These enourmous amounts of data are typically handled by machine learning techniques to classify into meaningful variables.
-The performance of these models is evaluated based on a gold standard, created by human annotators.
-Having a high level of agreement between these annotators is key, but recent research suggests personal characteristics of annotators, like political ideology or knowledge, interfere.
-We show in two pre-registered experiments that XXX.
-Thereby [contribution]."
+Human annotation represents a crucial instrument in the toolkit of the social sciences.
+To address concerns regarding the validity and reliability of annotation tasks, we establish strict standards.
+Due to the democratization of data and the advances in NLP more data can be analyzed or classified, these standards are more important than ever: An algorithm trained on biased data will reproduce and often exacerbate bias.
+Our tools to create valid and reliable annotation data does currently not allow for dealing with different perspectives of annotators.
+In two pre-registered experiments in the United States and the Netherlands, we show that personal characteristics of annotators, like political ideology or knowledge, interfere with annotators judgement of political stances.
+Our results show that to improve annotated data for automated text analyses, and for stance detection models in particular, we need to critically evaluate how we create our gold standards."
 keywords: "Experiment, Annotation Bias, Ideology, Measurinig Political Position, Text-as-Data, Political Knowledge"
@@
 mainfont: cochineal
-fontsize: 11pt
+fontsize: 12pt
 params:
   anonymous: ""
-doublespacing: TRUE
 endnote: no
@@
 
+\doublespacing
 # Introduction
-
 The behavior of citizens and the study of them by social scientists rely heavily on key information.
@@
 The democratization of data and advent of computational social science has paved the way for new possibilities of gathering and analyzing textual data [for a recent overview, see @vanatteveldt2022computational].
-In particular, advances in Natural Language Processing (NLP) have made it possible to automatically analyze large quantities of data using machine learning [e.g., see [ADD REF, Bender does not make sense here] @wei2023overview].
+In particular, advances in Natural Language Processing (NLP) have made it possible to automatically analyze large quantities of data using machine learning [e.g., see @grimmer2022text; @wei2023overview].
 To determine the validity of such large scale computational analyses, we rely on "gold standard" data, created by human annotators.
@@
 While the crowd offers a solution to some of these issues [@van2021validity], the main underlying assumption that there is one correct interpretation for every input example remains untouched.
-However, especially for more complex coding tasks different research fields have different interpretations of what counts as *correct* solution depending on the main goal of annotation: Do we want to create a clean benchmark data set that follows strict linguistic rules or one that reflects how statements are being interpreted by "real" readers that do not necessarily follow theoretical definitions on which elements are needed for a specified stance?
+However, especially for more complex coding tasks different research fields have different interpretations of what counts as *correct* solution depending on the main goal of annotation.
+Do we want to create a clean benchmark data set that follows strict linguistic rules or one that reflects how statements are being interpreted by "real" readers that do not necessarily follow theoretical definitions on which elements are needed for a specified stance?
 Trends of polarization have shown that people do interpret information according to their ideological position [e.g., @rekker2022understanding; @lee2021more].
-*Does this mean that some are right and others are wrong?
-Or is there an ideological difference in the ground truth?* These questions present a fundamental challenge to the main way of working when collecting gold standard data, as we operate from the baseline assumption that disagreement among the annotators should be avoided or reduced.
+*Does this mean that some are right and others are wrong? Or is there an ideological difference in the ground truth?* 
+These questions present a fundamental challenge to the main way of working when collecting gold standard data, as we operate from the baseline assumption that disagreement among the annotators should be avoided or reduced.
 Typically, when specific cases continuously cause disagreement, more instructions are added to limit interpretations (for recent innovations to improve annotation, see for example: @barbera2021automated; @struthers2020bridging; @winter2020online; @debell2013harder; @benoit2016crowd, @ying2022topics).
@@
 
-In this paper, we build upon the NLP literature on disagreement -- or bias -- in annotation [e.g., see @shen2021sounds; @geva2019we; @sommerauer2020penguin; @plank2014learning] and so-called perspectivism [@cabitza2023toward; @havens2022beyond] -- i.e.\ the adoption of methods that integrate the opinions and perspectives of the human subjects involved in the knowledge representation step of the machine learning processes [@cabitza2023toward].
+In this paper, we build upon the NLP literature on disagreement -- or bias -- in annotation [e.g., see @shen2021sounds; @geva2019we; @sommerauer2020penguin; @plank2014learning] and so-called perspectivism [@cabitza2023toward; @havens2022beyond] -- i.e.
+the adoption of methods that integrate the opinions and perspectives of the human subjects involved in the knowledge representation step of the machine learning processes [@cabitza2023toward].
 This literature puts forward that disagreement can occur because of differences in ideological position or political knowledge [@shen2021sounds; @alkiek2022classification; @joseph2021mis].
@@
 To do so, we have fielded two high-powered pre-registered experiments (see [here](link) and [here](link)) in the Netherlands -- a low-level polarized country -- and in the U.S.
--- a high-level polarized country testing the effect of ideological distance between the annotator and the political actor in the text (H1), the effect of overinterpretation based on political knowledge or ideological engagement (H2), and an offered solution of masking the political actor to mitigate the effects of ideology and knowledge (H3).[^1]
+-- a high-level polarized country testing the effect of ideological distance between the annotator and the political actor in the text (H1), the effect of overinterpretation based on political knowledge or ideological engagement (H2), and an offered solution of masking the political actor to mitigate the effects of ideology and knowledge (H3).^[The data and research compendium is published on the [[main author's github page]](https://github.com/MarikenvdVelden/bias-experiment) -- annonymized for the review process.]
 In the experiments, we vary the level of specification with which a political actor takes a stance -- a declarative sentence versus a sentence where with some knowledge on politics, the stance might be inferred -- as well as whether the political actor is shown or masked with putting `[ACTOR]` instead of the political party.
@@
 English is not only the most dominant language for computational text analysis in the social science [@baden2022three; @dolinsky2023threegaps], crowd-coders do not need to be first language speakers, given the dominance of English in our daily lives.
-This is different for Dutch, it is a language spoken by a smaller community, typically first language speakers, yet still an often-enough researched case in computational text analysis in the social science [@baden2022three; Dolinsky et al.
-2023].
-
-[^1]: The data and research compendium is published on the [main author's github page](https://github.com/MarikenvdVelden/bias-experiment) -- annonymized for the review process.
+This is different for Dutch, it is a language spoken by a smaller community, typically first language speakers, yet still an often-enough researched case in computational text analysis in the social science [@baden2022three; @dolinsky2023threegaps].
 
@@
 Generating large data-sets has become one of the main drivers of progress in natural language understanding.
-In studies of political communication, the most familiar annotation tasks involve identifying basic concepts.
+In studies of political communication, the most familiar annotation tasks involve identifying theoretical concepts.
 This includes noting the topic of the text, the position of the actor or the tone of the text.
-A recent study [@van2021validity] demonstrates that using crowd-coding platforms is a good way to collect such large data-sets.
-However, having only a few workers annotate the majority of text of interest has raised concerns about data diversity and models' ability to generalize beyond the crowd-workers: In a series of experiments, @geva2019we show that often models do not generalize well to annotations from annotators that did not contribute to the training set, suggesting that annotator bias should be monitored during data-set creation.
+Crowdcoding seems suitable for this task [@van2021validity].
+Yet, having only a few workers annotate the majority of text of interest has raised concerns about data diversity and models' ability to generalize beyond the crowd-workers.
+In a series of experiments, @geva2019we show that often models do not generalize well to annotations from annotators that did not contribute to the training set, suggesting that annotator bias should be monitored during data-set creation.
 One such potential bias, especially in times of increasing polarization [@iyengar2019origins; @gidron2019toward; @boxell2022cross], is based on ideological position of the annotator.
@@
 This challenges our way of data collection.
-We are interested in the effect of e.g.\ elite communication.
+We are interested in the effect of e.g. elite communication.
 To study this, we allow for heterogeneous treatment effects in experimental work.
@@
 The question then is: *Whose perspective is being recorded in these datasets, and then later in the models trained on these datasets?* Framing arbitrary representations in data as "bias" misses the political character of datasets: There is no neutral data and no apolitical standpoint from where we can call out bias.
-Datasets are always "a worldview" [26] and, as such, data always remains biased." (Miceli et al., 2022, p.
-5).
+Datasets are always "a worldview" and, as such, data always remains biased" [@miceli2022studying p.5]. 
 This is key to social scientist in general, and those studying political text in particular, since several tasks of interest are intrinsically societal, with answers that differ based on the make-up of the worldview of annotators.
 The answer of the annotators in turn influence how machine learning examples classify new models. 
-For instance, hate speech detection and abuse detection is one NLP task where the race and gender of annotators influences annotators and model performance [@gordon2022, @larimore2021reconsidering, @waseem2016].
-Language is inherently connected to society and culture: @shenDarling analyze sentiment analysis, and find that human annotators lead to certain perspectives on sentiment being recorded, and that notably that African American English dialects are often misunderstood by such models.
+For instance, hate speech and abuse detection are NLP tasks where race and gender of annotators influences both annotators and model performance [@gordon2022; @larimore2021reconsidering; @waseem2016].
+Language is inherently connected to society and culture: @shenDarling analyze sentiment analysis, and find that human annotators lead to certain perspectives on sentiment being recorded; notably African American English dialects are often misunderstood by such models.
 
@@
 @pavlick2019 find that for the logical coherence task Natural Language Inference, annotators have several valid interpretations that are not reflected in one ground truth label.
-They call for new training paradigms that can reflect "the full range of possible human inferences"[@pavlick2019, p. 688].
+They call for new training paradigms that can reflect "the full range of possible human inferences"[@pavlick2019, p.688].
 Recent approaches in NLP have sought to explicitly incorporate disagreement and diversity in training data annotations.
@@
 However, to our knowledge this phenomenon has not yet been tested in a controlled experimental setting with manipulations in the texts.
-
-To test whether people with different ideological backgrounds as well as their political knowledge might experience that position differently, challenging our ground-truth assumption in data annotation, we propose the following hypothesis:
+To test whether people with different ideological backgrounds experience a political position differently, challenging our ground-truth assumption in data annotation, we propose the following hypothesis:
 
@@
 A lack of explicitness in the annotated text is one of the main causes of the disagreements in earlier literature.
-@thornjakobsen2022 deduce that annotator bias comes from a process known as the affect heuristic [@slovic2007affect ]: making a decision based on the emotional response related to your own personal attitude towards the discussed topic, especially when the text is relatively ambiguous.
+@thornjakobsen2022 deduce that annotator bias comes from a process known as the affect heuristic [@slovic2007affect]: Making a decision based on the emotional response related to your own personal attitude towards the discussed topic, especially when the text is relatively ambiguous.
 We therefore expect that the strength of H1a is conditional on this ambiguity:
@@
 
-We will test these hypotheses both with a strict and a lenient interpretation of stance to illustrate whether decisions made in the research process on what constitutes a stance influence the results.
 In a next step, we ask: *If these biases exists, how can we alleviate them?* There have been several previous approaches to solve biased annotations, especially where it concerns political or societal aspects.
@@
 For the Austrian National Elections, @ennser2018impact already demonstrated that showing party labels impact annotators' assessment of the party position.
-So, one solution is masking:
+So, one solution is masking. 
+We therefore test whether masking reduces potential differentiation incited by ideoloigcal postion or political knowledge with the following hypotheses:
 
@@
 ## Pre-registered Results
-
 First, we test whether there is an ideological bias in interpreting stances (H1a), and if this bias increases for those that are further away from the ideological position of the political actor in the under-specified condition (H1b).
@@
 We also see that for a lenient interpretation of stances, people are quite likely to over interpret a position as a stance.
-**Being correct about the stance does not when masking the political actor in both cases**.
+Being correct about the stance does not decrease when masking the political actor in both cases.
 Yet, the chance of being correct decreases statistically significantly when the sentence is underspecified.
@@
 # Discussion
-
-In this paper we set out to critically examine annotation bias in the classification of political stances. 
-In recent years there have been groundbreaking advances in the field of natural language processing that make it increasingly possible to automatically classify stances in texts on a large scale, which holds great potential for communication research.
-However, research in the field of communication and political science gives reason for concern that the human annotated gold standard data that is used to train and validate these stance classification models suffers from systematic biases.
-This appears especially likely in crowd-coded data, where disagreement is mostly resolved with majority voting rather then calibrating coders through more explicit coding rules.
-We show in two pre-registered crowd-coding experiments that ideological differences can systematically affect coding decisions, and reflect on how the field of communication should deal with this form of annotation bias.
+Human annotation represents a crucial instrument in the toolkit of the social sciences.
+In order to address concerns regarding the validity and reliability of annotation tasks, we establish strict standards.
+Especially in these days, where due to the democratization of data and the advances in NLP more data can be analyzed or classified, these standards are more important than ever.
+After all, an algorithm trained on biased data will reproduce and often exacerbate bias [e.g. see @prost2019debiasing].
+It is key to realize that the underlying assumption to create valid and reliable annotated data is that there is only one correct interpretation for every input example, and that disagreement between the annotators is something that needs to be dealt with at all costs [@aroyo2015truth].
+In times of increasing factual opinion polarization [@rekker2022understanding; @lee2021more], it is crucial to ask *who's perspective is being recorded in these datasets*, given that no neutral data and no apolitical standpoint from where we can call out bias: "Datasets are always 'a worldview' and, as such, data always remains biased" [@miceli2022studying, p.5].
+Our tools to create valid and reliable annotation data does currently not allow for dealing with different perspectives. 
+In this paper, we examined annotation bias in the classification of political stances using two pre-registered crowd-coding experiments in the Netherlands and the United States.
+We used both a strict and a lenient interpretation of stance to illustrate whether decisions made in the research process on what constitutes a stance influences the results.
 
 First, we tested whether ideological bias affects the interpretation of a party's stance by analyzing whether a person's ideological distance to a party affects how they classify the party's stance. 
-Although we did not find a main effect of ideological distance (H1a), we did find evidence that the effect of ideological distance is greater for underspecified sentences.
-This implies that ideological distance can indeed influence how annotators infer stance, but only if the sentence leaves enough room for interpretation.
-Second, we looked at differences in terms of overinterpretation, which we define as coding stance in an underspecified sentence (i.e. in which the stance is not explicitly made clear).
-In other words, when annotators inferred the stance of the party on the issue based on prior knowledge or beliefs about the party.
+Although we did not find a main effect of ideological distance (H1a), we did find evidence that the effect of ideological distance is greater for underspecified sentences (H1b) in the US, depending on how one operationalizes stances.
+THe exploratory analyses showed that these findings where robust agaist differnt specifications of ideological distance.
+This conditional finding is important, because machine translation is on the rise [@licht2023cross; @de2018no]: Social scientists use this technique for example to comparatively analyze political stances.
+Currently, most scholars using text analysis rely on English text or annotated data, translating the original source text to English [@baden2022three].
+Our results show that the way a political stance is defined and the context, potentially the level of polarization, affects whether annotators induced bias.
+Hence, naively using US annotated data and auto-translating your source text could have downstream consequences for the classification of your stances.
+
+Second, we looked at differences in terms of overinterpretation, which we define as coding stance in an underspecified sentence (i.e.
+in which the stance is not explicitly made clear) using both a strict (computational linguistics) and more lenient interpretation of what a stance is.
+In other words, we tested when annotators inferred the stance of the party on the issue based on prior knowledge or beliefs about the party.
 We expected that overinterpretation is more likely if the ideological distance of the coder to the party is greater (H2a) and if the coder has more political knowledge (H2b). 
@@
 While this does not support the hypothesis in terms of the direction of the bias, it does imply that ideological distance can affect how stance is processed.
-Finally, we investigated whether annotation biases due to ideological distance from the party (H3a) and political knowledge (H3b) could be alleviated by masking the political party.
-Here we again found no convincing evidence.
-
-Although most of our hypotheses were not supported, our findings for H1a and H2b in particular pose important implications.
+Regarding political knowledge (H2b), our results showed again country differences as well as that the way one defines a stance matters.
+Our findings, for H2b in particular, pose important implications.
 For sentences with clear political stances on an issue (e.g., immigration should be made more difficult) annotators generally agreed on the stance that is expressed.
-But if the sentence refers to the issue without making a clear stance (e.g., many immigrants are crossing our borders) we do see a clear effect of ideological distance.
+But if the sentence refers to the issue without making a clear stance (e.g., many immigrants are crossing our borders) we do see a clear effect of political knowledge.
 This indicates that for annotation bias to occur, there has to be sufficient room for different interpretations. 
@@
 When a party talks about a political issue without taking an explicit stance, annotators need to fill in the gaps and make their own inferences about the party's stance on this issue.
-Our results imply that political ideology is indeed a relevant dimension in shaping how an annotator perceives political stances.
+Our results imply that political knowledge is indeed a relevant dimension in shaping how an annotator perceives political stances.
 
-This means that to improve stance detection models, we need to critically evaluate how we create our gold standards.
+Finally, we investigated whether annotation biases due to ideological distance from the party (H3a) and political knowledge (H3b) could be alleviated by masking the political party.
+Here we again found country differences as well as variation based on how a stance was defined.
+We showed that masking has no effect to resolve bias in the US context, but it does help in a less polarized context like the Netherlands. 
+Given that the US is one of the front runners in terms of polarization, it is important to realize that this affects our annotators, and therefore annotated data, too.
+Biases seems to be stronger and less easy to be resolved.
+
+Our results imply that to improve annotated data for automated text analyses, and for stance detection models in particular, we need to critically evaluate how we create our gold standards.
 In the best case scenario, ignoring personal differences on dimensions such as political ideology only adds noise to the data.
@@
 In reality, gold standards are often the product of a handful of expert coders, that typically study or work at universities.
-Even if a good amount of crowd coders is used, it cannot blindly be assumed that this provides a diverse and balanced reflection of political ideologies in society, as we also observed in our samples (Table 2).
+Even if a good amount of crowd coders is used, it cannot blindly be assumed that this provides a diverse and balanced reflection of political ideologies in society, as we also observed in our samples (Tables \ref{tab:profilenl} and \ref{tab:profileus}).
 Our results indicate that an annotation team that is skewed towards one end of the political spectrum could indeed result in a biased gold standard.
@@
 
-## Recommendations
-
-TBA
-
 \newpage
-
+\singlespacing
 # References

### 2023-07-13 report/draft.Rmd
new version

@@
 Due to the democratization of data and the advances in NLP more data can be analyzed or classified, these standards are more important than ever: An algorithm trained on biased data will reproduce and often exacerbate bias.
-Our tools to create valid and reliable annotation data does currently not allow for dealing with different perspectives of annotators.
-In two pre-registered experiments in the United States and the Netherlands, we show that personal characteristics of annotators, like political ideology or knowledge, interfere with annotators judgement of political stances.
+Our tools to create valid and reliable annotation data currently do not allow for dealing with different perspectives of annotators.
+In two pre-registered experiments in the United States and the Netherlands, we show that personal characteristics of annotators, like political ideology or knowledge, interfere with annotators' judgement of political stances.
 Our results show that to improve annotated data for automated text analyses, and for stance detection models in particular, we need to critically evaluate how we create our gold standards."
@@
 \doublespacing
+
+<!--Word count: `r as.integer(sub("(\\d+).+$", "\\1", system(sprintf("wc -w %s", knitr::current_input()), intern = TRUE))) - 20`-->
+
 # Introduction

### 2023-07-14 report/draft.Rmd
new version
minor textual changes

@@
 # Introduction
-The behavior of citizens and the study of them by social scientists rely heavily on key information.
+The behavior of citizens and the study thereof by social scientists both rely heavily on key information.
 Classical theories of political communication, such as agenda setting or framing [e.g.
@@
 Our study has been approved by the [Research Ethics Review Committee](https://fsw.vu.nl/nl/onderzoek/research-ethics-review/index.aspx) of the *Vrije Universiteit Amsterdam* (see the approval [here](https://github.com/MarikenvdVelden/bias-experiment/blob/master/docs/2022-3-30-59.pdf)).
-To ensure good quality of our data, one attention check (discussed in more detail in [Section 3.3](#attention-checks)) is included [@berinsky2014separating].
+To ensure good quality of our data, one attention check (discussed in more detail in **OA XX** is included [@berinsky2014separating].
 
@@
 
-***Experimental Conditions.*** Respondents are randomly assigned to either view a political party as an actor, or a masked condition, where they see `X` as an actor; simultaneously, respondents see either a fully specified sentence or a underspecified sentence, in which one needs additional information to interpret the position on an actor.
+***Experimental Conditions.*** Respondents are randomly assigned to either view a political party as an actor, or a masked condition, where they see `X` as an actor; simultaneously, respondents see either a fully specified sentence or a underspecified sentence, in which one needs additional information outside of the text to determine an actor's position.
 Table \ref{tab:conditions} gives an overview of the variations in treatment in the surveys.
@@
 # Discussion
-Human annotation represents a crucial instrument in the toolkit of the social sciences.
-In order to address concerns regarding the validity and reliability of annotation tasks, we establish strict standards.
+
+Human annotation is a crucial instrument in the toolkit of the social sciences.
+In order to safeguard the validity and reliability of annotation tasks, strict standards need to be upheld.
 Especially in these days, where due to the democratization of data and the advances in NLP more data can be analyzed or classified, these standards are more important than ever.

### 2023-07-15 report/draft.Rmd
Minor last changes

--- 2023-07-13 report/draft.Rmd
@@
 To address concerns regarding the validity and reliability of annotation tasks, we establish strict standards.
-Due to the democratization of data and the advances in NLP more data can be analyzed or classified, these standards are more important than ever: An algorithm trained on biased data will reproduce and often exacerbate bias.
-Our tools to create valid and reliable annotation data currently do not allow for dealing with different perspectives of annotators.
+Due to the democratization of data and the advances in NLP more data can be analyzed or classified, making these standards are more important than ever: An algorithm trained on biased data will reproduce and often exacerbate bias.
+Our tools to create valid and reliable annotation data does currently not allow for dealing with different perspectives of annotators.
 In two pre-registered experiments in the United States and the Netherlands, we show that personal characteristics of annotators, like political ideology or knowledge, interfere with annotators' judgement of political stances.
@@
 
-<!--Word count: `r as.integer(sub("(\\d+).+$", "\\1", system(sprintf("wc -w %s", knitr::current_input()), intern = TRUE))) - 20`-->
+# Introduction
 
-# Introduction
 The behavior of citizens and the study of them by social scientists rely heavily on key information.
@@
 To determine the validity of such large scale computational analyses, we rely on "gold standard" data, created by human annotators.
-It follows that the validity of these analyses hinges on the quality of these golden standards.
+It follows that the validity of these analyses hinges on the quality of these gold standards.
 If a gold standard contains biases, i.e., systematic errors, it can foster bias in any downstream analysis.
@@
 Most data collection efforts to create gold standards assume that there is only one correct interpretation for every input example, and that disagreement between the annotators is something that needs to be dealt with at all costs [@aroyo2015truth].
-A recent study  by @van2021validity demonstrates that using crowd-coding platforms is a good way to collect such golden standards without too much disagreement.
+A recent study by @van2021validity demonstrates that using crowd-coding platforms is a good way to collect such gold standards without too much disagreement.
 These platforms have also been used for experiments, and the quality of the respondents has been a center of attention [@coppock2019generalizing; @clifford2015samples; @huff2015these; @berinsky2014separating].
@@
 Trends of polarization have shown that people do interpret information according to their ideological position [e.g., @rekker2022understanding; @lee2021more].
-*Does this mean that some are right and others are wrong? Or is there an ideological difference in the ground truth?* 
-These questions present a fundamental challenge to the main way of working when collecting gold standard data, as we operate from the baseline assumption that disagreement among the annotators should be avoided or reduced.
+*Does this mean that some are right and others are wrong?
+Or is there an ideological difference in the ground truth?* These questions present a fundamental challenge to the main way of working when collecting gold standard data, as we operate from the baseline assumption that disagreement among the annotators should be avoided or reduced.
 Typically, when specific cases continuously cause disagreement, more instructions are added to limit interpretations (for recent innovations to improve annotation, see for example: @barbera2021automated; @struthers2020bridging; @winter2020online; @debell2013harder; @benoit2016crowd, @ying2022topics).
@@
 To do so, we have fielded two high-powered pre-registered experiments (see [here](link) and [here](link)) in the Netherlands -- a low-level polarized country -- and in the U.S.
--- a high-level polarized country testing the effect of ideological distance between the annotator and the political actor in the text (H1), the effect of overinterpretation based on political knowledge or ideological engagement (H2), and an offered solution of masking the political actor to mitigate the effects of ideology and knowledge (H3).^[The data and research compendium is published on the [[main author's github page]](https://github.com/MarikenvdVelden/bias-experiment) -- annonymized for the review process.]
+-- a high-level polarized country testing the effect of ideological distance between the annotator and the political actor in the text (H1), the effect of overinterpretation based on political knowledge or ideological engagement (H2), and an offered solution of masking the political actor to mitigate the effects of ideology and knowledge (H3).[^1]
 In the experiments, we vary the level of specification with which a political actor takes a stance -- a declarative sentence versus a sentence where with some knowledge on politics, the stance might be inferred -- as well as whether the political actor is shown or masked with putting `[ACTOR]` instead of the political party.
@@
 
-Our results demonstrate that overall sentences where with some knowledge on politics the stance might be inferred are really difficult for the crowd to annotate -- people overinterpret the position using their own knowledge of the world.
+[^1]: The data and research compendium is published on the [[main author's github page]](https://github.com/MarikenvdVelden/bias-experiment) -- anonymized for the review process.
+
+Our results demonstrate that overall sentences where with some knowledge on politics the stance might be inferred are more likely to be overinterpreted by the crowd, inferring a position where none is explicitly given.
 This is problematic as these sentences are very common in political text -- like legislative debates or speeches -- as well as media reports.
@@
 
-Generating large data-sets has become one of the main drivers of progress in natural language understanding.
+Generating large data sets has become one of the main drivers of progress in natural language understanding.
 In studies of political communication, the most familiar annotation tasks involve identifying theoretical concepts.
@@
 Yet, having only a few workers annotate the majority of text of interest has raised concerns about data diversity and models' ability to generalize beyond the crowd-workers.
-In a series of experiments, @geva2019we show that often models do not generalize well to annotations from annotators that did not contribute to the training set, suggesting that annotator bias should be monitored during data-set creation.
+In a series of experiments, @geva2019we show that often models do not generalize well to annotations from annotators that did not contribute to the training set, suggesting that annotator bias should be monitored during data set creation.
 One such potential bias, especially in times of increasing polarization [@iyengar2019origins; @gidron2019toward; @boxell2022cross], is based on ideological position of the annotator.
@@
 An additional reason to monitor the annotators' ideological position as a potential source of annotator bias is that a recent study in NLP showed that experiential factors influence the consistency of how political ideologies are perceived [@shen2021sounds].
-Their finding challenges the "ground-truth" assumption we as researcher make that a position for example is either left-leaning or right-wing leaning.
+Their finding challenges the "ground-truth" assumption we as researcher make that a position for example is either left-leaning or right-leaning.
 People with different ideological backgrounds might experience that position differently.
@@
 
-Another aspect to consider is that disagreement can be informative for the concept under measure - sometimes agreement can be used to validate hypotheses about how universal the perceptions of such concepts are [@sommerauer2020would].
+Another aspect to consider is that disagreement can be informative for the concept under measure -- sometimes agreement can be used to validate hypotheses about how universal the perceptions of such concepts are [@sommerauer2020would].
 Additional doubts on smoothing out disagreement in annotation have focussed on the lack of diversity when only annotating with one label or annotator, leading to a homogeneity especially in subjective and social tasks [@geva2019we] such as hatespeech detection or political affiliation classification.
-The question then is: *Whose perspective is being recorded in these datasets, and then later in the models trained on these datasets?* Framing arbitrary representations in data as "bias" misses the political character of datasets: There is no neutral data and no apolitical standpoint from where we can call out bias.
-Datasets are always "a worldview" and, as such, data always remains biased" [@miceli2022studying p.5]. 
+The question then is: *Whose perspective is being recorded in these datasets, and then later in the models trained on these datasets?* Framing arbitrary representations in data as "bias" misses the political character of data sets: There is no neutral data and no apolitical standpoint from where we can call out bias.
+Data sets are always "a worldview" and, as such, data always remains biased" [@miceli2022studying p.5].
 This is key to social scientist in general, and those studying political text in particular, since several tasks of interest are intrinsically societal, with answers that differ based on the make-up of the worldview of annotators.
-The answer of the annotators in turn influence how machine learning examples classify new models. 
+The answers of the annotators in turn influence how machine learning examples classify new models.
 For instance, hate speech and abuse detection are NLP tasks where race and gender of annotators influences both annotators and model performance [@gordon2022; @larimore2021reconsidering; @waseem2016].
@@
 @troiano2021emotion [p.2] also note how for a complex concept such as the "emotion" of a text, the annotator can make several assumptions during the annotation process on what is wanted, that are all valid and may or may not be useful in different contexts: "It is possible to assess one's own emotion after reading the text, to reconstruct the affective state of the writers who produced it, to guess the reaction that they intended to elicit in the readers, and so on." @shen2021sounds find that political ideology is not an inherent concept in many texts, but rather dependent on who is asked to annotate and their perceptions and background.
-Extralinguistic factors, such as annotator's own political ideology and also knowledge, influenced annotation and in turn model performance.
+Extralinguistic factors, such as annotators' own political ideology and also knowledge, influenced annotation and in turn model performance.
 @thornjakobsen2022 specifically analyze a task related to stance detection, argumentative sentence detection, on these extralinguistic factors, and find an effect of gender and political leaning on annotations and also model performance.
@@
 So, one solution is masking.
-We therefore test whether masking reduces potential differentiation incited by ideoloigcal postion or political knowledge with the following hypotheses:
+We therefore test whether masking reduces potential differentiation incited by ideological position or political knowledge with the following hypotheses:
 
@@
 These effects are, however, going in the other direction than hypothesized.
-For those who are ideologically further away from the party, they are less likely to be wrong than those who are close too the party, as shown in the right-hand panel of Figure \ref{fig:results-h1}.
+Those who are ideologically further away from the party are less likely to be wrong than those who are close to the party, as shown in the right-hand panel of Figure \ref{fig:results-h1}.
 For the Dutch case, using the strict interpretation, the difference is about `20%` -- from a coefficient of `-0.6` to `-0.4`.
@@
 For those issues, we see that almost everyone interprets the sentence correct (in blue and red).
-We also see that for a lenient interpretation of stances, people are quite likely to over interpret a position as a stance.
+We also see that for a lenient interpretation of stances, people are quite likely to overinterpret a position as a stance.
 Being correct about the stance does not decrease when masking the political actor in both cases.
@@
 That means there is a 15% increase in having it correct.
-The difference for over-interpreting is smaller between revealed and masked political actor -- shown in the right-hand panel of Figure **OA.XX** -- yet also statistically significant.
-Compared to `85%` over-interpreting the sentence as a stance, in the masking solution "only" `80%` over-interprets the sentence as a stance.
+The difference for overinterpreting is smaller between revealed and masked political actor -- shown in the right-hand panel of Figure **OA.XX** -- yet also statistically significant.
+Compared to `85%` overinterpreting the sentence as a stance, in the masking solution "only" `80%` over-interprets the sentence as a stance.
 In the recommendation section, we will reflect on the masking solution for under-specified sentences.
@@
 # Discussion
+
 Human annotation represents a crucial instrument in the toolkit of the social sciences.
 In order to address concerns regarding the validity and reliability of annotation tasks, we establish strict standards.
-Especially in these days, where due to the democratization of data and the advances in NLP more data can be analyzed or classified, these standards are more important than ever.
+Especially in times where due to the democratization of data and the advances in NLP more data can be analyzed or classified, these standards are more important than ever.
 After all, an algorithm trained on biased data will reproduce and often exacerbate bias [e.g. see @prost2019debiasing].
 It is key to realize that the underlying assumption to create valid and reliable annotated data is that there is only one correct interpretation for every input example, and that disagreement between the annotators is something that needs to be dealt with at all costs [@aroyo2015truth].
-In times of increasing factual opinion polarization [@rekker2022understanding; @lee2021more], it is crucial to ask *who's perspective is being recorded in these datasets*, given that no neutral data and no apolitical standpoint from where we can call out bias: "Datasets are always 'a worldview' and, as such, data always remains biased" [@miceli2022studying, p.5].
+In times of increasing factual opinion polarization [@rekker2022understanding; @lee2021more], it is crucial to ask *whose perspective is being recorded in these datasets*, given that there is no neutral data and no apolitical standpoint from where we can call out bias: "Datasets are always 'a worldview' and, as such, data always remains biased" [@miceli2022studying, p.5].
 Our tools to create valid and reliable annotation data does currently not allow for dealing with different perspectives.
@@
 Although we did not find a main effect of ideological distance (H1a), we did find evidence that the effect of ideological distance is greater for underspecified sentences (H1b) in the US, depending on how one operationalizes stances.
-THe exploratory analyses showed that these findings where robust agaist differnt specifications of ideological distance.
+The exploratory analyses showed that these findings where robust against different specifications of ideological distance.
 This conditional finding is important, because machine translation is on the rise [@licht2023cross; @de2018no]: Social scientists use this technique for example to comparatively analyze political stances.
@@
 When a party talks about a political issue without taking an explicit stance, annotators need to fill in the gaps and make their own inferences about the party's stance on this issue.
-Our results imply that political knowledge is indeed a relevant dimension in shaping how an annotator perceives political stances.
+Considering that in many political documents and specifically reporting on political news statements are not made explicit, underspecified sentences can be assumed to rather be the rule than the exception.
+Our results therefore imply that political knowledge is indeed a relevant dimension in shaping how an annotator perceives political stances.
 
@@
 Even if a good amount of crowd coders is used, it cannot blindly be assumed that this provides a diverse and balanced reflection of political ideologies in society, as we also observed in our samples (Tables \ref{tab:profilenl} and \ref{tab:profileus}).
+Especially in times where an increasing amount of crowd coding answers are based on large language models such as ChatGPT which (by and large) simulate the same ideological position over and over again [@veselovsky2023artificial], reflecting on the composition and quality of annotators and how it influences gold standards becomes more than just a methodological exercise.
 Our results indicate that an annotation team that is skewed towards one end of the political spectrum could indeed result in a biased gold standard.
@@
 But if the study involves making inferences about how the audience interprets political stances, for example in media effects research, then a ground-truth definition of political stance is problematic.
-To make accurate inferences about how a citizen subjectively perceived political content, and consequently how this might have affected behavior such as voting, we need to learn more about how shared personal characteristics relate to shared modes of perspective.
+To make accurate inferences about how a citizen subjectively perceives political content, and consequently how this might have affected behavior such as voting, we need to learn more about how shared personal characteristics relate to shared modes of perspective.
 
+```{=tex}
 \newpage
 \singlespacing
+```
 # References

### 2023-07-15 report/draft.Rmd

@@
 The behavior of citizens and the study of them by social scientists rely heavily on key information.
+<!--Word count: `r as.integer(sub("(\\d+).+$", "\\1", system(sprintf("wc -w %s", knitr::current_input()), intern = TRUE))) - 20`-->
+
+# Introduction
+The behavior of citizens and the study thereof by social scientists both rely heavily on key information.
 Classical theories of political communication, such as agenda setting or framing [e.g.
@@
 A recent study by @van2021validity demonstrates that using crowd-coding platforms is a good way to collect such gold standards without too much disagreement.
+A recent study by @van2021validity demonstrates that using crowd-coding platforms is a good way to collect such golden standards without too much disagreement.
 These platforms have also been used for experiments, and the quality of the respondents has been a center of attention [@coppock2019generalizing; @clifford2015samples; @huff2015these; @berinsky2014separating].
@@
 Especially in times where due to the democratization of data and the advances in NLP more data can be analyzed or classified, these standards are more important than ever.
+Human annotation is a crucial instrument in the toolkit of the social sciences.
+In order to safeguard the validity and reliability of annotation tasks, strict standards need to be upheld.
+Especially in these days, where due to the democratization of data and the advances in NLP more data can be analyzed or classified, these standards are more important than ever.
 After all, an algorithm trained on biased data will reproduce and often exacerbate bias [e.g. see @prost2019debiasing].

### 2023-07-16 report/draft.Rmd
fixed references'

@@
 The democratization of data and advent of computational social science has paved the way for new possibilities of gathering and analyzing textual data [for a recent overview, see @vanatteveldt2022computational].
-In particular, advances in Natural Language Processing (NLP) have made it possible to automatically analyze large quantities of data using machine learning [e.g., see @grimmer2022text; @wei2023overview].
+In particular, advances in Natural Language Processing (NLP) have made it possible to automatically analyze large quantities of data using machine learning [e.g., see @grimmer2022text; @akyurek-etal-2020-multi, @piskorski2023semeval].
 To determine the validity of such large scale computational analyses, we rely on "gold standard" data, created by human annotators.
@@
 Typically, when specific cases continuously cause disagreement, more instructions are added to limit interpretations (for recent innovations to improve annotation, see for example: @barbera2021automated; @struthers2020bridging; @winter2020online; @debell2013harder; @benoit2016crowd, @ying2022topics).
-However, work in computational linguistics has shown that increased annotation instructions do not increase quality [@parmar2022don].
+However, work in computational linguistics has shown that increased annotation instructions do not increase quality [@parmar2023don].
 This leaves us between a rock and a hard place.
@@
 If the stance is clearly specified, then there is more of a ground truth, and the opinions and perspectives of the annotator therefore matter less.
-It is when the stance is underspecified that perspectivism [@basile2021toward; @havens2022beyond] comes into play.
+It is when the stance is underspecified that perspectivism [@cabitza2023toward; @havens2022beyond] comes into play.
 When a party talks about a political issue without taking an explicit stance, annotators need to fill in the gaps and make their own inferences about the party's stance on this issue.

### 2023-07-17 report/OnlineAppendix.Rmd
update manuscript and add OA

---
output: 
  stevetemplates::article:
    number_sections: TRUE
    toc: yes
    dev: cairo_pdf
title: "ONLINE APPENDIX: Whose Truth is it Anyway? An Experiment on Annotation Bias in Times of Factual Opinion Polarization"
bibliography: references.bib
biblio-style: apsr
geometry: margin=1in
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
endnotes: FALSE
mainfont: cochineal
sansitup: FALSE
params:
  anonymous: ""
  doublespacing: ""
  removetitleabstract: ""
appendix: TRUE
appendixletter: "OA"
pandocparas: FALSE
anonymous: "`r params$anonymous`"
doublespacing: "`r params$doublespacing`"
removetitleabstract: "`r params$removetitleabstract`"
linkcolor: black
#nocite: |
#  @cartersmith2020fml, @andersetal2020bbgb
header-includes: 
   - \usepackage{array}
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \floatsetup[table]{capposition=top}
   - \usepackage{booktabs}
   - \usepackage{longtable}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
   - \usepackage{tabu}
   - \usepackage[para,online,flushleft]{threeparttable}
   - \usepackage{threeparttablex}
   - \usepackage[normalem]{ulem}
   - \usepackage{makecell}
   - \usepackage{xcolor}
   - \usepackage{caption}
   - \usepackage{graphicx}
   - \usepackage{epstopdf}
   - \usepackage{siunitx}
   - \usepackage{multirow}
   - \usepackage{hhline}
   - \usepackage{calc}
   - \usepackage{tabularx}
   - \usepackage{fontawesome}
   - \usepackage{amsthm}
   - \usepackage{csquotes}
   - \LTcapwidth=.95\textwidth
   - \linespread{1.05}
   - \usepackage{hyperref}
   - \usepackage{setspace}
   - \renewcommand{\thepage}{OA-\arabic{page}}
---

```{r setup, include=FALSE}
## include this at top of your RMarkdown file for pretty output
## make sure to have the printr package installed: install.packages('printr')
library(knitr)
opts_chunk$set(echo = FALSE,
                      message=FALSE, warning=FALSE,
               fig.pos = "H", out.extra = "")
options(kableExtra.latex.load_packages = FALSE)
```

```{r "data"}
here::i_am("report/OnlineAppendix.Rmd")
source(here::here("src/lib/functions.R"))
load(here("data/intermediate/cleaned_data_nl.RData"))
load(here("data/intermediate/cleaned_data_us.RData"))
```

\newpage
# A-1. Power Analysis
As detailed in our pre-analysis plan, we conduct multi-level regressions for the four sentences.
The manipulation, i.e. the masking condition vs. the political actor, as well as the specification of the sentence (i.e. _fully specified_ or _underspecified_) as well as respondents ideological distance to the party and their levels of political knowledge are the main independent variables.

To calculate power for the hypotheses, the R package `DeclareDesign` is used [@blair2019declaring].
The effect sizes are between $b = 0.2$ and $b = 0.1$ -- i.e. a small effect visualized by the purple and blue lines in Figure \ref{fig:power1}.
The hypothesis are directional, Figure \ref{fig:power1} therefore displays one-tailed tests with $\alpha = 0.05$.
The power analysis shows that testing the hypotheses requires a sample size of $2,000$ participants, who all annotate the four sentences, i.e. $8,000$ participants in total, (x-axis) to reach very high levels of power, i.e. $>95\%$, (indicated by the black dashed line in Figure \ref{fig:power}).

```{r power, fig.align = 'center', out.width = "100%", fig.cap = "\\label{fig:power1}Power Analysis", fig.topcaption=TRUE}
knitr::include_graphics(here::here("docs/pap_files/figure-latex/power-1.pdf"))
```

\newpage
# A-2. Data Collection: Questionnaire & Stimulus Material

## Questionnaire

**_Dependent Variables_**.
We rely on whether or not a party's (implied) stance is coded according to the party’s position (H1 and H3) as well as whether or not the statement is coded as a stance at all (H3). For each issue, we ask the respondent `what is according to the sentence above the position of [ACTOR]?`, with the answer categories: `in favor`, `against`, `no stance`, `don't know`.

&nbsp;

**_Control Variables_**.
As control variables, the following _demographics_ are measured post-treatment: gender, age, education, geographical region, level of urbanization, employment, and income.
For the analysis, only variables that are unbalanced over the experimental conditions will be included.

- _Gender_ is measured as `sex`. The answer categories are `Male` (value of `1`), `Female` (value of `0`), and `No answer` (value of `999`).

- _Age_ is measured using 6 categories: `17 or younger`, `18--29`, `30--39`, `40--49`, `50--59`, `60--74`.

- _Education_ is measured as the highest successfully completed level of education, recoded into four categories: `low`, `middle`, `high`, and `none`.
We created dummy variables for each level of education with the lowest category as base category.

- _Employment_ Respondents were asked which category of employment -- `Full-time employed`, `Part-time employed`, `Entrepreneur`, `Unemployed and searching for a job`, `Unemployed and not searching for a job or incapacitated`, `Housewife/Househusband or else`, `Retired`, `Student or full-time education` -- applied most to them.

- _Income_ Respondents were questioned on their monthly income in bins of 500 Euro's or Dolllars -- `500 or less`, `501-1000`, `1001-1500`, `1501-2000`, `2001-2500`, `2501-3000`, `3001-3500`, `3501-4000`, `4501-7500`, `7501 or more` -- as well as giving them the options of `won't  say` and `don't know`.

- _Geographical region_ is measured using the _Nielsen districs_, dividing the Netherlands into 1) the 3 major cities plus suburbs, Amsterdam (plus Diemen, Ouder-Amstel, Landsmeer, Amstelveen), Rotterdam (plus Schiedam, Capelle aan den IJssel, Krimpen aan den IJssel, Nederlek, Ridderkerk, Barendrecht, Albrandswaard) and The Hague (plus Leidschendam, Voorburg, Rijswijk, Wassenaar, Wateringen); 2) West (Noord-Holland, Zuid-Holland and Utrecht (excluding the major cities and their suburbs; 3) North (Groningen, Friesland and Drenthe), 4) East (Overijssel, Gelderland and Flevoland); and South (Zeeland, Noord-Brabant and Limburg). In the American case, we use the states.

In addition, pre-treatment, respondents' ideological position, political knowledge, vote recall, and position on the issues migration, climate, tax, and EU are measured.
Those variables will only be included in the analyses if balance checks indicate they are necessary.
Moreover, the variables will be used to explore heterogeneous relationships.

- _Ideological position_ is measured using an 11-point scale ranging from left (`0`) to right (`10`) in the Dutch case, since this is common in the national election studies (DPES). In the American case, we use the ANES standard liberal-conservative 7 point scale from extremely conservative to extremely liberal.

- _Vote Recall_ Respondents were asked which party they voted for in the 2021 parliamentary elections in the Dutch case, since party id is not a hellpful measure in this context. The options were 1) all parties that were elected into parliament -- `Bij1`, `BoerBurgerBeweging`, `CDA`, `ChristenUnie`, `D66`, `Denk`, `Forum for Democracy`, `JA21`, `GroenLinks`, `PvdA`, `Animal Rights Party`, `PVV`, `SGP`, `SP`, `VOLT`,`VVD`, `50Plus Party` -- 2) another party; 3) blanco vote; and 4) a `Don't know` option.

- _Party ID_ For the American case, respondents where asked whether they think of themselves as a Democrat, a Republican, an independent, or
something else.

- _Political knowledge_ is measured with six items from the DPES for the Dutch study and four items from the ANES for the American study.

- _Position on migration_ is measured by asking people whether or not there are too many immigrants in the Netherlands using a 5-point Likert-scale (`fully disagree`, `disagree`, `neutral`, `agree`, `fully agree`).

- _Position on climate_ is measured by asking people whether or not the climate crisis is exaggerated using a 5-point Likert-scale (`fully disagree`, `disagree`, `neutral`, `agree`, `fully agree`).

- _Position on tax_ is measured by asking people whether or not the  tax rate for the highest earners should go up using a 5-point Likert-scale (`fully disagree`, `disagree`, `neutral`, `agree`, `fully agree`).

- _Position on the EU_ is measured by asking people whether or not membership in the European Union has been especially bad for the Netherlands so far using a 5-point Likert-scale (`fully disagree`, `disagree`, `neutral`, `agree`, `fully agree`).

- _Position on Foreign Policy_ is measured by asking people whether or not say the U.S. needs to consider military build-up in the Pacific Ocean using a 5-point Likert-scale (`fully disagree`, `disagree`, `neutral`, `agree`, `fully agree`).

**_Attention Check_**.
We include one attention checks in the survey.
This is asked just before respondents enter the round of the experimental treatments.
The attention checks are taken from @berinsky2014separating and adapted to the Dutch context by the authors -- for the American data we use the original item.
If a respondents fails the first attention check, a warning appears and the respondent can only continue with the survey once the respondent has correctly answered the question correctly.

_When a big news story breaks people often go online to get up-to-the-minute details on what is going on. We want to know which websites people trust to get this information. We also want to know if people are paying attention to the question. To show that you have read this much, please ignore the question and select Volkskrant and Metro as your two answers. When there is a big news story, which is the one news website you would visit first? (Pleaseonly choose one). Eight (Dutch) news outlets are provided to choose from._ 
Respondents pass the attention check if they select **de Volkskrant** and **Metro**.

## Experimental Protocol
The study is conducted online and in Dutch for the Dutch data as well as in English for the American data.
Participants are told that they are taking part in a survey to get an overview of how people form their views on politics.
After reading an informed consent message participants are forwarded to the main questionnaire (or the survey will be terminated if they do not agree to the consent form).

First, participants complete a set of pre-treatment variables (i.e. vote recall, issue positions of the issues used in the experiment (Climate, EU/ Foreign Policy, Immigration, and Tax), ideology, and political knowledge).
This block ends with the attention check included in this survey.
When participants fail this attention check, a warning appears asking them to read the question again carefully  and  to  answer  again.
Therefter, participants see the experimental stimuli.
The stimuli in the experiment are to show respondents fully specified or underspecified sentences on the issue position of a political actor or a masked political actor (see Section Treatment for more details).

## Stimulus Material
Respondents are randomly assigned to either view a political party as an actor, or a masked condition, where they see `X` as an actor; simultaneously, respondents see either a fully specified sentence or a underspecified sentence, in which one needs additional information to interpret the positon on an actor.
Table \ref{tab:conditions} gives an overview of the variations in treatment  in the survey as well as their English translations.

```{r conditions, topcaption=TRUE}
d <- tibble(Condition = c(rep("Specified", 4),
                          rep("Underspecified", 4)), 
            `Wording ENG` = c("[PVV/X] says immigration should be made harder.",
                              "[GreenLeft/X] says nitrogen emissions need to be reduced.",
                              "[Labour Party/X] says tax rate should go up for highest earners.",
                              "[Forum for Democracy/X] says that membership in the European Union has been especially bad for the Netherlands so far.",
                              "[PVV/X] says many immigrants are coming this way.",
                              "[GreenLeft/X] says nitrogen policy must be different.",
                              "[Labour Party/X] says tax system must be changed.",
                              "[Forum for Democracy/X] says the Netherlands should have a different role in the European Union."),
            `Wording NL` = c("[PVV/X] zegt dat immigratie moeilijker gemaakt moet worden.",
                             "[GroenLinks/X] zegt dat stikstofuitstoot meer tegengegaan moet worden.",
                             "[PvdA/X] zegt dat het belastingtarief voor de hoogste inkomens omhoog moet.",
                             "[Forum voor Democratie/X] zegt dat het lidmaatschap van de Europese Unie tot nu toe vooral slecht geweest voor Nederland is.",
                             "[PVV/X] zegt dat veel immigranten deze kant op komen.",
                             "[GroenLinks/X] zegt dat het stikstofbeleid anders moet.",
                             "[PvdA/X] zegt dat het belastingstelsel moet worden aangepast.",
                             "[Forum voor Democratie/X] zegt dat Nederland een andere rol in de Europese Unie moet hebben."))

kbl(d, booktabs =T, caption = "\\label{tab:conditions}Survey Questions - Experimental Conditions") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "7cm") %>%
  column_spec(3, width = "7cm")
```

\newpage

# A-3. Balance Checkes

Figure \ref{fig:balance} below shows that the data is balanced on all variables.
As described in the Pre-Analysis Plan, we will therefore add no co-variates to the analyses as controls.

```{r "balance", out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:balance} Balance Test"}
knitr::include_graphics(here("report/figures", "balance-checks-1.png"))
```

\newpage

# A-4. Methods of Analyses

To test our hypotheses, we will conduct a multilevel model, with respondents clustered in issues, see Equation \ref{eq:pooled1}.
Using the pooled data we will estimate a within groups fixed effects model.
We have conducted a balance test based on demographics (age, gender, education, geographical region, level of urbanization, employment, and income), vote choice in the 2021 parliamentary elections, ideological self-placement, political knowledge, and positions on the issues, using the `cobalt` R package [@greifer2021].
This balance test indicated that none of the variables are unbalanced over the experimental groups, and therefore, as pre-registered, will not be added to the regression formula.
$Y\hat{Y}_{r, i, t}$ in Equation \ref{eq:pooled1} denotes the evaluation of a stance by respondent $r$, during issue $i$ and at experimental round $t$ -- ranging from round 1 to round 4.
The standard errors are clustered at the individual level.

```{=tex}
\begin{multline}\label{eq:pooled1}
    \hat{stance correct}_{r, i, t} =\beta_{0} + \beta_{1}masked_{r, i, t} +
    \beta_{2}specification_{r, i, t} + 
    \beta_{3}ideological distance to party_{r, i, t} +\\
    \beta_{4}political knowledge_{r, i, t} +
    \alpha_{i} + \gamma_{t} + \varepsilon_{r, i, t}
\end{multline}
```

\newpage

# A-5. Descriptive Analyses

Tables \ref{tab:profilenl} and \ref{tab:profileus} demonstrate the average profile of respondents who annotate correctly and incorrectly (where respondents who annotated some stances correctly and some incorrectly are weighted by proportion (in)correct).
In terms of demographics, there is not much of a difference.
Yet, people who are incorrectly identifying stances are more left-wing oriented compared to those who are correct -- i.e. an average score of four for those who are incorrect vs. an average score of five for those who are correct.
For other positions on issues or political knowledge, we do not see a difference in averages between those who are correctly and incorrectly identifying stances.
This profile is quite similar for the lenient interpretation of what a stance is.

```{r "profilenl-strict"}
load(here("data/intermediate/cleaned_data_nl.RData"))
load(here("data/intermediate/cleaned_data_us.RData"))
source(here("src/analysis/profile_resp.R"))
kbl(e1_nlp, booktabs =T, caption = "\\label{tab:profilenl}Profile Dutch Stance Annotators") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = "7cm") %>%
  column_spec(2, width = "7cm")
```

```{r "profileus"}
kbl(e1_us_nlp, booktabs =T, caption = "\\label{tab:conditions_us}Profile American Stance Annotators") %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = F, fixed_thead = T, position = "center") %>%
  column_spec(1, width = "7cm") %>%
  column_spec(2, width = "7cm")
```

\newpage

# A-6. Baseline Analyses

Looking at the effect of the experimental conditions on the four dependent variables -- correctly identifying a stance and over-interpreting a stance for both a strict and a lenient interpretation of stance -- Figure \ref{fig:results-base} visualizes the baseline.
The left-hand panel demonstrates the effect of the two experimental conditions for correctly identifying the stance in the Dutch case.
The right-hand panel does so for the American case.
On average, many respondents in both cases (respectively `85%` in the Dutch case and `65%` in the American case) correctly interpreted the stance using either the lenient or strict interpretation (respectively in blue and red) -- as indicated by the intercept.
When we mask the political actor -- i.e. instead of mentioning the party, we put "X" -- we see that this does on average not improve correctly interpreting the stance significantly in neither the Dutch or the American case.
Additionally, we do see that the level of specification of a sentence has a significant effect.
If a sentence is not fully specified, it has a substantive negative effect on the likelihood to correctly interpret the stance in both the lenient and strict interpretation of a stance.
These effects are substantial in both the American and Dutch case, with coefficients varying between `-0.2` and `-0.5`.
This indicates that compared to a fully specified sentence, between `20%` and `50%` of the respondents are more likely to be incorrect when the sentence is under-specified -- that is when the sentence does not state a clear position, but mentions the issue.
Looking at the other dependent variable, whether they interpreted the sentence as a stance or not, we see that almost nobody overinterprets a stance in the strict interpretation in either the Dutch or American case.
Yet, they do overinterpret a stance in the lenient interpretation.
Moreover, if people see an X compared to a political actor, they are statistically significantly more likely to interpret the sentence as a stance in its strict interpretation.
Yet, a coefficient of `0.02` (i.e. `2%`) is a very small effect.
For the condition of specification level, however, we see that compared to a fully specified sentence, people seeing an under-specified sentence are much more likely to interpret the sentence as a stance in the strict interpretation: an increase of `0.83`.
This indicates that people do not excel in this task without any instruction.
Using the lenient interpretation, however, people seem less likely to annotate the sentence as a stance.
In the pre-registered section, we demonstrate the tests of the hypotheses, and afterwards, we discuss some explorations of the data to show the robustness of our findings, the visualizations thereof are displayed in OA A-3.

```{r results-baseline, out.width = "95%", fig.align = 'center', fig.cap = "\\label{fig:results-base} Baseline Results of Experimental Conditions"}
knitr::include_graphics(here("report/figures", "baseline-1.png"))
```

\newpage

# A-7. Robustness Analyses

To check the robustness of our findings, Figure \ref{fig:results-exp1} demonstrates the analyses for each issue separately.
The different colors visualize the different dependent variables.
We do not see much variation between issues *Tax*, *EU/Foreign Policy*, and *Environment*.
For those issues, we see that almost everyone interprets the sentence correctly (in blue and red).
We also see that for a lenient interpretation of stances, people are quite likely to overinterpret a position as a stance.
Being correct about the stance does not decrease when masking the political actor in both cases.
Yet, the chance of being correct decreases statistically significantly when the sentence is underspecified.
The same holds for overinterpreting for the lenient interpretation, but the opposite is true for the strict interpretation; there overinterpretation is more likely with underspecified sentences.
Looking at *Immigration*, we see a different pattern.
We see that masking does not increase the likelihood of being correct, but does increase the likelihood of overinterpretation regardless of how one defines a stance.
Underspecified sentences are less likely to be correctly identified and more likely to be overinterpreted regardless of the definition of a stance.
So, while there are some differences in effect sizes between the issues, the overall findings are not driven by a single issue.

```{r results-exploration1, out.height = "85%",out.width = "95%",, fig.align = 'center', fig.cap = "\\label{fig:results-exp1} Exploration: Issue Specific Analyses", fig.pos="H", fig.show="hold"}
knitr::include_graphics(c(here("report/figures", "issues-1.png")))
```

In addition to issue-specific analyses, we also explore an interaction between treatments, visualized in Figure \ref{fig:results-exp2} for both dependent variables.
This shows that masking is of help when sentences are under-specified.
In the left-hand panel of Figure \ref{fig:results-exp2}, it demonstrates that for under-specified sentences, people are less likely to incorrectly identify a sentence as a stance when the actor is masked (coefficient of `-0.30`) than when an actor is revealed (coefficient of `-0.45`).
That means there is a 15% increase in having it correct.
The difference for overinterpreting is smaller between revealed and masked political actors -- shown in the right-hand panel of Figure \ref{fig:results-exp2} -- yet also statistically significant.
Compared to `85%` overinterpreting the sentence as a stance, in the masking solution "only" `80%` over-interprets the sentence as a stance.
In the recommendation section, we will reflect on the masking solution for under-specified sentences.

```{r results-exploration2, out.height = "65%",out.width = "95%",, fig.align = 'center', fig.cap = "\\label{fig:results-exp2} Exploration: Interaction bewtween Treatments", fig.pos="H", fig.show="hold"}
knitr::include_graphics(c(here("report/figures", "exploration-1.png")))
```

Lastly, we explore three different ways of measuring ideological distance and an alternative for political knowledge in the American case.
First, we measured ideological bias by looking at the ideology of the respondents -- not in relation to the political actor revealed, visualized in the upper panels of Figure \ref{fig:results-exp3}.
Secondly, we measured ideological bias by looking at whether the respondent is congruent or not with the issue position in the sentence, visualized in  the lower panels of Figure \ref{fig:results-exp3}.

```{r results-exploration3, out.height = "85%",out.width = "95%",, fig.align = 'center', fig.cap = "\\label{fig:results-exp3} Exploration: Interaction bewtween Treatments", fig.pos="H", fig.show="hold"}
knitr::include_graphics(c(here("report/figures", "exploration-3.png")))
```

And thirdly, we measured ideological bias by looking at whether the person voted for the party displayed in the sentence, visualized in the upper-rows of Figure \ref{fig:results-exp4}. 
In none of the analyses, we find evidence for ideological bias.
Also for the alternative measurement of political knowledge in the US, we find the same results as reported in the main analyses.

```{r results-exploration4, out.height = "85%",out.width = "95%",, fig.align = 'center', fig.cap = "\\label{fig:results-exp4} Exploration: Interaction bewtween Treatments", fig.pos="H", fig.show="hold"}
knitr::include_graphics(c(here("report/figures", "exploration-4.png")))
```
 \newpage

# References

### 2023-07-17 report/draft.Rmd
fixed a typo
update manuscript and add OA

@@
 Due to the democratization of data and the advances in NLP more data can be analyzed or classified, making these standards are more important than ever: An algorithm trained on biased data will reproduce and often exacerbate bias.
-Our tools to create valid and reliable annotation data does currently not allow for dealing with different perspectives of annotators.
+Our tools to create valid and reliable annotation data currently do not allow for dealing with different perspectives of annotators.
 In two pre-registered experiments in the United States and the Netherlands, we show that personal characteristics of annotators, like political ideology or knowledge, interfere with annotators' judgement of political stances.
@@
 
-# Introduction
-
-The behavior of citizens and the study of them by social scientists rely heavily on key information.
 <!--Word count: `r as.integer(sub("(\\d+).+$", "\\1", system(sprintf("wc -w %s", knitr::current_input()), intern = TRUE))) - 20`-->
@@
 Classical theories of political communication, such as agenda setting or framing [e.g.
-@vanaelst2016political; @lecheler2019news], formulate that political information drives opinion formation and participation in politics -- from voting to protests [for overviews hereof across countries, see @pfetsch2013comparing].
-The democratization of data and advent of computational social science has paved the way for new possibilities of gathering and analyzing textual data [for a recent overview, see @vanatteveldt2022computational].
-In particular, advances in Natural Language Processing (NLP) have made it possible to automatically analyze large quantities of data using machine learning [e.g., see @grimmer2022text; @akyurek-etal-2020-multi, @piskorski2023semeval].
+@vanaelst2016political; @lecheler2019news], formulate that political information drives opinion formation and participation in politics -- from voting to protests.
+The democratization of data and advent of computational social science has paved the way for new possibilities of gathering and analyzing textual data [for recent overviews, see @vanatteveldt2022computational; @grimmer2022text].
+In particular, advances in Natural Language Processing (NLP) have made it possible to automatically analyze large quantities of data using machine learning [e.g., see @akyurek-etal-2020-multi; @piskorski2023semeval].
 To determine the validity of such large scale computational analyses, we rely on "gold standard" data, created by human annotators.
@@
 Most data collection efforts to create gold standards assume that there is only one correct interpretation for every input example, and that disagreement between the annotators is something that needs to be dealt with at all costs [@aroyo2015truth].
-A recent study by @van2021validity demonstrates that using crowd-coding platforms is a good way to collect such gold standards without too much disagreement.
 A recent study by @van2021validity demonstrates that using crowd-coding platforms is a good way to collect such golden standards without too much disagreement.
@@
 For example, coders are more likely to report a left-wing party to be pro-immigration and a populist right-wing party to be against based on the exact same sentence.
-*Is this actually bias or a diversity of view points?
-And how big of a problem does this bias/diversity of view points generate for researchers relying on golden standard data?* This question also taps into newer developments related to data annotations: Increasingly, Large Language Models such as ChatGPT are being used by researchers [@gilardi2023chatgpt] as well as annotators themselves [@veselovsky2023artificial] to "simulate" human respondents for annotation tasks.
+*Is this actually bias or a diversity of viewpoints? And how big of a problem does this bias/diversity of viewpoints generate for researchers relying on golden standard data?* 
+This question also taps into newer developments related to data annotations.
+Increasingly, Large Language Models such as ChatGPT are being used by researchers [@gilardi2023chatgpt] as well as annotators themselves [@veselovsky2023artificial] to "simulate" human respondents for annotation tasks.
 These models draw among other things from the gold standard data sets that have been created by researchers, perpetuating the specific biases that are present in them and not allowing for any further disagreement or diversity of viewpoints.
@@
 There is a long history of text annotation in studies analyzing political text.
-While this yields lots of experiences as to how to train coders so that we get reliable hand-coded data, the procedure is expensive, protracted, and sometimes does not even get us the quality of data we need [@weber18].
+While this yields lots of experiences as to how to train coders to get reliable hand-coded data, the procedure is expensive, protracted, and sometimes does not even get us the quality of data we need [@weber18].
 While the crowd offers a solution to some of these issues [@van2021validity], the main underlying assumption that there is one correct interpretation for every input example remains untouched.
-However, especially for more complex coding tasks different research fields have different interpretations of what counts as *correct* solution depending on the main goal of annotation.
+However, especially for more complex coding tasks, research fields differ in their interpretations of what counts as a *correct* solution depending on the main goal of annotation.
 Do we want to create a clean benchmark data set that follows strict linguistic rules or one that reflects how statements are being interpreted by "real" readers that do not necessarily follow theoretical definitions on which elements are needed for a specified stance?
+
 Trends of polarization have shown that people do interpret information according to their ideological position [e.g., @rekker2022understanding; @lee2021more].
@@
 This leaves us between a rock and a hard place.
-Is there a potential bias in annotators that we should account for?
+*Is there a potential bias in annotators that we should account for?*
 
@@
 the adoption of methods that integrate the opinions and perspectives of the human subjects involved in the knowledge representation step of the machine learning processes [@cabitza2023toward].
-This literature puts forward that disagreement can occur because of differences in ideological position or political knowledge [@shen2021sounds; @alkiek2022classification; @joseph2021mis].
-This allows us to test the extent to which disagreement takes place, for what type of stances, as well as gives us directions on how to deal with the diversity in conceptions and the political heterogeneity that nowadays potentially occurs in our sample of annotators.
-To do so, we have fielded two high-powered pre-registered experiments (see [here](link) and [here](link)) in the Netherlands -- a low-level polarized country -- and in the U.S.
--- a high-level polarized country testing the effect of ideological distance between the annotator and the political actor in the text (H1), the effect of overinterpretation based on political knowledge or ideological engagement (H2), and an offered solution of masking the political actor to mitigate the effects of ideology and knowledge (H3).[^1]
+<!--This literature puts forward that disagreement can occur because of differences in ideological position or political knowledge [@shen2021sounds; @alkiek2022classification; @joseph2021mis].-->
+We test the extent to which disagreement takes place, for what type of stances, as well as gives us directions on how to deal with the diversity in conceptions and the political heterogeneity.
+<!-- that nowadays potentially occurs in our sample of annotators.-->
+We use two high-powered pre-registered experiments (see [here](link) and [here](link)) in the Netherlands -- a low-level polarized country -- and in the U.S. -- a high-level polarized country.
+We test possible biases stemming from ideology and political knowledge for correctly and overinterpreting stances, as well as whether masking he political actor to mitigates these biases.[^1]
+<!--effect of ideological distance between the annotator and the political actor in the text (H1), the effect of overinterpretation based on political knowledge or ideological engagement (H2), and an offered solution of masking the political actor to mitigate the effects of ideology and knowledge (H3).-->
 In the experiments, we vary the level of specification with which a political actor takes a stance -- a declarative sentence versus a sentence where with some knowledge on politics, the stance might be inferred -- as well as whether the political actor is shown or masked with putting `[ACTOR]` instead of the political party.
 We do this for four different political issues: *Environment*, *Immigration*, *Tax Policy*, and *EU* (for the Dutch case) or *Foreign Policy* (for the American case).
-This country selection does not only allow us to showcase the scope conditions of disagreement due to different levels of political heterogeneity, but also differentiates between languages.
-English is not only the most dominant language for computational text analysis in the social science [@baden2022three; @dolinsky2023threegaps], crowd-coders do not need to be first language speakers, given the dominance of English in our daily lives.
+This country selection not only allow us to showcase the scope conditions of disagreement due to different levels of political heterogeneity, but also differentiates between languages.
+English is not only the most dominant language for computational text analysis in social science [@baden2022three; @dolinsky2023threegaps], crowd-coders do not need to be first language speakers, given the dominance of English in our daily lives.
 This is different for Dutch, it is a language spoken by a smaller community, typically first language speakers, yet still an often-enough researched case in computational text analysis in the social science [@baden2022three; @dolinsky2023threegaps].
@@
 This is problematic as these sentences are very common in political text -- like legislative debates or speeches -- as well as media reports.
-Moreover, our results also demonstrate that for these disagreements in the crowd to occur, the level of polarization needs to be high.
-We do find support for our hypotheses in the American context, but not in the Dutch context -- except for the situation where masking overcomes political knowledge.
-In this case, we do find support in the Dutch experiment, but not in the American one.
-Our findings thus underline the importance of taking disagreement seriously for the creation of gold standard text -- the bread-and-butter of all machine learning endeavours.
+<!--Moreover, our results also demonstrate that for these disagreements in the crowd to occur, the context is important, as we find differences in disagreements between the American and Dutch context.-->
+Our findings underline the importance of taking disagreement seriously for the creation of gold standard text -- the bread-and-butter of all machine learning endeavors.
 We should look beyond the majority vote and model it in the data, because if an algorithm is trained on biased data from disagreeing annotators, it will reproduce and often exacerbate that bias when it is applied to new data [e.g., see @prost2019debiasing].
-To be able to model these characteristic of annotators, we should survey the characteristics of annotators when using the crowd (see Webb-Williams et al.
+To be able to model these annotators' characteristics, we should survey the characteristics of annotators when using the crowd (see Webb-Williams et al.
 2023 for a similar argument, yet different annotator characteristics).
@@
 
-Generating large data sets has become one of the main drivers of progress in natural language understanding.
-In studies of political communication, the most familiar annotation tasks involve identifying theoretical concepts.
+Generating large data sets has become one of the main drivers of progress in NLP.
+Studying political texts, the most familiar annotation tasks involve identifying theoretical concepts.
 This includes noting the topic of the text, the position of the actor or the tone of the text.
 Crowdcoding seems suitable for this task [@van2021validity].
-Yet, having only a few workers annotate the majority of text of interest has raised concerns about data diversity and models' ability to generalize beyond the crowd-workers.
+Yet, having only a few workers annotating the majority of texts has raised concerns about data diversity and models' ability to generalize beyond the crowd-workers.
 In a series of experiments, @geva2019we show that often models do not generalize well to annotations from annotators that did not contribute to the training set, suggesting that annotator bias should be monitored during data set creation.
-One such potential bias, especially in times of increasing polarization [@iyengar2019origins; @gidron2019toward; @boxell2022cross], is based on ideological position of the annotator.
+One such potential bias, especially in times of increasing polarization [@iyengar2019origins; @gidron2019toward; @boxell2022cross], is based on the ideological position of the annotator.
 Given that annotators on crowd-coding platforms tend to be younger and hold more liberal values than the general public [@clifford2015samples; @huff2015these], this could potentially hamper the data diversity and generalizability of the model.
 An additional reason to monitor the annotators' ideological position as a potential source of annotator bias is that a recent study in NLP showed that experiential factors influence the consistency of how political ideologies are perceived [@shen2021sounds].
-Their finding challenges the "ground-truth" assumption we as researcher make that a position for example is either left-leaning or right-leaning.
+Their finding challenges the "ground-truth" assumption we as researchers make that a position for example is either left-leaning or right-leaning.
 People with different ideological backgrounds might experience that position differently.
-This challenges our way of data collection.
-We are interested in the effect of e.g. elite communication.
-To study this, we allow for heterogeneous treatment effects in experimental work.
-This indicates that we often do not assume that the treatment, often using text, has the same effect for different partisans.
+This challenges our way of data collection: When we are interested in the effect of e.g. elite communication, we often allow for heterogeneous treatment effects in experimental work.
+This indicates that we do not assume that the treatment, most commonly using text, has the same effect for different partisans.
 Yet, at the same time, we forget or ignore that knowledge when creating large data-sets for our machine learning models.
 
-The field of Natural Language Processing often works on automatically classifying texts on labels of concepts such as stance, sentiment, and political orientation.
-These models are trained on data created by human annotators.
-Often, this process has a final step where disagreements and differences in annotations are leveled by aggregating, averaging, or in other ways coming to a consensus on one label for one example, which is then seen as "ground truth" [@aroyo2015].
+The field of Natural Language Processing also works on automatically classifying texts on labels of concepts such as stance, sentiment, and political orientation.
+These models are trained on data created by human annotators, including a final step where disagreements and differences in annotations are leveled by aggregating, averaging, or other ways of coming to a consensus on one label for one example -- i.e.
+a "ground truth" [@aroyo2015].
 Differences from this ground truth label are seen as errors that need to be removed or sorted out.
 Models then learn to predict labels for new examples based on this ground truth.
-However, since several years there is some discussion on how realistic it is to have one label, especially for subjective or complex concepts and/or texts with multiple interpretations.
-@aroyo2015 describes succinctly how the predominant annotation procedures for classification models run into myths such as "disagreement is bad" and "one annotation is enough".
-@plank2014learning adds to this that underlying ambiguity and linguistic complexity should be considered for disagreement in annotations: not all linguistic examples are created equal.
-Disagreement even occurs in seemingly objective tasks such as Part of Speech tagging [@fornaciari2021; @plank2014learning].
+Recently, there is some discussion on how realistic it is to have just one label for each example, especially for subjective or complex concepts and/or texts with multiple interpretations.^[Disagreement even occurs in seemingly objective tasks such as Part of Speech tagging [@fornaciari2021; @plank2014learning].]
+Annotation procedures for classification models run into myths such as "disagreement is bad" and "one annotation is enough" [@aroyo2015]. 
+Ambiguity and linguistic complexity should be considered for disagreement in annotations [@plank2014learning]: Not all linguistic examples are created equal.
+Moreover, disagreement can be informative for the researcher. 
+It can for instance be used to validate hypotheses about how universal the perceptions of such concepts are [@sommerauer2020would].
 
-Another aspect to consider is that disagreement can be informative for the concept under measure -- sometimes agreement can be used to validate hypotheses about how universal the perceptions of such concepts are [@sommerauer2020would].
-Additional doubts on smoothing out disagreement in annotation have focussed on the lack of diversity when only annotating with one label or annotator, leading to a homogeneity especially in subjective and social tasks [@geva2019we] such as hatespeech detection or political affiliation classification.
-The question then is: *Whose perspective is being recorded in these datasets, and then later in the models trained on these datasets?* Framing arbitrary representations in data as "bias" misses the political character of data sets: There is no neutral data and no apolitical standpoint from where we can call out bias.
+Additionally, doubts on smoothing out disagreement in annotation have focused on the lack of diversity.
+When only annotating with one label or annotator, we get homogeneity, which is undesirable in subjective and social tasks [@geva2019we] such as hate speech detection or political affiliation classification.
+*Whose perspective is being recorded in these datasets, and then later in the models trained on these datasets?* 
+Framing arbitrary representations in data as "bias" misses the political character of data sets: There is no neutral data and no apolitical standpoint from which we can call out bias.
 Data sets are always "a worldview" and, as such, data always remains biased" [@miceli2022studying p.5]. 
-This is key to social scientist in general, and those studying political text in particular, since several tasks of interest are intrinsically societal, with answers that differ based on the make-up of the worldview of annotators.
+<!--This is key to social scientist in general, and those studying political text in particular, since several tasks of interest are intrinsically societal, with answers that differ based on the make-up of the worldview of annotators.-->
 The answers of the annotators in turn influence how machine learning examples classify new models. 
@@
 
-Most recently, new annotation paradigms have gone one step further by asking whether we are modelling the task, or the annotator [@geva2019we].
-@pavlick2019 find that for the logical coherence task Natural Language Inference, annotators have several valid interpretations that are not reflected in one ground truth label.
-They call for new training paradigms that can reflect "the full range of possible human inferences"[@pavlick2019, p.688].
-Recent approaches in NLP have sought to explicitly incorporate disagreement and diversity in training data annotations.
-@rottger2022two introduce the idea of an explicitly subjective annotation paradigm existing in addition to one focussed on one label and "ground truth".
+Most recently, new annotation paradigms have gone one step further by asking whether we are modeling the task, or the annotator [@geva2019we].
+Because the logical coherence task Natural Language Inference showed that annotators have several valid interpretations that are not reflected in one ground truth label, @pavlick2019 call for new training paradigms reflecting "the full range of possible human inferences" (p.688).
+Recent approaches in NLP have sought to explicitly incorporate disagreement and diversity in training data annotations, such an explicitly subjective annotation paradigm [@rottger2022two], "perspectivism" [@cabitza2023toward], and "jury learning" [@gordon2022].
+<!--@rottger2022two introduce the idea of an explicitly subjective annotation paradigm existing in addition to one focused on one label and "ground truth".
 Such a subjective annotation paradigm can be used for purposes where the goal is finding diverse perspectives on the task or concepts, and for models to model more accurately how humans interpret a task or text.
@@
 In both approaches, demographic and other individual aspects of the annotator are explicitly mentioned and highlighted as having an influence on classification performance, but is used as an asset rather than as an error.
-
-@troiano2021emotion [p.2] also note how for a complex concept such as the "emotion" of a text, the annotator can make several assumptions during the annotation process on what is wanted, that are all valid and may or may not be useful in different contexts: "It is possible to assess one's own emotion after reading the text, to reconstruct the affective state of the writers who produced it, to guess the reaction that they intended to elicit in the readers, and so on." @shen2021sounds find that political ideology is not an inherent concept in many texts, but rather dependent on who is asked to annotate and their perceptions and background.
-Extralinguistic factors, such as annotators' own political ideology and also knowledge, influenced annotation and in turn model performance.
-@thornjakobsen2022 specifically analyze a task related to stance detection, argumentative sentence detection, on these extralinguistic factors, and find an effect of gender and political leaning on annotations and also model performance.
-However, to our knowledge this phenomenon has not yet been tested in a controlled experimental setting with manipulations in the texts.
+<!--"@troiano2021emotion [p.2] also note how for a complex concept such as the "emotion" of a text, the annotator can make several assumptions during the annotation process on what is wanted, that are all valid and may or may not be useful in different contexts.
+It is possible to assess one's own emotion after reading the text, to reconstruct the affective state of the writers who produced it, to guess the reaction that they intended to elicit in the readers, and so on." -->
+Political ideology is not an inherent concept in many texts, but rather dependent on who is asked to annotate and their perceptions and background [@shen2021sounds].
+Extra-linguistic factors, such as annotators' own political ideology and also knowledge, influenced annotation and in turn model performance [@thornjakobsen2022].
+<!--@thornjakobsen2022 specifically analyze a task related to stance detection, argumentative sentence detection, on these extra-linguistic factors, and find an effect of gender and political leaning on annotations and also model performance.
+However, to our knowledge this phenomenon has not yet been tested in a controlled experimental setting with manipulations in the texts.-->
 To test whether people with different ideological backgrounds experience a political position differently, challenging our ground-truth assumption in data annotation, we propose the following hypothesis:
@@
 
-As noted by @plank2014, some linguistic examples are ambiguous, and open to multiple interpretations even for seemingly objective tasks such as part of speech tagging, where annotators have to distinguish parts of speech such as nouns and verbs.
+<!--As noted by , some linguistic examples are ambiguous, and open to multiple interpretations even for seemingly objective tasks such as part of speech tagging, where annotators have to distinguish parts of speech such as nouns and verbs [@plank2014].-->
 A complex concept such as political ideology is much more likely to lead to multiple interpretations.
@@
 
-In a next step, we ask: *If these biases exists, how can we alleviate them?* There have been several previous approaches to solve biased annotations, especially where it concerns political or societal aspects.
+In a next step, we ask: *If these biases exist, how can we alleviate them?* There have been several previous approaches to solve biased annotations, especially where it concerns political or societal aspects.
 @geva2019we introduce an approach where training set annotators are separated from annotators annotating data sets that evaluate the models, to ensure the evaluation is not simply accurate at replicating the original annotators, but can generalize to new annotators' judgements.
@@
 We have conducted the survey experiments in the Netherlands in May 2022 and in the United States in January 2023.
-Both samples, recruited through [KiesKompas](https://www.kieskompas.nl/en/) and [Prolific](https://www.prolific.co/) for respectively the Dutch and American case, consist of 3,000 participants (based on the power analysis presented in our [online compendium](LINK)) of 18 years and older.
+Both samples, recruited through [KiesKompas](https://www.kieskompas.nl/en/) and [Prolific](https://www.prolific.co/) for respectively the Dutch and American case, consist of 3,000 participants (based on the power analysis presented in our [online compendium](LINK), and  Online Appendix (OA) A-1) of 18 years and older.
 Both survey companies works with non-random opt-in respondents.
-Therefore, we measured many demographic background variables, and balance checks have been conducted to demonstrate whether certain categories are over represented in a certain experimental group.
-Our study has been approved by the [Research Ethics Review Committee](https://fsw.vu.nl/nl/onderzoek/research-ethics-review/index.aspx) of the *Vrije Universiteit Amsterdam* (see the approval [here](https://github.com/MarikenvdVelden/bias-experiment/blob/master/docs/2022-3-30-59.pdf)).
-To ensure good quality of our data, one attention check (discussed in more detail in **OA XX** is included [@berinsky2014separating].
+Therefore, we measured many demographic background variables, and balance checks have been conducted to demonstrate whether certain categories are over represented in a certain experimental group (see OA A-3).
+Our study has been approved by the Research Ethics Review Committee of the researchers' institution.
+To ensure good quality of our data, one attention check (discussed in more detail in OA A-2) is included [@berinsky2014separating].
 
@@
 
-***Experimental Conditions.*** Respondents are randomly assigned to either view a political party as an actor, or a masked condition, where they see `X` as an actor; simultaneously, respondents see either a fully specified sentence or a underspecified sentence, in which one needs additional information outside of the text to determine an actor's position.
+***Experimental Conditions.*** Respondents are randomly assigned to either view a political party as an actor, or a masked condition, where they see `X` as an actor; simultaneously, respondents see either a fully specified sentence or an underspecified sentence, in which one needs additional information outside of the text to determine an actor's position.
 Table \ref{tab:conditions} gives an overview of the variations in treatment in the surveys.
@@
 Using the strict interpretation, respondents are correct if they say `no stance` for the underspecified sentences and `against`, `in favor`, `in favor`, and `against` to the specified sentences one to four.
-Using a more lenient interpretation, respondents could say either `in favor` or `against` as well for underspecified sentence two to four.
+Using a more lenient interpretation, respondents could say either `in favor` or `against` as well for underspecified sentences two to four.
 
 ***Moderating Covariates.*** *Ideological position* is measured using an 11-point scale ranging from left (`0`) to right (`10`).
-*Political knowledge* is measured with six items from the Dutch Parliamentary Election Studies for the Dutch sample, and the three items from the American National Election Studies.[\^]:The questionnaire can be found in OA **XX**.
+*Political knowledge* is measured with six items from the Dutch Parliamentary Election Studies for the Dutch sample, and the three items from the American National Election Studies.
+^[The questionnaire can be found in OA A-1].
 
@@
 
-## Method
+<!--## Method
 
@@
 Using the pooled data we will estimate a within groups fixed effects model.
-We have conducted a balance test based on demographics (age, gender, education, geographical region, level of urbanness, employment, and income), vote choice in the 2021 parliamentary elections, ideological self-placement, political knowledge, and positions on the issues, using the `cobalt` R package [@greifer2021].
+We have conducted a balance test based on demographics (age, gender, education, geographical region, level of urbanization, employment, and income), vote choice in the 2021 parliamentary elections, ideological self-placement, political knowledge, and positions on the issues, using the `cobalt` R package [@greifer2021].
 This balance test indicated that none of the variables are unbalanced over the experimental groups, and therefore, as pre-registered, will not be added to the regression formula.
@@
 ```
+-->
+
 # Results
@@
 To answer whether there is an ideological or knowledge-based annotation bias, we have conducted a two-by-two experiment.
-Tables \ref{tab:profilenl} and \ref{tab:profileus} demonstrate the average profile of respondents who annotate correctly and incorrectly (where respondents who annotated some stances correctly and some incorrectly are weighted by proportion (in)correct).
+To test our hypotheses, we will conduct a multilevel model, with respondents clustered in issues (for more details, see OA A-4).
+OA A-5 describes the average profile of the respondents who annotate correctly and incorrectly (where respondents who annotated some stances correctly and some incorrectly are weighted by proportion (in)correct).
+OA A-6 displays the effect of the experimental conditions on the four dependent variables -- correctly identifying a stance and over-interpreting a stance for both a strict and a lenient interpretation of stance.
+OA A-7 demonstrates the exploratory analyses that test the robustness of our results presented below.
+<!--Tables \ref{tab:profilenl} and \ref{tab:profileus} demonstrate the average profile of respondents who annotate correctly and incorrectly (where respondents who annotated some stances correctly and some incorrectly are weighted by proportion (in)correct).
 In terms of demographics, there is not much of a difference.
 Yet, people who are incorrectly identifying stances are more left-wing oriented compared to those who are correct -- i.e.
-an average score of 4 for those who are incorrect vs.
-an average score of 5 for those who are correct.
+an average score of four for those who are incorrect vs.
+an average score of five for those who are correct.
 For other positions on issues or political knowledge, we do not see a difference in averages between those who are correctly and incorrectly identifying stances.
@@
 
-Looking at the effect of the experimental conditions on the four dependent variables -- 1) correctly identifying a stance; and 2) over-interpreting a stance for both a strict and a lenient interpretation of stance -- Figure \ref{fig:results-base} visualizes the baseline.
+Looking at the effect of the experimental conditions on the four dependent variables -- correctly identifying a stance and over-interpreting a stance for both a strict and a lenient interpretation of stance -- Figure \ref{fig:results-base} visualizes the baseline.
 The left-hand panel demonstrates the effect of the two experimental conditions for correctly identifying the stance in the Dutch case.
@@
 Using the lenient interpretation, however, people seem less likely to annotate the sentence as a stance.
-In the pre-registered section, we demonstrate the tests of the hypotheses, and afterwards, we discuss some explorations of the data to show the robustness of our findings, the visualizations thereof are displayed in OA **XX**.
+In the pre-registered section, we demonstrate the tests of the hypotheses, and afterwards, we discuss some explorations of the data to show the robustness of our findings, the visualizations thereof are displayed in OA A-3.
 
@@
 ```
-
-## Pre-registered Results
+-->
 
@@
 
-## Exploratory Results
+<!--## Exploratory Results
 
@@
 We do not see much variation between issues *Tax*, *EU/Foreign Policy*, and *Environment*.
-For those issues, we see that almost everyone interprets the sentence correct (in blue and red).
+For those issues, we see that almost everyone interprets the sentence correctly (in blue and red).
 We also see that for a lenient interpretation of stances, people are quite likely to overinterpret a position as a stance.
@@
 That means there is a 15% increase in having it correct.
-The difference for overinterpreting is smaller between revealed and masked political actor -- shown in the right-hand panel of Figure **OA.XX** -- yet also statistically significant.
+The difference for overinterpreting is smaller between revealed and masked political actors -- shown in the right-hand panel of Figure **OA.XX** -- yet also statistically significant.
 Compared to `85%` overinterpreting the sentence as a stance, in the masking solution "only" `80%` over-interprets the sentence as a stance.
@@
 
-```{r results-exploration1, out.height = "85%",out.width = "95%",, fig.align = 'center', fig.cap = "\\label{fig:results-exp1} Exploration: Issue Speficific Analyses", fig.pos="H", fig.show="hold"}
+```{r results-exploration1, out.height = "85%",out.width = "95%",, fig.align = 'center', fig.cap = "\\label{fig:results-exp1} Exploration: Issue Specific Analyses", fig.pos="H", fig.show="hold"}
 knitr::include_graphics(c(here("report/figures", "issues-1.png")))
 ```
+-->
 
@@
 Especially in times where due to the democratization of data and the advances in NLP more data can be analyzed or classified, these standards are more important than ever.
-Human annotation is a crucial instrument in the toolkit of the social sciences.
-In order to safeguard the validity and reliability of annotation tasks, strict standards need to be upheld.
-Especially in these days, where due to the democratization of data and the advances in NLP more data can be analyzed or classified, these standards are more important than ever.
 After all, an algorithm trained on biased data will reproduce and often exacerbate bias [e.g. see @prost2019debiasing].
 It is key to realize that the underlying assumption to create valid and reliable annotated data is that there is only one correct interpretation for every input example, and that disagreement between the annotators is something that needs to be dealt with at all costs [@aroyo2015truth].
-In times of increasing factual opinion polarization [@rekker2022understanding; @lee2021more], it is crucial to ask *whose perspective is being recorded in these datasets*, given that there is no neutral data and no apolitical standpoint from where we can call out bias: "Datasets are always 'a worldview' and, as such, data always remains biased" [@miceli2022studying, p.5].
+In times of increasing factual opinion polarization [@rekker2022understanding; @lee2021more], it is crucial to ask *whose perspective is being recorded in these datasets*, given that there is no neutral data and no apolitical standpoint from which we can call out bias: "Datasets are always 'a worldview' and, as such, data always remains biased" [@miceli2022studying, p.5].
 Our tools to create valid and reliable annotation data does currently not allow for dealing with different perspectives.
@@
 Although we did not find a main effect of ideological distance (H1a), we did find evidence that the effect of ideological distance is greater for underspecified sentences (H1b) in the US, depending on how one operationalizes stances.
-The exploratory analyses showed that these findings where robust against different specifications of ideological distance.
+The exploratory analyses (see OA A-7) showed that these findings where robust against different specifications of ideological distance.
 This conditional finding is important, because machine translation is on the rise [@licht2023cross; @de2018no]: Social scientists use this technique for example to comparatively analyze political stances.
@@
 Overall, our findings did not support these hypotheses.
-With the more lenient interpretation of overinterpretation, we did find support for H2a for the Dutch context, but we found an opposite effect for the US.
-While this does not support the hypothesis in terms of the direction of the bias, it does imply that ideological distance can affect how stance is processed.
+Yet, they imply that ideological distance can affect how stance is processed.
+<!--With the more lenient interpretation of overinterpretation, we did find support for H2a for the Dutch context, but we found an opposite effect for the US.
+While this does not support the hypothesis in terms of the direction of the bias, it does imply that ideological distance can affect how stance is processed.-->
 Regarding political knowledge (H2b), our results showed again country differences as well as that the way one defines a stance matters.
-Our findings, for H2b in particular, pose important implications.
+<!--Our findings, for H2b in particular, pose important implications.-->
 For sentences with clear political stances on an issue (e.g., immigration should be made more difficult) annotators generally agreed on the stance that is expressed.
@@
 When a party talks about a political issue without taking an explicit stance, annotators need to fill in the gaps and make their own inferences about the party's stance on this issue.
-Considering that in many political documents and specifically reporting on political news statements are not made explicit, underspecified sentences can be assumed to rather be the rule than the exception.
+Considering that in many political texts, specifically reporting on political news, stances are not made explicit, underspecified sentences can be assumed to rather be the rule than the exception.
 Our results therefore imply that political knowledge is indeed a relevant dimension in shaping how an annotator perceives political stances.
@@
 Given that the US is one of the front runners in terms of polarization, it is important to realize that this affects our annotators, and therefore annotated data, too.
-Biases seems to be stronger and less easy to be resolved.
+Biases seem to be stronger and less easy to resolve.
 
@@
 However, this is only the case if a sufficiently large sample of annotators is randomly drawn from the population.
-In reality, gold standards are often the product of a handful of expert coders, that typically study or work at universities.
-Even if a good amount of crowd coders is used, it cannot blindly be assumed that this provides a diverse and balanced reflection of political ideologies in society, as we also observed in our samples (Tables \ref{tab:profilenl} and \ref{tab:profileus}).
+In reality, gold standards are often the product of a handful of expert coders, who typically study or work at universities.
+Even if a good amount of crowd coders is used, it cannot blindly be assumed that this provides a diverse and balanced reflection of political ideologies in society, as we also observed in our samples.
 Especially in times where an increasing amount of crowd coding answers are based on large language models such as ChatGPT which (by and large) simulate the same ideological position over and over again [@veselovsky2023artificial], reflecting on the composition and quality of annotators and how it influences gold standards becomes more than just a methodological exercise.
@@
 But if the study involves making inferences about how the audience interprets political stances, for example in media effects research, then a ground-truth definition of political stance is problematic.
-To make accurate inferences about how a citizen subjectively perceives political content, and consequently how this might have affected behavior such as voting, we need to learn more about how shared personal characteristics relate to shared modes of perspective.
+To make accurate inferences about how a citizen subjectively perceived political content, and consequently how this might have affected behavior such as voting, we need to learn more about how shared personal characteristics relate to shared modes of perspective.
 

### 2023-07-25 report/draft.Rmd
updates

@@
 anonymous: TRUE
-abstract: "The behavior of citizens and the study of them by social scientists rely heavily on key information.
-Human annotation represents a crucial instrument in the toolkit of the social sciences.
+abstract: "Information shapes citizens' political decision-making.
+This process is amply studied by social scientists, who have human annotation as a crucial instrument in their toolkit.
 To address concerns regarding the validity and reliability of annotation tasks, we establish strict standards.
@@
 # Introduction
-The behavior of citizens and the study thereof by social scientists both rely heavily on key information.
-Classical theories of political communication, such as agenda setting or framing [e.g.
+Information shapes citizens' political decision-making. 
+This process is amply studied by social scientists: Classical theories of political communication, such as agenda setting or framing [e.g.
 @vanaelst2016political; @lecheler2019news], formulate that political information drives opinion formation and participation in politics -- from voting to protests.
-The democratization of data and advent of computational social science has paved the way for new possibilities of gathering and analyzing textual data [for recent overviews, see @vanatteveldt2022computational; @grimmer2022text].
+To gather and analyze the information citizens receive, the democratization of data and advent of computational social science has paved the way for new possibilities [for recent overviews, see @vanatteveldt2022computational; @grimmer2022text].
 In particular, advances in Natural Language Processing (NLP) have made it possible to automatically analyze large quantities of data using machine learning [e.g., see @akyurek-etal-2020-multi; @piskorski2023semeval].
@@
 While this yields lots of experiences as to how to train coders to get reliable hand-coded data, the procedure is expensive, protracted, and sometimes does not even get us the quality of data we need [@weber18].
-While the crowd offers a solution to some of these issues [@van2021validity], the main underlying assumption that there is one correct interpretation for every input example remains untouched.
+While crowd-coding offers a solution to some of these issues [@van2021validity], the main underlying assumption that there is one correct interpretation for every input example remains untouched.
 However, especially for more complex coding tasks, research fields differ in their interpretations of what counts as a *correct* solution depending on the main goal of annotation.
-Do we want to create a clean benchmark data set that follows strict linguistic rules or one that reflects how statements are being interpreted by "real" readers that do not necessarily follow theoretical definitions on which elements are needed for a specified stance?
+Do we want to create a clean benchmark data set that follows strict linguistic rules?
+Or do we want one that reflects how statements are being interpreted by "real" readers?
 
@@
 Or is there an ideological difference in the ground truth?* These questions present a fundamental challenge to the main way of working when collecting gold standard data, as we operate from the baseline assumption that disagreement among the annotators should be avoided or reduced.
-Typically, when specific cases continuously cause disagreement, more instructions are added to limit interpretations (for recent innovations to improve annotation, see for example: @barbera2021automated; @struthers2020bridging; @winter2020online; @debell2013harder; @benoit2016crowd, @ying2022topics).
+Typically, when specific cases continuously cause disagreement, more instructions are added to limit interpretations (for recent innovations to improve annotation, see for example: @barbera2021automated; @struthers2020bridging; @winter2020online; @debell2013harder; @benoit2016crowd; @ying2022topics).
 However, work in computational linguistics has shown that increased annotation instructions do not increase quality [@parmar2023don].
@@
 <!--This literature puts forward that disagreement can occur because of differences in ideological position or political knowledge [@shen2021sounds; @alkiek2022classification; @joseph2021mis].-->
-We test the extent to which disagreement takes place, for what type of stances, as well as gives us directions on how to deal with the diversity in conceptions and the political heterogeneity.
+We test the extent to which disagreement takes place and for what type of stances.
+In addition, we provide guidance on how to deal with diversity in conceptions and political heterogeneity when creating gold standard data.
 <!-- that nowadays potentially occurs in our sample of annotators.-->
@@
 
-Our results demonstrate that overall sentences where with some knowledge on politics the stance might be inferred are more likely to be overinterpreted by the crowd, inferring a position where none is explicitly given.
+Overall, our results demonstrate that sentences where with some knowledge on politics the stance might be inferred are more likely to be overinterpreted, inferring a position where none is explicitly given.
 This is problematic as these sentences are very common in political text -- like legislative debates or speeches -- as well as media reports.

### 2023-07-26 report/draft.Rmd
add typo

@@
 - affiliation: Dep. of Communciation Science, Vrije Universiteit Amsterdam
-thanks: "Corresponding author: MACGvdV, Replication files are available on the author's Github account (https://github.com/MarikenvdVelden/bias-experiment); Author contributions: a) designed the study: MACGvdV, MR, WvA, AF, FL, & KW; b) conducted the study: MACGvdV, FL & MR; c) data cleaning & analysis: MACGvdV; d) writing of the paper: MACGvdV, MR, WvA, AF, FL, & KW"
+thanks: "Corresponding author: MACGvdV, Replication files are available on the author's Github account (https://github.com/MarikenvdVelden/bias-experiment); Author contributions: a) designed the study: MACGvdV, MR, AF, FL, WvA, & KW; b) conducted the study: MACGvdV, FL & MR; c) data cleaning & analysis: MACGvdV; d) writing of the paper: MACGvdV, MR, AF, FL, & KW"
 anonymous: TRUE

### 2024-05-23 report/draft.Rmd
update draft after SI comments

@@
 thanks: "Corresponding author: MACGvdV, Replication files are available on the author's Github account (https://github.com/MarikenvdVelden/bias-experiment); Author contributions: a) designed the study: MACGvdV, MR, AF, FL, WvA, & KW; b) conducted the study: MACGvdV, FL & MR; c) data cleaning & analysis: MACGvdV; d) writing of the paper: MACGvdV, MR, AF, FL, & KW"
-anonymous: TRUE
+anonymous: FALSE
 abstract: "Information shapes citizens' political decision-making.
@@
 It follows that the validity of these analyses hinges on the quality of these gold standards.
-If a gold standard contains biases, i.e., systematic errors, it can foster bias in any downstream analysis.
+Most data collection efforts to create gold standards assume that there is only one correct interpretation for every input example.
+Disagreement between the annotators is considered a bias -- i.e.
+a systematic deviation between the "true" value of a theoretical concept in a population and how it is observed by the coder -- and thereby something that needs to be dealt with at all costs [@aroyo2015truth].
+If a gold standard contains biases, it can foster bias in any downstream analysis.
 
-Most data collection efforts to create gold standards assume that there is only one correct interpretation for every input example, and that disagreement between the annotators is something that needs to be dealt with at all costs [@aroyo2015truth].
-A recent study by @van2021validity demonstrates that using crowd-coding platforms is a good way to collect such golden standards without too much disagreement.
-These platforms have also been used for experiments, and the quality of the respondents has been a center of attention [@coppock2019generalizing; @clifford2015samples; @huff2015these; @berinsky2014separating].
-While it has been argued that this data is of similar quality to data from a random sample of the population [@coppock2019generalizing], others have demonstrated that these online platforms are populated by people that are "unlike" the general public, being younger and holding more liberal values [@clifford2015samples; @huff2015these].
-If the latter is true, this might cause a problem for the coding of political texts: @ennser2018impact report that coders of political texts incorporate prior beliefs about parties' issue stances into their coding decisions.
+Crowd-coding platforms have been shown to be reliable platforms to collect a golden standard [@van2021validity].
+Yet, from experimental studies [@clifford2015samples; @huff2015these; @berinsky2014separating] we know that these online platforms are populated by people that are "unlike" the general public, being younger and holding more liberal values.
+This could mean that we introduce _sampling bias_ into the creation of the gold standard if we stricly consider only one position to be true.
+@ennser2018impact report that coders of political texts incorporate prior beliefs about parties' issue stances into their coding decisions.
 The authors find that party labels cue coders to a stance.
 For example, coders are more likely to report a left-wing party to be pro-immigration and a populist right-wing party to be against based on the exact same sentence.
-*Is this actually bias or a diversity of viewpoints? And how big of a problem does this bias/diversity of viewpoints generate for researchers relying on golden standard data?* 
+*Is this actually bias or a diversity of viewpoints?*
+*And how big of a problem is this bias/diversity of viewpoints?*
+
+There is a long history of text annotation in studies analyzing political text.
+Yet, the main underlying assumption that there is one correct interpretation for every input example remains untouched.
+However, especially for more complex coding tasks, research fields differ in their interpretations of what counts as a *correct* solution depending on the main goal of annotation.
+Do we want to create a clean benchmark data set that follows strict linguistic rules?
+Or do we want one that reflects how statements are being interpreted by "real" readers?
 This question also taps into newer developments related to data annotations. 
@@
 
-There is a long history of text annotation in studies analyzing political text.
-While this yields lots of experiences as to how to train coders to get reliable hand-coded data, the procedure is expensive, protracted, and sometimes does not even get us the quality of data we need [@weber18].
-While crowd-coding offers a solution to some of these issues [@van2021validity], the main underlying assumption that there is one correct interpretation for every input example remains untouched.
-However, especially for more complex coding tasks, research fields differ in their interpretations of what counts as a *correct* solution depending on the main goal of annotation.
-Do we want to create a clean benchmark data set that follows strict linguistic rules?
-Or do we want one that reflects how statements are being interpreted by "real" readers?
-
 Trends of polarization have shown that people do interpret information according to their ideological position [e.g., @rekker2022understanding; @lee2021more].
-*Does this mean that some are right and others are wrong?
-Or is there an ideological difference in the ground truth?* These questions present a fundamental challenge to the main way of working when collecting gold standard data, as we operate from the baseline assumption that disagreement among the annotators should be avoided or reduced.
+Does this mean that some are right and others are wrong? 
+Or is there an ideological difference in the ground truth?
+These questions present a fundamental challenge to the main way of working when collecting gold standard data, as we operate from the baseline assumption that disagreement among the annotators should be avoided or reduced.
 Typically, when specific cases continuously cause disagreement, more instructions are added to limit interpretations (for recent innovations to improve annotation, see for example: @barbera2021automated; @struthers2020bridging; @winter2020online; @debell2013harder; @benoit2016crowd; @ying2022topics).
@@
 We use two high-powered pre-registered experiments (see [here](link) and [here](link)) in the Netherlands -- a low-level polarized country -- and in the U.S. -- a high-level polarized country.
-We test possible biases stemming from ideology and political knowledge for correctly and overinterpreting stances, as well as whether masking he political actor to mitigates these biases.[^1]
+We test possible biases stemming from ideology and political knowledge for correctly and overinterpreting stances.
+Moreover, based on the results of @ennser2018impact, we test whether masking the political actor to mitigates these biases.^[The data and research compendium is published on the [[main author's github page]](X) -- ommitted for the review process.]
+Another offered solution by Webb-Williams et al.
+(2023) is to collect information on political positions of the annotators to potentially re-weight responses based on population proportions to create composite measures, which is the focus of our H1 and H2.
 <!--effect of ideological distance between the annotator and the political actor in the text (H1), the effect of overinterpretation based on political knowledge or ideological engagement (H2), and an offered solution of masking the political actor to mitigate the effects of ideology and knowledge (H3).-->
@@
 
-[^1]: The data and research compendium is published on the [[main author's github page]](https://github.com/MarikenvdVelden/bias-experiment) -- anonymized for the review process.
-
 Overall, our results demonstrate that sentences where with some knowledge on politics the stance might be inferred are more likely to be overinterpreted, inferring a position where none is explicitly given.
 This is problematic as these sentences are very common in political text -- like legislative debates or speeches -- as well as media reports.
-<!--Moreover, our results also demonstrate that for these disagreements in the crowd to occur, the context is important, as we find differences in disagreements between the American and Dutch context.-->
+Moreover, our results also demonstrate that for these disagreements in the crowd to occur, the context is important, as we find differences in disagreements between the American and Dutch context.
 Our findings underline the importance of taking disagreement seriously for the creation of gold standard text -- the bread-and-butter of all machine learning endeavors.

### 2024-05-23 src/analysis/robustness_nocenter.R
add additional robustness checks

@@ -1,0 +1,67 @@
+dl <- dl |> 
+  mutate(nostalgia = factor(nostalgia),
+         scapegoat = factor(scapegoat)) |> 
+  filter(E2_2 != 3)
+
+d_in <- dl |> 
+  filter(group_type == "In-group")
+d_out <- dl |> 
+  filter(group_type == "Out-group")
+h1a_e <- broom.mixed::tidy(lmer(sentiment ~  nostalgia + factor(urbanisation) +
+                                (1 | groups), data= d_in)) |> 
+  mutate(var = "Nostalgic Message",
+         hyp = "In-groups",
+         upper = (estimate + (std.error * 1.56)),
+         lower = (estimate - (std.error * 1.56))) |> 
+  filter(term=="nostalgia1") |> 
+  dplyr::select(var, hyp, estimate, upper, lower)
+h1b_e <- broom.mixed::tidy(lmer(sentiment ~  scapegoat + factor(urbanisation) +
+                                (1 | groups), data= d_in)) |> 
+  mutate(var = "Scapegoating Message",
+         hyp = "In-groups",
+         upper = (estimate + (std.error * 1.56)),
+         lower = (estimate - (std.error * 1.56))) |> 
+  filter(term=="scapegoat1") |> 
+  dplyr::select(var, hyp, estimate, upper, lower)
+
+h2a_e <- broom.mixed::tidy(lmer(sentiment ~  nostalgia + factor(urbanisation) +
+                                (1 | groups), data= d_out)) |> 
+  mutate(var = "Nostalgic Message",
+         hyp = "Out-groups",
+         upper = (estimate + (std.error * 1.56)),
+         lower = (estimate - (std.error * 1.56))) |> 
+  filter(term=="nostalgia1") |> 
+  dplyr::select(var, hyp, estimate, upper, lower)
+h2b_e <- broom.mixed::tidy(lmer(sentiment ~  scapegoat + factor(urbanisation) +
+                                (1 | groups), data= d_out)) |> 
+  mutate(var = "Scapegoating Message",
+         hyp = "Out-groups",
+         upper = (estimate + (std.error * 1.56)),
+         lower = (estimate - (std.error * 1.56))) |> 
+  filter(term=="scapegoat1") |> 
+  dplyr::select(var, hyp, estimate, upper, lower)
+
+p1_h1_e <- h1a_e |> 
+  add_case(h1b_e) |> 
+  add_case(h2a_e) |> 
+  add_case(h2b_e) |> 
+  mutate(var = factor(var,
[…]

### 2024-05-23 report/draft.Rmd
update draft after SI comments

@@
 - affiliation: Dep. of Communciation Science, Vrije Universiteit Amsterdam
+- name: Felicia Loecherbach
+- affiliation: Amsterdam School of Communication Research, University of Amsterdam
+- name: Kasper Welbers
+- affiliation: Dep. of Communciation Science, Vrije Universiteit Amsterdam
 - name: Myrthe Reuver
+- affiliation: Computational Linguistics & Text Mining Lab, Vrije Universiteit Amsterdam
+- name: Antse Fokkens
 - affiliation: Computational Linguistics & Text Mining Lab, Vrije Universiteit Amsterdam
@@
 - affiliation: Dep. of Communciation Science, Vrije Universiteit Amsterdam
-- name: Antse Fokkens
-- affiliation: Computational Linguistics & Text Mining Lab, Vrije Universiteit Amsterdam
-- name: Felicia Loecherbach
-- affiliation: Center for Social Media & Politics, New York University
-- name: Kasper Welbers
-- affiliation: Dep. of Communciation Science, Vrije Universiteit Amsterdam
+
 thanks: "Corresponding author: MACGvdV, Replication files are available on the author's Github account (https://github.com/MarikenvdVelden/bias-experiment); Author contributions: a) designed the study: MACGvdV, MR, AF, FL, WvA, & KW; b) conducted the study: MACGvdV, FL & MR; c) data cleaning & analysis: MACGvdV; d) writing of the paper: MACGvdV, MR, AF, FL, & KW"

### 2024-05-29 report/draft.Rmd
Changes for preparation full submission

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE,
-                      message=FALSE, warning=FALSE,
-                      fig.pos = "h", out.extra = "")
+knitr::opts_chunk$set(
+  echo = FALSE,
+  message = FALSE, warning = FALSE,
+  fig.pos = "h", out.extra = ""
+)
 options(kableExtra.latex.load_packages = FALSE)
@@
 
-Crowd-coding platforms have been shown to be reliable platforms to collect a golden standard [@van2021validity].
+Crowd-coding platforms have been shown to be reliable platforms to collect a gold standard [@van2021validity].
 Yet, from experimental studies [@clifford2015samples; @huff2015these; @berinsky2014separating] we know that these online platforms are populated by people that are "unlike" the general public, being younger and holding more liberal values.
@@
 
+In this context, bias would occur if there is one objectively correct solution to a coding task and coders systematically get the answer *wrong* due to prior beliefs.
+Opposed to this, diversity of viewspoints rather indicates that there are actually multiple ways of assessing the situation which all can be (in some way, depending on the siutation) correct.
 There is a long history of text annotation in studies analyzing political text.
@@
 Do we want to create a clean benchmark data set that follows strict linguistic rules?
-Or do we want one that reflects how statements are being interpreted by "real" readers?
+Or do we want one that reflects how statements are being interpreted by "real" readers, how the perception of different types of respondents differs for the same stimulus?
+This could potentially affect downstream tasks as manual stance annotations are further used for e.g.
+sentiment analysis or frame analysis.
+Are we modelling the sentiment and frames of a particular section of the population while the results would change with a different sample?
 This question also taps into newer developments related to data annotations. 
@@
 We should look beyond the majority vote and model it in the data, because if an algorithm is trained on biased data from disagreeing annotators, it will reproduce and often exacerbate that bias when it is applied to new data [e.g., see @prost2019debiasing].
+We thus rather opt for modelling a diversity of viewpoints instead of choosing for (potentially) biased solving of disagreements.
 To be able to model these annotators' characteristics, we should survey the characteristics of annotators when using the crowd (see Webb-Williams et al.
@@
 Yet, having only a few workers annotating the majority of texts has raised concerns about data diversity and models' ability to generalize beyond the crowd-workers.
-In a series of experiments, @geva2019we show that often models do not generalize well to annotations from annotators that did not contribute to the training set, suggesting that annotator bias should be monitored during data set creation.
+In a series of experiments, [@geva2019we] show that often models do not generalize well to annotations from annotators that did not contribute to the training set, suggesting that annotator bias should be monitored during data set creation.
 One such potential bias, especially in times of increasing polarization [@iyengar2019origins; @gidron2019toward; @boxell2022cross], is based on the ideological position of the annotator.
@@
 ```{r conditions, topcaption=TRUE}
-d <- tibble(`Condition` = c(rep("Specified", 4),
-                          rep("Underspecified", 4)), 
-            `US Experiment` = c("[Republicans/X] say immigration should be made more difficult.",
+d <- tibble(
+  `Condition` = c(
+    rep("Specified", 4),
+    rep("Underspecified", 4)
+  ),
+  `US Experiment` = c(
+    "[Republicans/X] say immigration should be made more difficult.",
     "[Democrats/X] say we need to put a tax on carbon emissions",
@@
     "[Democrats/X] say the tax system should be implemented differently.",
-                              "[Republicans/ X] say there should be a different military presence in the Pacific Ocean."),
-            `NL Experiment` = c("[PVV/X] says immigration should be made harder.",
+    "[Republicans/ X] say there should be a different military presence in the Pacific Ocean."
+  ),
+  `NL Experiment` = c(
+    "[PVV/X] says immigration should be made harder.",
     "[GreenLeft/X] says nitrogen emissions need to be reduced.",
@@
     "[Labour Party/X] says tax system must be changed.",
-                             "[Forum for Democracy/X] says the Netherlands should have a different role in the European Union."))
+    "[Forum for Democracy/X] says the Netherlands should have a different role in the European Union."
+  )
+)
 
-kbl(d, booktabs =T, caption = "\\label{tab:conditions}Survey Questions - Experimental Conditions") %>%
-  kable_styling(latex_options = c("striped", "hold_position"),
-                full_width = F, fixed_thead = T, position = "center") %>%
+kbl(d, booktabs = T, caption = "\\label{tab:conditions}Survey Questions - Experimental Conditions") %>%
+  kable_styling(
+    latex_options = c("striped", "hold_position"),
+    full_width = F, fixed_thead = T, position = "center"
+  ) %>%
   column_spec(1, width = "3cm") %>%
@@
 source(here("src/analysis/profile_resp.R"))
-kbl(e1_nlp, booktabs =T, caption = "\\label{tab:profilenl}Profile Dutch Stance Annotators") %>%
-  kable_styling(latex_options = c("striped", "hold_position"),
-                full_width = F, fixed_thead = T, position = "center") %>%
+kbl(e1_nlp, booktabs = T, caption = "\\label{tab:profilenl}Profile Dutch Stance Annotators") %>%
+  kable_styling(
+    latex_options = c("striped", "hold_position"),
+    full_width = F, fixed_thead = T, position = "center"
+  ) %>%
   column_spec(1, width = "7cm") %>%
@@
 ```{r "profileus"}
-kbl(e1_us_nlp, booktabs =T, caption = "\\label{tab:conditions_us}Profile American Stance Annotators") %>%
-  kable_styling(latex_options = c("striped", "hold_position"),
-                full_width = F, fixed_thead = T, position = "center") %>%
+kbl(e1_us_nlp, booktabs = T, caption = "\\label{tab:conditions_us}Profile American Stance Annotators") %>%
+  kable_styling(
+    latex_options = c("striped", "hold_position"),
+    full_width = F, fixed_thead = T, position = "center"
+  ) %>%
   column_spec(1, width = "7cm") %>%
@@
 It is key to realize that the underlying assumption to create valid and reliable annotated data is that there is only one correct interpretation for every input example, and that disagreement between the annotators is something that needs to be dealt with at all costs [@aroyo2015truth].
-In times of increasing factual opinion polarization [@rekker2022understanding; @lee2021more], it is crucial to ask *whose perspective is being recorded in these datasets*, given that there is no neutral data and no apolitical standpoint from which we can call out bias: "Datasets are always 'a worldview' and, as such, data always remains biased" [@miceli2022studying, p.5].
+In times of increasing factual opinion polarization [@rekker2022understanding; @lee2021more], it is crucial to ask *whose perspective is being recorded in these datasets*, given that there is no neutral data and no apolitical standpoint from which we can call out bias.
 Our tools to create valid and reliable annotation data does currently not allow for dealing with different perspectives.

### 2024-07-15 src/data-processing/clean_data_wva.R
Add WIP wva cleaning code

@@ -68,6 +68,7 @@ d_nl <- d_nl_raw |>
                      `7` = "Retired",                           
                      `8` = "Student",
                      .missing = "NA", .default = "NA"),
+
          PT1 = recode(PT1,
                       `1` = "VVD",
                       `2` = "D66",                        
@@ -257,7 +258,6 @@ d_us <- d_us_raw %>%
 # Combine NL and US data and compute attention
 
 
-
 # Code experimental conditions
 conditions = tribble(
   ~specified, ~gold, ~party_rile,~sentence, 
@@ -286,6 +286,7 @@ conditions = tribble(
 # Make a 'pivot_spec' to pivot this in one operation
 pivot_spec <- map(c("actor", "specified", "issue", "S"), 
                   ~tibble(.name=str_c(., 1:4), .value=., round=1:4)) |> list_rbind()
+
 
 d <- bind_rows(d_nl, d_us) |>
   mutate(across(Attention1_1:Attention1_9, ~recode(., .missing = 0, .default = 1)),

### 2024-07-15 src/data-processing/clean_data_wva.R
New version of data cleaning

@@ -1,0 +1,321 @@
+library(haven)
+library(tidyverse)
+library(here)
+library(qualtRics)
+
+# Load raw data
+
+read_file(here("data/raw-private/qualtrics_api_key.txt")) |>
+  trimws() |>
+  qualtrics_api_credentials("fra1.qualtrics.com")
+d_us_raw <- fetch_survey(surveyID = "SV_86XBcsFUJka9eke", 
+                         verbose = TRUE, force_request = T,
+                         label = FALSE, convert = FALSE)
+write_rds(d_us_raw, here("data/raw-private/d_us.rds"))
+
+d_nl_raw <- read_sav(here("data/raw-private/NLOZ2206_Mariken_Bias-experiment.sav"))
+d_us_raw <- read_rds(here("data/raw-private/d_us.rds"))
+
+
+# Clean and select NL survey columns
+
+
+d_nl <- d_nl_raw |>
+  zap_labels() %>% 
+  filter(age_cont>17) %>% 
+  select(id = ResponseId,
+         D1 = age_cont, #age - continues
+         D2 = sex_rec, #gender (2 categories)
+         D3 = edu_3_rec, #education 3 categories
+         D4 = geo_nielsen5_rec, #location (Nielen V)
+         D5, #income: no income (1), <500 (2), 501-1000 (3), 1001-1500 (4); 1501-2000 (5); 2001-2500 (6); 2501-3000 (7); 3001-3500 (8); 3501-4000 (9); 4001-4500 (10); 4501-7500 (11); >7501 (12)
+         D6, #job
+         PT1, #vote recall
+         PT2, #issue position: migration: 1 = neg, 5 = pos
+         PT3, #issue position: climate: 1 = neg, 5 = pos
+         PT4, #issue position: tax: 1 = neg, 5 = pos
+         PT5, #issue position: EU: 1 = neg, 5 = pos
+         PT6, #rile -self: 0 = left, 10 = right
+         PT7_1_3, PT7_1_4, PT7_1_5, PT7_1_16, #coalition partners (polknow)
+         PT7_2, #minister finance (polknow)
+         PT7_3, #2nd largest party (polknow)
+         PT7_4, #electing 1st chamber (polknow)
+         PT7_5, #who makes laws (polknow)
+         PT7_6, #coonstitutional court (polknow)
+         actor1:actor4,
+         issue1:issue4,
+         specified1:specified4,
+         S1:S4,
+         Attention1_1:Attention1_9) |>
+  mutate(D2 = recode(D2, `1` = "Male",
[…]

### 2024-07-16 src/analysis/exploratie.R
Exploratiory graph

@@ -1,0 +1,58 @@
+d |> group_by(country, round, specified, sentence, stance) |>
+  summarize(PT6=mean(PT6, na.rm=T)) |>
+  ggplot(aes(x=stance, y=sentence, label=round(PT6, 1), fill=PT6)) + geom_tile() + geom_text()
+
+
+
+d2 <- d |> 
+  mutate(mask=if_else(actor == "X", "masked", "notmasked")) |>
+  filter(!is.na(stance), stance != "dontknow") |>
+  group_by(country, round, specified, sentence, gold, mask, stance) |>
+  summarize(PT6=mean(PT6, na.rm=T), n=n()) |>
+  mutate(p=n/sum(n)) |>
+  mutate(sentence=glue::glue("[{country}{round} {gold}] {sentence}")) 
+  #pivot_wider(names_from="stance", values_from=PT6) |>
+ggplot(d2, aes(y=sentence, x=PT6)) + 
+  geom_point(aes(size=p, color=stance), alpha=.6) +
+  geom_text(data=filter(d2, p>.43), aes(label=round(p*100, 0)), size=3)+
+  facet_grid(rows=vars(specified), cols=vars(mask), scales = "free_y") + 
+  theme_minimal() + 
+  theme(legend.position="none") + 
+  xlab("Self-reported Left-Right")
+
+
+  geom_segment(aes(x=pro, xend=anti)) +
+  geom_point(aes(x=pro, size=n), color="blue") +
+  geom_point(aes(x=anti, size=n), color="red") +
+  geom_point(aes(x=neutral, size=n), color="orange") +
+  facet_grid(rows=vars(specified), scales = "free") + 
+  theme_minimal() + 
+  theme(legend.position="none") + 
+  xlab("Self-reported Left-Right")
+
+d |>
+  mutate(sentence=glue::glue("[{country}{round} {gold}] {sentence}")) |>
+  filter(!is.na(stance), stance != "dontknow") |>
+  ggplot() + 
+  geom_jitter(aes(x=PT6, y=sentence, color=stance), alpha=.25, width=.1) +
+  theme_minimal() 
+  
+
+  ggridges::geom_density_ridges(aes(x=PT6, y=sentence, fill=stance), alpha=.5, bandwidth=.15) +
+  facet_grid(rows=vars(specified),scales = "free") 
+  #facet_grid(rows=vars(specified), cols=vars(stance), scales = "free") 
+
+
+d |> 
[…]

### 2024-07-16 src/analysis/Analysis.Rmd
update data

@@
 ```{r "data for analyses"}
-load(here("data/intermediate/cleaned_data_nl.RData"))
-load(here("data/intermediate/cleaned_data_us.RData"))
+load(here("data/intermediate/cleaned_data_joined.RData"))
 ```
@@
 ```{r "balance-checks", fig.height=12, fig.width=8}
-source(here("src/analysis/balance-test.R"))
-df
+#source(here("src/analysis/balance-test.R"))
+#df
 ```
@@
 b
-rm(b_b, b1, b2, d, df, df_nl, df_us, tmp)
+rm(b_b, b1, b2, df, df_nl, df_us, tmp)
 ```
@@
 b_i
-rm(b1, b2, d, df, df_nl, df_us, tmp)
+rm(b1, b2, df, df_nl, df_us, tmp, issues_nl, issues_us, i)
 ```
@@
 p1b
-rm(df, df_nl, df_us, h1a, h1b, tmp)
+
+(p1b_1+p1b_2)/(p1b_3 + p1b_4) + 
+  plot_layout(guides = "collect") & theme(legend.position = 'bottom')
+rm(df, df_nl, df_us, h1a, h1b, p1b_1, 
+   pred1b_1, pred1b_2, pred1b_3, pred1b_4, tmp)
 ```
@@
 p2c2
-rm(df, df_nl, df_us, h2a, h2a_2, h2b, h2b_2, tmp, tmpp)
+
+(p2a_1+p2a_2)/(p2a_3 + p2a_4) + 
+  plot_layout(guides = "collect") & theme(legend.position = 'bottom')
+
+(p2b_1+p2b_2)/(p2b_3 + p2b_4) + 
+  plot_layout(guides = "collect") & theme(legend.position = 'bottom')
+rm(df, df_nl, df_us, h2a, h2a_2, h2b, h2b_2, tmp, tmpp,
+   pred2a_1, pred2a_2, pred2a_3, pred2a_4, 
+   pred2b_1, pred2b_2, pred2b_3, pred2b_4)
 ```
@@
 p3c2
-rm(df, df_nl, df_us, h3a, h3a_2, h3b, h3b_2,h3b_bs, h3b_2bs, tmp, tmpp)
+
+(p3a_1+p3a_2)/(p3a_3 + p3a_4) + 
+  plot_layout(guides = "collect") & theme(legend.position = 'bottom')
+
+(p3b_1+p3b_2)/(p3b_3 + p3b_4) + 
+  plot_layout(guides = "collect") & theme(legend.position = 'bottom')
+rm(df, df_nl, df_us, h3a, h3a_2, h3b, 
+   h3b_2,h3b_bs, h3b_2bs, tmp, tmpp,
+   pred3a_1, pred3a_2, pred3a_3, pred3a_4, 
+   pred3b_1, pred3b_2, pred3b_3, pred3b_4)
 ```
@@
 p_e1b
+```
 
+```{r "explor2", eval=FALSE}
 source(here("src/analysis/profile_resp.R"))

### 2024-07-16 src/data-processing/prepare-data.Rmd
update data

--- 2023-05-26 src/data-processing/prepare-data.Rmd
@@
 ```{r "get & clean data", results="hide"}
-#Clean Dutch Data
+#Load Dutch Data
 d_nl <- read_sav(here("data/raw-private/NLOZ2206_Mariken_Bias-experiment.sav"))
-source(here("src/data-processing/clean_data_nl.R"))
 
-#Clean US Data
+#Load US Data
 d_us <- fetch_survey(surveyID = "SV_86XBcsFUJka9eke", 
@@
                     label = FALSE, convert = FALSE)
-source(here("src/data-processing/clean_data_us.R"))
+#Clean data
+source(here("src/data-processing/clean_data.R"))
 ```
@@
 ```{r "save data"}
-save(d_nl, file = here("data/intermediate/cleaned_data_nl.RData"))
-save(d_us, file = here("data/intermediate/cleaned_data_us.RData"))
+save(d, file = here("data/intermediate/cleaned_data_joined.RData"))
 ```
@@
 ### Dependent Variable by Issue
-```{r "Dependent Variable", echo=F, fig.width=15, fig.height=8}
-p1 <- d_nl %>%
-  group_by(stance,issue,specification, masking) %>% 
+```{r "Dependent Variable", echo=F, fig.width=15, fig.height=10}
+p1a <- d |> 
+  filter(country == "NL") |> 
+  mutate(issue2 = recode(issue,
+                         `tax on carbon emissions` = "Climate",
+                         `a wealth tax for the richest Americans` = "Tax",
+                         `the tax system` = "Tax",
+                         `military presence in the Pacific Ocean` = "EU/Foreign Policy",
+                         `military build-up in the Pacific Ocean` = "EU/Foreign Policy",
+                         `immigranten` = "Immigration",
+                         `immigratie` = "Immigration",
+                         `het tegengaan van stikstofuitstoot` = "Climate",
+                         `het stikstofbeleid` = "Climate",
+                         `het verhogen van het belastingstarief voor de hoogste inkomens` = "Tax",
+                         `het belastingstelsel` = "Tax",
+                         `de  rol van Nederland in de Europese Unie` = "EU/Foreign Policy",
+                         `het lidmaatschap van de Europese Unie voor Nederland` = "EU/Foreign Policy")) |> 
+  group_by(stance_nlp,issue2,specified, masking, country) %>% 
   count() %>% 
-  mutate(stance = ifelse(stance==0, "Not Correctly identified stance",
-                         "Correctly identified stance"),
-         specification = ifelse(specification == "Specified", 
+  drop_na(stance_nlp) |> 
+  mutate(stance_nlp = recode(stance_nlp,
+                             `correct` = "Correct",
+                             `incorrect` = "Incorrect"),
+         specified = ifelse(specified == "Specified", 
                                 "Condition: Specified Sentence", "Underspecified Sentence"),
-         masking = ifelse(masking == "Masked", "Actor Masked", "Party Revealed")) %>% 
-  ggplot(aes(y = issue, x = n, fill = stance)) +
+         masking = ifelse(masking == 1, "Actor Masked", "Party Revealed")) %>% 
+  ggplot(aes(y = issue2, x = n, fill = stance_nlp)) +
   geom_col() +
-  facet_grid(masking~specification) +
+  facet_grid(masking~specified) +
   labs(x = "", y = "",
-       title = "Dutch Data") +
+       title = "NL Data - Strict Interpretation") +
   #scale_x_continuous(labels = scales::percent) +
@@
 
-p2 <- d_us %>%
-  group_by(stance,issue,specification, masking) %>% 
+p1b <- d |> 
+  filter(country == "NL") |> 
+  mutate(issue2 = recode(issue,
+                         `tax on carbon emissions` = "Climate",
+                         `a wealth tax for the richest Americans` = "Tax",
+                         `the tax system` = "Tax",
+                         `military presence in the Pacific Ocean` = "EU/Foreign Policy",
+                         `military build-up in the Pacific Ocean` = "EU/Foreign Policy",
+                         `immigranten` = "Immigration",
+                         `immigratie` = "Immigration",
+                         `het tegengaan van stikstofuitstoot` = "Climate",
+                         `het stikstofbeleid` = "Climate",
+                         `het verhogen van het belastingstarief voor de hoogste inkomens` = "Tax",
+                         `het belastingstelsel` = "Tax",
+                         `de  rol van Nederland in de Europese Unie` = "EU/Foreign Policy",
+                         `het lidmaatschap van de Europese Unie voor Nederland` = "EU/Foreign Policy")) |> 
+  group_by(stance_ss,issue2,specified, masking, country) %>% 
   count() %>% 
-  mutate(stance = ifelse(stance==0, "Not Correctly identified stance",
-                         "Correctly identified stance"),
-         specification = ifelse(specification == "Specified", 
+  drop_na(stance_ss) |> 
+  mutate(stance_ss = recode(stance_ss,
+                             `correct` = "Correct",
+                             `incorrect` = "Incorrect"),
+         specified = ifelse(specified == "Specified", 
                                 "Condition: Specified Sentence", "Underspecified Sentence"),
-         masking = ifelse(masking == "Masked", "Actor Masked", "Party Revealed")) %>% 
-  ggplot(aes(y = issue, x = n, fill = stance)) +
+         masking = ifelse(masking == 1, "Actor Masked", "Party Revealed")) %>% 
+  ggplot(aes(y = issue2, x = n, fill = stance_ss)) +
   geom_col() +
-  facet_grid(masking~specification) +
+  facet_grid(masking~specified) +
   labs(x = "", y = "",
-       title = "US Data") +
+       title = "NL Data - Lenient Interpretation") +
   #scale_x_continuous(labels = scales::percent) +
@@
 
-p1 + p2 + 
+p2a <- d |> 
+  filter(country == "US") |> 
+  mutate(issue2 = recode(issue,
+                         `tax on carbon emissions` = "Climate",
+                         `a wealth tax for the richest Americans` = "Tax",
+                         `the tax system` = "Tax",
+                         `military presence in the Pacific Ocean` = "EU/Foreign Policy",
+                         `military build-up in the Pacific Ocean` = "EU/Foreign Policy",
+                         `immigranten` = "Immigration",
+                         `immigratie` = "Immigration",
+                         `het tegengaan van stikstofuitstoot` = "Climate",
+                         `het stikstofbeleid` = "Climate",
+                         `het verhogen van het belastingstarief voor de hoogste inkomens` = "Tax",
+                         `het belastingstelsel` = "Tax",
+                         `de  rol van Nederland in de Europese Unie` = "EU/Foreign Policy",
+                         `het lidmaatschap van de Europese Unie voor Nederland` = "EU/Foreign Policy")) |> 
+  group_by(stance_nlp,issue2,specified, masking, country) %>% 
+  count() %>% 
+  drop_na(stance_nlp) |> 
+  mutate(stance_nlp = recode(stance_nlp,
+                             `correct` = "Correct",
+                             `incorrect` = "Incorrect"),
+         specified = ifelse(specified == "Specified", 
+                                "Condition: Specified Sentence", "Underspecified Sentence"),
+         masking = ifelse(masking == 1, "Actor Masked", "Party Revealed")) %>% 
+  ggplot(aes(y = issue2, x = n, fill = stance_nlp)) +
+  geom_col() +
+  facet_grid(masking~specified) +
+  labs(x = "", y = "", 
+       title = "US Data - Strict Interpretation") +
+  #scale_x_continuous(labels = scales::percent) +
+  theme_ipsum() +
+  scale_fill_manual(values = fig_cols[c(3, 2)]) +
+  theme(legend.position = "bottom",
+        legend.title = element_blank())
+
+p2b <- d |> 
+  filter(country == "US") |> 
+  mutate(issue2 = recode(issue,
+                         `tax on carbon emissions` = "Climate",
+                         `a wealth tax for the richest Americans` = "Tax",
+                         `the tax system` = "Tax",
+                         `military presence in the Pacific Ocean` = "EU/Foreign Policy",
+                         `military build-up in the Pacific Ocean` = "EU/Foreign Policy",
+                         `immigranten` = "Immigration",
+                         `immigratie` = "Immigration",
+                         `het tegengaan van stikstofuitstoot` = "Climate",
+                         `het stikstofbeleid` = "Climate",
+                         `het verhogen van het belastingstarief voor de hoogste inkomens` = "Tax",
+                         `het belastingstelsel` = "Tax",
+                         `de  rol van Nederland in de Europese Unie` = "EU/Foreign Policy",
+                         `het lidmaatschap van de Europese Unie voor Nederland` = "EU/Foreign Policy")) |> 
+  group_by(stance_ss,issue2,specified, masking, country) %>% 
+  count() %>% 
+  drop_na(stance_ss) |> 
+  mutate(stance_ss = recode(stance_ss,
+                             `correct` = "Correct",
+                             `incorrect` = "Incorrect"),
+         specified = ifelse(specified == "Specified", 
+                                "Condition: Specified Sentence", "Underspecified Sentence"),
+         masking = ifelse(masking == 1, "Actor Masked", "Party Revealed")) %>% 
+  ggplot(aes(y = issue2, x = n, fill = stance_ss)) +
+  geom_col() +
+  facet_grid(masking~specified) +
+  labs(x = "", y = "", 
+       title = "US Data - Lenient Interpretation") +
+  #scale_x_continuous(labels = scales::percent) +
+  theme_ipsum() +
+  scale_fill_manual(values = fig_cols[c(3, 2)]) +
+  theme(legend.position = "bottom",
+        legend.title = element_blank())
+
+(p2a + p2b)/ (p1a + p1b) +
   plot_layout(guides = "collect") & theme(legend.position = 'bottom')
-
 ```
@@
 ### Covariates
-```{r "Independent Variables", echo=F, message=FALSE, warning=FALSE,fig.width=10, fig.height=10}
-rm(p1,p2)
-p <- d_nl %>% select(distance, PT7, PT6) %>% 
+```{r "Independent Variables", echo=F, message=FALSE, warning=FALSE,fig.width=6, fig.height=6}
+rm(p1a, p1b, p2a, p2b)
+p <- d %>% 
+  filter(country=="NL") |> 
+  select(distance, PT7, PT6) %>% 
   pivot_longer(cols = everything(),
@@
                        `distance` = "Ideological Distance \n(> = More Distance)",
-                       `PT6` = "Ideological Self Placement \n(0 = Left, 10 = Right)",
-                       `PT7` = "Political Knowledge \n(0 = Low, 6 = High)"),
+                       `PT6` = "Ideological Self Placement \n(0 = Left/Liberal, 1 = Right/Conservative)",
+                       `PT7` = "Political Knowledge \n(0 = Low, 1 = High)"),
          country = "The Netherlands")
 
-p2 <- d_us %>% 
+p2 <-d %>% 
+  filter(country=="US") |> 
   select(distance, PT7, PT7b, PT6) %>% 
+  mutate(PT7b = PT7b / 5) |> 
   pivot_longer(cols = everything(),
                values_to = "score") %>% 
+  drop_na(score) |> 
   group_by(name) %>% 
@@
                        `distance` = "Ideological Distance \n(> = More Distance)",
-                       `PT6` = "Ideological Self Placement \n(0 = Extremely Liberal, 6 = Extremely Conservative)",
-                       `PT7` = "Political Knowledge \n(0 = Low, 3 = High)",
-                       `PT7b` = "Bullshit Receptivity \n(0 = Low, 5 = High)"),
+                       `PT6` = "Ideological Self Placement \n(0 = Left/Liberal, 1 = Right/Conservative)",
+                       `PT7` = "Political Knowledge \n(0 = Low, 1 = High)",
+                       `PT7b` = "Bullshit Receptivity \n(0 = Low, 1 = High)"),
          country = "United States")
@@
 ### Correlations Matrix
-```{r "Correlations Matrix", echo=F, message=F, warning=F, fig.width=15, fig.height=15}
-d_nl %>%
-  drop_na() %>% 
+```{r "Correlations Matrix", echo=F, message=F, warning=F, fig.width=8, fig.height=8}
+d %>%
+  filter(country == "NL") |> 
   select(`Age` = D1,
@@
          `Stance` = stance) %>%
+  drop_na() %>% 
   ggcorrmat(
@@
 
-d_us %>%
-  drop_na() %>% 
+d %>%
+  filter(country == "US") |> 
   select(`Age` = D1,
@@
          `PID`  = PT1,
-         `PID Strength` = PT1b,
+         #`PID Strength` = PT1b,
          `Position: Immigration`  = PT2,
@@
          `Position: Tax`  = PT4,
-         `Position: EU`  = PT5,
+         `Position: Foreign Policy`  = PT5,
          `Ideology`  = PT6,
@@
          `Stance` = stance) %>%
+    drop_na() %>% 
   ggcorrmat(

### 2024-07-16 src/analysis/balance-test.R

@@ -1,10 +1,12 @@
 ## NL Data
-covs <- d_nl %>%
-  mutate(treat = if_else(masking == "Masked" & specification == "Specified", 1,
-                 if_else(masking == "Masked" & specification == "Underspecified", 2,
-                 if_else(masking == "Party" & specification == "Specified", 3, 4)))) %>%
+covs <- d |> 
+  filter(country == "NL") |> 
+  mutate(treat = if_else(masking == 1 & specified == "Specified", 1,
+                 if_else(masking == 1 & specified == "Underspecified", 2,
+                 if_else(masking == 0 & specified == "Specified", 3, 4)))) %>%
   select(treat, D1:D6, PT1:PT7, attention) %>%
-  filter(PT1 != "NA", D6 != "NA")
+  filter(PT1 != "NA", D6 != "NA") |> 
+  drop_na()
 
 balanced <-bal.tab(treat ~ D1 + factor(D2) + factor(D3) + 
                      factor(D4) +D5 + factor(D6) + factor(PT1) + PT2 + 
@@ -58,17 +60,19 @@ df_nl <- balanced %>%
 rm(covs, balanced)
 
 ## US Data
-covs <- d_us %>%
-  mutate(treat = if_else(masking == "Masked" & specification == "Specified", 1,
-                         if_else(masking == "Masked" & specification == "Underspecified", 2,
-                                 if_else(masking == "Party" & specification == "Specified", 3, 4)))) %>%
+covs <- d |> 
+  filter(country == "US") |> 
+  mutate(treat = if_else(masking == 1 & specified == "Specified", 1,
+                 if_else(masking == 1 & specified == "Underspecified", 2,
+                 if_else(masking == 0 & specified == "Specified", 3, 4)))) %>%
   select(treat, D1:D6, PT1:PT7b, attention) %>%
   filter(PT1 != "NA", D6 != "NA", D2!= "NA",
-         D3 != "NA")
+         D3 != "NA") |> 
+  drop_na()
 
 balanced <-bal.tab(treat ~ D1 + factor(D2) + factor(D3) + 
                      factor(D4) +D5 + factor(D6) + factor(PT1) +
-                     factor(PT1b) + PT2 + PT3 + PT4 + PT5 + 
+                     PT2 + PT3 + PT4 + PT5 + 
                      PT6 + PT7 + PT7b + attention,
                    data = covs,
                    thresholds = c(m = 0.05))[[1]] 
[…]

### 2024-07-16 src/analysis/baseline.R

@@ -1,19 +1,19 @@
-df_us <- d_us %>% 
-  mutate(country = "United States") %>% 
-  select(stance_nlp, stance_ss, masking, specification, interpretation_nlp, interpretation_ss, country, issue)
-df_nl <- d_nl %>% 
-  mutate(country = "The Netherlands") %>% 
-  select(stance_nlp, stance_ss, masking, specification, interpretation_nlp, interpretation_ss, country, issue)
+#df_us <- d_us %>% 
+#  mutate(country = "United States") %>% 
+#  select(stance_nlp, stance_ss, masking, specification, interpretation_nlp, interpretation_ss, country, issue)
+#df_nl <- d_nl %>% 
+#  mutate(country = "The Netherlands") %>% 
+#  select(stance_nlp, stance_ss, masking, specification, interpretation_nlp, interpretation_ss, country, issue)
 
-d <- df_nl %>% 
-  add_case(df_us)
+#d <- df_nl %>% 
+#  add_case(df_us)
 
-b1 <- broom.mixed::tidy(lmer(stance_nlp ~ masking + specification + 
+b1 <- broom.mixed::tidy(lmer(stance_nlp2 ~ masking + specified + 
                          (country | issue), data= d)) %>%
   mutate(y = "Y: Correctly Interpreting Stance (Strict Interpretation)",
          type = "Pooled Analysis")
 
-tmp <- broom.mixed::tidy(lmer(stance_ss ~ masking + specification + 
+tmp <- broom.mixed::tidy(lmer(stance_ss2 ~ masking + specified + 
                                (country | issue), data= d)) %>%
   mutate(y = "Y: Correctly Interpreting Stance (Lenient Interpretation)",
          type = "Pooled Analysis")
@@ -21,15 +21,17 @@ tmp <- broom.mixed::tidy(lmer(stance_ss ~ masking + specification +
 b1 <- b1 %>% 
   add_case(tmp)
 
-tmp <- broom.mixed::tidy(lmer(stance_nlp ~ masking + specification + 
-                                (1 | issue), data= df_us)) %>%
+df_us <- d |> 
+  filter(country == "US")
+tmp <-  broom.mixed::tidy(lmer(stance_nlp2 ~ masking + specified + 
+                                (1 | issue), data = df_us)) %>%
   mutate(y = "Y: Correctly Interpreting Stance (Strict Interpretation)",
          type = "United States")
 
 b1 <- b1 %>% 
   add_case(tmp)
 
[…]

### 2024-07-16 src/analysis/exploration.R

@@ -1,14 +1,15 @@
-df_us <- d_us %>% 
-  select(stance_nlp, stance_ss, interpretation_nlp, interpretation_ss, masking, specification, distance, PT7, issue)
-df_nl <- d_nl %>% 
-  select(stance_nlp, stance_ss, interpretation_nlp, interpretation_ss, masking, specification, distance, PT7, issue)
+df_us <- d |> 
+  filter(country == "US") |> 
+  select(stance_nlp2, stance_ss2, interpret_nlp2, interpret_ss2, masking, specified, distance, PT7, issue)
+df_nl <- d |> 
+  filter(country == "NL") |> 
+  select(stance_nlp2, stance_ss2, interpret_nlp2, interpret_ss2, masking, specified, distance, PT7, issue)
 
-df <- d_us %>% 
-  mutate(b = specification,
-         a = recode(masking,
-                    `Masked` = 1,
-                    `Party` = 0))
-b3_a <- lmer(stance_nlp ~ b * a +
+df <- df_us %>% 
+  mutate(b = specified,
+         a = as.numeric(masking)) 
+
+b3_a <- lmer(stance_nlp2 ~ b * a +
                  (1 | issue), data= df)
 b3_a2 <- tidy(b3_a) %>% 
   mutate(condition = "Condition: Underspecified Sentence",
@@ -20,7 +21,7 @@ b3_a <- tidy(margins::margins(b3_a, variables = "b",
          y = "Y: Correctly Identifying Stance (Strict Interpretation)",
          type = "United States")
 
-tmp <- lmer(stance_ss ~ b * a + (1 | issue), data= df)
+tmp <- lmer(stance_ss2 ~ b * a + (1 | issue), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(condition = "Condition: Underspecified Sentence",
          y = "Y: Correctly Identifying Stance (Lenient Interpretation)",
@@ -36,12 +37,10 @@ tmp <- tidy(margins::margins(tmp, variables = "b",
 b3_a <- b3_a %>% 
   add_case(tmp)
 
-df <- df_nl %>% 
-  mutate(b = specification,
-         a = recode(masking,
-                    `Masked` = 1,
-                    `Party` = 0))
-tmp <- lmer(stance_nlp ~ b * a + (1 | issue), data= df)
+df <- df_us %>% 
+  mutate(b = specified,
[…]

### 2024-07-16 src/analysis/h1.R

@@ -1,9 +1,11 @@
-df_us <- d_us %>% 
-  select(stance_nlp, stance_ss, masking, specification, distance, PT7, issue)
-df_nl <- d_nl %>% 
-  select(stance_nlp, stance_ss, masking, specification, distance, PT7, issue)
+df_us <- d |>  
+  filter(country == "US") |> 
+  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
+df_nl <- d |> 
+  filter(country == "NL") |> 
+  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
 
-h1a <- broom.mixed::tidy(lmer(stance_nlp ~ masking + specification + 
+h1a <- broom.mixed::tidy(lmer(stance_nlp2 ~ masking + specified + 
                                 distance + PT7 +
                                 (1 | issue), data= df_us)) %>% 
   mutate(hyp = "H1a",
@@ -11,7 +13,7 @@ h1a <- broom.mixed::tidy(lmer(stance_nlp ~ masking + specification +
          y = "Interpreting Stance Correctly (Strict Interpretation)") %>% 
   filter(term %in% c("distance"))
 
-tmp <- broom.mixed::tidy(lmer(stance_ss ~ masking + specification + 
+tmp <- broom.mixed::tidy(lmer(stance_ss2 ~ masking + specified + 
                                 distance + PT7 +
                                 (1 | issue), data= df_us)) %>% 
   mutate(hyp = "H1a",
@@ -22,7 +24,7 @@ tmp <- broom.mixed::tidy(lmer(stance_ss ~ masking + specification +
 h1a <- h1a %>% 
   add_case(tmp)
 
-tmp <- broom.mixed::tidy(lmer(stance_nlp ~ masking + specification + 
+tmp <- broom.mixed::tidy(lmer(stance_nlp2 ~ masking + specified + 
                                 distance + PT7 +
                                 (1 | issue), data= df_nl)) %>% 
   mutate(hyp = "H1a",
@@ -33,7 +35,7 @@ tmp <- broom.mixed::tidy(lmer(stance_nlp ~ masking + specification +
 h1a <- h1a %>% 
   add_case(tmp)
 
-tmp <- broom.mixed::tidy(lmer(stance_ss ~ masking + specification + 
+tmp <- broom.mixed::tidy(lmer(stance_ss2 ~ masking + specified + 
[…]

### 2024-07-16 src/analysis/h2.R

@@ -1,26 +1,28 @@
-df_us <- d_us %>% 
-  select(stance_nlp, stance_ss, interpretation_nlp, interpretation_ss,
-         masking, specification, distance, PT7, PT7b, issue)
-df_nl <- d_nl %>% 
-  select(stance_nlp, stance_ss, interpretation_nlp, interpretation_ss,
-         masking, specification, distance, PT7, issue)
+df_us <- d |>  
+  filter(country == "US") |>
+  select(stance_nlp2, stance_ss2, interpret_nlp2, interpret_ss2,
+         masking, specified, distance, PT7, PT7b, issue)
+df_nl <- d |>  
+  filter(country == "NL") |>
+  select(stance_nlp2, stance_ss2, interpret_nlp2, interpret_ss2,
+         masking, specified, distance, PT7, issue)
 
 df <- df_us %>% 
-  mutate(b = specification,
+  mutate(b = specified,
          a = distance)
-h2a <- lmer(interpretation_nlp ~ masking + b * a +
+h2a <- lmer(interpret_nlp2 ~ masking + b * a +
               PT7 + (1 | issue), data= df)
 h2a_2 <- tidy(h2a) %>% 
   mutate(hyp = "H2a: Ideological Distance",
          type = "United States",
          y = "Overinterpreting Stance (Strict Interpretation)")
 h2a <- tidy(margins::margins(h2a, variables = "b",
-                             at = list("a" = 0:7))) %>% 
+                             at = list("a" = 0:1))) %>% 
   mutate(hyp = "H2a: Ideological Distance",
          type = "United States",
          y = "Overinterpreting Stance (Strict Interpretation)")
 
-tmp <- lmer(interpretation_ss ~ masking + b * a +
+tmp <- lmer(interpret_ss2 ~ masking + b * a +
               PT7 + (1 | issue), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(hyp = "H2a: Ideological Distance",
@@ -29,7 +31,7 @@ tmpp <- tidy(tmp) %>%
 h2a_2 <- h2a_2 %>% 
   add_case(tmpp) 
 tmp <- tidy(margins::margins(tmp, variables = "b",
-                             at = list("a" = 0:7))) %>%
+                             at = list("a" = 0:1))) %>%
   mutate(hyp = "H2a: Ideological Distance",
          type = "United States",
[…]

### 2024-07-16 src/analysis/h3.R

@@ -1,158 +1,148 @@
-df_us <- d_us %>% 
-  select(stance_nlp, stance_ss, masking, specification, distance, PT7, PT7b, issue)
-df_nl <- d_nl %>% 
-  select(stance_nlp, stance_ss, masking, specification, distance, PT7, issue)
+df_us <- d |>  
+  filter(country == "US") |>
+  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, PT7b, issue)
+df_nl <- d |>  
+  filter(country == "NL") |> 
+  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
 
 df <- df_us %>% 
-  mutate(b = recode(masking,
-                    `Masked` = 1,
-                    `Party` = 0),
-         b= factor(b),
+  mutate(b= factor(masking),
          a = distance)
-h3a <- lmer(stance_nlp ~ specification + b * a +
+h3a <- lmer(stance_nlp2 ~ specified + b * a +
               PT7 + (1 | issue), data= df)
 h3a_2 <- tidy(h3a) %>% 
   mutate(hyp = "H3a: Ideological Distance",
          type = "United States",
-         y = "Overinterpreting Stance (Strict Interpretation)")
+         y = "Correctly Interpreting Stance (Strict Interpretation)")
 h3a <- broom.mixed::tidy(margins::margins(h3a, variables = "b",
-                             at = list("a" = 0:7))) %>% 
+                             at = list("a" = 0:1))) %>% 
   mutate(hyp = "H3a: Ideological Distance",
          type = "United States",
-         y = "Overinterpreting Stance (Strict Interpretation)")
+         y = "Correctly Interpreting Stance (Strict Interpretation)")
 
-tmp <- lmer(stance_ss ~ specification + b * a +
+tmp <- lmer(stance_ss2 ~ specified + b * a +
               PT7 + (1 | issue), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(hyp = "H3a: Ideological Distance",
          type = "United States",
-         y = "Overinterpreting Stance (Lenient Interpretation)")
+         y = "Correctly Interpreting Stance (Lenient Interpretation)")
 h3a_2 <- h3a_2 %>% 
   add_case(tmpp)
 tmp <- tidy(margins::margins(tmp, variables = "b",
-                             at = list("a" = 0:7))) %>% 
+                             at = list("a" = 0:1))) %>% 
[…]

### 2024-07-16 src/analysis/issue_specific.R

@@ -1,22 +1,37 @@
-df_us <- d_us %>% 
-  mutate(country = "United States") %>% 
-  select(stance_nlp, stance_ss, masking, specification, interpretation_nlp, interpretation_ss, country, issue)
-df_nl <- d_nl %>% 
-  mutate(country = "The Netherlands") %>% 
-  select(stance_nlp, stance_ss, masking, specification, interpretation_nlp, interpretation_ss, country, issue)
+df_us <- d %>% 
+  filter(country == "US") %>% 
+  select(stance_nlp2, stance_ss2, masking, specified, interpret_nlp2, interpret_ss2, country, issue) |> 
+  mutate(issue2 = recode(issue,
+                         `tax on carbon emissions` = "Climate",
+                         `a wealth tax for the richest Americans` = "Tax",
+                         `the tax system` = "Tax",
+                         `military presence in the Pacific Ocean` = "EU/Foreign Policy",
+                         `military build-up in the Pacific Ocean` = "EU/Foreign Policy"))
+df_nl <- d %>% 
+  filter(country == "NL") %>% 
+  select(stance_nlp2, stance_ss2, masking, specified, interpret_nlp2, interpret_ss2, country, issue) |> 
+  mutate(issue2 = recode(issue,
+                         `immigranten` = "Immigration",
+                         `immigratie` = "Immigration",
+                         `het tegengaan van stikstofuitstoot` = "Climate",
+                         `het stikstofbeleid` = "Climate",
+                         `het verhogen van het belastingstarief voor de hoogste inkomens` = "Tax",
+                         `het belastingstelsel` = "Tax",
+                         `de  rol van Nederland in de Europese Unie` = "EU/Foreign Policy",
+                         `het lidmaatschap van de Europese Unie voor Nederland` = "EU/Foreign Policy"))
 
-issues_us <- unique(df_us$issue)
-issues_nl <- unique(df_nl$issue)
+issues_us <- unique(df_us$issue2)
+issues_nl <- unique(df_nl$issue2)
 
 for(i in 1:length(issues_us)){
   df <- df_us %>% filter(issue==issues_us[i])
   if(i==1){
-    b1 <- broom::tidy(lm(stance_nlp ~ masking + specification , data= df)) %>%
[…]

### 2024-07-16 src/data-processing/clean_data.R

@@ -68,7 +68,6 @@ d_nl <- d_nl |>
                      `7` = "Retired",                           
                      `8` = "Student",
                      .missing = "NA", .default = "NA"),
-
          PT1 = recode(PT1,
                       `1` = "VVD",
                       `2` = "D66",                        
@@ -259,6 +258,7 @@ d_us <- d_us  |>
 # Combine NL and US data and compute attention
 
 
+
 # Code experimental conditions
 conditions = tribble(
   ~specified, ~gold, ~party_rile,~sentence, 
@@ -287,7 +287,6 @@ conditions = tribble(
 # Make a 'pivot_spec' to pivot this in one operation
 pivot_spec <- map(c("actor", "specified", "issue", "S"), 
                   ~tibble(.name=str_c(., 1:4), .value=., round=1:4)) |> list_rbind()
-
 
 d <- bind_rows(d_nl, d_us) |>
   mutate(across(Attention1_1:Attention1_9, ~recode(., .missing = 0, .default = 1)),

### 2024-07-16 src/analysis/exploratie.R
Improve sentence display, faceting

@@ -2,57 +2,46 @@ d |> group_by(country, round, specified, sentence, stance) |>
   summarize(PT6=mean(PT6, na.rm=T)) |>
   ggplot(aes(x=stance, y=sentence, label=round(PT6, 1), fill=PT6)) + geom_tile() + geom_text()
 
+actors = d |> select(sentence, actor)  |> unique() |> filter(actor != "X")
 
 
 d2 <- d |> 
   mutate(mask=if_else(actor == "X", "masked", "notmasked")) |>
   filter(!is.na(stance), stance != "dontknow") |>
   group_by(country, round, specified, sentence, gold, mask, stance) |>
-  summarize(PT6=mean(PT6, na.rm=T), n=n()) |>
+  summarize(m=mean(PT6, na.rm=T), n=n(), sd=sd(PT6, na.rm=T)) |>
+  inner_join(actors) |>
   mutate(p=n/sum(n)) |>
-  mutate(sentence=glue::glue("[{country}{round} {gold}] {sentence}")) 
-  #pivot_wider(names_from="stance", values_from=PT6) |>
-ggplot(d2, aes(y=sentence, x=PT6)) + 
+  mutate(sentence=if_else(!str_detect(sentence, "Euro"), sentence,
+                          "zegt dat [de EU] slecht geweest voor Nederland is."
+                          )) |>
+  mutate(sentence=glue::glue("[{actor}/X] {sentence} ")) |>
+  mutate(sentence = fct_reorder(sentence, round))
+  
+
+ggplot(d2, aes(y=sentence, x=m)) + 
   geom_point(aes(size=p, color=stance), alpha=.6) +
-  geom_text(data=filter(d2, p>.43), aes(label=round(p*100, 0)), size=3)+
-  facet_grid(rows=vars(specified), cols=vars(mask), scales = "free_y") + 
+  geom_text(data=filter(d2, stance==gold), aes(label=round(p*100, 0)), size=3)+
+  ggh4x::facet_nested(specified + country ~ mask , scales="free_y")+ 
   theme_minimal() + 
-  theme(legend.position="none") + 
-  xlab("Self-reported Left-Right")
+  scale_size(guide="none") + 
+  theme(legend.position="bottom") + 
+  xlab("Self-reported Left-Right") + 
+  ylab("") +
+  theme(legend.title = element_blank(), 
+        strip.background = element_rect(color = "black", size = .5))
 
-
[…]

### 2024-07-19 src/analysis/Analysis.Rmd
update analyses

@@
 ```{r "balance-checks", fig.height=12, fig.width=8}
-#source(here("src/analysis/balance-test.R"))
-#df
+source(here("src/analysis/balance-test.R"))
+df
 ```
@@
 
+```{r "explore-profiles", fig.height=7, fig.width=8}
+source(here("src/analysis/profiles.R"))
+prof1 
+prof2
+```
+
 ```{r "explor2", eval=FALSE}

### 2024-07-19 src/analysis/profiles.R
update analyses

@@ -1,0 +1,56 @@
+d |> group_by(country, round, specified, sentence, stance) |>
+  summarize(PT6=mean(PT6, na.rm=T)) |>
+  ggplot(aes(x=stance, y=sentence, label=round(PT6, 1), fill=PT6)) + geom_tile() + geom_text()
+
+
+
+d2 <- d |> 
+  mutate(mask=if_else(actor == "X", "Masked", "Party")) |>
+  filter(!is.na(stance), stance != "dontknow") |>
+  group_by(country, round, specified, sentence, gold, mask, stance) |>
+  summarize(PT6=mean(PT6, na.rm=T), n=n()) |>
+  mutate(p=n/sum(n)) |>
+  mutate(sentence=glue::glue("[{country}{round} {gold}] {sentence}")) 
+#pivot_wider(names_from="stance", values_from=PT6) |>
+prof1 <- ggplot(d2, aes(y=sentence, x=PT6)) + 
+  geom_point(aes(size=p, color=stance), alpha=.6) +
+  geom_text(data=filter(d2, p>.43), aes(label=round(p*100, 0)), size=3)+
+  facet_grid(rows=vars(specified), cols=vars(mask), scales = "free_y") + 
+  theme_minimal() + 
+  theme(legend.position="none") + 
+  xlab("Self-reported Left-Right")
+
+
+#geom_segment(aes(x=pro, xend=anti)) +
+#  geom_point(aes(x=pro, size=n), color="blue") +
+#  geom_point(aes(x=anti, size=n), color="red") +
+#  geom_point(aes(x=neutral, size=n), color="orange") +
+#  facet_grid(rows=vars(specified), scales = "free") + 
+#  theme_minimal() + 
+#  theme(legend.position="none") + 
+#  xlab("Self-reported Left-Right")
+
+prof2 <- d |>
+  mutate(sentence=glue::glue("[{country}{round} {gold}] {sentence}")) |>
+  filter(!is.na(stance), stance != "dontknow") |>
+  ggplot() + 
+  geom_jitter(aes(x=PT6, y=sentence, color=stance), alpha=.25, width=.1) +
+  theme_minimal() 
+
+
+#ggridges::geom_density_ridges(aes(x=PT6, y=sentence, fill=stance), alpha=.5, bandwidth=.15) +
+#  facet_grid(rows=vars(specified),scales = "free") 
+#facet_grid(rows=vars(specified), cols=vars(stance), scales = "free") 
+
+
[…]

### 2024-07-24 report/draft.Rmd
updated manuscript

@@
 <!--Word count: `r as.integer(sub("(\\d+).+$", "\\1", system(sprintf("wc -w %s", knitr::current_input()), intern = TRUE))) - 20`-->
+
+\newpage
 
@@
 
+Before discussing the results of our preregistered analysis, we will describe which profiles, based on ideology (see Figure \ref{fig:profile1}) and political knowledge (see Figure \ref{fig:profile2}).
+OA A-5 gives even more detail; it describes the average profile of the respondents who annotate correctly and incorrectly (where respondents who annotated some stances correctly and some incorrectly are weighted by proportion (in)correct).
+Figure \ref{fig:profile1} demonstrates that in both countries people positioned on the ideological left of the spectrum are more likely to correctly code the specified sentences (upper-panel) to our gold standard.
+In general, we see that for the specified sentences, the coders where almost always in line with our gold standard.
+This picture is a bit different for the underspecified sentences. 
+In principle, all these sentences where defined so that they are linguistacally neutral (no stance).
+Yet, we see that all cases but the sentence "many immigrants are crossing the border" when the party is masked, a stance is interpreted. 
+Except for the sentences "there should be a different military presence in the Pacific Ocean" and "carbon emissions should be implemented differently" when the Democrats are shown, all sentences are interpreted as negative.
+What is interesting is that in the underspecified sentences, which are likely to also be in political information we have annotators for, there is quite a bit of variation on how one codes it depending on their ideological position.
+Looking at the profiles based on political knowledge, Figure \ref{fig:profile2} demonstrates that in the US, people with lower levels of political knowledge are more likely to be correct, whereas in the Netherlands, people with higher levels of political knowledge are more likely to be correct.
+This indicates that biases based on political knowledge can be context dependent.
+For the rest, Figure \ref{fig:profile2} paints a similar picture as \ref{fig:profile1}: Respondents where to high degrees correct based on the gold standard in the specified sentences, but interpreted the underspecified sentences overwhelmingly as a negative stance.
+
+```{r prof1, fig.align = 'center', fig.cap = "\\label{fig:profile1} Profile Ideological Distance", fig.pos="H", fig.show="hold",out.width = "90%"}
+knitr::include_graphics(here("report/figures", "explore-profiles-1.png"))
+```
+
+```{r prof2, fig.align = 'center', fig.cap = "\\label{fig:profile2} Profile Political Knowledge", fig.pos="H", fig.show="hold",out.width = "90%"}
+knitr::include_graphics(here("report/figures", "explore-profiles-2.png"))
+```
+
 To answer whether there is an ideological or knowledge-based annotation bias, we have conducted a two-by-two experiment.
 To test our hypotheses, we will conduct a multilevel model, with respondents clustered in issues (for more details, see OA A-4).
-OA A-5 describes the average profile of the respondents who annotate correctly and incorrectly (where respondents who annotated some stances correctly and some incorrectly are weighted by proportion (in)correct).
 OA A-6 displays the effect of the experimental conditions on the four dependent variables -- correctly identifying a stance and over-interpreting a stance for both a strict and a lenient interpretation of stance.
@@
 First, we test whether there is an ideological bias in interpreting stances (H1a), and if this bias increases for those that are further away from the ideological position of the political actor in the under-specified condition (H1b).
-Figure \ref{fig:results-h1} demonstrates on the left-hand panel the regression coefficients and on the right-hand panel the average marginal effects of the interaction between ideological distance and under-specified sentences.
+Figure \ref{fig:results-h1} demonstrates on the regression coefficients and the predicted effects.
 The upper-left panel of Figure \ref{fig:results-h1} demonstrates the coefficient of ideological distances for the likelihood of interpreting the stance correctly.
-There is a negligible positive effect -- a coefficient of `0.002` -- that is borderline significant.
+There is a negligible positive effect -- a coefficient of `0.002` -- that is borderline significant in the Netherlands, when looking at the strict interpretation of a stance.
 Substantially, this means that there is no effect of ideological distance for correctly interpreting the stance in either the American or Dutch case.
 Hence, no ideological bias found, thus no support for our H1a.
-Looking at the lower-left, and right-hand panel of Figure \ref{fig:results-h1}, we see a small but significant effect of the interaction between ideological distance and the under-specified condition for both interpretations in the American case, and only for the strict interpretation of a stance in the Dutch case.
-These effects are, however, going in the other direction than hypothesized.
-Those who are ideologically further away from the party are less likely to be wrong than those who are close to the party, as shown in the right-hand panel of Figure \ref{fig:results-h1}.
-For the Dutch case, using the strict interpretation, the difference is about `20%` -- from a coefficient of `-0.6` to `-0.4`.
-In the American case, the slope is even steeper, with a difference of about `40%`.
-We thus find evidence for H1b.
-This demonstrates that even in times of heightened polarization, ideological annotation bias can be a concern.
+Looking at the lower-left panel of Figure \ref{fig:results-h1}, we see a small but significant effect of the interaction between ideological distance and the under-specified condition for the strict interpretations in the Dutch case (hypothesized direction), and a small negative effect for the lenient interpretation of a stance in the US case (other direction as hypothesized).
+Those who are ideologically further away from the party are less likely to be wrong than those who are close to the party in the Netherlands, as shown in the right-hand panel of Figure \ref{fig:results-h1}.
+For the Dutch case, using the strict interpretation, the difference is about `10%` -- from a coefficient of `0.3` to `0.4`.
+In the American case, the slope for the lenient interpretation goes down as predicted, but not that much, with a difference of about `5%`.
+This could indicate that in times of heightened polarization, ideological annotation bias can be a concern.
+We thus find mixed evidence for H1b.
 
@@
 We measure this with an interaction between the experimental condition of specification and the variable of interest.
-Following @brambor2006understanding, we visualize the average marginal effects for both interactions to enhance interpretation in the right-hand panel of Figure \ref{fig:results-h2}.
-Figure \ref{fig:results-h2} demonstrates the average marginal effects for both interaction effects: In the upper-panel the results for H2a, and in the bottom-panel, the results for H2b.
-In both countries, we see that the different interpretations of stance have opposite effects: A strict interpretation increases the chance of overinterpreting, but this is not exacerbated by ideological distance or political knowledge.
-Using a more lenient interpretation, we see that this decreases the chance of overinterpreting.
+<!-- We visualize the predicted effects for both interactions to enhance interpretation in the right-hand panel of Figure \ref{fig:results-h2}.
+Figure \ref{fig:results-h2} demonstrates the predicted marginal effects for both interaction effects: In the upper-panel the results for H2a, and in the bottom-panel, the results for H2b.-->
+In both countries, we see that the different interpretations of stance have opposite effects: A strict interpretation decreases the chance of overinterpreting, and this is exacerbated by ideological distance but not by political knowledge.
+Using a more lenient interpretation, we see that this increases the chance of overinterpreting.
 While this is not further diminished by political knowledge, ideological distance does further diminish the change in the American case -- i.e. against expectation H2b.
@@
 ```{r results-h1, fig.align = 'center', fig.cap = "\\label{fig:results-h1} Ideological Distance", fig.pos="H", fig.show="hold",out.width = "49%", out.height="40%"}
-knitr::include_graphics(c(here("report/figures", "h1-1.png"), here("report/figures", "h1-2.png")))
+knitr::include_graphics(c(here("report/figures", "h1-1.png"), here("report/figures", "h1-3.png")))
 ```
 ```{r results-h2, fig.align = 'center', fig.cap = "\\label{fig:results-h2} Results Level of Sentence Specification", fig.pos="H", fig.show="hold",out.width = "49%", out.height="40%"}
-knitr::include_graphics(c(here("report/figures", "h2-1.png"), here("report/figures", "h2-2.png")))
+knitr::include_graphics(here("report/figures", "h2-1.png"))
 ```
@@
 We test these hypotheses with an interaction between the condition masking and the variables of interest and Figure \ref{fig:results-h3} demonstrates the regression coefficients.
-The upper-left panel of Figure \ref{fig:results-h3} shows, against expectation, a slight increase in the Dutch case: Those that are further away from the masked political actor are more likely to incorrectly interpret the stance.
-The decrease in slope is however negligible: From approximately `0.1%` to `0.2%` of the respondents being incorrect.
-There is no effect found in the American case.
+There is no effect found in the American or Dutch case regarding ideological distance.
 The same goes for the interaction between masking and political knowledge, on the bottom-panel of Figure \ref{fig:results-h3}.
-There is no significant effect found in the US, but a very small negative effect in the Dutch case.
+There is no significant effect found in the US nor int the Netherlands.
 This means that masking of political actors does not help to correctly interpret a sentence as a stance -- i.e. no support for H3a and H3b.

### 2024-07-24 src/analysis/Analysis.Rmd
last cleanups in analysis files

@@
 p_e1b
+p_e1c
 ```
 
-```{r "explore-profiles", fig.height=7, fig.width=8}
+```{r "explore-profiles", fig.height=10, fig.width=15}
 source(here("src/analysis/profiles.R"))
@@
 
-```{r "explor2", eval=FALSE}
+```{r "explor2", eval=FALSE, echo=FALSE}
 source(here("src/analysis/profile_resp.R"))

### 2024-07-24 src/analysis/exploration.R

@@ -477,7 +477,7 @@ h2b_e_pid <- tidy(margins::margins(h2b_e_pid, variables = "b",
                                at = list("a" = c("Forum voor Democratie","GroenLinks", "PvdA","PVV")))) %>% 
   mutate(hyp = "Party Voted For/ PID",
          condition = "Condition: Underspecified Sentence",
-         y = "Y: Overinterpreting Stance (Strict Interpretation",
+         y = "Y: Overinterpreting Stance (Strict Interpretation)",
          type = "The Netherlands")
 
 tmp <- lmer(interpret_ss2 ~ masking + b * a +
@@ -485,7 +485,7 @@ tmp <- lmer(interpret_ss2 ~ masking + b * a +
 tmpp <- tidy(tmp) %>% 
   mutate(hyp = "Party Voted For/ PID",
          condition = "Condition: Underspecified Sentence",
-         y = "Y: Overinterpreting Stance (Strict Interpretation)",
+         y = "Y: Overinterpreting Stance (Lenient Interpretation)",
          type = "The Netherlands")
 h2b_e_pid2 <- h2b_e_pid2 %>% 
   add_case(tmpp)
@@ -493,7 +493,7 @@ tmp <- tidy(margins::margins(tmp, variables = "b",
                                 at = list("a" = c("Forum voor Democratie","GroenLinks", "PvdA","PVV")))) %>% 
   mutate(hyp = "Party Voted For/ PID",
          condition = "Condition: Underspecified Sentence",
-         y = "Y: Overinterpreting Stance (Strict Interpretation",
+         y = "Y: Overinterpreting Stance (Lenient Interpretation)",
          type = "The Netherlands")
 h2b_e_pid <- h2b_e_pid %>% 
   add_case(tmp)
@@ -962,7 +962,7 @@ h2b_e_pid <- tidy(margins::margins(h2b_e_pid, variables = "b",
                                    at = list("a" = c("Democrat","Republican")))) %>% 
   mutate(hyp = "Party Voted For/ PID",
          condition = "Condition: Underspecified Sentence",
-         y = "Y: Overinterpreting Stance (Strict Interpretation",
+         y = "Y: Overinterpreting Stance (Strict Interpretation)",
          type = "United States")
 
 tmp <- lmer(interpret_ss2 ~ masking + b * a +
@@ -970,7 +970,7 @@ tmp <- lmer(interpret_ss2 ~ masking + b * a +
[…]

### 2024-07-24 src/analysis/issue_specific.R

@@ -6,7 +6,8 @@ df_us <- d %>%
                          `a wealth tax for the richest Americans` = "Tax",
                          `the tax system` = "Tax",
                          `military presence in the Pacific Ocean` = "EU/Foreign Policy",
-                         `military build-up in the Pacific Ocean` = "EU/Foreign Policy"))
+                         `military build-up in the Pacific Ocean` = "EU/Foreign Policy",
+                         `immigration` = "Immigration"))
 df_nl <- d %>% 
   filter(country == "NL") %>% 
   select(stance_nlp2, stance_ss2, masking, specified, interpret_nlp2, interpret_ss2, country, issue) |> 

### 2024-07-24 src/analysis/profiles.R

@@ -1,7 +1,6 @@
-d |> group_by(country, round, specified, sentence, stance) |>
-  summarize(PT6=mean(PT6, na.rm=T)) |>
-  ggplot(aes(x=stance, y=sentence, label=round(PT6, 1), fill=PT6)) + geom_tile() + geom_text()
-
+# |> group_by(country, round, specified, sentence, stance) |>
+#  summarize(PT6=mean(PT6, na.rm=T)) |>
+#  ggplot(aes(x=stance, y=sentence, label=round(PT6, 1), fill=PT6)) + geom_tile() + geom_text()
 
 
 d2 <- d |> 
@@ -16,26 +15,30 @@ prof1 <- ggplot(d2, aes(y=sentence, x=PT6)) +
   geom_point(aes(size=p, color=stance), alpha=.6) +
   geom_text(data=filter(d2, p>.43), aes(label=round(p*100, 0)), size=3)+
   facet_grid(rows=vars(specified), cols=vars(mask), scales = "free_y") + 
-  theme_minimal() + 
-  theme(legend.position="none") + 
-  xlab("Self-reported Left-Right")
+  theme_ipsum() + 
+  scale_colour_manual(values = fig_cols[c(2,1,3)]) +
+  theme(legend.position="none",
+        legend.title = element_blank()) + 
+  labs(y = "", x ="Ideological Position")
 
-
-#geom_segment(aes(x=pro, xend=anti)) +
-#  geom_point(aes(x=pro, size=n), color="blue") +
-#  geom_point(aes(x=anti, size=n), color="red") +
-#  geom_point(aes(x=neutral, size=n), color="orange") +
-#  facet_grid(rows=vars(specified), scales = "free") + 
-#  theme_minimal() + 
-#  theme(legend.position="none") + 
-#  xlab("Self-reported Left-Right")
-
-prof2 <- d |>
-  mutate(sentence=glue::glue("[{country}{round} {gold}] {sentence}")) |>
+d3 <- d |> 
+  mutate(mask=if_else(actor == "X", "Masked", "Party")) |>
   filter(!is.na(stance), stance != "dontknow") |>
-  ggplot() + 
-  geom_jitter(aes(x=PT6, y=sentence, color=stance), alpha=.25, width=.1) +
-  theme_minimal() 
+  group_by(country, round, specified, sentence, gold, mask, stance) |>
[…]

### 2024-07-24 report/draft.Rmd
Updated figures, small changes intro and abtract

@@
 abstract: "Information shapes citizens' political decision-making.
-This process is amply studied by social scientists, who have human annotation as a crucial instrument in their toolkit.
-To address concerns regarding the validity and reliability of annotation tasks, we establish strict standards.
-Due to the democratization of data and the advances in NLP more data can be analyzed or classified, making these standards are more important than ever: An algorithm trained on biased data will reproduce and often exacerbate bias.
-Our tools to create valid and reliable annotation data currently do not allow for dealing with different perspectives of annotators.
+This process is amply studied by social scientists, for whom human annotation is a crucial instrument in their toolkit.
+To address concerns regarding the validity and reliability of annotation tasks, benchmark data sets are created using rigourous annotation procedures.
+Due to the democratization of data and the advances in NLP more data can be analyzed or classified, making these benchmarks more important than ever: An algorithm trained on biased data will reproduce and often exacerbate bias.
+Current tools to create valid and reliable annotation data do not allow for dealing with different perspectives of annotators.
 In two pre-registered experiments in the United States and the Netherlands, we show that personal characteristics of annotators, like political ideology or knowledge, interfere with annotators' judgement of political stances.
@@
 For example, coders are more likely to report a left-wing party to be pro-immigration and a populist right-wing party to be against based on the exact same sentence.
-*Is this actually bias or a diversity of viewpoints?*
+This brings us to the main question addressed in this paper:
+*Is systematic disagreement actually bias or a diversity of viewpoints?*
 *And how big of a problem is this bias/diversity of viewpoints?*
@@
 In this context, bias would occur if there is one objectively correct solution to a coding task and coders systematically get the answer *wrong* due to prior beliefs. 
-Opposed to this, diversity of viewspoints rather indicates that there are actually multiple ways of assessing the situation which all can be (in some way, depending on the siutation) correct.
+Opposed to this, diversity of viewspoints rather indicates that there are actually multiple ways of assessing the situation which all can be (in some way, depending on the situation) correct. 
 There is a long history of text annotation in studies analyzing political text.
@@
 However, especially for more complex coding tasks, research fields differ in their interpretations of what counts as a *correct* solution depending on the main goal of annotation.
-Do we want to create a clean benchmark data set that follows strict linguistic rules?
+Do we want to create a clean benchmark data set that stays as close to the manifest meaning of the text as possible?
 Or do we want one that reflects how statements are being interpreted by "real" readers, how the perception of different types of respondents differs for the same stimulus? 
-This could potentially affect downstream tasks as manual stance annotations are further used for e.g.
-sentiment analysis or frame analysis.
-Are we modelling the sentiment and frames of a particular section of the population while the results would change with a different sample?
-This question also taps into newer developments related to data annotations. 
-Increasingly, Large Language Models such as ChatGPT are being used by researchers [@gilardi2023chatgpt] as well as annotators themselves [@veselovsky2023artificial] to "simulate" human respondents for annotation tasks.
-These models draw among other things from the gold standard data sets that have been created by researchers, perpetuating the specific biases that are present in them and not allowing for any further disagreement or diversity of viewpoints.
+This could potentially affect downstream tasks as manual stance annotations are further used for e.g. sentiment analysis or frame analysis. 
+<!-- Are we modelling the sentiment and frames of a particular section of the population while the results would change with a different sample?  -->
+<!-- This question also taps into newer developments related to data annotations.  -->
+<!-- Increasingly, Large Language Models such as ChatGPT are being used by researchers [@gilardi2023chatgpt] as well as annotators themselves [@veselovsky2023artificial] to "simulate" human respondents for annotation tasks.
+-->
+<!-- These models draw among other things from the gold standard data sets that have been created by researchers, perpetuating the specific biases that are present in them and not allowing for any further disagreement or diversity of viewpoints.
+-->
 
@@
 Typically, when specific cases continuously cause disagreement, more instructions are added to limit interpretations (for recent innovations to improve annotation, see for example: @barbera2021automated; @struthers2020bridging; @winter2020online; @debell2013harder; @benoit2016crowd; @ying2022topics).
-However, work in computational linguistics has shown that increased annotation instructions do not increase quality [@parmar2023don].
-This leaves us between a rock and a hard place.
+However, work in computational linguistics has shown that increased annotation instructions do not always increase quality [@parmar2023don].
+This leaves us between a rock and a hard place:
 *Is there a potential bias in annotators that we should account for?*
@@
 <!--effect of ideological distance between the annotator and the political actor in the text (H1), the effect of overinterpretation based on political knowledge or ideological engagement (H2), and an offered solution of masking the political actor to mitigate the effects of ideology and knowledge (H3).-->
-In the experiments, we vary the level of specification with which a political actor takes a stance -- a declarative sentence versus a sentence where with some knowledge on politics, the stance might be inferred -- as well as whether the political actor is shown or masked with putting `[ACTOR]` instead of the political party.
-We do this for four different political issues: *Environment*, *Immigration*, *Tax Policy*, and *EU* (for the Dutch case) or *Foreign Policy* (for the American case).
+In the experiments, we vary the level of specification with which a political actor takes a stance 
+<!-- -- a declarative sentence versus a sentence where with some knowledge on politics, the stance might be inferred --  -->
+as well as whether the political actor is shown or masked. 
+<!-- with putting `[ACTOR]` instead of the political party. -->
+<!-- We do this for four different political issues: *Environment*, *Immigration*, *Tax Policy*, and *EU* (for the Dutch case) or *Foreign Policy* (for the American case). -->
 This country selection not only allow us to showcase the scope conditions of disagreement due to different levels of political heterogeneity, but also differentiates between languages.

### 2024-07-24 src/analysis/profiles.R

@@ -1,23 +1,34 @@
 # |> group_by(country, round, specified, sentence, stance) |>
 #  summarize(PT6=mean(PT6, na.rm=T)) |>
 #  ggplot(aes(x=stance, y=sentence, label=round(PT6, 1), fill=PT6)) + geom_tile() + geom_text()
+library(tidyverse)
 
+actors <- d |> select(sentence, actor) |> filter(actor != "X") |> unique() |>
+  mutate(actor = if_else(actor == "Forum voor Democratie", "FvD", actor))
 
+load(here::here("data/intermediate/cleaned_data_joined.RData"))
+fig_cols = c("#0C5BB0CC", "#EE0011CC", "#15983DCC", "#EC579ACC", "#FA6B09CC", "#149BEDCC", "#A1C720CC", "#FEC10BCC", "#16A08CCC")
 d2 <- d |> 
   mutate(mask=if_else(actor == "X", "Masked", "Party")) |>
   filter(!is.na(stance), stance != "dontknow") |>
   group_by(country, round, specified, sentence, gold, mask, stance) |>
   summarize(PT6=mean(PT6, na.rm=T), n=n()) |>
   mutate(p=n/sum(n)) |>
-  mutate(sentence=glue::glue("[{country}{round} {gold}] {sentence}")) 
+  left_join(actors) |>
+  mutate(sentence=glue::glue("{actor}/X {sentence}")) |>
+  mutate(sentence=str_replace(sentence, "het lidmaatschap van de Europese Unie", "[de EU]"))
 #pivot_wider(names_from="stance", values_from=PT6) |>
-prof1 <- ggplot(d2, aes(y=sentence, x=PT6)) + 
+prof1 <- 
+  ggplot(d2, aes(y=sentence, x=PT6)) + 
   geom_point(aes(size=p, color=stance), alpha=.6) +
-  geom_text(data=filter(d2, p>.43), aes(label=round(p*100, 0)), size=3)+
-  facet_grid(rows=vars(specified), cols=vars(mask), scales = "free_y") + 
-  theme_ipsum() + 
+  geom_text(data=filter(d2, p>.43), aes(label=round(p*100, 0)), size=3.5)+
+  ggh4x::facet_nested(specified + country ~ mask, scales = "free_y") +
+  #facet_grid(rows=vars(specified), cols=vars(mask), scales = "free_y") + 
+  hrbrthemes::theme_ipsum(base_size = 16) + 
   scale_colour_manual(values = fig_cols[c(2,1,3)]) +
[…]

### 2025-09-03 src/analysis/Analysis.Rmd
Updated figures, small changes intro and abtract
update Compendium, final submission

@@
 ```{r "data for analyses"}
-load(here("data/intermediate/cleaned_data_joined.RData"))
+load(here::here("data/intermediate/cleaned_data_joined.RData"))
 ```
@@
 source(here("src/analysis/balance-test.R"))
-df
+df #Figure OA3
 ```
 
-## Baseline Model
-```{r "baseline", fig.align="center", fig.height=4, fig.width=8}
-source(here("src/analysis/baseline.R"))
-b
-rm(b_b, b1, b2, df, df_nl, df_us, tmp)
-```
-
-```{r "issues", fig.align="center", fig.height=16, fig.width=8}
-source(here("src/analysis/issue_specific.R"))
-b_i
-rm(b1, b2, df, df_nl, df_us, tmp, issues_nl, issues_us, i)
-```
-
-## Ideological Bias hypothesis
-```{r "h1", fig.align="center", fig.height=8, fig.width=8}
-source(here("src/analysis/h1.R"))
-p1a
-p1b
-
-(p1b_1+p1b_2)/(p1b_3 + p1b_4) + 
-  plot_layout(guides = "collect") & theme(legend.position = 'bottom')
-rm(df, df_nl, df_us, h1a, h1b, p1b_1, 
-   pred1b_1, pred1b_2, pred1b_3, pred1b_4, tmp)
-```
-
-## Overinterpretation Hypotheses
-```{r "h2", fig.align="center", fig.height=8, fig.width=8}
-source(here("src/analysis/h2.R"))
-p2a
-p2b
-p2c1
-p2c2
-
-(p2a_1+p2a_2)/(p2a_3 + p2a_4) + 
-  plot_layout(guides = "collect") & theme(legend.position = 'bottom')
-
-(p2b_1+p2b_2)/(p2b_3 + p2b_4) + 
-  plot_layout(guides = "collect") & theme(legend.position = 'bottom')
-rm(df, df_nl, df_us, h2a, h2a_2, h2b, h2b_2, tmp, tmpp,
-   pred2a_1, pred2a_2, pred2a_3, pred2a_4, 
-   pred2b_1, pred2b_2, pred2b_3, pred2b_4)
-```
-
-## Masking Hypothesis
-```{r "h3", fig.align="center", fig.height=8, fig.width=8}
-source(here("src/analysis/h3.R"))
-p3a
-p3b
-p3c1
-p3c2
-
-(p3a_1+p3a_2)/(p3a_3 + p3a_4) + 
-  plot_layout(guides = "collect") & theme(legend.position = 'bottom')
-
-(p3b_1+p3b_2)/(p3b_3 + p3b_4) + 
-  plot_layout(guides = "collect") & theme(legend.position = 'bottom')
-rm(df, df_nl, df_us, h3a, h3a_2, h3b, 
-   h3b_2,h3b_bs, h3b_2bs, tmp, tmpp,
-   pred3a_1, pred3a_2, pred3a_3, pred3a_4, 
-   pred3b_1, pred3b_2, pred3b_3, pred3b_4)
-```
-
-## Exploration
-```{r "exploration", fig.height=7, fig.width=8}
-source(here("src/analysis/exploration.R"))
-p_b1
-p_b2
-p_e1a
-p_e1b
-p_e1c
-```
-
+#Descriptives: Profiles
 ```{r "explore-profiles", fig.height=10, fig.width=15}
 source(here("src/analysis/profiles.R"))
-prof1 
-prof2
+prof1 #Figure 2
+prof2 #Figure 3
 ```
@@
 ```
+
+## Ideological Bias hypothesis (H1)
+```{r "h1", fig.align="center", fig.height=8, fig.width=8}
+source(here("src/analysis/h1.R"))
+p1a #Figure 4a
+p1b #Figure 4b
+
+(p1b_1+p1b_2)/(p1b_3 + p1b_4) + 
+  plot_layout(guides = "collect") & theme(legend.position = 'bottom') #Figure 4b
+rm(df, df_nl, df_us, h1a, h1b, p1b_1, 
+   pred1b_1, pred1b_2, pred1b_3, pred1b_4, tmp)
+```
+
+## Overinterpretation Hypotheses (H2)
+```{r "h2", fig.align="center", fig.height=8, fig.width=8}
+source(here("src/analysis/h2.R"))
+p2a #Figure 5
+p2b
+p2c1
+p2c2
+
+(p2a_1+p2a_2)/(p2a_3 + p2a_4) + 
+  plot_layout(guides = "collect") & theme(legend.position = 'bottom')
+
+(p2b_1+p2b_2)/(p2b_3 + p2b_4) + 
+  plot_layout(guides = "collect") & theme(legend.position = 'bottom')
+rm(df, df_nl, df_us, h2a, h2a_2, h2b, h2b_2, tmp, tmpp,
+   pred2a_1, pred2a_2, pred2a_3, pred2a_4, 
+   pred2b_1, pred2b_2, pred2b_3, pred2b_4)
+```
+
+## Masking Hypothesis (H3)
+```{r "h3", fig.align="center", fig.height=8, fig.width=8}
+source(here("src/analysis/h3.R"))
+p3a #Figure 6
+p3b
+p3c1
+p3c2
+
+(p3a_1+p3a_2)/(p3a_3 + p3a_4) + 
+  plot_layout(guides = "collect") & theme(legend.position = 'bottom')
+
+(p3b_1+p3b_2)/(p3b_3 + p3b_4) + 
+  plot_layout(guides = "collect") & theme(legend.position = 'bottom')
+rm(df, df_nl, df_us, h3a, h3a_2, h3b, 
+   h3b_2,h3b_bs, h3b_2bs, tmp, tmpp,
+   pred3a_1, pred3a_2, pred3a_3, pred3a_4, 
+   pred3b_1, pred3b_2, pred3b_3, pred3b_4)
+```
+
+## Exploration
+
+### Issues effect
+```{r "issues", fig.align="center", fig.height=16, fig.width=8}
+source(here("src/analysis/issue_specific.R"))
+b_i #Figure OA.4
+rm(b1, b2, df, df_nl, df_us, tmp, issues_nl, issues_us, i)
+```
+
+### Baseline Model
+```{r "baseline", fig.align="center", fig.height=4, fig.width=8}
+source(here("src/analysis/baseline.R"))
+b #Figure OA.5
+rm(b_b, b1, b2, df, df_nl, df_us, tmp)
+```
+
+### Different ways of measuring ideological distance
+```{r "exploration", fig.height=7, fig.width=8}
+source(here("src/analysis/exploration.R")) #Figures OA.6
+p_b1
+p_b2
+p_e1a
+p_e1b
+p_e1c
+```

### 2025-09-03 src/data-processing/prepare-data.Rmd
update Compendium, final submission

@@
 
-```{r "get & clean data", results="hide"}
+```{r "get & clean data", results="hide", eval=FALSE}
 #Load Dutch Data
@@
 ## Save Data for Analysis
-```{r "save data"}
+```{r "save data", eval=FALSE}
 save(d, file = here("data/intermediate/cleaned_data_joined.RData"))
+rm(d_nl, d_us)
 ```
@@
 ```{r "Dependent Variable", echo=F, fig.width=15, fig.height=10}
+load(here("data/intermediate/cleaned_data_joined.RData"))
 p1a <- d |> 
@@
 rm(p1a, p1b, p2a, p2b)
+
 p <- d %>% 

### 2025-09-03 src/analysis/profiles.R

@@ -1,7 +1,6 @@
 # |> group_by(country, round, specified, sentence, stance) |>
 #  summarize(PT6=mean(PT6, na.rm=T)) |>
 #  ggplot(aes(x=stance, y=sentence, label=round(PT6, 1), fill=PT6)) + geom_tile() + geom_text()
-library(tidyverse)
 
 actors <- d |> select(sentence, actor) |> filter(actor != "X") |> unique() |>
   mutate(actor = if_else(actor == "Forum voor Democratie", "FvD", actor))

### 2025-09-05 src/analysis/baseline.R
update Compendium, final submission

@@ -113,8 +113,8 @@ b <- b1 %>%
                        `masking1` = "Condition: Masked Political Actor",
                        `specifiedUnderspecified` = "Condition: Underspecified Sentence")) %>% 
   ggplot(aes(y = term, x = estimate,
-             xmin = estimate -1.56*std.error,
-             xmax = estimate +1.56*std.error,
+             xmin = estimate -1.645*std.error,
+             xmax = estimate +1.645*std.error,
              color = y)) +
   geom_point(position = position_dodge(.5)) +
   geom_errorbar(position = position_dodge(.5), width = 0) +
@@ -138,8 +138,8 @@ b_b <- b1 %>%
                        `masking1` = "Condition:Masked Political Actor",
                        `specifiedUnderspecified` = "Condition: Underspecified Sentence")) %>% 
   ggplot(aes(y = term, x = estimate,
-             xmin = estimate -1.56*std.error,
-             xmax = estimate +1.56*std.error,
+             xmin = estimate -1.645*std.error,
+             xmax = estimate +1.645*std.error,
              color = y)) +
   geom_point(position = position_dodge(.5)) +
   geom_errorbar(position = position_dodge(.5), width = 0) +

### 2025-09-05 src/analysis/exploration.R

@@ -147,8 +147,8 @@ p_b1 <- b3_a2 %>%
                              `bUnderspecified` = "Condition: Underspecified Sentence",
                          `bUnderspecified:a` = "Interaction: Masking * specified")) %>% 
   ggplot(aes(y = `term`, x = estimate,
-             xmin = estimate -1.56*std.error,
-             xmax = estimate +1.56*std.error,
+             xmin = estimate -1.645*std.error,
+             xmax = estimate +1.645*std.error,
              color = y)) +
   geom_point(position = position_dodge(.5)) +
   geom_errorbar(position = position_dodge(.5), width = 0) +
@@ -168,8 +168,8 @@ p_b2 <- b3_a %>%
                              `0` = "Political Actor Revealed",
                              `1` = "Political Actor Masked")) %>% 
   ggplot(aes(y = `at.value`, x = estimate,
-             xmin = estimate -1.56*std.error,
-             xmax = estimate +1.56*std.error,
+             xmin = estimate -1.645*std.error,
+             xmax = estimate +1.645*std.error,
              color = y)) +
   geom_point(position = position_dodge(.5)) +
   geom_errorbar(position = position_dodge(.5), width = 0) +

### 2025-09-05 src/analysis/h1.R

@@ -109,8 +109,8 @@ p1a <- h1a %>%
                                   "Ideological Distance from Party",
                                   "Condition: Underspecified Sentence",
                                   "Condition: Political Actor Revealed")), 
-         lower = estimate - 1.56 * std.error,
-         upper = estimate + 1.56 * std.error) %>% 
+         lower = estimate - 1.645 * std.error,
+         upper = estimate + 1.645 * std.error) %>% 
     ggplot(aes(x = estimate, y = term,
              xmin = lower, xmax = upper,
              color = y)) +

### 2025-09-05 src/analysis/h2.R

@@ -160,8 +160,8 @@ p2a <- h2a_2  %>%
                        "Interaction: Underspecified * Distance", term),
          term = ifelse(term == "bUnderspecified:a" & hyp != "H2a: Ideological Distance",
                        "Interaction: Underspecified * Knowledge", term), 
-         lower = estimate - 1.56 * std.error,
-         upper = estimate + 1.56 * std.error) %>% 
+         lower = estimate - 1.645 * std.error,
+         upper = estimate + 1.645 * std.error) %>% 
   ggplot(aes(x = estimate, y = term,
              xmin = lower, xmax = upper,
              color = y)) +

### 2025-09-05 src/analysis/h3.R

@@ -160,8 +160,8 @@ p3a <- h3a_2  %>%
                        "Interaction: Masking * Distance", term),
          term = ifelse(term == "b1:a" & hyp != "H3a: Ideological Distance",
                        "Interaction: Masking * Knowledge", term), 
-         lower = estimate - 1.56 * std.error,
-         upper = estimate + 1.56 * std.error) %>% 
+         lower = estimate - 1.645 * std.error,
+         upper = estimate + 1.645 * std.error) %>% 
   ggplot(aes(x = estimate, y = term,
              xmin = lower, xmax = upper,
              color = y)) +
@@ -180,8 +180,8 @@ p3a <- h3a_2  %>%
 p3b <- h3a  %>% 
   add_case(h3b) %>% 
   ggplot(aes(x = `at.value`, y = estimate,
-             ymin = (estimate - (1.65 * `std.error`)),
-             ymax = (estimate + (1.65 * `std.error`)),
+             ymin = (estimate - (1.645 * `std.error`)),
+             ymax = (estimate + (1.645 * `std.error`)),
              color = y, fill = y)) +
   geom_ribbon(alpha = 0.2, position = position_dodge(.2)) +
   geom_line(position = position_dodge(.2)) +
@@ -239,8 +239,8 @@ p3c1 <- h3b_2bs %>%
          term = ifelse(term == "b1", "Condition: Masked Actor", term),
          term = ifelse(term == "b1:a",
                        "Interaction: Masking * Bullshit Receptivity", term),
-         lower = estimate - 1.56 * std.error,
-         upper = estimate + 1.56 * std.error,
+         lower = estimate - 1.645 * std.error,
+         upper = estimate + 1.645 * std.error,
          hyp = factor(hyp,
                       levels = c("H3b: Political Knowledge",
                                  "Exploration: Bullshit Receptivity"))) %>% 

### 2025-09-05 src/analysis/issue_specific.R

@@ -132,8 +132,8 @@ b_i <- b1 %>%
          issue = ifelse(issue=="Foreign Policy" & type == "The Netherlands",
                         "EU", issue)) %>% 
   ggplot(aes(y = term, x = estimate,
-             xmin = estimate -1.56*std.error,
-             xmax = estimate +1.56*std.error,
+             xmin = estimate -1.645*std.error,
+             xmax = estimate +1.645*std.error,
              color = y)) +
   geom_point(position = position_dodge(.5)) +
   geom_errorbar(position = position_dodge(.5), width = 0) +

### 2025-09-09 src/analysis/baseline.R
update Compendium, final submission

@@ -8,13 +8,13 @@
 #d <- df_nl %>% 
 #  add_case(df_us)
 
-b1 <- broom.mixed::tidy(lmer(stance_nlp2 ~ masking + specified + 
-                         (country | issue), data= d)) %>%
+b1 <- broom.mixed::tidy(lmer(stance_nlp2 ~ masking + specified + issue +
+                         (country | sentetence), data= d)) %>%
   mutate(y = "Y: Correctly Interpreting Stance (Strict Interpretation)",
          type = "Pooled Analysis")
 
 tmp <- broom.mixed::tidy(lmer(stance_ss2 ~ masking + specified + 
-                               (country | issue), data= d)) %>%
+                               issue + (country | sentetence), data= d)) %>%
   mutate(y = "Y: Correctly Interpreting Stance (Lenient Interpretation)",
          type = "Pooled Analysis")
 
@@ -58,12 +58,12 @@ b1 <- b1 %>%
   add_case(tmp)
 
 b2 <- broom.mixed::tidy(lmer(interpret_nlp2 ~ masking + specified + 
-                               (country | issue), data= d)) %>%
+                               issue + (country | sentetence), data= d)) %>%
   mutate(y = "Y: Overinterpreting Stance (Strict Interpretation)",
          type = "Pooled Analysis")
 
 tmp <- broom.mixed::tidy(lmer(interpret_ss2 ~ masking + specified + 
-                                   (country | issue), data= d)) %>%
+                                   issue + (country | sentetence), data= d)) %>%
   mutate(y = "Y: Overinterpreting Stance (Lenient Interpretation)",
          type = "Pooled Analysis")
 

### 2025-09-09 src/analysis/exploration.R

@@ -10,7 +10,7 @@ df <- df_us %>%
          a = as.numeric(masking)) 
 
 b3_a <- lmer(stance_nlp2 ~ b * a +
-                 (1 | issue), data= df)
+                 issue + (1 | sentence), data= df)
 b3_a2 <- tidy(b3_a) %>% 
   mutate(condition = "Condition: Underspecified Sentence",
          y = "Y: Correctly Identifying Stance (Strict Interpretation)",
@@ -21,7 +21,7 @@ b3_a <- tidy(margins::margins(b3_a, variables = "b",
          y = "Y: Correctly Identifying Stance (Strict Interpretation)",
          type = "United States")
 
-tmp <- lmer(stance_ss2 ~ b * a + (1 | issue), data= df)
+tmp <- lmer(stance_ss2 ~ b * a + issue + (1 | sentence), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(condition = "Condition: Underspecified Sentence",
          y = "Y: Correctly Identifying Stance (Lenient Interpretation)",
@@ -40,7 +40,7 @@ b3_a <- b3_a %>%
 df <- df_us %>% 
   mutate(b = specified,
          a = as.numeric(masking))
-tmp <- lmer(stance_nlp2 ~ b * a + (1 | issue), data= df)
+tmp <- lmer(stance_nlp2 ~ b * a + issue + (1 | sentence), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(condition = "Condition: Underspecified Sentence",
          y = "Y: Correctly Identifying Stance (Strict Interpretation)",
@@ -56,7 +56,7 @@ tmp <- tidy(margins::margins(tmp, variables = "b",
 b3_a <- b3_a %>% 
   add_case(tmp)
 
-tmp <- lmer(stance_ss2 ~ b * a + (1 | issue), data= df)
+tmp <- lmer(stance_ss2 ~ b * a + issue + (1 | sentence), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(condition = "Condition: Underspecified Sentence",
          y = "Y: Correctly Identifying Stance (Lenient Interpretation)",
@@ -77,7 +77,7 @@ df <- df_us %>%
   mutate(b = specified,
          a = as.numeric(masking))
 b3_b <- lmer(interpret_nlp2 ~ b * a +
-               (1 | issue), data= df)
+               issue + (1 | sentence), data= df)
[…]

### 2025-09-09 src/analysis/h1.R

@@ -6,16 +6,16 @@ df_nl <- d |>
   select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
 
 h1a <- broom.mixed::tidy(lmer(stance_nlp2 ~ masking + specified + 
-                                distance + PT7 +
-                                (1 | issue), data= df_us)) %>% 
+                                distance + PT7 + issue +
+                                (1 | sentence), data= df_us)) %>% 
   mutate(hyp = "H1a",
          type = "United States",
          y = "Interpreting Stance Correctly (Strict Interpretation)") %>% 
   filter(term %in% c("distance"))
 
 tmp <- broom.mixed::tidy(lmer(stance_ss2 ~ masking + specified + 
-                                distance + PT7 +
-                                (1 | issue), data= df_us)) %>% 
+                                distance + PT7 + issue +
+                                (1 | sentence), data= df_us)) %>% 
   mutate(hyp = "H1a",
          type = "United States",
          y = "Interpreting Stance Correctly (Lenient Interpretation)") %>% 
@@ -25,8 +25,8 @@ h1a <- h1a %>%
   add_case(tmp)
 
 tmp <- broom.mixed::tidy(lmer(stance_nlp2 ~ masking + specified + 
-                                distance + PT7 +
-                                (1 | issue), data= df_nl)) %>% 
+                                distance + PT7 + issue +
+                                (1 | sentence), data= df_nl)) %>% 
   mutate(hyp = "H1a",
          type = "The Netherlands",
          y = "Interpreting Stance Correctly (Strict Interpretation)") %>% 
@@ -37,7 +37,7 @@ h1a <- h1a %>%
 
 tmp <- broom.mixed::tidy(lmer(stance_ss2 ~ masking + specified + 
                                 distance + PT7 +
-                                (1 | issue), data= df_nl)) %>% 
+                                issue +  (1 | sentence), data= df_nl)) %>% 
   mutate(hyp = "H1a",
          type = "The Netherlands",
          y = "Interpreting Stance Correctly (Lenient Interpretation)") %>% 
@@ -49,7 +49,7 @@ rm(tmp)
 
 h1b <- broom.mixed::tidy(lmer(stance_nlp2 ~ masking + 
                                 specified * distance +
[…]

### 2025-09-09 src/analysis/h2.R

@@ -11,7 +11,7 @@ df <- df_us %>%
   mutate(b = specified,
          a = distance)
 h2a <- lmer(interpret_nlp2 ~ masking + b * a +
-              PT7 + (1 | issue), data= df)
+              PT7 + issue + (1 | sentence), data= df)
 h2a_2 <- tidy(h2a) %>% 
   mutate(hyp = "H2a: Ideological Distance",
          type = "United States",
@@ -23,7 +23,7 @@ h2a <- tidy(margins::margins(h2a, variables = "b",
          y = "Overinterpreting Stance (Strict Interpretation)")
 
 tmp <- lmer(interpret_ss2 ~ masking + b * a +
-              PT7 + (1 | issue), data= df)
+              PT7 + issue + (1 | sentence), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(hyp = "H2a: Ideological Distance",
          type = "United States",
@@ -44,7 +44,7 @@ df <- df_nl %>%
          a = distance)
 
 tmp <- lmer(interpret_nlp2 ~ masking + b * a +
-              PT7 + (1 | issue), data= df)
+              PT7 + issue + (1 | sentence), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(hyp = "H2a: Ideological Distance",
          type = "The Netherlands",
@@ -61,7 +61,7 @@ h2a <- h2a %>%
   add_case(tmp)
 
 tmp <- lmer(interpret_ss2 ~ masking + b * a +
-              PT7 + (1 | issue), data= df)
+              PT7 + issue + (1 | sentence), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(hyp = "H2a: Ideological Distance",
          type = "The Netherlands",
@@ -81,7 +81,7 @@ df <- df_us %>%
   mutate(b = specified,
          a = PT7)
 h2b <- lmer(interpret_nlp2 ~ masking + b * a +
-              distance + (1 | issue), data= df)
+              distance + issue + (1 | sentence), data= df)
 h2b_2 <- tidy(h2b) %>% 
   mutate(hyp = "H2b: Political Knowledge",
          type = "United States",
@@ -93,7 +93,7 @@ h2b <- tidy(margins::margins(h2b, variables = "b",
          y = "Overinterpreting Stance (Strict Interpretation)")
 
 tmp <- lmer(interpret_ss2 ~ masking + b * a +
-              distance + (1 | issue), data= df)
[…]

### 2025-09-09 src/analysis/h3.R

@@ -9,7 +9,7 @@ df <- df_us %>%
   mutate(b= factor(masking),
          a = distance)
 h3a <- lmer(stance_nlp2 ~ specified + b * a +
-              PT7 + (1 | issue), data= df)
+              PT7 + issue + (1 | sentence), data= df)
 h3a_2 <- tidy(h3a) %>% 
   mutate(hyp = "H3a: Ideological Distance",
          type = "United States",
@@ -21,7 +21,7 @@ h3a <- broom.mixed::tidy(margins::margins(h3a, variables = "b",
          y = "Correctly Interpreting Stance (Strict Interpretation)")
 
 tmp <- lmer(stance_ss2 ~ specified + b * a +
-              PT7 + (1 | issue), data= df)
+              PT7 + issue + (1 | sentence), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(hyp = "H3a: Ideological Distance",
          type = "United States",
@@ -42,7 +42,7 @@ df <- df_nl %>%
          a = distance)
 
 tmp <- lmer(stance_nlp2 ~ specified + b * a +
-              PT7 + (1 | issue), data= df)
+              PT7 + issue + (1 | sentence), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(hyp = "H3a: Ideological Distance",
          type = "The Netherlands",
@@ -60,7 +60,7 @@ h3a <- h3a %>%
 
 
 tmp <- lmer(stance_ss2 ~ specified + b * a +
-              PT7 + (1 | issue), data= df)
+              PT7 + issue + (1 | sentence), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(hyp = "H3a: Ideological Distance",
          type = "The Netherlands",
@@ -81,7 +81,7 @@ df <- df_us %>%
   mutate(b= factor(masking),
          a = PT7)
 h3b <- lmer(stance_nlp2 ~ specified + b * a +
-              distance + (1 | issue), data= df)
+              distance + issue + (1 | sentence), data= df)
 h3b_2 <- tidy(h3b) %>% 
   mutate(hyp = "H3b: Political Knowledge",
          type = "United States",
@@ -93,7 +93,7 @@ h3b <- tidy(margins::margins(h3b, variables = "b",
          y = "Correctly Interpreting Stance (Strict Interpretation)")
 
 tmp <- lmer(stance_ss2 ~ specified + b * a +
-              distance + (1 | issue), data= df)
+              distance + issue + (1 | sentence), data= df)
[…]

### 2025-09-09 check.R
Changed filter to include sentence, unified issues
and corrected typo in "baseline.R"
Checks are documented in check.R.
Analysis.Rmd technically reproduces results. Not checked all substantive results in detail.

@@ -1,0 +1,66 @@
+# Data and functions
+source(here::here("src/lib/functions.R")) # edited to remove qualtrics stuff
+load(here::here("data/intermediate/cleaned_data_joined.RData"))
+
+# From h1.R
+# For US
+df_us <- d |>  
+  filter(country == "US") |> 
+  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
+df_us |> 
+  names() # does not include "sentence" 
+
+# Get complete data for US
+d |>  
+  filter(country == "US") |> 
+  distinct(issue) # Duplicated issues
+
+# Fix duplicated issues 
+df_us2 <- d |>  
+  filter(country == "US") |> 
+  mutate(issue2 = case_when(
+    str_detect(issue, "carbon") ~ "carbon",
+    str_detect(issue, "tax") ~ "tax",
+    str_detect(issue, "military") ~ "military",
+    str_detect(issue, "immigration") ~ "immigration"
+  ))
+df_us2 |> 
+  distinct(issue2, issue) |>
+  arrange(issue2) # check: 4 unique issues
+
+# Fit model from H1.R
+lmer(stance_nlp2 ~ masking + specified + 
+       distance + PT7 + issue2 +
+       (1 | sentence), data=df_us2) |> 
+  summary(correlation = FALSE) # seems fine
+
+# For NL
+df_nl <- d |> 
+  filter(country == "NL") |> 
+  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
+df_nl |> 
+  names() # does not include "sentence" 
+
+# Get complete data for NL
+d |>  
+  filter(country == "NL") |> 
+  distinct(issue) # Duplicated issues
+
+# Fix duplicated issues 
+df_nl2 <- d |>  
+  filter(country == "NL") |> 
+  mutate(issue2 = case_when(
+    str_detect(issue, "immigra") ~ "immigratie",
+    str_detect(issue, "stikstof") ~ "stikstof",
+    str_detect(issue, "belasting") ~ "belasting",
+    str_detect(issue, "Europese Unie") ~ "EU"
+  ))
+df_nl2 |> 
+  distinct(issue2, issue) |>
+  arrange(issue2) # check: 4 unique issues
+
+# Fit model from H1.R
+lmer(stance_nlp2 ~ masking + specified + 
+       distance + PT7 + issue2 +
+       (1 | sentence), data=df_nl2) |> 
+  summary(correlation = FALSE) # seems fine

### 2025-09-09 src/analysis/baseline.R

@@ -9,12 +9,12 @@
 #  add_case(df_us)
 
 b1 <- broom.mixed::tidy(lmer(stance_nlp2 ~ masking + specified + issue +
-                         (country | sentetence), data= d)) %>%
+                         (country | sentence), data= d)) %>%
   mutate(y = "Y: Correctly Interpreting Stance (Strict Interpretation)",
          type = "Pooled Analysis")
 
 tmp <- broom.mixed::tidy(lmer(stance_ss2 ~ masking + specified + 
-                               issue + (country | sentetence), data= d)) %>%
+                               issue + (country | sentence), data= d)) %>%
   mutate(y = "Y: Correctly Interpreting Stance (Lenient Interpretation)",
          type = "Pooled Analysis")
 
@@ -58,12 +58,12 @@ b1 <- b1 %>%
   add_case(tmp)
 
 b2 <- broom.mixed::tidy(lmer(interpret_nlp2 ~ masking + specified + 
-                               issue + (country | sentetence), data= d)) %>%
+                               issue + (country | sentence), data= d)) %>%
   mutate(y = "Y: Overinterpreting Stance (Strict Interpretation)",
          type = "Pooled Analysis")
 
 tmp <- broom.mixed::tidy(lmer(interpret_ss2 ~ masking + specified + 
-                                   issue + (country | sentetence), data= d)) %>%
+                                   issue + (country | sentence), data= d)) %>%
   mutate(y = "Y: Overinterpreting Stance (Lenient Interpretation)",
          type = "Pooled Analysis")
 

### 2025-09-09 src/analysis/exploration.R

@@ -1,9 +1,26 @@
-df_us <- d |> 
+# df_us <- d |>  
+#   filter(country == "US") |> 
+#   select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
+# df_nl <- d |> 
+#   filter(country == "NL") |> 
+#   select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
+
+df_us <- d |>  
   filter(country == "US") |> 
-  select(stance_nlp2, stance_ss2, interpret_nlp2, interpret_ss2, masking, specified, distance, PT7, issue)
-df_nl <- d |> 
+  mutate(issue = case_when(
+    str_detect(issue, "carbon") ~ "carbon",
+    str_detect(issue, "tax") ~ "tax",
+    str_detect(issue, "military") ~ "military",
+    str_detect(issue, "immigration") ~ "immigration"
+  ))
+df_nl <- d |>  
   filter(country == "NL") |> 
-  select(stance_nlp2, stance_ss2, interpret_nlp2, interpret_ss2, masking, specified, distance, PT7, issue)
+  mutate(issue = case_when(
+    str_detect(issue, "immigra") ~ "immigratie",
+    str_detect(issue, "stikstof") ~ "stikstof",
+    str_detect(issue, "belasting") ~ "belasting",
+    str_detect(issue, "Europese Unie") ~ "EU"
+  ))
 
 df <- df_us %>% 
   mutate(b = specified,

### 2025-09-09 src/analysis/h1.R

@@ -1,9 +1,26 @@
+# df_us <- d |>  
+#   filter(country == "US") |> 
+#   select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
+# df_nl <- d |> 
+#   filter(country == "NL") |> 
+#   select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
+
 df_us <- d |>  
   filter(country == "US") |> 
-  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
-df_nl <- d |> 
+  mutate(issue = case_when(
+    str_detect(issue, "carbon") ~ "carbon",
+    str_detect(issue, "tax") ~ "tax",
+    str_detect(issue, "military") ~ "military",
+    str_detect(issue, "immigration") ~ "immigration"
+  ))
+df_nl <- d |>  
   filter(country == "NL") |> 
-  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
+  mutate(issue = case_when(
+    str_detect(issue, "immigra") ~ "immigratie",
+    str_detect(issue, "stikstof") ~ "stikstof",
+    str_detect(issue, "belasting") ~ "belasting",
+    str_detect(issue, "Europese Unie") ~ "EU"
+  ))
 
 h1a <- broom.mixed::tidy(lmer(stance_nlp2 ~ masking + specified + 
                                 distance + PT7 + issue +

### 2025-09-09 src/analysis/h2.R

@@ -1,11 +1,26 @@
+# df_us <- d |>  
+#   filter(country == "US") |> 
+#   select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
+# df_nl <- d |> 
+#   filter(country == "NL") |> 
+#   select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
+
 df_us <- d |>  
-  filter(country == "US") |>
-  select(stance_nlp2, stance_ss2, interpret_nlp2, interpret_ss2,
-         masking, specified, distance, PT7, PT7b, issue)
+  filter(country == "US") |> 
+  mutate(issue = case_when(
+    str_detect(issue, "carbon") ~ "carbon",
+    str_detect(issue, "tax") ~ "tax",
+    str_detect(issue, "military") ~ "military",
+    str_detect(issue, "immigration") ~ "immigration"
+  ))
 df_nl <- d |>  
-  filter(country == "NL") |>
-  select(stance_nlp2, stance_ss2, interpret_nlp2, interpret_ss2,
-         masking, specified, distance, PT7, issue)
+  filter(country == "NL") |> 
+  mutate(issue = case_when(
+    str_detect(issue, "immigra") ~ "immigratie",
+    str_detect(issue, "stikstof") ~ "stikstof",
+    str_detect(issue, "belasting") ~ "belasting",
+    str_detect(issue, "Europese Unie") ~ "EU"
+  ))
 
 df <- df_us %>% 
   mutate(b = specified,

### 2025-09-09 src/analysis/h3.R

@@ -1,9 +1,26 @@
+# df_us <- d |>  
+#   filter(country == "US") |> 
+#   select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
+# df_nl <- d |> 
+#   filter(country == "NL") |> 
+#   select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
+
 df_us <- d |>  
-  filter(country == "US") |>
-  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, PT7b, issue)
+  filter(country == "US") |> 
+  mutate(issue = case_when(
+    str_detect(issue, "carbon") ~ "carbon",
+    str_detect(issue, "tax") ~ "tax",
+    str_detect(issue, "military") ~ "military",
+    str_detect(issue, "immigration") ~ "immigration"
+  ))
 df_nl <- d |>  
   filter(country == "NL") |> 
-  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
+  mutate(issue = case_when(
+    str_detect(issue, "immigra") ~ "immigratie",
+    str_detect(issue, "stikstof") ~ "stikstof",
+    str_detect(issue, "belasting") ~ "belasting",
+    str_detect(issue, "Europese Unie") ~ "EU"
+  ))
 
 df <- df_us %>% 
   mutate(b= factor(masking),

### 2025-09-12 src/data-processing/prepare-data.Rmd
update Compendium, final submission

@@
 
-```{r "get & clean data", results="hide", eval=FALSE}
+```{r "get & clean data", results="hide"}
 #Load Dutch Data
@@
 #Load US Data
-d_us <- fetch_survey(surveyID = "SV_86XBcsFUJka9eke", 
-                    verbose = TRUE, force_request = T,
-                    label = FALSE, convert = FALSE)
+d_us <- readRDS(here("data/raw-private/d_us.rds"))
+#d_us <- fetch_survey(surveyID = "SV_86XBcsFUJka9eke", 
+#                    verbose = TRUE, force_request = T,
+#                    label = FALSE, convert = FALSE)
 #Clean data

### 2025-09-12 src/analysis/baseline.R

@@ -9,12 +9,12 @@
 #  add_case(df_us)
 
 b1 <- broom.mixed::tidy(lmer(stance_nlp2 ~ masking + specified + issue +
-                         (country | sentetence), data= d)) %>%
+                         (country | sentence2), data= d)) %>%
   mutate(y = "Y: Correctly Interpreting Stance (Strict Interpretation)",
          type = "Pooled Analysis")
 
 tmp <- broom.mixed::tidy(lmer(stance_ss2 ~ masking + specified + 
-                               issue + (country | sentetence), data= d)) %>%
+                               issue + (country | sentence2), data= d)) %>%
   mutate(y = "Y: Correctly Interpreting Stance (Lenient Interpretation)",
          type = "Pooled Analysis")
 
@@ -58,12 +58,12 @@ b1 <- b1 %>%
   add_case(tmp)
 
 b2 <- broom.mixed::tidy(lmer(interpret_nlp2 ~ masking + specified + 
-                               issue + (country | sentetence), data= d)) %>%
+                               issue + (country | sentence2), data= d)) %>%
   mutate(y = "Y: Overinterpreting Stance (Strict Interpretation)",
          type = "Pooled Analysis")
 
 tmp <- broom.mixed::tidy(lmer(interpret_ss2 ~ masking + specified + 
-                                   issue + (country | sentetence), data= d)) %>%
+                                   issue + (country | sentence2), data= d)) %>%
   mutate(y = "Y: Overinterpreting Stance (Lenient Interpretation)",
          type = "Pooled Analysis")
 

### 2025-09-12 src/analysis/exploration.R

@@ -1,16 +1,16 @@
 df_us <- d |> 
   filter(country == "US") |> 
-  select(stance_nlp2, stance_ss2, interpret_nlp2, interpret_ss2, masking, specified, distance, PT7, issue)
+  select(stance_nlp2, stance_ss2, interpret_nlp2, interpret_ss2, masking, specified, distance, PT7, issue, sentence2)
 df_nl <- d |> 
   filter(country == "NL") |> 
-  select(stance_nlp2, stance_ss2, interpret_nlp2, interpret_ss2, masking, specified, distance, PT7, issue)
+  select(stance_nlp2, stance_ss2, interpret_nlp2, interpret_ss2, masking, specified, distance, PT7, issue, sentence2)
 
 df <- df_us %>% 
   mutate(b = specified,
          a = as.numeric(masking)) 
 
 b3_a <- lmer(stance_nlp2 ~ b * a +
-                 issue + (1 | sentence), data= df)
+                 issue + (1 | sentence2), data= df)
 b3_a2 <- tidy(b3_a) %>% 
   mutate(condition = "Condition: Underspecified Sentence",
          y = "Y: Correctly Identifying Stance (Strict Interpretation)",
@@ -21,7 +21,7 @@ b3_a <- tidy(margins::margins(b3_a, variables = "b",
          y = "Y: Correctly Identifying Stance (Strict Interpretation)",
          type = "United States")
 
-tmp <- lmer(stance_ss2 ~ b * a + issue + (1 | sentence), data= df)
+tmp <- lmer(stance_ss2 ~ b * a + issue + (1 | sentence2), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(condition = "Condition: Underspecified Sentence",
          y = "Y: Correctly Identifying Stance (Lenient Interpretation)",
@@ -40,7 +40,7 @@ b3_a <- b3_a %>%
 df <- df_us %>% 
   mutate(b = specified,
          a = as.numeric(masking))
-tmp <- lmer(stance_nlp2 ~ b * a + issue + (1 | sentence), data= df)
+tmp <- lmer(stance_nlp2 ~ b * a + issue + (1 | sentence2), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(condition = "Condition: Underspecified Sentence",
          y = "Y: Correctly Identifying Stance (Strict Interpretation)",
[…]

### 2025-09-12 src/analysis/h1.R

@@ -1,13 +1,13 @@
 df_us <- d |>  
   filter(country == "US") |> 
-  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
+  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue, sentence2)
 df_nl <- d |> 
   filter(country == "NL") |> 
-  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
+  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue, sentence2)
 
 h1a <- broom.mixed::tidy(lmer(stance_nlp2 ~ masking + specified + 
                                 distance + PT7 + issue +
-                                (1 | sentence), data= df_us)) %>% 
+                                (1 | sentence2), data= df_us)) %>% 
   mutate(hyp = "H1a",
          type = "United States",
          y = "Interpreting Stance Correctly (Strict Interpretation)") %>% 
@@ -15,7 +15,7 @@ h1a <- broom.mixed::tidy(lmer(stance_nlp2 ~ masking + specified +
 
 tmp <- broom.mixed::tidy(lmer(stance_ss2 ~ masking + specified + 
                                 distance + PT7 + issue +
-                                (1 | sentence), data= df_us)) %>% 
+                                (1 | sentence2), data= df_us)) %>% 
   mutate(hyp = "H1a",
          type = "United States",
          y = "Interpreting Stance Correctly (Lenient Interpretation)") %>% 
@@ -26,7 +26,7 @@ h1a <- h1a %>%
 
 tmp <- broom.mixed::tidy(lmer(stance_nlp2 ~ masking + specified + 
                                 distance + PT7 + issue +
-                                (1 | sentence), data= df_nl)) %>% 
+                                (1 | sentence2), data= df_nl)) %>% 
   mutate(hyp = "H1a",
          type = "The Netherlands",
          y = "Interpreting Stance Correctly (Strict Interpretation)") %>% 
@@ -37,7 +37,7 @@ h1a <- h1a %>%
 
 tmp <- broom.mixed::tidy(lmer(stance_ss2 ~ masking + specified + 
                                 distance + PT7 +
-                                issue +  (1 | sentence), data= df_nl)) %>% 
+                                issue +  (1 | sentence2), data= df_nl)) %>% 
   mutate(hyp = "H1a",
[…]

### 2025-09-12 src/analysis/h2.R

@@ -1,17 +1,17 @@
 df_us <- d |>  
   filter(country == "US") |>
   select(stance_nlp2, stance_ss2, interpret_nlp2, interpret_ss2,
-         masking, specified, distance, PT7, PT7b, issue)
+         masking, specified, distance, PT7, PT7b, issue, sentence2)
 df_nl <- d |>  
   filter(country == "NL") |>
   select(stance_nlp2, stance_ss2, interpret_nlp2, interpret_ss2,
-         masking, specified, distance, PT7, issue)
+         masking, specified, distance, PT7, issue, sentence2)
 
 df <- df_us %>% 
   mutate(b = specified,
          a = distance)
 h2a <- lmer(interpret_nlp2 ~ masking + b * a +
-              PT7 + issue + (1 | sentence), data= df)
+              PT7 + issue + (1 | sentence2), data= df)
 h2a_2 <- tidy(h2a) %>% 
   mutate(hyp = "H2a: Ideological Distance",
          type = "United States",
@@ -23,7 +23,7 @@ h2a <- tidy(margins::margins(h2a, variables = "b",
          y = "Overinterpreting Stance (Strict Interpretation)")
 
 tmp <- lmer(interpret_ss2 ~ masking + b * a +
-              PT7 + issue + (1 | sentence), data= df)
+              PT7 + issue + (1 | sentence2), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(hyp = "H2a: Ideological Distance",
          type = "United States",
@@ -44,7 +44,7 @@ df <- df_nl %>%
          a = distance)
 
 tmp <- lmer(interpret_nlp2 ~ masking + b * a +
-              PT7 + issue + (1 | sentence), data= df)
+              PT7 + issue + (1 | sentence2), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(hyp = "H2a: Ideological Distance",
          type = "The Netherlands",
@@ -61,7 +61,7 @@ h2a <- h2a %>%
   add_case(tmp)
 
 tmp <- lmer(interpret_ss2 ~ masking + b * a +
-              PT7 + issue + (1 | sentence), data= df)
+              PT7 + issue + (1 | sentence2), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(hyp = "H2a: Ideological Distance",
          type = "The Netherlands",
@@ -81,7 +81,7 @@ df <- df_us %>%
   mutate(b = specified,
          a = PT7)
[…]

### 2025-09-12 src/analysis/h3.R

@@ -1,15 +1,15 @@
 df_us <- d |>  
   filter(country == "US") |>
-  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, PT7b, issue)
+  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, PT7b, issue, sentence2)
 df_nl <- d |>  
   filter(country == "NL") |> 
-  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue)
+  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue, sentence2)
 
 df <- df_us %>% 
   mutate(b= factor(masking),
          a = distance)
 h3a <- lmer(stance_nlp2 ~ specified + b * a +
-              PT7 + issue + (1 | sentence), data= df)
+              PT7 + issue + (1 | sentence2), data= df)
 h3a_2 <- tidy(h3a) %>% 
   mutate(hyp = "H3a: Ideological Distance",
          type = "United States",
@@ -21,7 +21,7 @@ h3a <- broom.mixed::tidy(margins::margins(h3a, variables = "b",
          y = "Correctly Interpreting Stance (Strict Interpretation)")
 
 tmp <- lmer(stance_ss2 ~ specified + b * a +
-              PT7 + issue + (1 | sentence), data= df)
+              PT7 + issue + (1 | sentence2), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(hyp = "H3a: Ideological Distance",
          type = "United States",
@@ -42,7 +42,7 @@ df <- df_nl %>%
          a = distance)
 
 tmp <- lmer(stance_nlp2 ~ specified + b * a +
-              PT7 + issue + (1 | sentence), data= df)
+              PT7 + issue + (1 | sentence2), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(hyp = "H3a: Ideological Distance",
          type = "The Netherlands",
@@ -60,7 +60,7 @@ h3a <- h3a %>%
 
 
 tmp <- lmer(stance_ss2 ~ specified + b * a +
-              PT7 + issue + (1 | sentence), data= df)
+              PT7 + issue + (1 | sentence2), data= df)
 tmpp <- tidy(tmp) %>% 
   mutate(hyp = "H3a: Ideological Distance",
          type = "The Netherlands",
@@ -81,7 +81,7 @@ df <- df_us %>%
   mutate(b= factor(masking),
[…]

### 2025-09-12 src/data-processing/clean_data.R

@@ -324,7 +324,8 @@ d <- bind_rows(d_nl, d_us) |>
     stance_nlp2 = ifelse(stance_nlp == "correct", 1, 0),
     stance_ss2 = ifelse(stance_ss == "correct", 1, 0),
     interpret_nlp2  = ifelse(interpret_nlp == "Correct", 0, 1),
-    interpret_ss2  = ifelse(interpret_ss == "Correct", 0, 1)
+    interpret_ss2  = ifelse(interpret_ss == "Correct", 0, 1),
+    sentence2 = id
   ) |> 
   select(-PT7_1, -PT1b)
 

### 2025-09-12 src/data-processing/clean_data.R
update Compendium, final submission

@@ -327,5 +327,20 @@ d <- bind_rows(d_nl, d_us) |>
     interpret_ss2  = ifelse(interpret_ss == "Correct", 0, 1),
     sentence2 = id
   ) |> 
-  select(-PT7_1, -PT1b)
+  select(-PT7_1, -PT1b) |> 
+  mutate(issue = case_when(
+    issue == "a wealth tax for the richest Americans" ~ "Tax Policy",
+    issue == "de  rol van Nederland in de Europese Unie" ~ "Foreign Policy",
+    issue == "het belastingstelsel" ~ "Tax Policy",
+    issue == "het lidmaatschap van de Europese Unie voor Nederland" ~ "Foreign Policy",
+    issue == "het stikstofbeleid" ~ "Climate Policy",
+    issue == "het tegengaan van stikstofuitstoot" ~ "Climate Policy",
+    issue == "het verhogen van het belastingstarief voor de hoogste inkomens" ~ "Tax Policy",
+    issue == "immigranten" ~ "Immigration Policy",
+    issue == "immigratie" ~ "Immigration Policy",
+    issue == "military build-up in the Pacific Ocean" ~ "Foreign Policy",
+    issue == "military presence in the Pacific Ocean" ~ "Foreign Policy",
+    issue == "tax on carbon emissions" ~ "Climate Policy",
+    issue == "the tax system" ~ "Tax Policy"
+  ))
 

### 2025-09-12 src/analysis/h1.R
Checked
and found issue with the issue variable

@@ -1,9 +1,21 @@
 df_us <- d |>  
   filter(country == "US") |> 
-  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue, sentence2)
+  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue, sentence2) |> 
+  mutate(issue = case_when(
+    str_detect(issue, "carbon") ~ "carbon",
+    str_detect(issue, "tax") ~ "tax",
+    str_detect(issue, "military") ~ "military",
+    str_detect(issue, "immigration") ~ "immigration"
+  ))
 df_nl <- d |> 
   filter(country == "NL") |> 
-  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue, sentence2)
+  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue, sentence2) |> 
+  mutate(issue = case_when(
+    str_detect(issue, "immigra") ~ "immigratie",
+    str_detect(issue, "stikstof") ~ "stikstof",
+    str_detect(issue, "belasting") ~ "belasting",
+    str_detect(issue, "Europese Unie") ~ "EU"
+  ))
 
 h1a <- broom.mixed::tidy(lmer(stance_nlp2 ~ masking + specified + 
                                 distance + PT7 + issue +

### 2025-09-12 src/analysis/h2.R

@@ -1,11 +1,23 @@
 df_us <- d |>  
   filter(country == "US") |>
   select(stance_nlp2, stance_ss2, interpret_nlp2, interpret_ss2,
-         masking, specified, distance, PT7, PT7b, issue, sentence2)
+         masking, specified, distance, PT7, PT7b, issue, sentence2) |> 
+  mutate(issue = case_when(
+    str_detect(issue, "carbon") ~ "carbon",
+    str_detect(issue, "tax") ~ "tax",
+    str_detect(issue, "military") ~ "military",
+    str_detect(issue, "immigration") ~ "immigration"
+  ))
 df_nl <- d |>  
   filter(country == "NL") |>
   select(stance_nlp2, stance_ss2, interpret_nlp2, interpret_ss2,
-         masking, specified, distance, PT7, issue, sentence2)
+         masking, specified, distance, PT7, issue, sentence2) |> 
+  mutate(issue = case_when(
+    str_detect(issue, "immigra") ~ "immigratie",
+    str_detect(issue, "stikstof") ~ "stikstof",
+    str_detect(issue, "belasting") ~ "belasting",
+    str_detect(issue, "Europese Unie") ~ "EU"
+  ))
 
 df <- df_us %>% 
   mutate(b = specified,

### 2025-09-12 src/analysis/h3.R

@@ -1,9 +1,21 @@
 df_us <- d |>  
   filter(country == "US") |>
-  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, PT7b, issue, sentence2)
+  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, PT7b, issue, sentence2) |> 
+  mutate(issue = case_when(
+    str_detect(issue, "carbon") ~ "carbon",
+    str_detect(issue, "tax") ~ "tax",
+    str_detect(issue, "military") ~ "military",
+    str_detect(issue, "immigration") ~ "immigration"
+  ))
 df_nl <- d |>  
   filter(country == "NL") |> 
-  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue, sentence2)
+  select(stance_nlp2, stance_ss2, masking, specified, distance, PT7, issue, sentence2) |> 
+  mutate(issue = case_when(
+    str_detect(issue, "immigra") ~ "immigratie",
+    str_detect(issue, "stikstof") ~ "stikstof",
+    str_detect(issue, "belasting") ~ "belasting",
+    str_detect(issue, "Europese Unie") ~ "EU"
+  ))
 
 df <- df_us %>% 
   mutate(b= factor(masking),

### 2025-09-12 src/lib/functions.R
Checked
and everything seems fine now

@@ -14,5 +14,5 @@ fig_cols <- yarrr::piratepal(palette = "basel",
                              trans = .2)
 fig_cols <- as.character(fig_cols[1:9])
 
-api_key_fn <- here("data/raw-private/qualtrics_api_key.txt")
-API <- read_file(api_key_fn) %>% trimws()
+# api_key_fn <- here("data/raw-private/qualtrics_api_key.txt")
+# API <- read_file(api_key_fn) %>% trimws()
